### 2022-08-30 intro.tex
initial draft of chapters

\documentclass[main.tex]{subfiles}
\begin{document}
\chapter{Intro}
\label{chapter:intro}
\section{Introduction to standard model of particle physics}
\section{LHC and comparison to experiments}
\section{Introduction to QCD}
\subsection{Lagrangian}
\subsection{Running coupling}
\section{Observables}
\section{Amplitudes/Matrix elements}
\section{Monte Carlo integration????}
\end{document}

### 2022-08-30 main.tex
initial draft of chapters

\documentclass[twoside,openright,frontopenright]{ip3thesis}
\input{preamble}
\usepackage{subfiles}

\begin{document}
\title{Emulation of high multiplicity matrix elements for electron-positron and hadron-hadron colliders}
% \subtitle{An optional subtitle}
\author{Henry Truong}
\researchgroup{Institute for Particle Physics Phenomenology}
\maketitlepage*

\begin{abstract}
%
	This is some abstract about this thesis.
%
\end{abstract}

\disableprotrusion
\tableofcontents*
\listoffigures
\listoftables
\enableprotrusion

\begin{declaration*}
%
	The work in this thesis is based on research carried out in the Department of
	Physics at Durham University. No part of this thesis has been
	submitted elsewhere for any degree or qualification.
%
\end{declaration*}

\begin{acknowledgements*}
%
	I would like to thank\ldots
%
\end{acknowledgements*}

\begin{epigraph*}
%
	This a place for some profound quote which sums up your work, feelings or
	life.
	\source{Some book}{A. Author}
%
\end{epigraph*}

%\dedicationtext{This thesis is dedicated to}
\begin{dedication*}
%
	Someone special
	\also[and]{that other guy}
%
\end{dedication*}

\cleardoublepage

\subfile{intro}
\subfile{qcd}
\subfile{ml}
\subfile{paper1}
\subfile{paper2}
\subfile{paper3}

\appendix
%\subfile{appendix1}
%\subfile{appendix2}

\bibliographystyle{JHEP}
\bibliography{bibliography}

\end{document}

### 2022-08-30 ml.tex
initial draft of chapters

\documentclass[main.tex]{subfiles}
\begin{document}
\chapter{Machine learning}
\label{chapter:ml}
\section{Introduction to machine learning}
\subsection{Supervised and unsupervised learning}
\section{Neural networks}
\subsection{Perceptron}
\subsection{Densely connected neural networks}
\subsection{Optimisation of weights}
\subsection{ML applications in HEP}
\end{document}

### 2022-08-30 qcd.tex
initial draft of chapters

\documentclass[main.tex]{subfiles}
\begin{document}
\chapter{Quantum chromodynamics}
\label{chapter:qcd}
\section{Fixed order perturbation theory}
\subsection{NLO QCD}
\section{Divergent structures}
\section{Infrared factorisation}
\subsection{Collinear limits}
\subsection{Soft limits}
\section{Subtraction}
\subsection{Catani-Seymour dipoles}
\subsection{Antenna subtraction}
\end{document}

### 2022-09-07 intro.tex
added intro to SM

@@
 \begin{document}
-\chapter{Intro}
+\chapter{Introduction}
+The standard model of particle physics (SM) is our current best working
+theory that describes the phenomenon we observe, for  example,
+ at collider experiments.
+The large hadron collider (LHC) is a high energy proton-proton
+collider experiment that allows us to study the fundamental particles
+laid out in the standard model.
+
+The interaction of these fundamental particles are described
+by scattering amplitudes. The standard model allows us to
+calculate these amplitudes, or matrix elements, from first
+principles. A key challenge in the past few decades has been
+to compute these matrix elements for increasing number of
+final state particles and for higher orders in perturbative
+expansations in fixed order calculations. Although there
+has been tremendous progress in this sector a complication
+that naturally arises is that the evaluation of these matrix
+elements becomes increasingly time consuming due to their
+algebraic complexity.
+
+The topic of this thesis is to apply modern machine learning
+techniques to reduce the time spent evaluating these
+complex matrix elements. Considering the physical properties
+of these matrix elements allows us to implement physics
+knowledge into these machine learning algorithms giving
+us higher levels of performance than would otherwise be
+achievable.
+
+The structure of this thesis is as follows: in Chapter~\ref{chapter:intro}
+I will recap the necessary concepts from the standard
+model, in particular, on quantum chromodynamics (QCD), the sector
+which governs the behaviour of quarks and gluons. Furthermore,
+the relationship between theory and experiment will be elaborated on.
+
+In Chapter~\ref{chapter:qcd} I will discuss the challenges that arise
+in fixed order perturbative calculations and the current methods
+that have been adopted in the community to deal with them. These
+methods also form the basis of our physics knowledge which we
+embed into the machine learning algorithm. This procedure is 
+described in detail in Chapter~\ref{chapter:fame1} for electron-positron
+annihilation for tree-level matrix element emulation.
+A similar philosophy is applied in Chapter~\ref{chapter:fame2}
+where next-to-leading order QCD k-factors are emulated.
+These two chapters discuss in detail the procedure for
+electron-positron annihilation, however, the most relevant
+experiment currently is the LHC which is colliding protons.
+The extension to hadron-hadron colliders is detailed in
+Chapter~\ref{chapter:fame3} where we explore using the
+emulator in a novel implementation to replace traditional
+matrix element providers in the event generator SHERPA.
+
+Finally, I will conclude the thesis in Chapter~\ref{chapter:conclusion}.
+
 \label{chapter:intro}
-\section{Introduction to standard model of particle physics}
-\section{LHC and comparison to experiments}
-\section{Introduction to QCD}
-\subsection{Lagrangian}
+\section{The Standard Model of particle physics}
+    The Standard Model of particle physics (SM) is our current
+    best working theory to describe all known elementary particles
+    as well as three of the four fundamental forces. Developed
+    predominantly in the latter half of the 20th century, it is
+    one of the most well tested theories we have in science today.
+    Some highlights include the highly precise predictions of the
+    anomalous magnetic moment of the electron, agreeing with experimental
+    measurements to more than 10 significant figures \cite{Aoyama:2017uqe}.
+    and the discovery of the Higgs boson in 2012 by the ATLAS \cite{ATLAS:2012yve}
+    and CMS experiments \cite{CMS:2012qbp} at the Large Hadron Collider
+    (LHC) which was theorised decades prior.
+
+    The SM is a gauge quantum field theory (QFT) where particles
+    are described as excitations in quantum fields. It can be specified
+    by the gauge group
+    \begin{equation}\label{eqn:SM_gauge}
+        \mathrm{SU}(3)_{C} \times \mathrm{SU}(2)_{L} \times \mathrm{U}(1)_{Y}\, ,
+    \end{equation}
+    where subscripts denote the charges of the gauge groups. The first gauge
+    group with colour charge $C$ describes the interactions of the
+    strong force within the theory of quantum chronodynamics (QCD), which we
+    will elaborate more on in the next section. The second
+    and third gauge groups represent the electroweak sector of the SM with
+    charges $L$ for left and $Y$ for hypercharge. Under electroweak
+    spontaneous symmetry breaking (EWSB), this product becomes
+    \begin{equation}\label{eqn:SM_SSB}
+        \mathrm{SU}(2)_{L} \times \mathrm{U}(1)_{Y} \xrightarrow{\mathrm{EWSB}} \mathrm{U}(1)_{\mathrm{EM}} \, ,
+    \end{equation}
+    giving rise to the electromagnetic and weak forces we observe.
+    The effect of gravity is considered to be negligible on the scales
+    considered in the SM and so it is not described.
+    Each of the three fundamental forces described by the SM is mediated
+    by the exchange of a gauge boson. The gauge bosons mediating the strong
+    and EM force are the massless gluon $g$ and massless photon $\gamma$,
+    respectively. For the weak force the gauge bosons are the $W^{\pm}$
+    and $Z^{0}$ bosons which attain a mass through the Higgs mechanism \cite{Englert:1964et,Higgs:1964pj,Guralnik:1964eu}
+    during EWSB, which elucidates the Higgs boson $H$.
+
+    The matter content of the SM consists of fermions which can be split
+    into quarks and leptons. Quarks are massive and experience the strong,
+    weak, and EM force. Leptons are defined by their lack of colour charge,
+    meaning they do not experience the strong force. Leptons can be separated
+    into charged leptons (electron $e$, muon $\mu$, tau $\tau$
+    and their antiparticles) which experience the weak and EM force, and 
+    neutrinos (electron neutrino $\nu_{e}$, muon neutrino $\nu_{\mu}$, tau
+    neutrino $\nu_{\tau}$ and their antiparticles) which only experience the
+    weak force. Neutrinos are massless in the SM but this is not a requirement
+    of the model. In fact they have been observed to have mass (cittation here).
+
+    The particle content of the SM is summarised in Figure \ref{fig:SM_particles}
+    showing the quarks, leptons and bosons along with their masses, charges
+    and spin.
+    \begin{figure}
+        \includegraphics[width=\linewidth]{introduction/sm_particles.pdf}
+        \caption{Particle content of the SM split into quarks, leptons,
+        gauge bosons and scalar bosons. The columns for the fermions depict
+        the three different generations. Masses, electric charge, and spin
+        are given for each particle. The yellow contours indicate the coupling
+        of bosons to fermions, illustrating the forces experienced by the
+        fermions inside the contour. Figure reference \cite{SM_figure}.}
+        \label{fig:SM_particles}
+    \end{figure}
+    The interactions of these fields are governed by the SM Lagrangian
+    \footnote{technically Lagrangian density but we use the terms
+    Lagrangian and Lagrangian density interchangeably}
+    which can be written as
+    \begin{equation}\label{eqn:L_SM}
+        \mathcal{L}_{\mathrm{SM}} = \mathcal{L}_{\mathrm{gauge}} + \mathcal{L}_{\mathrm{fermion}} + \mathcal{L}_{\mathrm{Higgs}} + \mathcal{L}_{\mathrm{Yukawa}} + \mathcal{L}_{\mathrm{GF}} + \mathcal{L}_{\mathrm{ghost}} \, ,
+    \end{equation}
+    where $\mathcal{L}_{\mathrm{gauge}}$ describes the gauge fields,
+    $\mathcal{L}_{\mathrm{fermion}}$ describes how fermions interact with
+    gauge fields as well as their kinetic terms,
+    $\mathcal{L}_{\mathrm{Higgs}}$ describes the Higgs field,
+    $\mathcal{L}_{\mathrm{Yukawa}}$ describes the interaction between the Higgs
+    field and fermions,
+    $\mathcal{L}_{\mathrm{GF}}$ is a gauge fixing term,
+    and $\mathcal{L}_{\mathrm{ghost}}$ is a ghost term.
+    The last two terms are required to remove unphysical degrees of freedom
+    when gauge fixing the theory.
+    All terms in the SM Lagrangian are invariant under local transformations
+    of the gauge group (\ref{eqn:SM_gauge}).
+
+    In the following section we will focus on the gauge, fermion, gauge-fixing
+    and ghost Lagrangian terms in the framework of QCD as that will be the
+    most relevant sector for this thesis. The remaining terms are discussed at
+    length in standard reference texts \cite{Peskin:1995ev,Schwartz:2014sze,Romao:2012pq}.
+
+\section{Introduction to quantum chromodynamics}
+\begin{itemize}
+    \item QCD is the quantum field theory describing the strong
+    interaction between quarks and gluons. It is a non-abelian
+    gauge theory with gauge group SU(3).
+    \item Write down the Lagrangian of QCD and define all the
+    terms (field strength tensor, quark fields, group generators,
+    structure constants, coupling constants.)
+    \item Begin to talk about how Feynman rules can be used
+    to represent the interactions that you can derive from
+    the Lagrangian.
+\end{itemize}
+    Quantum chronodynamics (QCD) is the sector of the SM that
+    describes the strong interaction. 
+    QCD is a non-abelian gauge theory with gauge group
+    $\mathrm{SU}(3)_{C}$ where the charge is named colour.
+    The gauge and fermion part of the QCD Lagrangian is
+    \begin{equation}\label{eqn:L_QCD}
+        \mathcal{L}_{\mathrm{QCD}} = -\dfrac{1}{4}F^{a}_{\mu\nu}F^{a, \, \mu\nu}
+        + \sum_{f} \bar{\psi}_{i}^{f}(\mathrm{i}\slashed{D}_{ij} - \delta_{ij}m_{f})\psi_{j}^{f} \, ,
+    \end{equation}
+    where twice repeated indices are summed over.
+    The fields $\psi_{i}^{(f)}$ ($\bar{\psi}_{i}^{f}$) are the fermions (antifermions),
+    representing quarks (antiquarks) with flavours $f$: 
+    up $u$, down $d$, strange $s$, charm $c$, top $t$, and bottom $b$,
+    with masses $m_{f}$. They transform under the fundamental
+    representation with indices $i, j \in \{1, 2, 3\}$, named colour indices.
+    The gauge fields $A^{a}_{\mu}$, corresponding to gluons, 
+    appear in the quark covariant derivative
+    \begin{equation}\label{eqn:covariant_deriv}
+        (D_{\mu})_{ij} = \delta_{ij}\partial_{\mu} - ig_{\mathrm{s}}T^{a}_{ij}A^{a}_{\mu} \, ,
+    \end{equation}
+    and the gauge field strength tensor
+    \begin{equation}\label{eqn:F_munu}
+        F^{a}_{\mu\nu} = \partial_{\mu}A^{a}_{\nu} - \partial_{\nu}A^{a}_{\mu} + g_{\mathrm{s}} f^{abc}A^{b}_{\mu}A^{c}_{\nu} \, .
+    \end{equation}
+    Gauge fields transform under the adjoint representation, meaning
+    the adjoint indices $a,b,c \in \{1,...,8\}$. $T^{a}_{ij}$ are the
+    group generators in a fundamental representation. In SU(3) it is
+    common to write the group generators as
+    \begin{equation}\label{eqn:group_generators}
+        T^{a}_{ij} = \dfrac{1}{2}\lambda^{a}_{ij}
+    \end{equation}
+    where $\lambda^{a}_{ij}$ are the Gell-Mann matrices \cite{Gell-Mann:1962yej}.
+    The generators of the group have to obey the Lie algebra
+    \begin{equation}\label{eqn:lie_algebra}
+        [T^{a}, T^{b}] = \mathrm{i}f^{abc}T^{c} \, ,
+    \end{equation}
+    where $f^{abc}$ are the structure constants of SU(3).
+    The gauge coupling of the group $g_{\mathrm{s}}$
+    is a dimensional free parameter of the theory, usually
+    written as
+    \begin{equation}\label{eqn:alpha_s}
+        \alpha_{\mathrm{s}} = \dfrac{g_{\mathrm{s}}^{2}}{4\pi} \, ,
+    \end{equation}
+    for reasons which will become clear when we introduce
+    scattering amplitudes.
+
 \subsection{Running coupling}
-\section{Observables}
-\section{Amplitudes/Matrix elements}
-\section{Monte Carlo integration????}
+\begin{itemize}
+    \item Strength of the coupling depends on the energy
+    that the process occurs at. This is known as the running
+    of the coupling.
+    \item The value of $\alpha_{s}$ is usually quoted
+    at the mass of the Z boson.
+    \item In QCD the evolution of $\alpha_{s}$ is governed
+    by the renormalisation group equation (Callan-Symanzik
+    $\beta$ function). The coefficients of this function
+    are known to many loops.
+    \item Asymptotic freedom is a property of non-abelian
+    gauge groups and this discovery won a Nobel prize.
+\end{itemize}
+\subsection{Scattering amplitudes}
+\begin{itemize}
+    \item S-matrix/Transition matrix
+    \item Make the link between Feynman diagrams and
+    scattering amplitudes/matrix elements. The sum
+    of all Feynman diagrams you can draw for a particular
+    process is the matrix element.
+    \item Explicitly state how the QCD Feynman rules are
+    derived and have the corresponding diagrams.
+    \item Explain how you can basically join these
+    Feynman rules into a diagram which represents a
+    convenient way to calculate interactions between
+    particles.
+    \item Introduce the concept of loop amplitudes here,
+    which naturally leads onto divergences.
+    \item UV divergences lead to the introduction of
+    dimensional regularisation and also to renormalisation
+    of fields to make them represent physical quantities.
+    The main scheme that people use is the modified minimal
+    subtraction scheme.
+\end{itemize}
+\section{Experiments and Phenomonology}
+\subsection{Observables}
+\begin{itemize}
+    \item Experiments do not measure single particles,
+    they measure hadrons which is far from the collision
+    of protons.
+    \item Factorisation of hard and soft scales leads
+    to the concepts of PDFs (DGLAP equation), partonic
+    cross-sections (Feynman diagrams), parton showers,
+    and hadronisation.
+    \item Link back to matrix elements and how they
+    are integrated to give you cross-sections which
+    can be compared with experiments.
+    \item Probably need to talk about infrared safe
+    observables and jets.
+\end{itemize}
+\section{Monte Carlo event generator overview}
+\subsection{Event unweighting}
+\begin{itemize}
+    \item MC generators are the main tools that
+    people use to compare with experimental data.
+    There are many groups that have their own tool.
+    \item The usage of MC tools came about because
+    the multi-dimensional phase-space integrals
+    are too complex to do analytically.
+    \item This doesn't mean that they only do
+    the integrals, the event generators basically
+    contain all the tools from phase-space generation
+    to matrix element evaluation, to parton showers
+    to hadronisation.
+    \item Have a look at Katherina thesis to see
+    how much I can skip because only really want to
+    introduce unweighting. Probably along the lines
+    of event weight is statistically sampled vs
+    what they measure in real life. Therefore we need
+    to unweight events.
+\end{itemize}
 \end{document}

### 2022-09-09 intro.tex
qcd, scattering amplitudes, cross-sections

@@
     where twice repeated indices are summed over.
-    The fields $\psi_{i}^{(f)}$ ($\bar{\psi}_{i}^{f}$) are the fermions (antifermions),
+    The fields $\psi_{i}^{f}$ ($\bar{\psi}_{i}^{f}$) are the fermions (antifermions),
     representing quarks (antiquarks) with flavours $f$: 
@@
     \end{equation}
-    for reasons which will become clear when we introduce
+    for reasons which will become clear when we discuss
     scattering amplitudes.
 
-\subsection{Running coupling}
-\begin{itemize}
-    \item Strength of the coupling depends on the energy
-    that the process occurs at. This is known as the running
-    of the coupling.
-    \item The value of $\alpha_{s}$ is usually quoted
-    at the mass of the Z boson.
-    \item In QCD the evolution of $\alpha_{s}$ is governed
-    by the renormalisation group equation (Callan-Symanzik
-    $\beta$ function). The coefficients of this function
-    are known to many loops.
-    \item Asymptotic freedom is a property of non-abelian
-    gauge groups and this discovery won a Nobel prize.
-\end{itemize}
-\subsection{Scattering amplitudes}
-\begin{itemize}
-    \item S-matrix/Transition matrix
-    \item Make the link between Feynman diagrams and
-    scattering amplitudes/matrix elements. The sum
-    of all Feynman diagrams you can draw for a particular
-    process is the matrix element.
-    \item Explicitly state how the QCD Feynman rules are
-    derived and have the corresponding diagrams.
-    \item Explain how you can basically join these
-    Feynman rules into a diagram which represents a
-    convenient way to calculate interactions between
-    particles.
-    \item Introduce the concept of loop amplitudes here,
-    which naturally leads onto divergences.
-    \item UV divergences lead to the introduction of
-    dimensional regularisation and also to renormalisation
-    of fields to make them represent physical quantities.
-    The main scheme that people use is the modified minimal
-    subtraction scheme.
-\end{itemize}
+    To remove unphysical degrees of freedom from the
+    theory we need to gauge-fix the theory
+    and add a ghost Lagrangian. The gauge-fixing term
+    in the $R_{\xi}$ gauges is written as
+    \begin{equation}\label{eqn:L_GF}
+        \mathcal{L}_{\mathrm{GF}} = -\dfrac{1}{2\xi}(\partial^{\mu}A_{\mu}^{a})^{2} \, ,
+    \end{equation}
+    where $\xi = 1$ corresponds to the Feynman-'t-Hooft gauge.
+    The ghost Lagrangian is most commonly written
+    in the Faddeev-Popov procedure as
+    \begin{equation}\label{eqn:L_ghost}
+        \mathcal{L}_{\mathrm{ghost}} = (\partial_{\mu}\bar{c}^{a})(\delta^{ac}\partial_{\mu} + g_{\mathrm{s}}f^{abc}A^{b}_{\mu})c^{c} \, ,
+    \end{equation}
+    where $c^{a}$ ($\bar{c}^{a}$) are Faddeev-Popov ghosts (anti-ghosts).
+    They are unphysical external states but must be
+    included in internal lines for a Lorentz invariant
+    perturbative gauge theory.
+
+    Now that we have written down the QCD Lagrangian,
+    we can examine the terms to write down how the
+    gauge bosons interact with the fermions. A convenient
+    way to do this is with Feynman rules, which can be
+    pieced together to form Feynman diagrams. We will see
+    that Feynman diagrams are a simple, but powerful, tool
+    to systematically build all the possible ways a process
+    can occur. This will be discussed further in the
+    scattering amplitude section.
+
+    Expanding (\ref{eqn:L_QCD}) and extracting the terms
+    that mix the gauge and fermions fields, we get an interaction Lagrangian
+    \begin{equation}\label{L_int}
+        \mathcal{L}_{\mathrm{int}} = g_{\mathrm{s}}A^{a}_{\mu}\bar{\psi}_{i}^{f}\gamma^{\mu}T^{a}_{ij}\psi_{j}^{f}-g_{\mathrm{s}}f^{abc}(\partial_{\mu}A^{a}_{\nu})A^{b,\,\mu}A^{c,\,\nu} -\dfrac{g_{\mathrm{s}}^{2}}{4}f^{abc}A^{b}_{\mu}A^{c}_{\nu}f^{ade}A^{d,\,\mu}A^{e,\,\nu} \, ,
+    \end{equation}
+    which we can interpret as follows: the first term
+    is an interaction between a gluon, a quark, and an
+    anti-quark, the second term is a triple gluon
+    self-interaction and the third term is a four gluon
+    self-interaction. These mixing terms can be recast
+    as Feynman rules as interaction vertices.
+
+    To connect vertices we require propagators, which
+    we can read off from the kinetic Lagrangian where
+    we collect terms from (\ref{eqn:L_QCD}) and (\ref{eqn:L_GF}), 
+    \begin{equation}\label{L_kin}
+        \mathcal{L}_{\mathrm{kin}} = -\dfrac{1}{4}(\partial_{\mu}A^{a}_{\nu}-\partial_{\nu}A^{a}_{\mu})^{2} - \dfrac{1}{2\xi}(\partial_{\mu}A^{a}_{\mu})^{2} + \bar{\psi}_{i}^{f}(\mathrm{i}\slashed{\partial}-m_{f})\psi_{i}^{f} \, ,
+    \end{equation}
+    where the first two terms corresponds to a gluon
+    propagator and the third term corresponds to
+    a quark propagator.
+
+    Note that we have only written down the vertices
+    and propagators for the physical gluons and quarks,
+    ghosts also have associated Feynman rules but we do
+    not collect those here. A full list of Feynman rules
+    in the SM can be found at Ref. \cite{Romao:2012pq}.
+
+    Now that we have the Feynman rules for the theory
+    we can move on to discuss scattering amplitudes
+    and cross-sections.
+
+\subsection{Scattering amplitudes and cross-sections}
+    In collider experiments we typically collide
+    two beams, consisting of bundles of energetic
+    particles, and analyse the resulting products
+    of these collisions. To model collisions in
+    a collider experiment, we look at the specific
+    case of two particles colliding into $n$ particles.
+    To mathematically describe these
+    scattering processes, scattering amplitudes are used.
+    For an initial-state
+    $|i\rangle$ and final-state $|f\rangle$
+    the scattering amplitude can be written as
+    \begin{equation}\label{eqn:scattering}
+        \langle f | S | i \rangle \, ,
+    \end{equation}
+    where the scattering matrix, or $S$-matrix
+    can be decomposed into an identity matrix
+    and a transfer matrix $\mathcal{T}$,
+    \begin{equation}\label{eqn:S_matrix}
+        S = \mathbbm{1} + \mathrm{i}\mathcal{T} \, .
+    \end{equation}
+    The $S$-matrix encodes all the information
+    about how the initial-state will evolve
+    into the final-state over time. By writing
+    the $S$-matrix in this way, all the interactions
+    are separated into the transfer matrix $\mathcal{T}$,
+    as the identity matrix describes the free theory.
+    By imposing a momentum conserving constraint
+    on the $S$-matrix, the matrix element,
+    $\mathcal{M}$, is defined in the following expression
+    \begin{equation}\label{eqn:matrix_element}
+        \langle f | \mathcal{T} | i \rangle = (2\pi)^{4} \delta^{4}\left(p_{a} + p_{b} - \sum_{f=1}^{n} p_{f}^{\mu}\right) \mathcal{M} \, .
+    \end{equation}
+    where we take $p_{a}$ and $p_{b}$ to be
+    the momenta of the two colliding
+    particles in the initial-state, and $p_{f}$ to be the
+    momenta of the $n$-particle final states.
+    The transition probability of this scattering process
+    is the modulus squared of the scattering amplitude
+    \begin{equation}\label{eqn:S_prob}
+        P = |\langle f | S | i \rangle|^{2} \propto |\langle f | \mathcal{M} | i \rangle 
+    \end{equation}
+    where $|\langle f | \mathcal{M} | i \rangle|^{2}  \equiv |\mathcal{M}|^{2}$
+    is the matrix element squared.
+
+    Matrix elements can be calculated from
+    first principles in QFT using Feynman diagrams.
+
+    Since detectors at collider experiments
+    have limited resolution, single scattering
+    events are not measured. Instead we consider
+    the cross-section, $\sigma$, which is a property
+    of the particles being scattered, and is independant
+    of the way the experiment is carried out.
+    The cross-section directly relates the
+    matrix elements which we calculate in our QFT,
+    to measurable observables at experiments.
+    In practice, it is more useful to consider the
+    differential cross-section, $\mathrm{d}\sigma$
+    where the cross-section can be differential
+    in quantities such as energies and angles.
+    This is because the final-state particles
+    will be produced with a range of, for example,
+    energies and momenta so the cross-section would
+    be vanishing for an exact momentum configuration.
+
+    The differential cross-section can be written as
+    \begin{equation}\label{eqn:dsigma}
+        \mathrm{d}\sigma = \dfrac{1}{\mathcal{F}}|\mathcal{M}_{2 \rightarrow n}|^{2} \mathrm{d}\Phi_{n} \, .
+    \end{equation}
+    $\mathcal{F}$, known as the flux factor
+    is
+    \begin{equation}\label{eqn:flux}
+        \mathcal{F} = 4\sqrt{(p_{a}p_{b})^{2} - m_{a}^{2}m_{b}^{2}} \xrightarrow[\mathrm{limit}]{\mathrm{massless}} 2(p_{a} + p_{b})^{2} = 2s_{ab} \, .
+    \end{equation}
+
+    $\mathrm{d}\Phi_{n}$ is the Lorentz-invariant
+    phase-space which contains all the possible
+    configurations of the $n$-particle final state.
+    Absorbing the momentum conserving $\delta$-function
+    from (\ref{eqn:matrix_element}),
+    we can write it as
+    % \begin{equation}\label{eqn:dlips}
+    %     \mathrm{d}\Phi_{n} = (2\pi)^{d}\delta^{(d)}\left(p_{a} + p_{b} - \sum_{f} p_{f}^{\mu}\right) \prod_{f=1}^{n} \dfrac{\mathrm{d}^{d}p_{f}}{(2\pi)^{d-1}}\delta(p_{f}^{2} - m_{f}^{2})\theta(E_{f}) \, ,
+    % \end{equation}
+    \begin{equation}\label{eqn:dlips_4d}
+        \mathrm{d}\Phi_{n} = (2\pi)^{4}\delta^{4}\left(p_{a} + p_{b} - \sum_{f=1}^{n} p_{f}^{\mu}\right) \prod_{f=1}^{n} \dfrac{\mathrm{d}^{4}p_{f}}{(2\pi)^{3}}\delta(p_{f}^{2} - m_{f}^{2})\theta(E_{f}) \, ,
+    \end{equation}
+    where the second $\delta$-function restricts
+    the final state particles to be on mass-shell
+    (on-shell). The step function selects only
+    the positive energy solution for this constraint.
+% \begin{itemize}
+%     \item Strength of the coupling depends on the energy
+%     that the process occurs at. This is known as the running
+%     of the coupling.
+%     \item The value of $\alpha_{s}$ is usually quoted
+%     at the mass of the Z boson.
+%     \item In QCD the evolution of $\alpha_{s}$ is governed
+%     by the renormalisation group equation (Callan-Symanzik
+%     $\beta$ function). The coefficients of this function
+%     are known to many loops.
+%     \item Asymptotic freedom is a property of non-abelian
+%     gauge groups and this discovery won a Nobel prize.
+% \end{itemize}
 \section{Experiments and Phenomonology}

### 2022-09-09 intro.tex
some notes to not forget

@@
 \section{Introduction to quantum chromodynamics}
-\begin{itemize}
-    \item QCD is the quantum field theory describing the strong
-    interaction between quarks and gluons. It is a non-abelian
-    gauge theory with gauge group SU(3).
-    \item Write down the Lagrangian of QCD and define all the
-    terms (field strength tensor, quark fields, group generators,
-    structure constants, coupling constants.)
-    \item Begin to talk about how Feynman rules can be used
-    to represent the interactions that you can derive from
-    the Lagrangian.
-\end{itemize}
     Quantum chronodynamics (QCD) is the sector of the SM that
@@
 
-\subsection{Scattering amplitudes and cross-sections}
+\section{Scattering amplitudes and cross-sections}
     In collider experiments we typically collide
@@
     the positive energy solution for this constraint.
-% \begin{itemize}
-%     \item Strength of the coupling depends on the energy
-%     that the process occurs at. This is known as the running
-%     of the coupling.
-%     \item The value of $\alpha_{s}$ is usually quoted
-%     at the mass of the Z boson.
-%     \item In QCD the evolution of $\alpha_{s}$ is governed
-%     by the renormalisation group equation (Callan-Symanzik
-%     $\beta$ function). The coefficients of this function
-%     are known to many loops.
-%     \item Asymptotic freedom is a property of non-abelian
-%     gauge groups and this discovery won a Nobel prize.
-% \end{itemize}
+    \begin{itemize}
+        \item Matrix elements are calculated at fixed order in perturbative QFT -> Need to mentioned what perturbative even is...
+        \item How can QCD be perturbative when $\alpha_{s}$ is large? running, beta functions, asymptotic freedom
+        \item Why don't we see quarks and gluons at colliders? Hard/soft factorisation, pdfs, DGLAP, confinement, infrared safe observables
+        \item Relate hadronic cross-section to partonic cross-section.
+    \end{itemize}
+
 \section{Experiments and Phenomonology}

### 2022-09-10 intro.tex
wrote about running coupling

@@
 \chapter{Introduction}
+\label{chapter:intro}
 The standard model of particle physics (SM) is our current best working
@@
 final state particles and for higher orders in perturbative
-expansations in fixed order calculations. Although there
+expansions in fixed order calculations. Although there
 has been tremendous progress in this sector a complication
 that naturally arises is that the evaluation of these matrix
-elements becomes increasingly time consuming due to their
+elements becomes increasingly time-consuming due to their
 algebraic complexity.
@@
 
-\label{chapter:intro}
 \section{The Standard Model of particle physics}
@@
     \begin{figure}
-        \includegraphics[width=\linewidth]{introduction/sm_particles.pdf}
+        % \includegraphics[width=\linewidth]{introduction/sm_particles.pdf}
         \caption{Particle content of the SM split into quarks, leptons,
@@
     describes the strong interaction. 
-    QCD is a non-abelian gauge theory with gauge group
+    QCD is a non-Abelian gauge theory with gauge group
     $\mathrm{SU}(3)_{C}$ where the charge is named colour.
@@
     \begin{equation}\label{eqn:alpha_s}
-        \alpha_{\mathrm{s}} = \dfrac{g_{\mathrm{s}}^{2}}{4\pi} \, ,
+        \alpha_{\mathrm{s}} = \dfrac{g_{\mathrm{s}}^{2}}{4\pi} \, .
     \end{equation}
-    for reasons which will become clear when we discuss
-    scattering amplitudes.
+    The coupling in QCD is dependent on the energy scale
+    of the process, $\mu$. The behaviour of $\alpha_{\mathrm{s}}$
+    is governed by the Callan-Symanzik \cite{Callan:1970yg,Symanzik:1970rt}
+    $\beta$-function
+    \begin{equation}\label{eqn:beta_fn}
+        \mu^{2} \dfrac{\partial \alpha_{\mathrm{s}}(\mu^{2})}{\partial \mu^{2}} = \beta(\alpha_{\mathrm{s}}) \, ,
+    \end{equation}
+    where the $\beta$-function can be written as a perturbative
+    expansion in $\alpha_{\mathrm{s}}$
+    \begin{equation}\label{eqn:beta_series}
+        -\beta(\alpha_{\mathrm{s}}) = \sum_{n=0}^{\infty}\dfrac{\alpha_{\mathrm{s}}^{n+2}}{(4\pi)^{n+1}} \beta_{n}\, .
+    \end{equation}
+    where the coefficients $\beta_{n}$ have been computed
+    up to fifth order \cite{Baikov:2016tgj,Luthe:2017ttg}.
+    The coefficients computed up to this point have been
+    strictly positive, meaning that $\alpha_{\mathrm{s}}$
+    decreases with increasing energy due to the minus sign
+    in (\ref{eqn:beta_series}), a property known as
+    asymptotic freedom \cite{Gross:1973id,Politzer:1973fx}.
+    The solution to (\ref{eqn:beta_fn}) to first order
+    is
+    \begin{equation}\label{eqn:1l_alpha}
+        \alpha_{\mathrm{s}}(\mu^{2}) = \dfrac{1}{\frac{\beta_{0}}{4\pi}\log(\frac{\mu^{2}}{\Lambda_{\mathrm{QCD}}^{2}})} \, ,
+    \end{equation}
+    where $\Lambda_{\mathrm{QCD}}$ is the QCD scale,
+    the scale at which $\alpha_{\mathrm{s}}$ becomes
+    large enough that perturbation theory breaks down.
+    The running of the coupling constant has been
+    observed experimentally as can be seen in Figure~\ref{fig:alpha_s_running}.
+    \begin{figure}
+        \includegraphics{alpha_s_running.pdf}
+        \caption{The running of the strong coupling constant $\alpha_{\mathrm{s}}$,
+        as determined by experiments, with QCD theory prediction in black.
+        Figure reference \cite{Workman:2022ynf}.}
+        \label{fig:alpha_s_running}
+    \end{figure}
 
@@
     \begin{equation}\label{eqn:matrix_element}
-        \langle f | \mathcal{T} | i \rangle = (2\pi)^{4} \delta^{4}\left(p_{a} + p_{b} - \sum_{f=1}^{n} p_{f}^{\mu}\right) \mathcal{M} \, .
+        \langle f | \mathcal{T} | i \rangle = (2\pi)^{4} \delta^{4}\left(p_{a} + p_{b} - \sum_{f=1}^{n} p_{f}^{\mu}\right) \mathcal{M} \, ,
     \end{equation}
@@
     the cross-section, $\sigma$, which is a property
-    of the particles being scattered, and is independant
+    of the particles being scattered, and is independent
     of the way the experiment is carried out.

### 2022-09-12 intro.tex
start writing about factorisation

@@
     The gauge coupling of the group $g_{\mathrm{s}}$
-    is a dimensional free parameter of the theory, usually
-    written as
+    is a dimensionless free parameter of the theory.
+    It is common to use the strong coupling constant
+    instead, 
     \begin{equation}\label{eqn:alpha_s}
@@
     \end{equation}
-    The coupling in QCD is dependent on the energy scale
-    of the process, $\mu$. The behaviour of $\alpha_{\mathrm{s}}$
+    The strong coupling constant, is not constant, but
+    is dependent on the energy scale of the process, $\mu$.
+    The behaviour of $\alpha_{\mathrm{s}}$
     is governed by the Callan-Symanzik \cite{Callan:1970yg,Symanzik:1970rt}
@@
     \end{figure}
+    Another consequence of the running coupling
+    is colour confinement, in which quarks and
+    gluons cannot exist as free particles. Due
+    to the increasing strength of coupling at low
+    energies, quarks and gluons are forced to
+    form composite, colourless particles, known as
+    hadrons. The most common example of a hadron
+    is the proton which is the origin of the name
+    Large Hadron Collider.
 
@@
     and cross-sections.
+
+\section{Factorisation theorem}
+    At the LHC, collisions occur between two protons,
+    which have constituent quarks and gluons.
+    The application of QCD to describe phenomenon
+    in these proton-proton collisions rests on the use of
+    the factorisation theorem which enables the
+    separation of `soft' and `hard' energy scales.
+    Within the proton, there are composite quarks
+    and gluons that are constantly absorbed and
+    emitted. However, during a high energy collision
+    of two protons, the constituent quarks and gluons are
+    effectively frozen due to the extremely high
+    energy of the collision.
+    The `soft' energy scales refers to the quarks
+    and gluons residing inside a proton and the `hard'
+    energy scale refers to the energetic collision.
+
+    The production of particles from a hadronic collision,
+    the hadronic cross-section can therefore be
+    written as
+    ...
 

### 2022-09-13 qcd.tex
update figures

@@
 \section{Divergent structures}
+\subsection{Running of the coupling constant}
+The behaviour of $\alpha_{\mathrm{s}}$
+    is governed by the Callan-Symanzik \cite{Callan:1970yg,Symanzik:1970rt}
+    $\beta$-function
+    \begin{equation}\label{eqn:beta_fn}
+        \mu^{2} \dfrac{\partial \alpha_{\mathrm{s}}(\mu^{2})}{\partial \mu^{2}} = \beta(\alpha_{\mathrm{s}}) \, ,
+    \end{equation}
+    where the $\beta$-function can be written as a perturbative
+    expansion in $\alpha_{\mathrm{s}}$
+    \begin{equation}\label{eqn:beta_series}
+        -\beta(\alpha_{\mathrm{s}}) = \sum_{n=0}^{\infty}\dfrac{\alpha_{\mathrm{s}}^{n+2}}{(4\pi)^{n+1}} \beta_{n}\, .
+    \end{equation}
+    where the coefficients $\beta_{n}$ have been computed
+    up to fifth order \cite{Baikov:2016tgj,Luthe:2017ttg}.
+    The coefficients computed up to this point have been
+    strictly positive, meaning that $\alpha_{\mathrm{s}}$
+    decreases with increasing energy due to the minus sign
+    in (\ref{eqn:beta_series}), a property known as
+    asymptotic freedom \cite{Gross:1973id,Politzer:1973fx}.
+    The solution to (\ref{eqn:beta_fn}) to first order
+    is
+    \begin{equation}\label{eqn:1l_alpha}
+        \alpha_{\mathrm{s}}(\mu^{2}) = \dfrac{1}{\frac{\beta_{0}}{4\pi}\log\left(\frac{\mu^{2}}{\Lambda_{\mathrm{QCD}}^{2}}\right)} \, ,
+    \end{equation}
+    where $\Lambda_{\mathrm{QCD}}$ is the QCD scale,
+    the scale at which $\alpha_{\mathrm{s}}$ becomes
+    large enough that perturbation theory breaks down.
+    The running of the coupling constant has been
+    observed experimentally as can be seen in Figure~\ref{fig:alpha_s_running}.
+    \begin{figure}
+        \includegraphics{qcd/alpha_s_running.pdf}
+        \caption{The running of the strong coupling constant $\alpha_{\mathrm{s}}$,
+        as determined by experiments, with QCD theory prediction in black.
+        Figure reference \cite{Workman:2022ynf}.}
+        \label{fig:alpha_s_running}
+    \end{figure}
+    Another consequence of the running coupling
+    is colour confinement, in which quarks and
+    gluons cannot exist as free particles. Due
+    to the increasing strength of coupling at low
+    energies, quarks and gluons are forced to
+    form composite, colourless particles, known as
+    hadrons. The most common example of a hadron
+    is the proton which is the origin of the name
+    Large Hadron Collider.
 \section{Infrared factorisation}

### 2022-09-13 intro.tex
words on factorisation and pdfs

--- 2022-09-09 intro.tex
@@
     \begin{equation}\label{eqn:SM_gauge}
-        \mathrm{SU}(3)_{C} \times \mathrm{SU}(2)_{L} \times \mathrm{U}(1)_{Y}\, ,
+        \mathrm{SU}(3)_{c} \times \mathrm{SU}(2)_{L} \times \mathrm{U}(1)_{Y}\, ,
     \end{equation}
@@
 \section{Introduction to quantum chromodynamics}
+    \subsection{The QCD Lagrangian}
     Quantum chronodynamics (QCD) is the sector of the SM that
     describes the strong interaction. 
-    QCD is a non-abelian gauge theory with gauge group
-    $\mathrm{SU}(3)_{C}$ where the charge is named colour.
+    QCD is a non-Abelian gauge theory with gauge group
+    $\mathrm{SU}(3)_{c}$ where the charge is named colour.
     The gauge and fermion part of the QCD Lagrangian is
@@
     \end{equation}
-    Gauge fields transform under the adjoint representation, meaning
+    Gauge fields transform under the adjoint representation, 
+    which is an $N_{c}^{2}-1$ dimensional representation, meaning
     the adjoint indices $a,b,c \in \{1,...,8\}$. $T^{a}_{ij}$ are the
@@
     The gauge coupling of the group $g_{\mathrm{s}}$
-    is a dimensional free parameter of the theory, usually
-    written as
+    is a dimensionless free parameter of the theory.
+    It is common to use the strong coupling constant
+    instead, 
     \begin{equation}\label{eqn:alpha_s}
-        \alpha_{\mathrm{s}} = \dfrac{g_{\mathrm{s}}^{2}}{4\pi} \, ,
+        \alpha_{\mathrm{s}} = \dfrac{g_{\mathrm{s}}^{2}}{4\pi} \, .
     \end{equation}
-    for reasons which will become clear when we discuss
-    scattering amplitudes.
+    The strong coupling constant, is in fact not constant, but
+    is dependent on the energy scale of the process, $\mu$. This
+    is due to the process of renormalisation which will be discussed
+    in section ...
 
@@
 
+    \subsection{QCD Feynman rules}
     Now that we have written down the QCD Lagrangian,
@@
 
+\section{Factorisation theorem}
+    At the LHC, collisions occur between two protons,
+    which have constituent quarks and gluons, collectively
+    named partons.
+    The application of QCD to describe phenomenon
+    in these proton-proton collisions rests on the use of
+    the factorisation theorem which enables the
+    separation of soft and hard energy scales.
+    Within the proton, there are composite quarks
+    and gluons that are constantly absorbed and
+    emitted on a timescale inversely proportional
+    to the mass of the proton (or hadron in general).
+    This fluctuation timescale is much longer
+    than the timescale of the interaction between
+    an incoming probing parton from another proton in a
+    highly energetic collision, like at the LHC.
+    During the collision, or the hard scattering,
+    the probe is able to interact with a
+    parton that is effectively frozen.
+    The probing parton knows nothing of the
+    proton being probed, except for the fact
+    that it collided with a parton carrying
+    a fraction of the proton momentum. The
+    distribution of partons within a parton
+    is process independent and a fundamental
+    property of the proton, meaning it can be
+    separated from the actual scattering between
+    the partons.
+
+    This heuristic argument motivates the form of
+    the factorisation equation, where the production of
+    particles from a hadronic collision, the hadronic
+    cross-section, can be
+    written as
+    \begin{equation}\label{eqn:hadronic_cs}
+        \sigma_{AB \rightarrow n} = \sum_{a, b}\int_{0}^{1} \mathrm{d}x_{a}\mathrm{d}x_{b} \; f_{a/A}(x_{a},\mu_{\mathrm{F}})f_{b/B}(x_{b}, \mu_{\mathrm{F}}) \; \hat{\sigma}_{ab \rightarrow n}(Q, \mu_{\mathrm{F}}, \mu_{\mathrm{R}}) + \mathcal{O}\left(\dfrac{\Lambda_{\mathrm{QCD}}}{Q}\right)\, ,
+    \end{equation}
+    where $a$ is the parton identified from hadron $A$,
+    and $b$ is the parton identified from hadron $B$.
+    This factorisation is not exact as indicated by the
+    correction term inversely proportional to characteristic
+    hard scale $Q$. However, for high energy collisions, where
+    $Q \gg \Lambda_{\mathrm{QCD}}$, this term and higher order
+    terms are suppresed.
+    In fact, the factorisation equation has only been
+    proven for a few specific cases (DIS etc, cite here).
+
+    $f_{a/h}(x, \mu)$ are parton distribution functions
+    (PDFs) which depend on the momenta fractions $x$ parton $a$
+    has with respect to its parent hadron $h$, at a scale $\mu$,
+    usually taken to be the factorisation scale $\mu_{\mathrm{F}}$,
+    the interface between soft and hard physics.
+    The interpretation of PDFs at leading order are
+    as probability distributions. $f_{a/h}(x, \mu)$ is
+    the probability of finding parton $a$ within $h$
+    carrying a momentum fraction $x$ at the energy scale $\mu$.
+    PDFs are non-perturbative objects that encapsulate
+    the soft effects in a scattering process below energy
+    $\mu_{\mathrm{F}}$. They are non-perturbative because
+    they are determined by fitting to experimental data,
+    instead of calculated in perturbation theory.
+    Multiple groups release PDF fits (cite groups here).
+    In practice, these PDF fits
+    are accessed through the LHAPDF interface (cite here).
+    The evolution of PDFs between factorisation scales is possible
+    through the use of the Dokshitzer-Gribov-Lipatov-Altarelli-Parisi (DGLAP)
+     equations (cite here).
+
+    $\hat{\sigma}_{ab \rightarrow n}(Q, \mu_{\mathrm{F}}, \mu_{\mathrm{R}})$
+    is the partonic cross-section which descibes the interaction of
+    partons $a$ and $b$ scattering into $n$ particles,
+    with the scattering occurs at energy scale $Q$.
+    Notice that the partonic cross-section depends on
+    both the factorisation scale, $\mu_{\mathrm{F}}$ and
+    the renormalisation scale, $\mu_{\mathrm{R}}$.
+    {\color{red} Need to mention how these scales
+    are unphysical but there is a residual dependence
+    on them at fixed-order calculations. Scale
+    variations are a conventional way to determine
+    the effects of higher order terms.
+    Renormalisation will be discussed in Chapter 2
+    when divergences are introduced.}
+
+    The discussion of how $\hat{\sigma}_{ab \rightarrow n}$
+    is computed is discussed in detail in the next section.
+
 \section{Scattering amplitudes and cross-sections}
+    {\color{red} Cross-sections here are all
+    partonic cross-sections. Need to talk about
+    how they can be calculated order by order
+    in perturbation theory with Feynman diagrams.}
+
     In collider experiments we typically collide
@@
     In practice, it is more useful to consider the
-    differential cross-section, $\mathrm{d}\sigma$
+    differential cross-section, $\mathrm{d}\sigma$,
     where the cross-section can be differential
@@
     the positive energy solution for this constraint.
-    \begin{itemize}
-        \item Matrix elements are calculated at fixed order in perturbative QFT -> Need to mentioned what perturbative even is...
-        \item How can QCD be perturbative when $\alpha_{s}$ is large? running, beta functions, asymptotic freedom
-        \item Why don't we see quarks and gluons at colliders? Hard/soft factorisation, pdfs, DGLAP, confinement, infrared safe observables
-        \item Relate hadronic cross-section to partonic cross-section.
-    \end{itemize}
 
-\section{Experiments and Phenomonology}
-\subsection{Observables}
-\begin{itemize}
-    \item Experiments do not measure single particles,
-    they measure hadrons which is far from the collision
-    of protons.
-    \item Factorisation of hard and soft scales leads
-    to the concepts of PDFs (DGLAP equation), partonic
-    cross-sections (Feynman diagrams), parton showers,
-    and hadronisation.
-    \item Link back to matrix elements and how they
-    are integrated to give you cross-sections which
-    can be compared with experiments.
-    \item Probably need to talk about infrared safe
-    observables and jets.
-\end{itemize}
 \section{Monte Carlo event generator overview}

### 2022-09-14 intro.tex
wrote: cross-section, matrix elements, divergences

@@
 
-\section{Introduction to quantum chromodynamics}
-    \subsection{The QCD Lagrangian}
+\section{Introduction to quantum chromodynamics}\label{sec:qcd}
+    \subsection{The QCD Lagrangian}\label{sec:L_qcd}
     Quantum chronodynamics (QCD) is the sector of the SM that
@@
     The strong coupling constant, is in fact not constant, but
-    is dependent on the energy scale of the process, $\mu$. This
+    is dependent on the energy scale of the process. This
     is due to the process of renormalisation which will be discussed
     in section ...
+    The implications of this is that at collider
+    experiments where collisions are extremely high energy,
+    the strong coupling constant becomes small, a
+    propery known as asymptotic freedom. A consequence
+    of this is that predictions made in QCD can be expressed
+    in the form of perturbative expansions in $\alpha_{\mathrm{s}}$.
 
@@
 
-    \subsection{QCD Feynman rules}
+    \subsection{QCD Feynman rules}\label{sec:qcd_feynman}
     Now that we have written down the QCD Lagrangian,
@@
 
-    Note that we have only written down the vertices
-    and propagators for the physical gluons and quarks,
+    {\color{red} Need to write down the Feynman
+    rules and draw diagrams. Best to put it in a
+    figure that we can refer back to.}
+
+    Note that we have only written down the vertices,
+    propagators, and external lines for the physical gluons and quarks,
     ghosts also have associated Feynman rules but we do
@@
 
-    Now that we have the Feynman rules for the theory
-    we can move on to discuss scattering amplitudes
-    and cross-sections.
+    From these Feynman rules, we identify that each
+    quark-antiquark-gluon and triple gluon vertices
+    are associated with one power of $\alpha_{\mathrm{s}}$,
+    and the four gluon vertex is associated with
+    two powers of $\alpha_{\mathrm{s}}$. This relationship
+    between vertices and $\alpha_{\mathrm{s}}$ allows
+    us to systematically build up Feynman diagrams
+    that correspond to a fixed order in $\alpha_{\mathrm{s}}$.
+    This point will be elaborated on in Section
+    \ref{sec:xs} where we discuss scattering amplitudes
+    and how they are related to Feynman diagrams.
 
-\section{Factorisation theorem}
+\section{Factorisation theorem}\label{sec:factorisation}
     At the LHC, collisions occur between two protons,
@@
     to the mass of the proton (or hadron in general).
-    This fluctuation timescale is much longer
+    The timescale of this fluctuation is much longer
     than the timescale of the interaction between
-    an incoming probing parton from another proton in a
-    highly energetic collision, like at the LHC.
+    a parton and a highly energetic probing parton
+    from another proton. This is the scenario
+    that occurs at collisions at the LHC.
     During the collision, or the hard scattering,
@@
     a fraction of the proton momentum. The
-    distribution of partons within a parton
+    distribution of partons within a proton
     is process independent and a fundamental
@@
     particles from a hadronic collision, the hadronic
-    cross-section, can be
-    written as
+    cross-section, can be written as
     \begin{equation}\label{eqn:hadronic_cs}
@@
     \end{equation}
-    where $a$ is the parton identified from hadron $A$,
-    and $b$ is the parton identified from hadron $B$.
+    where $a$ is the parton from hadron $A$,
+    and $b$ is the parton from hadron $B$.
+    The sum over initial-state partons $a$ and $b$
+    indicates the sum over flavours, which depends
+    on the hadron composition in general.
     This factorisation is not exact as indicated by the
@@
     PDFs are non-perturbative objects that encapsulate
-    the soft effects in a scattering process below energy
+    the soft effects of the scattering which occur below energy
     $\mu_{\mathrm{F}}$. They are non-perturbative because
@@
 
+    The momenta fractions, $x$, also link the
+    squared centre-of-mass energies of the hadronic collision,
+    denoted as $s$, to the partonic equivalent as
+    \begin{equation}\label{eqn:E_cm_partonic}
+        \hat{s} = x_{a}x_{b}s \, .
+    \end{equation}
+
     $\hat{\sigma}_{ab \rightarrow n}(Q, \mu_{\mathrm{F}}, \mu_{\mathrm{R}})$
@@
     partons $a$ and $b$ scattering into $n$ particles,
-    with the scattering occurs at energy scale $Q$.
+    where the scattering occurs at energy scale $Q$,
+    which is often taken to the $\hat{s}$.
     Notice that the partonic cross-section depends on
     both the factorisation scale, $\mu_{\mathrm{F}}$ and
-    the renormalisation scale, $\mu_{\mathrm{R}}$.
+    the renormalisation scale, $\mu_{\mathrm{R}}$
+    (see \ref{chapter:qcd}).
     {\color{red} Need to mention how these scales
@@
 
-    The discussion of how $\hat{\sigma}_{ab \rightarrow n}$
-    is computed is discussed in detail in the next section.
+    The discussion so far has been focused hadron-hadron
+    initiated scattering, however, a similar argument
+    can be made about electron-positron annihilation,
+    where the electron-positron annihilates to form a
+    photon ($Z$ boson) which fluctuates into a quark-antiquark
+    pair. For a sufficiently high energy collision,
+    these interactions can be calculated in perturbation
+    theory.
+    Since electrons and positrons are not composite
+    particles, there is no requirement for PDFs,
+    therefore (\ref{eqn:hadronic_cs}) simplifies
+    to just the partonic cross-section, which
+    will be computed in perturbation theory.
 
-\section{Scattering amplitudes and cross-sections}
-    {\color{red} Cross-sections here are all
-    partonic cross-sections. Need to talk about
-    how they can be calculated order by order
-    in perturbation theory with Feynman diagrams.}
+    The details of how $\hat{\sigma}_{ab \rightarrow n}$
+    is computed is discussed in the next section.
 
-    In collider experiments we typically collide
+\section{Scattering amplitudes and cross-sections}\label{sec:xs}
+    As mentioned in the previous section, at
+    collider experiments we typically collide
     two beams, consisting of bundles of energetic
     particles, and analyse the resulting products
-    of these collisions. To model collisions in
+    of these collisions. The predictions we make
+    from our theoretical model are the partonic
+    cross-sections. To arrive at an expression
+    for the partonic cross-section, we need to
+    first consider the hard scattering of particles.
+    To model the hard collisions in
     a collider experiment, we look at the specific
@@
     \begin{equation}\label{eqn:S_prob}
-        P = |\langle f | S | i \rangle|^{2} \propto |\langle f | \mathcal{M} | i \rangle 
+        P = |\langle f | S | i \rangle|^{2} \propto | \langle f | \mathcal{M} | i \rangle |^{2}
     \end{equation}
     where $|\langle f | \mathcal{M} | i \rangle|^{2}  \equiv |\mathcal{M}|^{2}$
-    is the matrix element squared.
-
-    Matrix elements can be calculated from
-    first principles in QFT using Feynman diagrams.
+    is the matrix element squared \footnote{from henceforth we refer to the matrix
+    element squared as matrix element}.
 
@@
     events are not measured. Instead we consider
-    the cross-section, $\sigma$, which is a property
+    the cross-section, $\sigma$, as introduced in
+    (\ref{eqn:hadronic_cs}). The cross-section is a property
     of the particles being scattered, and is independent
     of the way the experiment is carried out.
-    The cross-section directly relates the
-    matrix elements which we calculate in our QFT,
+    Since the hadronic cross-section factorises
+    the soft energy effects into the non-perturbative PDFs,
+    which are determined once and for all for all
+    processes, the object of interest now becomes the
+    partonic cross-section, $\hat{\sigma}$, which
+    encodes the scattering information of the specific
+    process considered.
+
+    Given that cross-sections are measurable
+    quantities, the partonic cross-section directly
+    relates the matrix elements which we calculate in our QFT,
     to measurable observables at experiments.
     In practice, it is more useful to consider the
-    differential cross-section, $\mathrm{d}\sigma$,
+    differential cross-section, $\mathrm{d}\hat{\sigma}$,
     where the cross-section can be differential
@@
 
-    The differential cross-section can be written as
+    The differential partonic cross-section can be written as
     \begin{equation}\label{eqn:dsigma}
-        \mathrm{d}\sigma = \dfrac{1}{\mathcal{F}}|\mathcal{M}_{2 \rightarrow n}|^{2} \mathrm{d}\Phi_{n} \, .
+        \mathrm{d}\hat{\sigma} = \dfrac{1}{\mathcal{F}}|\mathcal{M}_{2 \rightarrow n}|^{2} \mathrm{d}\Phi_{n} \, .
     \end{equation}
-    $\mathcal{F}$, known as the flux factor
-    is
+    $\mathcal{F}$, the flux factor is
     \begin{equation}\label{eqn:flux}
-        \mathcal{F} = 4\sqrt{(p_{a}p_{b})^{2} - m_{a}^{2}m_{b}^{2}} \xrightarrow[\mathrm{limit}]{\mathrm{massless}} 2(p_{a} + p_{b})^{2} = 2s_{ab} \, .
+        \mathcal{F} = 4\sqrt{(p_{a}p_{b})^{2} - m_{a}^{2}m_{b}^{2}} \, ,
     \end{equation}
+    which in the massless limit reduces to
+    \begin{equation}\label{eqn:massless_flux}
+        \mathcal{F} = 2(p_{a} + p_{b})^{2} = 2\hat{s} \, .
+    \end{equation}
+    It is common to use the massless limit because
+    the centre-of-mass energy of a collider
+    experiment is much greater than the mass
+    of the colliding particles.
 
@@
     the positive energy solution for this constraint.
+    This gives an intuitive picture of the partonic
+    cross-section as the probability of a $2\rightarrow n$
+    process occurring, summed over all the possible
+    valid final-state configurations.
+
+    From (\ref{eqn:S_prob}), it becomes clear
+    that the matrix element is the object which
+    relates back to the Lagrangian of the theory
+    as it encodes the interactions of the particles.
+    We saw in Section \ref{sec:qcd_feynman} that
+    interactions could be codified into Feynman
+    rules which are pieced together to construct
+    Feynman diagrams. In this picture, matrix elements
+    are exactly the sum over all Feynman diagrams
+    \footnote{technically the interference of
+    Feynman diagrams as we are working on the level
+    of $|\mathcal{M}|^{2}$}
+    for a particular process. However, there are
+    an infinite number of Feynman diagrams you can
+    draw as there is no constraints on how the
+    vertices and propagators can be connected.
+    Fortunately because $\alpha_{\mathrm{s}}$
+    is small in high energy collisions, we can expand
+    the matrix element as a perturbative series in
+    $\alpha_{\mathrm{s}}$ to write
+    \begin{equation}\label{eqn:matrix_element}
+        |\mathcal{M}|^{2} =  \alpha_{\mathrm{s}}^{m} |\mathcal{M}|^{2}_{\mathrm{LO}} + \alpha_{\mathrm{s}}^{m+1} |\mathcal{M}|^{2}_{\mathrm{NLO}} + \alpha_{\mathrm{s}}^{m+2} |\mathcal{M}|^{2}_{\mathrm{NNLO}} + \mathcal{O}(\alpha_{\mathrm{s}}^{m+3}) \, ,
+    \end{equation}
+    where $m$ corresponds to the powers of $\alpha_{\mathrm{s}}$
+    in the simplest Feynman diagrams you can draw
+    for the process of interest.
+    Since each term of this expansion is at a fixed-order
+    in $\alpha_{\mathrm{s}}$, it is possible to systematically
+    compute the diagrams which contribute at the given order of $\alpha_{\mathrm{s}}$.
+    The set of diagrams that contribute at the lowest
+    order in $\alpha_{\mathrm{s}}$ are $|\mathcal{M}|^{2}_{\mathrm{LO}}$,
+    where LO stands for leading order. The next term
+    in the expansion corresponds to the leading order
+    diagrams with an additional loop, or external leg,
+    which contributes an extra factor of $\alpha_{\mathrm{s}}$.
+    This term is dubbed NLO for next-to-leading order
+    in $\alpha_{\mathrm{s}}$, and the following term
+    next-to-next-to-leading order has again an additional
+    loop or leg. In general, we have terms
+    N$^{k}$LO where each additional power of $\alpha_{\mathrm{s}}$
+    increases complixity of the computations, however,
+    because the expansion $\alpha_{\mathrm{s}}$
+    is small, each additional term contribute less.
+    This means that the first terms dominate the expansion,
+    and it is sufficient to terminate the series after
+    a few terms to reach an acceptable level of accuracy.
+    A further discussion of this point will be given
+    in Chapter~\ref{chapter:qcd} where we break down
+    the computation of each term in (\ref{eqn:matrix_element}).
+
+    In summary, the partonic cross-section of a collision
+    reduces to the computation of the matrix elements
+    up to a fixed-order in the coupling expansion,
+    which is then integrated over the valid phase-space of
+    final-state particle configurations. For hadronic
+    collisions the partonic cross-section also needs to
+    be convoluted with the PDFs to obtain the hadronic
+    cross-sections.
+
+    With the introduction of matrix elements and
+    cross-sections complete, we will move the discussion onto
+    Monte Carlo event generators (MCEG). These event
+    generators are multi-purpose tools that
+    incorporate many aspects of the theory discussed
+    so far (and to be discussed) for practical use.
+    Most importantly, they are
+    used to compare theoretical predictions to
+    experimental measurements.
 

### 2022-09-14 qcd.tex
wrote: cross-section, matrix elements, divergences

@@
 \begin{document}
-\chapter{Quantum chromodynamics}
+\chapter{Quantum chromodynamics in practice}
 \label{chapter:qcd}
-\section{Fixed order perturbation theory}
-\subsection{NLO QCD}
+    In the previous chapter we introduced the basic concepts
+    of QCD and setup the framework of calculating
+    cross-sections. Cross-sections are the most relevant
+    quantities to compute as they can be measured
+    experimentally, and are a property of the particles
+    being collided, rather than depending on the
+    specifics of the experimental procedure.
+    This provides a bridge to compare
+    theoretical predictions and experiment.
+    From (\ref{eqn:dsigma}, the partonic cross-section
+    is the matrix element integrated over the relevant
+    final-state phase-space, normalised by a flux factor.
+    The matrix elements are
+    calculated order-by-order in the strong coupling
+    $\alpha_{\mathrm{s}}$, (\ref{eqn:matrix_element}),
+    where each order corresponds to a set of Feynman diagrams
+    that have the appropriate number of vertices,
+    as each QCD vertex brings along a factor of $\alpha_{\mathrm{s}}$.
+
+    In this chapter we will discuss the problems that
+    arise when evaluating these Feynman diagrams and outline
+    some solutions that have been widely adopted to circumvent
+    these issues to provide real-world applicable predictions.
+    This chapter forms the theoretical foundations on which
+    the work of this thesis is built upon.
+
+    We will consider the divergent nature of matrix
+    elements and introduce the most widely used techniques
+    to tame these singularities.
+
 \section{Divergent structures}
+    We have seen that matrix element computations
+    result in evaluating Feynman diagrams. From
+    the QCD Feynman rules (ref here), it becomes
+    clear that integrating over all momentum modes
+    will give rise to singularities. The divergences
+    associated with high energy modes are ultraviolet
+    (UV) divergences. On the opposite end of the energy
+    spectrum, low energy modes also give rise to 
+    infrared (IR) divergences.
+
+    The de-facto methods to alleviate these
+    divergences are through regularisation
+    for UV divergences, and through subtraction
+    schemes for IR divergences. Both of these
+    methods will be discussed in this section.
+
+\subsection{Ultraviolet divergences}
+    In loop diagrams we have to evaluate integrals of the form
+    \begin{equation}
+        \int_{0}^{\Lambda} \dfrac{\mathrm{d}^{4}k}{(2\pi)^{4}}\dfrac{1}{k^{2}(k+p)^{2}} \sim \log{\Lambda} \, ,
+    \end{equation}
+    where a cutoff scale $\Lambda$ has been introduced to
+    capture the divergence as the loop momenta
+    $k \rightarrow \infty$. It is clear that the
+    integral divergences in this high energy limit,
+    hence the name ultraviolet divergence.
+\subsection{Renormalisation}
 \subsection{Running of the coupling constant}
@@
     Large Hadron Collider.
-\section{Infrared factorisation}
-\subsection{Collinear limits}
-\subsection{Soft limits}
+\subsection{Infrared divergences}
+\subsection{Factorisation of matrix elements}
+\subsubsection{Soft limits}
+\subsubsection{Collinear limits}
+
 \section{Subtraction}

### 2022-09-15 intro.tex
wrote: UV, DR, renormalisation, alpha running

@@
     \begin{equation}\label{eqn:covariant_deriv}
-        (D_{\mu})_{ij} = \delta_{ij}\partial_{\mu} - ig_{\mathrm{s}}T^{a}_{ij}A^{a}_{\mu} \, ,
+        (D_{\mu})_{ij} = \delta_{ij}\partial_{\mu} - ig_{s}T^{a}_{ij}A^{a}_{\mu} \, ,
     \end{equation}
@@
     \begin{equation}\label{eqn:F_munu}
-        F^{a}_{\mu\nu} = \partial_{\mu}A^{a}_{\nu} - \partial_{\nu}A^{a}_{\mu} + g_{\mathrm{s}} f^{abc}A^{b}_{\mu}A^{c}_{\nu} \, .
+        F^{a}_{\mu\nu} = \partial_{\mu}A^{a}_{\nu} - \partial_{\nu}A^{a}_{\mu} + g_{s} f^{abc}A^{b}_{\mu}A^{c}_{\nu} \, .
     \end{equation}
@@
     where $f^{abc}$ are the structure constants of SU(3).
-    The gauge coupling of the group $g_{\mathrm{s}}$
+    The generators are normalised by convention to be
+    \begin{equation}\label{eqn:generator_normalisation}
+        \mathrm{Tr}(T^{a}T^{b}) = \delta^{ab}T_{R} \, , \quad T_{R} = \dfrac{1}{2} \, .
+    \end{equation}
+    This normalisation sets the values of the Casimirs
+    of the group as
+    \begin{align}\label{eqn:casimirs}
+        \begin{split}
+            T^{a}_{ij}T^{a}_{jk} &= \delta{ik}C_{F} \, , \quad C_{F} = \dfrac{N_{c}^2-1}{2N_{c}} \, , \\
+            f^{abc}f^{abd} &= \delta^{cd}C_{A} \, , \quad C_{A} = N_{c} \, ,
+        \end{split}
+    \end{align}
+    where $C_{F}=4/3$ and $C_{A}=3$ for QCD.
+    They are collectively referred to as colour factors.
+
+    The gauge coupling of the group $g_{s}$
     is a dimensionless free parameter of the theory.
@@
     \begin{equation}\label{eqn:alpha_s}
-        \alpha_{\mathrm{s}} = \dfrac{g_{\mathrm{s}}^{2}}{4\pi} \, .
+        \alpha_{s} = \dfrac{g_{s}^{2}}{4\pi} \, .
     \end{equation}
@@
     of this is that predictions made in QCD can be expressed
-    in the form of perturbative expansions in $\alpha_{\mathrm{s}}$.
+    in the form of perturbative expansions in $\alpha_{s}$.
 
@@
     \begin{equation}\label{eqn:L_ghost}
-        \mathcal{L}_{\mathrm{ghost}} = (\partial_{\mu}\bar{c}^{a})(\delta^{ac}\partial_{\mu} + g_{\mathrm{s}}f^{abc}A^{b}_{\mu})c^{c} \, ,
+        \mathcal{L}_{\mathrm{ghost}} = (\partial_{\mu}\bar{c}^{a})(\delta^{ac}\partial_{\mu} + g_{s}f^{abc}A^{b}_{\mu})c^{c} \, ,
     \end{equation}
@@
     \begin{equation}\label{L_int}
-        \mathcal{L}_{\mathrm{int}} = g_{\mathrm{s}}A^{a}_{\mu}\bar{\psi}_{i}^{f}\gamma^{\mu}T^{a}_{ij}\psi_{j}^{f}-g_{\mathrm{s}}f^{abc}(\partial_{\mu}A^{a}_{\nu})A^{b,\,\mu}A^{c,\,\nu} -\dfrac{g_{\mathrm{s}}^{2}}{4}f^{abc}A^{b}_{\mu}A^{c}_{\nu}f^{ade}A^{d,\,\mu}A^{e,\,\nu} \, ,
+        \mathcal{L}_{\mathrm{int}} = g_{s}A^{a}_{\mu}\bar{\psi}_{i}^{f}\gamma^{\mu}T^{a}_{ij}\psi_{j}^{f}-g_{s}f^{abc}(\partial_{\mu}A^{a}_{\nu})A^{b,\,\mu}A^{c,\,\nu} -\dfrac{g_{s}^{2}}{4}f^{abc}A^{b}_{\mu}A^{c}_{\nu}f^{ade}A^{d,\,\mu}A^{e,\,\nu} \, ,
     \end{equation}
@@
     quark-antiquark-gluon and triple gluon vertices
-    are associated with one power of $\alpha_{\mathrm{s}}$,
+    are associated with one power of $\alpha_{s}$,
     and the four gluon vertex is associated with
-    two powers of $\alpha_{\mathrm{s}}$. This relationship
-    between vertices and $\alpha_{\mathrm{s}}$ allows
+    two powers of $\alpha_{s}$. This relationship
+    between vertices and $\alpha_{s}$ allows
     us to systematically build up Feynman diagrams
-    that correspond to a fixed order in $\alpha_{\mathrm{s}}$.
+    that correspond to a fixed order in $\alpha_{s}$.
     This point will be elaborated on in Section
@@
     \begin{equation}\label{eqn:hadronic_cs}
-        \sigma_{AB \rightarrow n} = \sum_{a, b}\int_{0}^{1} \mathrm{d}x_{a}\mathrm{d}x_{b} \; f_{a/A}(x_{a},\mu_{\mathrm{F}})f_{b/B}(x_{b}, \mu_{\mathrm{F}}) \; \hat{\sigma}_{ab \rightarrow n}(Q, \mu_{\mathrm{F}}, \mu_{\mathrm{R}}) + \mathcal{O}\left(\dfrac{\Lambda_{\mathrm{QCD}}}{Q}\right)\, ,
+        \sigma_{AB \rightarrow n} = \sum_{a, b}\int_{0}^{1} \mathrm{d}x_{a}\mathrm{d}x_{b} \; f_{a/A}(x_{a},\mu_{F})f_{b/B}(x_{b}, \mu_{F}) \; \hat{\sigma}_{ab \rightarrow n}(Q, \mu_{F}, \mu_{R}) + \mathcal{O}\left(\dfrac{\Lambda_{\mathrm{QCD}}}{Q}\right)\, ,
     \end{equation}
@@
     has with respect to its parent hadron $h$, at a scale $\mu$,
-    usually taken to be the factorisation scale $\mu_{\mathrm{F}}$,
+    usually taken to be the factorisation scale $\mu_{F}$,
     the interface between soft and hard physics.
@@
     the soft effects of the scattering which occur below energy
-    $\mu_{\mathrm{F}}$. They are non-perturbative because
+    $\mu_{F}$. They are non-perturbative because
     they are determined by fitting to experimental data,
@@
 
-    $\hat{\sigma}_{ab \rightarrow n}(Q, \mu_{\mathrm{F}}, \mu_{\mathrm{R}})$
+    $\hat{\sigma}_{ab \rightarrow n}(Q, \mu_{F}, \mu_{R})$
     is the partonic cross-section which descibes the interaction of
@@
     Notice that the partonic cross-section depends on
-    both the factorisation scale, $\mu_{\mathrm{F}}$ and
-    the renormalisation scale, $\mu_{\mathrm{R}}$
+    both the factorisation scale, $\mu_{F}$ and
+    the renormalisation scale, $\mu_{R}$
     (see \ref{chapter:qcd}).
@@
     vertices and propagators can be connected.
-    Fortunately because $\alpha_{\mathrm{s}}$
+    Fortunately because $\alpha_{s}$
     is small in high energy collisions, we can expand
     the matrix element as a perturbative series in
-    $\alpha_{\mathrm{s}}$ to write
+    $\alpha_{s}$ to write
     \begin{equation}\label{eqn:matrix_element}
-        |\mathcal{M}|^{2} =  \alpha_{\mathrm{s}}^{m} |\mathcal{M}|^{2}_{\mathrm{LO}} + \alpha_{\mathrm{s}}^{m+1} |\mathcal{M}|^{2}_{\mathrm{NLO}} + \alpha_{\mathrm{s}}^{m+2} |\mathcal{M}|^{2}_{\mathrm{NNLO}} + \mathcal{O}(\alpha_{\mathrm{s}}^{m+3}) \, ,
+        |\mathcal{M}|^{2} =  \alpha_{s}^{m} |\mathcal{M}|^{2}_{\mathrm{LO}} + \alpha_{s}^{m+1} |\mathcal{M}|^{2}_{\mathrm{NLO}} + \alpha_{s}^{m+2} |\mathcal{M}|^{2}_{\mathrm{NNLO}} + \mathcal{O}(\alpha_{s}^{m+3}) \, ,
     \end{equation}
-    where $m$ corresponds to the powers of $\alpha_{\mathrm{s}}$
+    where $m$ corresponds to the powers of $\alpha_{s}$
     in the simplest Feynman diagrams you can draw
@@
     Since each term of this expansion is at a fixed-order
-    in $\alpha_{\mathrm{s}}$, it is possible to systematically
-    compute the diagrams which contribute at the given order of $\alpha_{\mathrm{s}}$.
+    in $\alpha_{s}$, it is possible to systematically
+    compute the diagrams which contribute at the given order of $\alpha_{s}$.
     The set of diagrams that contribute at the lowest
-    order in $\alpha_{\mathrm{s}}$ are $|\mathcal{M}|^{2}_{\mathrm{LO}}$,
+    order in $\alpha_{s}$ are $|\mathcal{M}|^{2}_{\mathrm{LO}}$,
     where LO stands for leading order. The next term
@@
     diagrams with an additional loop, or external leg,
-    which contributes an extra factor of $\alpha_{\mathrm{s}}$.
+    which contributes an extra factor of $\alpha_{s}$.
     This term is dubbed NLO for next-to-leading order
-    in $\alpha_{\mathrm{s}}$, and the following term
+    in $\alpha_{s}$, and the following term
     next-to-next-to-leading order has again an additional
     loop or leg. In general, we have terms
-    N$^{k}$LO where each additional power of $\alpha_{\mathrm{s}}$
+    N$^{k}$LO where each additional power of $\alpha_{s}$
     increases complixity of the computations, however,
-    because the expansion $\alpha_{\mathrm{s}}$
+    because the expansion $\alpha_{s}$
     is small, each additional term contribute less.

### 2022-09-15 qcd.tex
wrote: UV, DR, renormalisation, alpha running

--- 2022-08-30 qcd.tex
@@
 \begin{document}
-\chapter{Quantum chromodynamics}
+\chapter{Quantum chromodynamics in practice}
 \label{chapter:qcd}
-\section{Fixed order perturbation theory}
-\subsection{NLO QCD}
+    In the previous chapter we introduced the basic concepts
+    of QCD and setup the framework of calculating
+    cross-sections. Cross-sections are the most relevant
+    quantities to compute as they can be measured
+    experimentally, and are a property of the particles
+    being collided, rather than depending on the
+    specifics of the experimental procedure.
+    This provides a bridge to compare
+    theoretical predictions and experiment.
+    From (\ref{eqn:dsigma}, the partonic cross-section
+    is the matrix element integrated over the relevant
+    final-state phase-space, normalised by a flux factor.
+    The matrix elements are
+    calculated order-by-order in the strong coupling
+    $\alpha_{s}$, (\ref{eqn:matrix_element}),
+    where each order corresponds to a set of Feynman diagrams
+    that have the appropriate number of vertices,
+    as each QCD vertex brings along a factor of $\alpha_{s}$.
+
+    In this chapter we will discuss the problems that
+    arise when evaluating these Feynman diagrams and outline
+    some solutions that have been widely adopted to circumvent
+    these issues to provide real-world applicable predictions.
+    This chapter forms the theoretical foundations on which
+    the work of this thesis is built upon.
+
+    We will consider the divergent nature of matrix
+    elements and introduce the most widely used techniques
+    to tame these singularities.
+
 \section{Divergent structures}
-\section{Infrared factorisation}
+    We have seen that matrix element computations
+    result in evaluating Feynman diagrams. From
+    the QCD Feynman rules (ref here), it becomes
+    clear that integrating over all momentum modes
+    will give rise to singularities. The divergences
+    associated with high energy modes are ultraviolet
+    (UV) divergences. On the opposite end of the energy
+    spectrum, low energy modes also give rise to 
+    infrared (IR) divergences.
+
+    The de-facto methods to alleviate these
+    divergences are through regularisation
+    for UV divergences, and through subtraction
+    schemes for IR divergences. Both of these
+    methods will be discussed in this section.
+
+\subsection{Ultraviolet divergences}
+    During intermediate steps of calculations,
+    such as trying to compute loop diagrams,
+    we have to evaluate integrals of the form
+    \begin{equation}\label{eqn:loop_integral}
+        \int_{0}^{\Lambda} \dfrac{\mathrm{d}^{4}\ell}{(2\pi)^{4}}\dfrac{1}{\ell^{2}(\ell+p)^{2}} \sim \log{\Lambda} \, ,
+    \end{equation}
+    where a cutoff scale $\Lambda$ has been introduced to
+    capture the divergence as the loop momenta
+    $k \rightarrow \infty$. It is clear that the
+    integral diverges in this high energy limit,
+    hence the name ultraviolet divergence.
+    This UV divergence cannot influence the
+    experimentally observable quantities at experiments,
+    which are finite, because the final observable
+    should not depend on the intermediate steps taken
+    to get there.
+    Infinities associated with UV divergences
+    are removed via the process of renormalisation
+    which introduces counterterms in the Lagrangian
+    to exactly cancel these divergences.
+
+    Before we introduce these counterterms, we
+    first consider the procedure of regularisation
+    which makes these infinities explicitly manifest.
+    The most widely adopted method of regularisation
+    is dimensional regularisation (DR) (cite here) which
+    is based on the observation that the integral
+    (\ref{eqn:loop_integral}), which is carried
+    out in $d=4$ space-time dimensions, would be finite
+    if $d < 4$.
+
+    In DR we define $d = 4 - 2\epsilon$ such that
+    after Feynman parametrisation (\ref{eqn:loop_integral})
+    becomes
+    \begin{equation}\label{eqn:dim_reg}
+        \int \dfrac{\mathrm{d}^{d}\ell}{(2\pi)^{d}}\dfrac{1}{\ell^{2}(\ell+p)^{2}} = \dfrac{(-p^{2})^{-\epsilon}}{(4\pi)^{2-\epsilon}}\dfrac{\Gamma(1-\epsilon)^{2}}{\Gamma(2-2\epsilon)}\Gamma(\epsilon) \, ,
+    \end{equation}
+    where the divergence for $d = 4$, or equivalently
+    $\epsilon \rightarrow 0$, is now captured in the
+    Gamma function, $\Gamma(\epsilon)$. This becomes clear once
+    we expand $\Gamma(\epsilon)$ around small $\epsilon$
+    \begin{equation}\label{eqn:gamma_laurent}
+        \Gamma(\epsilon) = \dfrac{1}{\epsilon} - \gamma_{E} + \dfrac{1}{2}\left(\gamma_{E}^{2} + \dfrac{\pi^{2}}{6}\right)\epsilon + \mathcal{O}(\epsilon^{2}) \, ,
+    \end{equation}
+    where $\gamma_{E} \approx 0.577$ is the Euler-Mascheroni
+    constant, meaning the divergence is now regularised as a pole in
+    $\epsilon$.
+
+\subsection{Renormalisation}
+    With the UV divergence regularised by dimensional
+    regularisation, it is now possible to construct the forms
+    of the counterterms order-by-order in $\alpha_{s}$ to
+    explicitly cancel them.
+
+    A subtlety with adjusting the dimension of the theory
+    from $d=4$ to arbitrary dimensions is that the mass dimension
+    of the Lagrangian has to change accordingly to retain a
+    dimensionless action. We can account for this by
+    introducing an arbitrary scale $\mu_{R}$, the renormalisation
+    scale, into the gauge coupling
+    constant to modify the mass dimension of the Lagrangian.
+    For QCD this is done via the modification
+    \begin{equation}\label{eqn:dimensionful_coupling}
+        g_{s} \rightarrow g_{s}\mu_{R}^{\epsilon} \, .
+    \end{equation}
+    where it is understood the original gauge coupling is
+    dimensionless. The order of $\mu_{R}$ is determined by examining
+    the dimensions of the gauge and fermion fields
+    in the Lagrangian (\ref{eqn:L_QCD}).
+    We see that this introduces a scale dependence
+    on the coupling constant $\alpha_{s}$, a point we will return
+    to in \ref{sec:alpha_running}.
+    % In a calculation,
+    % intermediate steps will have a dependence on $\mu$,
+    % however, once the $\epsilon \rightarrow 0$ limit
+    % is taken the scale dependence drops out.
+
+    The inclusion of counterterms into the theory can be
+    thought of as renormalisations of the the bare fields
+    and couplings in (\ref{eqn:L_QCD}) to restore predictive
+    power to the theory. The renormalised fields and couplings
+    can be given as (cite here)
+    \begin{align}\label{eqn:renormlised_fields}
+        \psi_{\mathrm{bare}} &= \sqrt{Z_{2}} \psi_{R} \, , \nonumber \\
+        A_{\mathrm{bare}}^{\mu} &= \sqrt{Z_{3}} A_{R}^{\mu} \, ,\\
+        g_{s,\mathrm{bare}} &= Z_{g} \mu_{R}^{\epsilon} g_{s,R} \, , \nonumber
+    \end{align}
+    where it conventional to define $Z_{1} = Z_{g}Z_{2}\sqrt{Z_{3}}$
+    such that we can set $Z_{n} = 1 + \delta_{n}$ for $n \in \{1, 2, 3\}$.
+    In this way we can write the Lagrangian as
+    \begin{equation}
+        \mathcal{L}_{\mathrm{renorm}} = \mathcal{L}_{\mathrm{bare}} + \mathcal{L}_{\mathrm{c.t.}} \, ,
+    \end{equation}
+    where the counterterms appearing in $\mathcal{L}_{\mathrm{c.t.}}$
+    are determined by calculating $\delta_{n}$. There is freedom
+    in the choice of $\delta_{n}$ which gives rise to
+    different regularisation schemes. The minimal subtraction
+    (MS) scheme subtracts only the epsilon pole appearing
+    in loop integrals (\ref{eqn:gamma_laurent}). However, the most
+    commonly used regularisation scheme is the modified minimal
+    subtraction ($\overline{\mathrm{MS}}$) scheme which subtracts the
+    epsilon pole plus a universal constant appearing in all
+    loop integrals.
+
+    In an all orders calculation of an observable,
+    there would be no dependence on the renormalisation
+    scale as it is a remnant of the regularisation prescription
+    However, because in perturbative QCD observables
+    are calculated order-by-order, there will be a residual
+    dependence on the renormalisation scale stemming from
+    the missing higher order terms. The conventional way
+    that the community estimates the uncertainty associated
+    with this residual dependence is by carrying out
+    a scale variation. This involves evaluating the
+    relevant quantities at the nominal renormalisation
+    scale and again with the scale varied by a factor of
+    two above and below.
+
+\subsection{Running of the coupling constant}\label{sec:alpha_running}
+    The bare coupling should not depend on the actual
+    value of the renormalisation scale,
+    \begin{equation}\label{eqn:d_g_bare_d_mu}
+        \dfrac{\mathrm{d}g_{\mathrm{bare}}}{\mathrm{d}\mu_{R}} = 0 \, ,
+    \end{equation}
+    or said another way, the renormalisation process
+    is independent of the renormalisation scale.
+
+    The behaviour of $\alpha_{s}$
+    is governed by the Callan-Symanzik \cite{Callan:1970yg,Symanzik:1970rt}
+    $\beta$-function
+    \begin{equation}\label{eqn:beta_fn}
+        \mu^{2} \dfrac{\partial \alpha_{s}(\mu^{2})}{\partial \mu^{2}} = \beta(\alpha_{s}) \, ,
+    \end{equation}
+    where the $\beta$-function can be written as a perturbative
+    expansion in $\alpha_{s}$
+    \begin{equation}\label{eqn:beta_series}
+        -\beta(\alpha_{s}) = \alpha_{s}\sum_{n=0}^{\infty}\left(\dfrac{\alpha_{s}}{4\pi}\right)^{n+1} \beta_{n}\, 
+    \end{equation}
+    where the coefficients $\beta_{n}$ have been computed
+    up to fourth order \cite{Baikov:2016tgj,Luthe:2017ttg}.
+    The solution to (\ref{eqn:beta_fn}) to first order
+    is
+    \begin{equation}\label{eqn:1l_alpha}
+        \alpha_{s}(\mu^{2}) = \dfrac{1}{\frac{\beta_{0}}{4\pi}\log\left(\frac{\mu^{2}}{\Lambda_{\mathrm{QCD}}^{2}}\right)} \, ,
+    \end{equation}
+    where $\Lambda_{\mathrm{QCD}}$ is the QCD scale,
+    the scale at which $\alpha_{s}$ becomes
+    large enough that perturbation theory breaks down,
+    and $\beta_{0} = (11C_{A} - \frac{4}{3}T_{R}n_{f})/6\pi$.
+    In QCD, $C_{A}=3$ and $T_{R}=1/2$, meaning for
+    $n_{f} < 33/2$ the sign of $\beta_{0}$ is positive,
+    which is what we have observed in nature with six
+    flavours of quarks. This behaviour of positive
+    $\beta_{n}$ is also the case in all coefficients
+    computed to date, meaning that $\alpha_{s}$
+    decreases with increasing energy due to the minus sign
+    in (\ref{eqn:beta_series}). This property is known as
+    asymptotic freedom \cite{Gross:1973id,Politzer:1973fx},
+    and is justification for treating QCD perturbatively.
+    The running of the coupling constant has been
+    observed experimentally as illustrated in Figure~\ref{fig:alpha_s_running}.
+    \begin{figure}
+        \includegraphics{qcd/alpha_s_running.pdf}
+        \caption{The running of the strong coupling constant $\alpha_{s}$,
+        as determined by experiments, with QCD theory prediction in black.
+        Figure reference \cite{Workman:2022ynf}.}
+        \label{fig:alpha_s_running}
+    \end{figure}
+
+    Another consequence of the running coupling
+    is colour confinement, in which quarks and
+    gluons cannot exist as free particles. Due
+    to the increasing strength of coupling at low
+    energies, quarks and gluons are forced to
+    form composite, colourless particles. This means
+    that at low energies, perturbative QCD
+    is not an accurate description of nature,
+    and the physics of hadrons cannot be predicted
+    from perturbation theory.
+\subsection{Infrared divergences}
+\section{Factorisation of matrix elements}
+\subsection{Soft limits}
 \subsection{Collinear limits}
-\subsection{Soft limits}
+
 \section{Subtraction}

### 2022-09-16 intro.tex
start writing IR divergences

@@
     are accessed through the LHAPDF interface (cite here).
+
     The evolution of PDFs between factorisation scales is possible
     through the use of the Dokshitzer-Gribov-Lipatov-Altarelli-Parisi (DGLAP)
-    equations (cite here).
+    equations (cite here)
+    \begin{equation}\label{eqn:DGLAP}
+        \mu_{F}^{2} \dfrac{\partial}{\partial \mu_{F}^{2}}
+        \begin{pmatrix}
+            f_{q/h}(x,\mu_{F}) \\
+            f_{g/h}(x,\mu_{F})
+        \end{pmatrix} \\
+        = \dfrac{\alpha_{s}(\mu_{F})}{2\pi} \int_{x}^{1} \dfrac{\mathrm{d}z}{z}
+        \begin{pmatrix}
+            P_{qq}(\frac{x}{z}) & P_{qg}(\frac{x}{z}) \\
+            P_{gq}(\frac{x}{z}) & P_{gg}(\frac{x}{z})
+        \end{pmatrix}
+        \begin{pmatrix}
+            f_{q/h}(x,\mu_{F}) \\
+            f_{g/h}(x,\mu_{F})
+        \end{pmatrix} \, ,
+    \end{equation}
+    where $P_{ba}(\dfrac{x}{z})$ are splitting functions
+    representing parton $b$ emitting a parton $a$ carrying
+    a momentum fraction $z$. These splitting functions are calculable
+    as a power series in $\alpha_{s}$ where they have been computed
+    up to three-loops (cite here). At leading order (cite here) they are
+    \begin{align}\label{eqn:AP_kernels}
+        P^{(0)}_{qq}(z) &= C_{F}\left[\dfrac{1+z^{2}}{(1-z)_{+}} + \dfrac{3}{2}\delta(1-z)\right] \\
+        P^{(0)}_{qg}(z) &= T_{R}\left[z^{2} + (1-z)^{2}\right] \\
+        P^{(0)}_{gq}(z) &= C_{F}\left[\dfrac{1+(1-z)^{2}}{z^{2}}\right] \\
+        P^{(0)}_{gg}(z) &= 2C_{A}\left[\dfrac{z}{(1-z)_{+}}+\dfrac{1-z}{z}+z(1-z)\right] + \left(\dfrac{11C_{A}-4T_{R}n_{f}}{6}\right)\delta(1-z)
+    \end{align}
+    where the `$+$'-prescription has been used to regulate divergences
+    at $z = 1$.
+    \begin{equation}\label{eqn:plus_prescription}
+        \int_{0}^{1} \mathrm{d}z \; [g(z)]_{+}f(z) = \int_{0}^{1}\mathrm{d}z \; g(z)(f(z)-f(1)) \, ,
+    \end{equation}
+    where the function $f(z)$ cancels the divergence if it is
+    sufficiently smooth at $z=1$.
+    At leading order, there is a physical interpretation
+    of the splitting kernels as the probability of finding a parton
+    $a$ in parton $b$ carrying a fraction $x$ of the parent parton.
 

### 2022-09-16 qcd.tex
start writing IR divergences

@@
 
-\section{Divergent structures}
+\section{Divergent structures}\label{sec:divergences}
     We have seen that matrix element computations
@@
 
-\subsection{Ultraviolet divergences}
+\subsection{Ultraviolet divergences}\label{sec:UV}
     During intermediate steps of calculations,
@@
 
-\subsection{Renormalisation}
+\subsection{Renormalisation}\label{sec:renormalisation}
     With the UV divergence regularised by dimensional
@@
     or said another way, the renormalisation process
-    is independent of the renormalisation scale.
+    is independent of the renormalisation scale. The
+    consequence of this is that $\alpha_{s}$ has to
+    depends on the renormalisation scale.
 
@@
     where the coefficients $\beta_{n}$ have been computed
-    up to fourth order \cite{Baikov:2016tgj,Luthe:2017ttg}.
+    up to $\beta_{4}$ \cite{Baikov:2016tgj,Luthe:2017ttg}.
     The solution to (\ref{eqn:beta_fn}) to first order
@@
     large enough that perturbation theory breaks down,
-    and $\beta_{0} = (11C_{A} - \frac{4}{3}T_{R}n_{f})/6\pi$.
-    In QCD, $C_{A}=3$ and $T_{R}=1/2$, meaning for
-    $n_{f} < 33/2$ the sign of $\beta_{0}$ is positive,
+    and $\beta_{0} = (11C_{A} - 4T_{R}n_{f})/3$.
+    In QCD, $C_{A}=3$ and $T_{R}=\frac{1}{2}$, meaning for
+    $n_{f} < \frac{33}{2}$ the sign of $\beta_{0}$ is positive,
     which is what we have observed in nature with six
@@
     is not an accurate description of nature,
-    and the physics of hadrons cannot be predicted
+    and the behaviour of hadrons cannot be predicted
     from perturbation theory.
 \subsection{Infrared divergences}
+    In Section~\ref{sec:UV} we saw how divergences
+    stemming from high energy behaviour would occur
+    in loop diagrams. On the other end of the energy
+    spectrum we also have divergences from low energy
+    modes. Since they are related to low energy
+    behaviours they are called IR divergences. There
+    are two cases in which these divergences arise:
+    in loop integrals (virtual) and when an emission
+    of an extra particle becomes soft or collinear
+    (real).
+
+    Virtual divergences can be understood by looking
+    at the integral (\ref{eqn:loop_integral})
+
 \section{Factorisation of matrix elements}

### 2022-09-20 qcd.tex
wrote: IR divergences

@@
     spectrum we also have divergences from low energy
-    modes. Since they are related to low energy
-    behaviours they are called IR divergences. There
+    modes. Since these divergences emerge from low energy
+    behaviour they are called infrared (IR) divergences. There
     are two cases in which these divergences arise:
     in loop integrals (virtual) and when an emission
-    of an extra particle becomes soft or collinear
-    (real).
+    of an extra particle (c.f. splitting functions
+    \ref{eqn:AP_kernels}) has vanishing energy (soft),
+    or becomes parallel to an external colourful line (collinear).
 
-    Virtual divergences can be understood by looking
-    at the integral (\ref{eqn:loop_integral})
+    Virtual IR divergences can be understood by looking
+    at an integral of the form
+    \begin{equation}\label{eqn:virtual_IR_integral}
+        \int \dfrac{\mathrm{d}^{4}\ell}{(2\pi)^{4}} \dfrac{1}{\ell^{2}(\ell+p_{1})^{2}(\ell-p_{2})^{2}} \, ,
+    \end{equation}
+    which is encountered when calculating the one-loop
+    virtual correction to the gluon-quark-antiquark vertex.
+    It is clear that the denominator vanishes when $\ell \rightarrow 0$
+    or when either $(\ell + p_{1})^{2}$ or $(\ell - p_{2})^{2} \rightarrow 0$.
+    These situations correspond to the gluon in the loop propagator
+    going soft, or collinear to the external quark or antiquark.
+    These divergences are regulated in DR to give
+    \begin{equation}\label{eqn:virtual_IR_DR}
+    \end{equation}
+    where we see a double $\epsilon$ pole corresponding to the
+    soft and collinear divergence.
 
-\section{Factorisation of matrix elements}
-\subsection{Soft limits}
-\subsection{Collinear limits}
+    Real IR divergences are manifest when integrating
+    matrix elements over the phase-space of external state momenta.
+    Matrix elements can be written as functions of Mandelstam variables
+    \begin{equation}\label{eqn:mandelstam}
+        s_{ij} = (p_{i} + p_{j})^{2} \xrightarrow[\mathrm{limit}]{\mathrm{massless}} 2 E_{i}E_{j}(1-\cos \theta_{ij}) \, ,
+    \end{equation}
+    where $p_{i}$ and $p_{j}$ are the 4-momenta of partons $i$ and $j$
+    in the hard scattering.
+    When either of the partons beecome soft, $E_{i,j}\rightarrow 0$,
+    or when they go collinear to each other, $\theta_{ij} \rightarrow 0$,
+    the matrix element will diverge if $s_{ij}$ appears
+    in the demoninator. These scenarios correspond to an external
+    particle becoming unresolved.
 
-\section{Subtraction}
-\subsection{Catani-Seymour dipoles}
-\subsection{Antenna subtraction}
+    In Section~\ref{sec:renormalisation}, we removed UV divergences
+    through the use of counterterms to renormalise the theory, making
+    predictions from the theory physical.
+    IR divergences, on the other hand, arise even after the theory
+    has been renormalised. The resolution to IR divergences is given by the
+    Kinoshita-Lee-Nauenberg (KLN) theorem (cite here) which states
+    that at each order of perturbation theory, the virtual and real
+    contributions exactly cancel out, leaving only finite corrections.
+    This cancellation occurs because the infrared pole structure is
+    identical in the real and virtual corrections, but with an opposite sign.
+
+    At a detector of a collider experiment, there is a finite
+    energy resolution of the calorimeter. This means that
+    it is not physically possible to observe arbitrarily soft
+    partons. It is also not possible to distinguish two partons
+    at arbitrarily small angles from one parton with the combined
+    momenta. These limitations along with the KLN theorem
+    means that any physical observable that we wish to predict
+    with fixed-order perturbation theory must be insensitive to the
+    emissions of soft or collinear partons. Any observable obeying
+    this criteria is an infrared and collinear (IRC) safe observable.
+    This leads to the concept of jets (cite here) which are collimated
+    particles combined according to some jet algorithm.
+
+    In DR, the singularities from soft and collinear divergences
+    are manifest as $\epsilon$ poles which makes the cancellation
+    simple between the real and virtual parts. However, in practice, 
+    the phase-space integrals of the matrix elements are rarely carried
+    out analytically due to the high dimensionality of the integral.
+    Instead they are done numerically with the use
+    of Monte Carlo methods, which requires the integral to be in
+    integer dimensions. Therefore, numerical techniques
+    have been developed to deal with these IR divergences as well. The
+    most common technique, subtraction, will be discussed in
+    more detail in Section~\ref{sec:subtraction}.
+
+\section{Factorisation of matrix elements}\label{sec:me_factorisation}
+
+\subsection{Soft limits}\label{sec:me_soft}
+\subsection{Collinear limits}\label{sec:me_collinear}
+
+\section{Subtraction}\label{sec:subtraction}
+\subsection{Catani-Seymour dipoles}\label{sec:CS_dipoles}
+\subsection{Antenna}\label{sec:antenna_functions}
 \end{document}

### 2022-09-20 intro.tex
add references

--- 2022-09-14 intro.tex
@@
     anomalous magnetic moment of the electron, agreeing with experimental
-    measurements to more than 10 significant figures \cite{Aoyama:2017uqe}.
+    measurements to more than 10 significant figures \cite{Aoyama:2017uqe},
     and the discovery of the Higgs boson in 2012 by the ATLAS \cite{ATLAS:2012yve}
@@
     group with colour charge $C$ describes the interactions of the
-    strong force within the theory of quantum chronodynamics (QCD), which we
+    strong force within the theory of quantum chromodynamics (QCD), which we
     will elaborate more on in the next section. The second
@@
     weak force. Neutrinos are massless in the SM but this is not a requirement
-    of the model. In fact they have been observed to have mass (cittation here).
+    of the model. In fact they have been observed to have mass \cite{Super-Kamiokande:1998kpq,SNO:2002tuh}.
 
@@
     \subsection{The QCD Lagrangian}\label{sec:L_qcd}
-    Quantum chronodynamics (QCD) is the sector of the SM that
+    Quantum chromodynamics (QCD) is the sector of the SM that
     describes the strong interaction. 
@@
     the strong coupling constant becomes small, a
-    propery known as asymptotic freedom. A consequence
+    property known as asymptotic freedom. A consequence
     of this is that predictions made in QCD can be expressed
@@
     $Q \gg \Lambda_{\mathrm{QCD}}$, this term and higher order
-    terms are suppresed.
+    terms are suppressed.
     In fact, the factorisation equation has only been
-    proven for a few specific cases (DIS etc, cite here).
+    proven for a few specific cases \cite{Collins:2011zzd,Collins:1987pm,Collins:1989gx,Amoroso:2022eow}.
 
@@
     instead of calculated in perturbation theory.
-    Multiple groups release PDF fits (cite groups here).
-    In practice, these PDF fits
-    are accessed through the LHAPDF interface (cite here).
+    For a review of PDFs and how they are determined
+    see Ref. \cite{Ethier:2020way,Jimenez-Delgado:2013sma}.
+    In practice, these PDF fits are accessed through the
+    LHAPDF interface \cite{Buckley:2014ana} which provides
+    PDF sets from multiple working groups \cite{PDF4LHCWorkingGroup:2022cjn,Hou:2019efy,Bailey:2020ooq,NNPDF:2021njg}.
     The evolution of PDFs between factorisation scales is possible
     through the use of the Dokshitzer-Gribov-Lipatov-Altarelli-Parisi (DGLAP)
-    equations (cite here).
+    equations \cite{Dokshitzer:1977sg,Gribov:1972ri,Lipatov:1974qm,Altarelli:1977zs}.
 
@@
     $\hat{\sigma}_{ab \rightarrow n}(Q, \mu_{\mathrm{F}}, \mu_{\mathrm{R}})$
-    is the partonic cross-section which descibes the interaction of
+    is the partonic cross-section which describes the interaction of
     partons $a$ and $b$ scattering into $n$ particles,
@@
     case of two particles colliding into $n$ particles.
-    To mathematically describe these
-    scattering processes, scattering amplitudes are used.
+    Scattering amplitudes are used to mathematically
+    describe these scattering processes.
     For an initial-state
@@
     the soft energy effects into the non-perturbative PDFs,
-    which are determined once and for all for all
+    which are determined once and for all
     processes, the object of interest now becomes the
@@
     N$^{k}$LO where each additional power of $\alpha_{\mathrm{s}}$
-    increases complixity of the computations, however,
+    increases complexity of the computations, however,
     because the expansion $\alpha_{\mathrm{s}}$
@@
     Monte Carlo event generators (MCEG). These event
-    generators are multi-purpose tools that
+    generators are multipurpose tools that
     incorporate many aspects of the theory discussed

### 2022-09-20 intro.tex

--- 2022-09-16 intro.tex
@@
     instead of calculated in perturbation theory.
-    Multiple groups release PDF fits (cite groups here).
-    In practice, these PDF fits
-    are accessed through the LHAPDF interface (cite here).
 
+    For a review of PDFs and how they are determined
+    see Ref. \cite{Ethier:2020way,Jimenez-Delgado:2013sma}.
+    In practice, these PDF fits are accessed through the
+    LHAPDF interface \cite{Buckley:2014ana} which provides
+    PDF sets from multiple working groups \cite{PDF4LHCWorkingGroup:2022cjn,Hou:2019efy,Bailey:2020ooq,NNPDF:2021njg}.
     The evolution of PDFs between factorisation scales is possible
     through the use of the Dokshitzer-Gribov-Lipatov-Altarelli-Parisi (DGLAP)
-    equations (cite here)
+    equations \cite{Dokshitzer:1977sg,Gribov:1972ri,Lipatov:1974qm,Altarelli:1977zs}.
     \begin{equation}\label{eqn:DGLAP}

### 2022-09-20 qcd.tex

@@
 \section{Factorisation of matrix elements}\label{sec:me_factorisation}
+    In the previous section we examined the IR divergences
+    that can arise in matrix elements, either explicitly
+    through loop diagrams, or implicitly when carrying
+    out the phase-space integral over the external state
+    momenta. In this section we will discuss how these IR
+    singularities are universal and can be factorised out
+    of the matrix element. This property is exploited
+    extensively in Sections ... where the research of this
+    thesis is presented.
 

### 2022-09-21 intro.tex
wrote: matrix element factorisation

@@
 
-    {\color{red} Need to write down the Feynman
-    rules and draw diagrams. Best to put it in a
-    figure that we can refer back to.}
+    \begin{figure}\label{fig:qcd_feynman_rules}
+        \includegraphics[width=\linewidth]{introduction/qcd_feynman_rules.pdf}
+        \caption{QCD Feynman rules.}
+    \end{figure}
 
+    In Figure~\ref{fig:qcd_feynman_rules} we have
+    collected the QCD Feynman rules.
     Note that we have only written down the vertices,
@@
     as a power series in $\alpha_{s}$ where they have been computed
-    up to three-loops (cite here). At leading order (cite here) they are
+    up to three-loops \cite{Moch:2004pa,Vogt:2004mw}.
+    At leading order \cite{Altarelli:1977zs} they are
     \begin{align}\label{eqn:AP_kernels}
@@
     loop or leg. In general, we have terms
-    N$^{k}$LO where each additional power of $\alpha_{s}$
+    N$^{\mathrm{k}}$LO where each additional power of $\alpha_{s}$
     increases complixity of the computations, however,
@@
     cross-sections complete, we will move the discussion onto
-    Monte Carlo event generators (MCEG). These event
+    Monte Carlo event generators. These event
     generators are multipurpose tools that

### 2022-09-21 qcd.tex
wrote: matrix element factorisation

@@
     theoretical predictions and experiment.
-    From (\ref{eqn:dsigma}, the partonic cross-section
+    From (\ref{eqn:dsigma}), the partonic cross-section
     is the matrix element integrated over the relevant
@@
     During intermediate steps of calculations,
-    such as trying to compute loop diagrams,
+    such as trying to compute loop diagrams seen in
+    the Figure~\ref{fig:bubble_diagram},
     we have to evaluate integrals of the form
     \begin{equation}\label{eqn:loop_integral}
-        \int_{0}^{\Lambda} \dfrac{\mathrm{d}^{4}\ell}{(2\pi)^{4}}\dfrac{1}{\ell^{2}(\ell+p)^{2}} \sim \log{\Lambda} \, ,
+        I_{\mathrm{UV}} = \int_{0}^{\Lambda} \dfrac{\mathrm{d}^{4}\ell}{(2\pi)^{4}}\dfrac{1}{\ell^{2}(\ell+p)^{2}} \sim \log{\Lambda} \, ,
     \end{equation}
@@
 
+    \begin{figure}
+        \includegraphics[scale=1.0]{qcd/bubble_diagram.pdf}
+        \caption{Bubble diagram}
+        \label{fig:bubble_diagram}
+    \end{figure}
+
     Before we introduce these counterterms, we
@@
     The most widely adopted method of regularisation
-    is dimensional regularisation (DR) (cite here) which
+    is dimensional regularisation (DR) \cite{tHooft:1972tcz} which
     is based on the observation that the integral
@@
     \begin{equation}\label{eqn:dim_reg}
-        \int \dfrac{\mathrm{d}^{d}\ell}{(2\pi)^{d}}\dfrac{1}{\ell^{2}(\ell+p)^{2}} = \dfrac{(-p^{2})^{-\epsilon}}{(4\pi)^{2-\epsilon}}\dfrac{\Gamma(1-\epsilon)^{2}}{\Gamma(2-2\epsilon)}\Gamma(\epsilon) \, ,
+        % \int \dfrac{\mathrm{d}^{d}\ell}{(2\pi)^{d}}\dfrac{1}{\ell^{2}(\ell+p)^{2}} = \dfrac{\mathrm{i}(-p^{2})^{-\epsilon}}{(4\pi)^{2-\epsilon}}\dfrac{\Gamma(1-\epsilon)^{2}}{\Gamma(2-2\epsilon)}\Gamma(\epsilon) \, ,
+        I_{\mathrm{UV}} = \int \dfrac{\mathrm{d}^{d}\ell}{(2\pi)^{d}}\dfrac{1}{\ell^{2}(\ell+p)^{2}} = \dfrac{\mathrm{i}}{(4\pi)^{2}}\left(\dfrac{-p^{2}}{4\pi}\right)^{-\epsilon}\dfrac{\Gamma(1-\epsilon)^{2}}{\Gamma(2-2\epsilon)}\Gamma(\epsilon) \, ,
     \end{equation}
@@
     power to the theory. The renormalised fields and couplings
-    can be given as (cite here)
+    can be given as (see e.g. Ref.~\cite{mangano1999introduction})
     \begin{align}\label{eqn:renormlised_fields}
@@
     from perturbation theory.
+
 \subsection{Infrared divergences}
@@
     of an extra particle (c.f. splitting functions
-    \ref{eqn:AP_kernels}) has vanishing energy (soft),
+    (\ref{eqn:AP_kernels})) has vanishing energy (soft),
     or becomes parallel to an external colourful line (collinear).
@@
     \begin{equation}\label{eqn:virtual_IR_integral}
-        \int \dfrac{\mathrm{d}^{4}\ell}{(2\pi)^{4}} \dfrac{1}{\ell^{2}(\ell+p_{1})^{2}(\ell-p_{2})^{2}} \, ,
+        I_{v} = \int \dfrac{\mathrm{d}^{4}\ell}{(2\pi)^{4}} \dfrac{1}{\ell^{2}(\ell+p_{1})^{2}(\ell-p_{2})^{2}} \, ,
     \end{equation}
     which is encountered when calculating the one-loop
-    virtual correction to the gluon-quark-antiquark vertex.
+    virtual correction to the gluon-quark-antiquark vertex
+    as shown in Figure~\ref{fig:vertex_correction}.
     It is clear that the denominator vanishes when $\ell \rightarrow 0$
@@
     \begin{equation}\label{eqn:virtual_IR_DR}
+        I_{v} = \dfrac{\mathrm{i}}{(4\pi)^{2}Q^{2}}\left(\dfrac{-Q^{2}}{4\pi}\right)^{-\epsilon}\dfrac{\Gamma(1+\epsilon)\Gamma(1-\epsilon)^{2}}{\Gamma(1-2\epsilon)}\left[\dfrac{1}{\epsilon^{2}}\right] \, ,
     \end{equation}
-    where we see a double $\epsilon$ pole corresponding to the
-    soft and collinear divergence.
+    where $Q = (p_{1} + p_{2})$. There is a double $\epsilon$ pole manifest,
+    corresponding to the soft and collinear divergence.
+
+    \begin{figure}
+        \includegraphics[scale=1.0]{qcd/vertex_correction.pdf}
+        \caption{Vertex correction}
+        \label{fig:vertex_correction}
+    \end{figure}
 
@@
     has been renormalised. The resolution to IR divergences is given by the
-    Kinoshita-Lee-Nauenberg (KLN) theorem (cite here) which states
+    Block-Nordsieck \cite{Bloch:1937pw,Yennie:1961ad} and Kinoshita-Lee-Nauenberg
+    (KLN) theorems \cite{Kinoshita:1962ur,Lee:1964is} which states
     that at each order of perturbation theory, the virtual and real
@@
     this criteria is an infrared and collinear (IRC) safe observable.
-    This leads to the concept of jets (cite here) which are collimated
+    This leads to the concept of jets \cite{Salam:2010nqg} which are collimated
     particles combined according to some jet algorithm.
@@
     momenta. In this section we will discuss how these IR
-    singularities are universal and can be factorised out
+    singularities are universal and how the divergences
+    associated with real emissions can be factorised out
     of the matrix element. This property is exploited
-    extensively in Sections ... where the research of this
-    thesis is presented.
+    extensively in Chapters~\ref{chapter:fame1}, \ref{chapter:fame2},
+    and \ref{chapter:fame3}
+    where the research of this thesis is presented.
+
+    The regions of phase-space in which real emission
+    matrix elements diverge are when the emission is
+    soft and/or collinear. It can be shown that in these
+    regions of phase-space, the matrix element factorises
+    into a process-independent singular factor, multiplied
+    by a matrix element with the unresolved parton
+    being absorbed by an emitting leg (reduced matrix
+    element). This factorisation, however, is not exact
+    in QCD where there are spin and colour correlations.
+
+    Schematically, the $(n+1)$-body matrix element
+    factorises as
+    \begin{equation}\label{eqn:me_factorisation}
+        |\mathcal{M}_{n+1}|^{2} \rightarrow S_{ijk} \otimes |\mathcal{M}_{n}|^{2} \, ,
+    \end{equation}
+    where $S_{ijk}$ is a universal singular factor capturing
+    the IR divergent behaviour and $|\mathcal{M}_{n}|^{2}$
+    is the reduced matrix element. The $\otimes$
+    represents the colour and spin correlations existing
+    between the singular function and the reduced matrix element.
+    This singular factor
+    is not unique and can be represented by different
+    approximations as long as they reproduce the
+    correct IR behaviour. In this thesis we discuss
+    two approximations: Catani-Seymour dipoles \cite{Catani:1996vz}
+    in Section~\ref{sec:CS_dipoles} and antenna functions \cite{Gehrmann-DeRidder:2005btv}
+    in Section~\ref{sec:antenna_functions}. The indices
+    of $S_{ijk}$ already hints at the fact that the singular
+    functions only depends on three partons, and not all
+    of the $n+1$ external states.
+
+    We will inspect the soft and collinear limits separately
+    to see how the matrix element factorises in these respective
+    limits.
+    The following section will follow the conventions
+    of Ref.~\cite{Catani:1996vz}, namely that the matrix
+    element for $n$ partons in the final-state is written as
+    \begin{equation}\label{eqn:CS_matrix_element}
+        |\mathcal{M}_{n}|^{2} = {}_{n}\langle 1, \ldots, n | 1, \ldots, n \rangle_{n}
+    \end{equation}
+    where $| 1, \ldots, n \rangle_{n}$ is a vector
+    in colour + helicity space (see Appendix~\ref{appendix:catani_seymour}
+    for a more thorough explanation on notation).
 
 \subsection{Soft limits}\label{sec:me_soft}
+    Consider a tree-level matrix element $|\mathcal{M}_{n+1}|^{2}$
+    with a final-state gluon $j$. Note that reduced matrix
+    elements associated with taking a quark soft have to vanish
+    due to violation of quark number, therefore only gluons
+    are considered.
+    The limit of the soft gluon with momentum
+    $p_{j}$ can be parametrised by
+    \begin{equation}\label{eqn:soft_parametrisation}
+        p_{j}^{\mu} = \lambda q^{\mu} \, , \quad \lambda \rightarrow 0 \, ,
+    \end{equation}
+    where $q^{\mu}$ is an arbitrary four-vector and $\lambda$
+    is a scale parameter. In this limit the matrix element
+    can be written as
+    \begin{equation}\label{eqn:soft_factorisation}
+        |\mathcal{M}_{n+1}|^{2} \rightarrow -\dfrac{1}{\lambda^{2}}8\pi \mu^{2\epsilon}\alpha_{s} \sum_{i}\dfrac{1}{p_{i}q}\sum_{k \neq i} \dfrac{p_{k}p_{i}}{(p_{i} + p_{k})q} {}_{n}\langle 1, \ldots, n+1 | \bm{T}_{k} \cdot \bm{T}_{i} | 1, \ldots, n+1 \rangle_{n} \, ,
+    \end{equation}
+    where terms less singular than $1 / \lambda^{2}$ have been
+    neglected. The indices $i$, $j$, and $k$ label partons involved
+    in the factorsation process: $i$ is the emitter parton,
+    $j$ is the emitted parton, and $k$ is a parton accounting
+    for the colour correlations. The scale $\mu$ can be identified as the
+    renormalisation scale, and $\bm{T_{i}}$ is the colour-charge
+    operator. It acts on the colour vectors to give factors of
+    $C_{A}$ and $C_{F}$, hence this factorisation is not exact.
+
+    The matrix elements in (\ref{eqn:soft_factorisation}) are
+    unambiguously defined only when momentum conservation is
+    fulfilled. This is only true in the strict $\lambda = 0$
+    limit. Away from the limit, care has to be taken to conserve
+    momentum conservation and keep the relevant partons on-shell.
+    This can be done through the use of momentum mappings
+    \cite{Catani:1996vz,Kosower:1997zr}.
+
 \subsection{Collinear limits}\label{sec:me_collinear}
+    For the collinear limit, consider partons $i$ and
+    $j$ in the matrix element $|\mathcal{M}_{n+1}|^{2}$.
+    Their momenta can be decomposed as
+    \begin{align}\label{eqn:collinear_momenta}
+        \begin{split}
+        p_{i}^{\mu} &= z p^{\mu} + k_{\perp}^{\mu} - \dfrac{k_{\perp}^{2}}{z} \dfrac{n^{\mu}}{2 p n} \, \\
+        p_{j}^{\mu} &= (1-z) p^{\mu} - k_{\perp}^{\mu} - \dfrac{k_{\perp}^{2}}{1-z}\dfrac{n^{\mu}}{2 p n} \, , 
+        \end{split}
+    \end{align}
+    where $p^{\mu}$ denotes the collinear direction
+    of the two partons, $k_{\perp}^{\mu}$ specifies
+    the transverse direction perpendicular to the
+    collinear direciton, and $n^{\mu}$ is an auxiliary
+    vector satisfying the conditions $n^{2} = 0$ and $k_{\perp} n = 0$.
+    $z$ is the fraction of momenta carried away from
+    the collinear momentum by parton $i$.
 
+    The collinear limit can be then be defined as
+    \begin{equation}\label{eqn:collinear limit}
+        2 p_{i} p_{j} = - \dfrac{k_{\perp}^{2}}{z(1-z)} \, , \quad k_{\perp} \rightarrow 0 \, .
+    \end{equation}
 \section{Subtraction}\label{sec:subtraction}

### 2022-09-23 appendix1.tex
wrote: matrix element factorisation
wrote: colour-ordered and one-loop factorisation

--- 2022-08-30 qcd.tex
@@
 \begin{document}
-\chapter{Quantum chromodynamics}
-\label{chapter:qcd}
-\section{Fixed order perturbation theory}
-\subsection{NLO QCD}
-\section{Divergent structures}
-\section{Infrared factorisation}
-\subsection{Collinear limits}
-\subsection{Soft limits}
-\section{Subtraction}
-\subsection{Catani-Seymour dipoles}
-\subsection{Antenna subtraction}
+\chapter{Catani-Seymour notation}
+\label{appendix:catani_seymour}
+    The notation for matrix elements given in Ref.~\cite{Catani:1996vz}
+    is repeated here for convenience.
+
+    Consider a tree-level matrix element with $n$ partons in
+    the final-state
+    \begin{equation}\label{eqn:CS_ME}
+        \mathcal{M}^{c_{1},\ldots,c_{n};s_{1},\ldots,s_{n}}_{n}(p_{1},\ldots,p_{n}) \, ,
+    \end{equation}
+    where $\{c_{1},\ldots,c_{n}\}$ are colour indices,
+    $\{s_{1},\ldots,s_{n}\}$ are spin indices,
+    and $\{p_{1},\ldots,p_{n}\}$ are momenta.
+
+    Introducing the basis $\{| c_{1},\ldots,c_{n} \rangle \otimes | s_{1},\ldots,s_{n} \rangle \}$
+    in colour + helicity space allows us to write
+    \begin{equation}\label{eqn:CS_hilbert}
+        \mathcal{M}^{c_{1},\ldots,c_{n};s_{1},\ldots,s_{n}}_{n}(p_{1},\ldots,p_{n}) \equiv \left(\langle c_{1},\ldots,c_{n} | \otimes \langle s_{1},\ldots,s_{n} | \right) | 1, \ldots,n \rangle_{n} \, ,
+    \end{equation}
+    where $|1, \ldots, n \rangle_{n}$ is a vector in colour + helicity space.
+    The indices inside the vector denote the spin and colour indices,
+    while the subscript denotes the multiplicity of the vector.
+    The matrix element squared, summed over final-state colours and spins,
+    can therefore be written as
+    \begin{equation}\label{eqn:CS_ME2}
+        |\mathcal{M}_{n}|^{2} = {}_{n} \langle 1,\ldots,n|1,\ldots,n\rangle_{n} \, .
+    \end{equation}
+
+    In the case of partons in the initial-state the colour + helicity vector
+    $|1, \ldots, n \rangle$ has to be normalised by a factor $\sqrt{n_{c}}$ for each
+    initial state parton. For hadron-hadron collisions this amounts to
+    \begin{equation}
+        | 1, \ldots, n \rangle_{n} \rightarrow \dfrac{1}{\sqrt{n_{c}(a)n_{c}(b)}} | 1, \ldots, n \rangle_{n} \, ,
+    \end{equation}
+    where $n_{c}(a)$ denotes the number of colour states of parton $a$.
+    Namely $n_{c}(q) = n_{c}(\bar{q}) = N_{c}$ and $n_{c}(g) = N_{c}^{2}-1$.
+
 \end{document}

### 2022-09-23 qcd.tex
wrote: colour-ordered and one-loop factorisation

@@
     observed experimentally as illustrated in Figure~\ref{fig:alpha_s_running}.
+
     \begin{figure}
@@
     \begin{equation}\label{eqn:me_factorisation}
-        |\mathcal{M}_{n+1}|^{2} \rightarrow S_{ijk} \otimes |\mathcal{M}_{n}|^{2} \, ,
+        |\mathcal{M}_{n+1}|^{2} \rightarrow \bm{S}_{ijk} \otimes |\mathcal{M}_{n}|^{2} \, ,
     \end{equation}
-    where $S_{ijk}$ is a universal singular factor capturing
+    where $\bm{S}_{ijk}$ is a universal singular factor capturing
     the IR divergent behaviour and $|\mathcal{M}_{n}|^{2}$
@@
     in Section~\ref{sec:antenna_functions}. The indices
-    of $S_{ijk}$ already hints at the fact that the singular
+    of $\bm{S}_{ijk}$ already hints at the fact that the singular
     functions only depends on three partons, and not all
@@
     The following section will follow the conventions
-    of Ref.~\cite{Catani:1996vz}, namely that the matrix
-    element for $n$ partons in the final-state is written as
+    of Ref.~\cite{Catani:1996vz}, namely that the colour and
+    helicity summed $n$-body matrix element written as
     \begin{equation}\label{eqn:CS_matrix_element}
@@
     \begin{equation}\label{eqn:soft_factorisation}
-        |\mathcal{M}_{n+1}|^{2} \rightarrow -\dfrac{1}{\lambda^{2}}8\pi \mu^{2\epsilon}\alpha_{s} \sum_{i}\dfrac{1}{p_{i}q}\sum_{k \neq i} \dfrac{p_{k}p_{i}}{(p_{i} + p_{k})q} {}_{n}\langle 1, \ldots, n+1 | \bm{T}_{k} \cdot \bm{T}_{i} | 1, \ldots, n+1 \rangle_{n} \, ,
+        |\mathcal{M}_{n+1}|^{2} \rightarrow -\dfrac{1}{\lambda^{2}}8\pi \mu^{2\epsilon}\alpha_{s} \sum_{i}\dfrac{1}{p_{i}q}\sum_{k \neq i} \dfrac{p_{k}p_{i}}{(p_{i} + p_{k})q} \; {}_{n}\langle 1, \ldots, n+1 | \bm{T}_{k} \cdot \bm{T}_{i} | 1, \ldots, n+1 \rangle_{n} \, ,
     \end{equation}
     where terms less singular than $1 / \lambda^{2}$ have been
-    neglected. The indices $i$, $j$, and $k$ label partons involved
+    neglected. The reduced matrix element is obtained by removing
+    the soft gluon. The indices $i$, $j$, and $k$ label partons involved
     in the factorsation process: $i$ is the emitter parton,
@@
     Their momenta can be decomposed as
-    \begin{align}\label{eqn:collinear_momenta}
+    \begin{align}\label{eqn:F_collinear_momenta}
         \begin{split}
@@
     the transverse direction perpendicular to the
-    collinear direciton, and $n^{\mu}$ is an auxiliary
+    collinear direciton ($p k_{\perp} = 0$), and $n^{\mu}$ is an auxiliary
     vector satisfying the conditions $n^{2} = 0$ and $k_{\perp} n = 0$.
@@
     The collinear limit can be then be defined as
-    \begin{equation}\label{eqn:collinear limit}
-        2 p_{i} p_{j} = - \dfrac{k_{\perp}^{2}}{z(1-z)} \, , \quad k_{\perp} \rightarrow 0 \, .
+    \begin{equation}\label{eqn:F_collinear limit}
+        2 p_{i} p_{j} = s_{ij} = - \dfrac{k_{\perp}^{2}}{z(1-z)} \, , \quad k_{\perp} \rightarrow 0 \, .
     \end{equation}
+    In this limit the matrix element can be written as
+    \begin{equation}\label{eqn:F_collinear_factorisation}
+        |\mathcal{M}_{n+1}|^{2} \rightarrow \dfrac{1}{p_{i}p_{j}}4\pi \mu^{2\epsilon}\alpha_{s} \; {}_{n} \langle 1, \ldots, n+1 | \hat{P}_{ij}(z, k_{\perp}) | 1, \ldots, n+1 \rangle_{n} \, ,
+    \end{equation}
+    where terms less singular than $k_{\perp}^{2}$ have been neglected.
+    This reduced matrix element is obtained by replacing the partons
+    $i$ and $j$ with a single parton $ij$ which carries the
+    momentum $p^{\mu}$ and suitable quantum numbers depending
+    on the partons $i$ and $j$. For example, if $i =$ quark and $j=$ gluon,
+    then $ij =$ quark, or if $i=$ quark and $j =$ antiquark, then $ij =$ gluon.
+    $\hat{P}_{ij}$ are the $d$-dimensional Altarelli-Parisi splitting
+    functions that depend on the momentum fraction $z$ for the splitting
+    $ij \rightarrow i + j$ and the transverse momentum $k_{\perp}$.
+    Each splitting function is matrix acting on the spin indices of $ij$.
+    Due to these spin correlations the matrix
+    element does not factorise from the splitting functions exactly.
+
+    $\hat{P}_{ij}$ become the more recognisable Altarelli-Parisi
+    splitting functions (\ref{eqn:AP_kernels}) once spin-averaged
+    and the $\epsilon \rightarrow 0$ limit is taken.
+
+    In general, it is possible for the case in which a final-state
+    parton $i$ becomes collinear with an initial-state parton $a$.
+    This is described by the splitting process $a \rightarrow ai + i$.
+    In this case their momenta can be decomposed as
+    \begin{align}\label{eqn:I_collinear_momenta}
+        \begin{split}
+        p_{i}^{\mu} &= (1-x)p_{a}^{\mu} + k_{\perp}^{\mu} - \dfrac{k_{\perp}^{2}}{1-x}\dfrac{n^{\mu}}{2p_{a} n} \, , \\
+        p_{a}^{\mu} &= x p_{a}^{\mu} \, ,
+        \end{split}
+    \end{align}
+    with the collinear limit defined as
+    \begin{equation}\label{eqn:I_collinear_limit}
+        2p_{i}p_{a} = s_{ia} = -\dfrac{k_{\perp}^{2}}{1-x} \, , \quad k_{\perp} \rightarrow 0 \, .
+    \end{equation}
+
+    The analogous expression of (\ref{eqn:F_collinear_factorisation})
+    for initial-state splitting is
+    \begin{equation}\label{eqn:I_collinear_factorisation}
+        |\mathcal{M}_{n+1}|^{2} \rightarrow \dfrac{1}{x} \dfrac{1}{p_{i}p_{a}} 4\pi \mu^{2\epsilon} \alpha_{s} \; {}_{n} \langle 1,\ldots,n+1;ai, \ldots | \hat{P}_{ai}(x, k_{\perp}) | 1,\ldots,n+1;ai,\ldots \rangle_{n} \, ,
+    \end{equation}
+    where the replacement of partons $a$ and $i$ have been made
+    explicit by the presence of the parton $ai$ in the colour + helicity vector.
+    The specific type of parton $ai$ depends upon the types of $a$ and $i$.
+
+    Similar to (\ref{eqn:soft_factorisation}), (\ref{eqn:F_collinear_factorisation})
+     and (\ref{eqn:I_collinear_factorisation}) are only
+    true in the strict collinear limit. Away from these
+    limits care has to be taken to conserve momenta
+    via the use of momenta mappings.
+
+\subsection{Factorisation of colour-ordered amplitudes}
+    The factorisation expressed in (\ref{eqn:soft_factorisation}),
+    (\ref{eqn:F_collinear_factorisation}), and (\ref{eqn:I_collinear_factorisation})
+    were not exact because of colour and spin correlations.
+
+    It can be shown that the factorisation of matrix
+    elements in the soft and collinear limits becomes
+    exact once the colour structure of the gauge group
+    is separated from the kinematics.
+
+    For the scope of this thesis it is sufficient
+    to consider colour decomposition of electron-positron
+    annihilation. In general any QCD amplitude can be
+    colour decomposed. Consider the process $e^{+}e^{-} \rightarrow q\bar{q} + n g$,
+    the amplitude can be written as a product of hadronic and
+    leptonic currents \cite{Campbell:1998nn} (on which this discussion is based on)
+    \begin{equation}\label{eqn:epem_amplitude}
+        \mathcal{M}(q_{1}, \bar{q}_{2}; 1, \ldots, n) = \hat{\mathcal{S}}_{\mu}^{n+2}(q_{1}; 1,\ldots,n; \bar{q}_{2})V^{\mu} \, ,
+    \end{equation}
+    where the leptonic current is $V^{\mu}$ and the hadronic
+    current is
+    \begin{equation}\label{eqn:hadronic_current}
+        \hat{\mathcal{S}}_{\mu}^{n+2}(q_{1};1,\ldots,n;\bar{q}_{2}) = \mathrm{i}eg_{s}^{n} \sum_{P(1,\ldots,n)}(T^{a_{1}} \ldots T^{a_{n}})_{c_{1}c_{2}}S_{\mu}(q_{1};1,\ldots,n;\bar{q}_{2}) \, .
+    \end{equation}
+    In this expression we have the electromagnetic gauge coupling
+    $e$ and the strong gauge coupling $g_{s}$ appearing. The colour
+    structure has been factored into a product of fundamental group
+    generators where the indices $a_{i} \in {1,\ldots, N_{c}^{2}-1}$
+    and $c_{i} \in {1,\ldots,N_{c}}$. This leaves the colour-ordered
+    partial amplitude $S_{\mu}(q_{1};1,\ldots,n;\bar{q}_{2})$ depending
+    only on kinematical variables, and where the gluons are emitted in an
+    ordered fashion from the quarks. The sum over $P(1,\ldots,n)$
+    represents a sum over all permutations of gluon emissions
+    which accounts for all Feynman diagrams and colour structures.
+
+    Upon squaring the amplitude (\ref{eqn:epem_amplitude}), we get
+    \begin{equation}\label{eqn:epem_me2}
+        \left|\hat{\mathcal{S}}_{\mu}^{n+2}V^{\mu}\right|^{2} = e^{2}\left(\dfrac{g_{s}^{2}N}{2}\right)^{n}\left(\dfrac{N_{c}^{2}-1}{N_{c}}\right) \sum_{P(1,\ldots,n)} \left(\left|S_{\mu}(q_{1};1,\ldots,n;\bar{q}_{2})V^{\mu}\right|^{2} + \mathcal{O}\left(\dfrac{1}{N_{c}^{2}}\right)\right) \, ,
+    \end{equation}
+    where the subleading colour terms have been omitted.
+
+    With the amplitude written in terms of the colour-ordered
+    partial amplitudes, it is now possible to factorise (\ref{eqn:epem_me2})
+    exactly in the soft and collinear limits. In the limit where gluon
+    $j$ is soft, we have
+    \begin{equation}\label{eqn:CO_soft_factorisation}
+        \left|S_{\mu}(q_{1};1,\ldots,i,j,k,\ldots,n;\bar{q}_{2})V^{\mu}\right|^{2} \rightarrow S_{ijk} \left|S_{\mu}(q_{1};1,\ldots,i,k,\ldots,n;\bar{q}_{2})V^{\mu}\right|^{2} \, ,
+    \end{equation}
+    where the factor
+    \begin{equation}\label{eqn:eikonal_factor}
+        S_{ijk} = 4\dfrac{s_{ik}}{s_{ij}s_{jk}} \, ,
+    \end{equation}
+    is the well-known eikonal factor. We see that the colour-ordered
+    amplitude on the RHS of (\ref{eqn:eikonal_factor}) has gluon $j$
+    removed but the ordering of all hard partons remain unchanged.
+
+    In the collinear limit where partons $i$ and $j$ become collinear
+    to form parton $k$, the matrix element factorises as
+    \begin{equation}\label{eqn:CO_collinear_factorisation}
+        \left|S_{\mu}(q_{1};1,\ldots,i,j,\ldots,n;\bar{q}_{2})V^{\mu}\right|^{2} \rightarrow \dfrac{2}{s_{ij}}P_{ij}(z) \left|S_{\mu}(q_{1};1,\ldots,k,\ldots,n;\bar{q}_{2})V^{\mu}\right|^{2} \, ,
+    \end{equation}
+    where $P_{ij}(z)$ are the spin-averaged Altarelli-Parisi splitting
+    functions (\ref{eqn:AP_kernels}) with the colour factors removed.
+    In this expression it is understood that the indices $i$, $j$, and $k$
+    have to be self-consistent for the different partonic splittings.
+    For partons which are not colour connected (are not neighbouring
+    partons in the colour ordered amplitude), there will be no singular
+    behaviour as $s_{ij} \rightarrow 0$.
+
+    In (\ref{eqn:CO_soft_factorisation}) and (\ref{eqn:CO_collinear_factorisation})
+    the colour-ordered amplitude factorises exactly into a
+    colour-ordered amplitude with one parton removed, multiplied
+    by a universal singular factor. This singular factor depends on the unresolved
+    parton and the two neighbouring hard particles. Interpreting this as
+    the two hard particles forming an antenna which radiates the unresolved
+    parton gives rise to the antenna functions which will be discussed
+    further in Section~\ref{sec:antenna_functions}.
+
+\subsection{One-loop matrix element factorisation}\label{sec:OL_factorisation}
+    The discussion of matrix element factorisation so
+    far has been focused on tree-level matrix elements.
+    It has been shown that one-loop colour-ordered amplitudes
+    also factorise in the soft and collinear limits. At the one-loop level,
+    there are new universal singular functions and the
+    factorisation formula are modified.
+
+    In the soft limit, a one-loop colour-ordered amplitude
+    factorises as (cite here)
+    \begin{equation}\label{eqn:1L_soft_factorisation}
+        M_{n+1}^{(1)}(\ldots, i, j, k, \ldots) \rightarrow S_{ijk}^{(0)} \, M_{n}^{1}(\ldots, i, k, \ldots) + S_{ijk}^{(1)}(\epsilon) \, M_{n}^{0}(\ldots, i, k, \ldots) \, ,
+    \end{equation}
+    where $S_{ijk}^{(0)}$ is the eikonal factor (\ref{eqn:eikonal_factor}),
+    and $S_{ijk}^{(1)}(\epsilon)$ is the one-loop soft radiation function (cite here).
+
+    A similar factorisation formula for the collinear limit is
+    \begin{equation}\label{eqn:1L_collinear_factorisation}
+        M_{n+1}^{(1)}(\ldots, i, j, \ldots) \rightarrow \dfrac{1}{s_{ij}} \left[ P_{ij}^{(0)}(z) \, M_{n}^{(1)}(\ldots, k, \ldots) + P_{ij}^{(1)}(z, \epsilon) \, M_{n}^{(0)}(\ldots, k, \ldots) \right] \, ,
+    \end{equation}
+    where $P_{ij}^{(0)}$ are the tree-level splitting functions
+    (\ref{eqn:AP_kernels}) and $P_{ij}^{(1)}$
+    are the one-loop splitting functions (cite here).
+
+    In both these formulae the structure is of the form:
+    tree-level splitting function multipled by a one-loop
+    amplitude, plus a one-loop splitting function multiplied
+    by a tree-level amplitude as diagrammatically shown
+    in Fig ...
+
+    In Section~\ref{sec:antenna_functions} we will give
+    a brief overview of antenna functions where these one-loop
+    factorisation formulae were used to obtain universal singular functions
+    at the one-loop level for squared matrix elements. These
+    antenna functions are then applied in the context of NLO
+    k-factor emulation in Chapter~\ref{chapter:fame2}.
+
 \section{Subtraction}\label{sec:subtraction}

### 2022-09-26 qcd.tex
wrote: subtraction

@@
 
-\subsection{Infrared divergences}
+\subsection{Infrared divergences}\label{sec:IR_divergences}
     In Section~\ref{sec:UV} we saw how divergences
@@
     \end{equation}
-    where $P_{ij}^{(0)}$ are the tree-level splitting functions
-    (\ref{eqn:AP_kernels}) and $P_{ij}^{(1)}$
+    where $P_{ij}^{(0)}(z)$ are the tree-level splitting functions
+    (\ref{eqn:AP_kernels}) and $P_{ij}^{(1)}(z, \epsilon)$
     are the one-loop splitting functions (cite here).
@@
 \section{Subtraction}\label{sec:subtraction}
+    In Section~\ref{sec:IR_divergences} we discussed the
+    structure of IR divergences and how the KLN theorem
+    necessitated the cancellation of IR divergences 
+    arising from the virtual corrections and real-emission
+    matrix elements once integrated over soft and collinear
+    regions of phase-space, at each order in perturbation
+    theory.
+
+    Over the past few decades there has been a vast
+    amount of research into devising methods to systematically
+    isolate these singularities such that it is possible
+    to numerically achieve predictions of physical
+    quantities that are finite. To tackle this task, three
+    main mathods have been proposed: phase-space slicing (cite here),
+    sector decomposition (cite here), and subtraction (cite here).
+
+    By now the method of choice at NLO QCD is subtraction,
+    and is the method we will focus on in this section. The main
+    idea behind subtraction methods is to define local counterterms
+    that exactly replicate the IR divergent behaviour of matrix
+    elements. While there is not one single subtraction scheme
+    that is universally used, the general form of a subtraction term
+    must fulfill the requirements of replicating the matrix element
+    behaviour in all IR singular limits, and be analytically integrable
+    over the regions of phase-space corresponding to these IR limits.
+
+    To illustrate the idea behind the subtraction method,
+    first consider the calculation of a LO partonic cross-section
+    \begin{equation}\label{eqn:LO_cs}
+        \sigma^{\mathrm{LO}} = \int \mathrm{d}\Phi_{n} \; \mathcal{B}_{n} \, ,
+    \end{equation}
+    where $\mathcal{B}_{n}$ is the Born (tree-level) matrix
+    element and $\Phi_{n}$ is the $n$-body phase-space.
+    At the next order in perturbation theory, we have the NLO
+    cross-section which recieves contributions from the real and virtual
+    corrections
+    \begin{equation}\label{eqn:NLO_cs}
+        \sigma^{\mathrm{NLO}} = \int \mathrm{d}\Phi_{n} \; \left[\mathcal{B}_{n} + \mathcal{V}_{n}\right] + \int \mathrm{d}\Phi_{n+1} \; \mathcal{R}_{n+1} \, ,
+    \end{equation}
+    where $\mathcal{V}_{n}$ is the virtual matrix element (renormalised
+    to remove UV divergences as described in Section~\ref{sec:UV})
+    which lives in the same $n$-body phase-space as the Born
+    matrix element. The real-emission matrix element, $\mathcal{R}_{n+1}$,
+    lives in the $(n+1)$-body phase-space, $\Phi_{n+1}$, due to the emission
+    of an additional external particle. The integrals over $\Phi_{n}$ and
+    $\Phi_{n+1}$ are separately divergent but their sum is finite. To
+    carry out a numerical calculation, it is therefore necessary
+    to regulate these divergences. In dimensional regularisation these
+    divergences are mapped to poles in $\epsilon$.
+
+    The motivation behind the subtraction method is that the
+    divergences in (\ref{eqn:NLO_cs}) can be cancelled upon the insertion
+    of a counterterm evaluated in $\Phi_{n+1}$ and an integrated counterterm
+    evaluated in $\Phi_{n}$ such that the condition
+    \begin{equation}\label{eqn:subtraction_term_condition}
+        \int \mathrm{d}\Phi_{n} \; \mathcal{I}_{n} - \int \mathrm{d}\Phi_{n+1} \; \mathcal{C}_{n+1} = 0 \, ,
+    \end{equation}
+    holds. Here we have made use of the fact that
+    $\Phi_{n+1} \rightarrow \Phi_{n}\Phi_{1}$ when using
+    an appropriate $3 \rightarrow 2$ momentum mapping such that
+    \begin{equation}\label{eqn:integrated_counterterm}
+        \mathcal{I}_{n} = \int \Phi_{1} \; \mathcal{C}_{n+1}
+    \end{equation}
+    where $\Phi_{1}$ is the one-parton phase-space leading
+    to $\epsilon$ poles once $\mathcal{C}_{n+1}$ is integrated over.
+    The counterterm $\mathcal{C}_{n+1}$ should be a proper
+    pointwise approximation of $\mathcal{R}_{n+1}$ to cancel all IR
+    divergences such that $\mathcal{R}_{n+1} - \mathcal{C}_{n+1}$ is finite.
+    Additionally, since the pole structure in $\mathcal{V}_{n}$ is identical
+    to $\mathcal{R}_{n+1}$, $\mathcal{V}_{n} + \mathcal{I}_{n}$ will be finite
+    by construction.
+    Inserting this into (\ref{eqn:NLO_cs}) we get
+    \begin{equation}\label{eqn:NLO_subtraction}
+        \sigma^{\mathrm{NLO}} = \int \mathrm{d}\Phi_{n} \left[\mathcal{B}_{n} + \mathcal{V}_{n} + \mathcal{I}_{n}\right] + \int \mathrm{d}\Phi_{n+1} \left[\mathcal{R}_{n+1} - \mathcal{C}_{n+1}\right] \, ,
+    \end{equation}
+    where each integral can now be carried out numerically
+    in integer dimensions by taking the $\epsilon \rightarrow 0$
+    limit. This is possible as the integrals are all separately
+    finite now.
+
+    The counterterm $\mathcal{C}_{n+1}$ has been kept general
+    but specific exmples of subtractions schemes include
+    Catani-Seymour (CS) (cite here), Frixione-Kunszt-Signer (FKS) (cite here),
+    and antenna subtraction (cite here).
+
+    In the next sections we will describe in detail two sets of functions
+    that are used to build subtraction terms: Catani-Seymour dipoles,
+    antenna functions. However, we will not construct the counterterms explicitly.
+    In other words, we are interested in the approximations of the matrix elements
+    in the soft and collinear limits as these are universal and can
+    be applied to any process. These functions will become instrumental
+    during our construction of matrix element emulators in Chapters
+    \ref{chapter:fame1}, \ref{chapter:fame2}, and \ref{chapter:fame3}.
+
 \subsection{Catani-Seymour dipoles}\label{sec:CS_dipoles}

### 2022-09-28 appendix1.tex
wrote: CS dipoles

@@
 \begin{document}
-\chapter{Catani-Seymour notation}
+\chapter{Catani-Seymour notation and dipole functions}
 \label{appendix:catani_seymour}
+    \section{Notation}
     The notation for matrix elements given in Ref.~\cite{Catani:1996vz}
@@
     Namely $n_{c}(q) = n_{c}(\bar{q}) = N_{c}$ and $n_{c}(g) = N_{c}^{2}-1$.
-
 \end{document}

### 2022-09-28 qcd.tex
wrote: CS dipoles

@@
     quantities that are finite. To tackle this task, three
-    main mathods have been proposed: phase-space slicing (cite here),
-    sector decomposition (cite here), and subtraction (cite here).
+    main mathods have been proposed: phase-space slicing \cite{Fabricius:1981sx,Kramer:1986mc,Giele:1991vf},
+    sector decomposition \cite{Binoth:2000ps,Binoth:2003ak}, and subtraction \cite{Ellis:1980nc}.
 
@@
     The counterterm $\mathcal{C}_{n+1}$ has been kept general
-    but specific exmples of subtractions schemes include
-    Catani-Seymour (CS) (cite here), Frixione-Kunszt-Signer (FKS) (cite here),
-    and antenna subtraction (cite here).
+    but specific exmples of automated subtractions schemes include
+    Catani-Seymour (CS) \cite{Catani:1996vz,Catani:2002hc},
+    Frixione-Kunszt-Signer (FKS) \cite{Frixione:1995ms,Frixione:1997np},
+    and antenna subtraction \cite{Campbell:1998nn,Kosower:1997zr,Kosower:2003bh}.
+
+    Beyond NLO QCD, the algorithms available have
+    not reached the maturity of the automated methods widely use
+    at NLO. However, this is an active area of
+    research. See (cite black book) for a review of methods
+    that have been applied successfuly in NNLO QCD and beyond.
 
@@
 \subsection{Catani-Seymour dipoles}\label{sec:CS_dipoles}
-\subsection{Antenna}\label{sec:antenna_functions}
+    Catani-Seymour dipoles introduced in Ref.~\cite{Catani:1996vz}
+    are process independent functions that reproduce the IR
+    singular behaviour of matrix elements. They depend on the
+    momenta and quantum numbers of three partons in the real-emission
+    phase-space. These three partons are identified by indices
+    $i$, $j$, and $k$ where $i$ is the emitting parton, $j$ is
+    the unresolved emitted parton, and $k$ is a spectator parton.
+    Since the dipoles only depend on three partons, to map out
+    all the singular limits, it will be required to construct all
+    permutations of dipole functions.
+
+    Dipole functions can be separated into four categories
+    depending on whether the emitter and spectator are in the
+    initial-state or the final-state, as illustrated in Figure~\ref{fig:CS_dipoles}.
+    Namely, there are final-final (FF) dipoles, final-initial dipoles (FI),
+    initial-final (IF) dipoles and initial-initial (II) dipoles where
+    the nomenclature is emitter-spectator dipole.
+    For electron-positron annihilations only FF dipoles are required,
+    but for any hadronic collision the inclusion of the remaining
+    dipoles are required to capture all IR singular behaviour.
+
+    In the following we will describe in detail the FF dipoles
+    but only give a brief description of the remaining dipoles as
+    the structure of the factorisation formula is essentially identical.
+
+    \begin{figure}
+        \includegraphics[width=\linewidth]{qcd/CS_dipoles.pdf}
+        \caption{Schematic diagrams of the four classes of Catani-Seymour
+        dipoles, $\mathcal{D}$. Each dipole consists of a composite particle
+        that decays into two partons, and a spectator that recoils
+        to conserve momentum. $\boldsymbol{V}$ represents the splitting
+        function within the dipole function which contains the divergent
+        behaviour. The solid black blob represents the hard scattering process,
+        with incoming and outgoing lines representing initial and final-state
+        partons, respectively.}
+        \label{fig:CS_dipoles}
+    \end{figure}
+
+    \subsubsection{Final-final dipoles}
+        The utilisation of dipoles is encapsulated in the
+        dipole factorisation formula where matrix elements in the limit
+        $p_{i}p_{j} \rightarrow 0$ can be written as
+        \begin{equation}\label{eqn:FF_factorisation}
+            {}_{n+1} \langle 1, \ldots, n+1 | | 1, \ldots, n+1 \rangle_{n+1} = \sum_{k\neq i,j}\mathcal{D}_{ij,k}(p_{1},\ldots,p_{n+1}) + \ldots \, ,
+        \end{equation}
+        where terms not singular in the limit $p_{i}p_{j} \rightarrow 0$
+        are omitted, and the dipole is
+        \begin{equation}\label{eqn:D_ijk}
+            \begin{split}
+                \mathcal{D}_{ij,k}(p_{1},\ldots,p_{n+1}) &= -\dfrac{1}{2p_{i}p_{j}} \\
+                {}_{n}\langle 1, &\ldots, \widetilde{ij}, \ldots, \tilde{k}, \ldots, n+1 | \dfrac{\boldsymbol{T}_{k} \cdot \boldsymbol{T}_{ij}}{\boldsymbol{T}_{ij}^{2}} \boldsymbol{V}_{ij,k} | 1, \ldots, \widetilde{ij}, \ldots, \tilde{k}, \ldots, n+1 \rangle_{n} \, ,
+            \end{split}
+        \end{equation}
+        where $\boldsymbol{T}_{i}$ are colour charges and $\boldsymbol{V}_{ij,k}$
+        is a matrix in the helicity space of the emitter embedding the
+        IR divergent behaviour.
+        The sum can be understood as summing over all possible
+        3 leg permutations to capture all the soft and collinear limits.
+        The reduced matrix element on the RHS of (\ref{eqn:D_ijk})
+        is obtained by replacing the partons $i$ and $j$ with
+        a single parton $\widetilde{ij}$, and replacing parton
+        $k$ with a parton $\tilde{k}$. $\tilde{k}$ has all the
+        same quantum numbers as $k$, where as the
+        partonic nature of $\widetilde{ij}$ depends on
+        the specific splitting process (c.f. Section~\ref{sec:me_collinear}).
+        The momenta of these particles are modified in the following way
+        \begin{equation}\label{eqn:FF_mapping}
+            \tilde{p}_{ij}^{\mu} = p_{i}^{\mu} + p_{j}^{\mu} - \dfrac{y_{ij,k}}{1-y_{ij,k}} p_{k}^{\mu} \, , \quad \tilde{p}_{k}^{\mu} = \dfrac{1}{1-y_{ij,k}}p_{k}^{\mu} \, ,
+        \end{equation}
+        where $y_{ij,k}$, the recoil parameter, is a dimensionless variable given as
+        \begin{equation}\label{eqn:y_ijk}
+            y_{ij,k} = \dfrac{p_{i}p_{j}}{p_{i}p_{j} + p_{i}p_{k} + p_{j}p_{k}} \, .
+        \end{equation}
+        Equations (\ref{eqn:FF_mapping}) and (\ref{eqn:y_ijk})
+        are a $2\rightarrow 3$ momenta mapping that maps
+        $p_{i} + p_{j} + p_{k} \rightarrow \tilde{p}_{ij} + \tilde{p}_{k}$.
+        The mapping ensures momentum conservation is maintained across all of phase-space and
+        all particles are kept on-shell:
+        \begin{align}\label{eqn:mapping_conditions}
+            \begin{split}
+                &p_{i}^{\mu} + p_{j}^{\mu} + p_{k}^{\mu} = \tilde{p}_{ij}^{\mu} + \tilde{p}_{k}^{\mu} \, , \\
+                &\tilde{p}_{ij}^{2} = \tilde{p}_{k}^{2} = 0 \, .
+            \end{split}
+        \end{align}
+
+        The matrices $\boldsymbol{V}_{ij,k}$ are functions of
+        $y_{ij,k}$ and the splitting variables $\tilde{z}$
+        \begin{equation}\label{eqn:zi}
+            \tilde{z}_{i} = \dfrac{p_{i}p_{k}}{p_{i}p_{k} + p_{j}p_{k}} \, , \quad \tilde{z}_{j} = \dfrac{p_{j}p_{k}}{p_{i}p_{k} + p_{j}p_{k}} = 1 - \tilde{z}_{i} \, ,
+        \end{equation}
+        which are analogous to the $z$ from Altarelli-Parisi splitting functions.
+        $\boldsymbol{V}_{ij,k}$ acts on the spin indices of the emitter and emitted
+        parton, but is independent of the type of the spectator. Making the dependence
+        on partons $i$ and $j$ explicit we list all the kernels here
+        \begin{align}\label{eqn:FF_V_ijk}
+            &\langle s | \boldsymbol{V}_{q_{i}g_{j},k} | s' \rangle = 8\pi \mu^{2\epsilon} \alpha_{s} C_{F} \left[\dfrac{2}{1-\tilde{z}_{i}(1-y_{ij,k})} - (1+\tilde{z}_{i}) - \epsilon(1-\tilde{z}_{i})\right] \delta_{ss'} \, , \nonumber \\
+            &\langle \mu | \boldsymbol{V}_{q_{i},\bar{q}_{j},k} | \nu \rangle = 8\pi \mu^{2\epsilon} \alpha_{s} T_{R} \left[-g^{\mu\nu} - \dfrac{2}{p_{i}p_{j}}(\tilde{z}_{i}p_{i}^{\mu}-\tilde{z}_{j}p_{j}^{\mu})(\tilde{z}_{i}p_{i}^{\nu}-\tilde{z}_{j}p_{j}^{\nu})\right] \, , \nonumber \\
+            \begin{split}
+                &\langle \mu | \boldsymbol{V}_{g_{i}g_{j},k} | \nu \rangle = 16\pi \mu^{2\epsilon} \alpha_{s} C_{A} \left[-g^{\mu\nu}\left(\dfrac{1}{1-\tilde{z}_{i}(1-y_{ij,k})} + \dfrac{1}{1-\tilde{z}_{j}(1-y_{ij,k})} - 2\right) \right.
+\\
+                &\hspace{5cm} \left. + \dfrac{(1-\epsilon)}{p_{i}p_{j}}(\tilde{z}_{i}p_{i}^{\mu}-\tilde{z}_{j}p_{j}^{\mu})(\tilde{z}_{i}p_{i}^{\nu}-\tilde{z}_{j}p_{j}^{\nu})\right] \, .
+            \end{split}
+        \end{align}
+
+        A key feature of these kernels and are that they smoothly interpolate
+        between the soft and collinear limits, meaning they do not double count
+        any limit. Additionally, they only contain divergences in the
+        $p_{i}p_{j} \rightarrow 0$ limit and not for any other pair of
+        momenta. In the soft and collinear limits, the dipole function
+        $\mathcal{D}_{ij,k}$ correctly reproduces the matrix element factorisation
+        behaviour seen in (\ref{eqn:soft_factorisation}) and (\ref{eqn:F_collinear_factorisation}),
+        repsectively. In particular, $\boldsymbol{V}_{ij,k}$ becomes
+        proportional to the eikonal factor and Altarelli-Parisi splitting
+        functions in these respective limits.
+
+        In this thesis we will be focussing on the emulation of
+        colour and spin-averaged matrix elements. Therefore the spin-indices
+        are not explicitly available. It is possible to average over the spin indices in (\ref{eqn:FF_V_ijk})
+        to obtain the spin-averaged splitting functions
+        \begin{align}\label{eqn:FF_avg_V_ijk}
+            &\langle \boldsymbol{V}_{q_{i}g_{j},k} \rangle = 8\pi \mu^{2\epsilon} \alpha_{s} C_{F} \left[\dfrac{2}{1-\tilde{z}_{i}(1-y_{ij,k})} - (1+\tilde{z}_{i}) - \epsilon(1-\tilde{z}_{i})\right] \, , \nonumber \\
+            &\langle \boldsymbol{V}_{q_{i}\bar{q}_{j},k} \rangle = \, 8\pi \mu^{2\epsilon} \alpha_{s} T_{R} \left[1-\dfrac{2\tilde{z}_{i}\tilde{z}_{j}}{1-\epsilon}\right] \, , \nonumber \\
+            &\langle \boldsymbol{V}_{g_{i}g_{j},k} \rangle = 16\pi \mu^{2\epsilon} \alpha_{s} C_{A} \left[\dfrac{1}{1-\tilde{z}_{i}(1-y_{ij,k})}+\dfrac{1}{1-\tilde{z}_{j}(1-y_{ij,k})}-2+\tilde{z}_{i}\tilde{z}_{j}\right] \, .
+        \end{align}
+        These spin-averaged functions will form the set of basis functions
+        that map out all the IR singular limits in matrix elements
+        in Chapter~\ref{chapter:fame1}.
+
+        \subsubsection{Final-initial dipoles}
+        For the case of final-state emitter and initial-state
+        spectator we have the dipole factorisation formula
+        \begin{equation}\label{eqn:FI_factorisation}
+            \begin{split}
+                {}_{n+1}\langle 1, \ldots, n+1; a, \ldots || 1, \ldots, n+1; a, \ldots \rangle_{n+1} = \sum_{k\neq i,j} &\mathcal{D}_{ij,k}(p_{1},\ldots,p_{n+1};p_{a},\ldots) \\
+                + &\mathcal{D}_{ij}^{a}(p_{1},\ldots,p_{n+1};p_{a},\ldots) + \ldots \, ,
+            \end{split}
+        \end{equation}
+        where the new dipole contribution appears on the second line. $a$ represents an initial-state parton.
+        The dipole term $\mathcal{D}_{ij}^{a}$ is given by
+        \begin{equation}\label{eqn:D_aij}
+            \begin{split}
+                \mathcal{D}_{ij}^{a}(p_{1}, \ldots, p_{n+1}; p_{a}, \ldots) &= -\dfrac{1}{2p_{i}p_{j}} \dfrac{1}{x_{ij,a}} \\
+                {}_{n} \langle 1, \ldots, \widetilde{ij}, &\ldots, n+1; \tilde{a}, \ldots | \dfrac{\boldsymbol{T}_{a} \cdot \boldsymbol{T}_{ij}}{\boldsymbol{T}_{ij}^{2}} \boldsymbol{V}_{ij}^{a} | 1, \ldots, \widetilde{ij}, \ldots, n+1; \tilde{a} , \ldots \rangle_{n} \, .
+            \end{split}
+        \end{equation}
+        where $\widetilde{ij}$ is the composite emitter in the
+        final-state, and the spectator $\tilde{a}$ is in the initial-state.
+        The momentum mapping for $p_{i} + p_{j} + p_{a} \rightarrow \tilde{p}_{ij} + \tilde{p}_{a}$ is given as
+        \begin{equation}\label{eqn:FI_mapping}
+            \tilde{p}_{ij}^{\mu} = p_{i}^{\mu} + p_{j}^{\mu} - (1-x_{ij,a})p_{a}^{\mu} \, , \quad \tilde{p}_{a}^{\mu} = x_{ij,a} \, p_{a}^{\mu} \, , \\
+        \end{equation}
+        such that the following momentum conservation and on-shell conditions are met
+        \begin{equation}\label{eqn:FI_mapping_conditions}
+            \begin{split}
+                &p_{i}^{\mu} + p_{j}^{\mu} - p_{a}^{\mu} = \tilde{p}_{ij}^{\mu} - \tilde{p}_{a}^{\mu} \, , \\
+                &\tilde{p}_{ij}^{2} = \tilde{p}_{a}^{2} = 0 \, .
+            \end{split}
+        \end{equation}
+        The recoil parameter $x_{ij,a}$ and splitting variables $\tilde{z}$ are given by
+        \begin{align}\label{eqn:x_ija}
+            &x_{ij,a} = \dfrac{p_{i}p_{a} + p_{j}p_{a} - p_{i}p_{j}}{(p_{i}p_{a} + p_{j}p_{a}} \, , \\
+            &\tilde{z}_{i} = \dfrac{p_{i}p_{a}}{p_{i}p_{a} + p_{j}p_{a}} \, , \quad \tilde{z}_{j} = \dfrac{p_{j}p_{k}}{p_{i}p_{k} + p_{j}p_{k}} = 1 - \tilde{z}_{1} \, .
+        \end{align}
+        The spin-averaged splitting functions are given as
+        \begin{align}\label{eqn:FI_V_aij}
+            &\langle \boldsymbol{V}_{q_{i}g_{j},a} \rangle = 8\pi \mu^{2\epsilon}\alpha_{s} C_{F} \left[\dfrac{2}{1 - \tilde{z}_{i} + (1-x_{ij,a})} - (1+\tilde{z}_{i}) - \epsilon(1-\tilde{z}_{i})\right] \, , \nonumber \\
+            &\langle \boldsymbol{V}_{q_{i}\bar{q}_{j},a} \rangle = 8\pi \mu^{2\epsilon}\alpha_{s} T_{R} \, \left[1-\dfrac{2\tilde{z}_{i}\tilde{z}_{j}}{1-\epsilon}\right] \, , \nonumber \\
+            &\langle \boldsymbol{V}_{g_{i}g_{j},a} \rangle = 16\pi \mu^{2\epsilon} \alpha_{s} C_{A} \left[\dfrac{1}{\tilde{z}_{j} + (1-x_{ij,a})} + \dfrac{1}{\tilde{z}_{i} + (1-x_{ij,a})} - 2 + \tilde{z}_{i}\tilde{z}_{j}\right] \, .
+        \end{align}
+
+        \subsubsection{Initial-final dipoles}
+        For the case of initial-state emitter and final-state spectator,
+        the dipole factorisation formula is given as
+        \begin{equation}\label{eqn:IF_factorisation}
+            {}_{n+1} \langle 1, \ldots, n+1; a || 1, \ldots, n+1; a \rangle_{n+1} = \sum_{k \neq i}\mathcal{D}_{k}^{ai}(p_{1},\ldots,p_{n+1};p_{a}) + \ldots \, ,
+        \end{equation}
+        where there is only one initial-state parton. The dipole is given by
+        \begin{equation}\label{eqn:D_aik}
+            \begin{split}
+            \mathcal{D}_{k}^{ai}(p_{1}, \ldots, p_{n+1};p_{a}) &= -\dfrac{1}{2p_{a}p_{i}}\dfrac{1}{x_{ik,a}} \\
+            {}_{n}\langle 1, \ldots, \tilde{k}, &\ldots, n+1 ; \widetilde{ai} | \dfrac{\boldsymbol{T}_{k} \cdot \boldsymbol{T}_{ai}}{\boldsymbol{T}_{ai}^{2}} \boldsymbol{V}_{k}^{ai} | 1, \ldots, \tilde{k}, \ldots, n+1; \widetilde{ai} \rangle_{n} \, .
+            \end{split}
+        \end{equation}
+        where the emitter is the initial-state parton $\widetilde{ai}$
+        and the spectator is the final-state parton $\tilde{k}$.
+        The momenta mapping for $p_{a} + p_{i} + p_{k} \rightarrow \tilde{p}_{ai} + \tilde{p}_{k}$ is
+        \begin{equation}\label{eqn:IF_mapping}
+            \tilde{p}_{ai}^{\mu} = x_{ik,a} \, p_{a}^{\mu} \, , \quad \tilde{p}_{k}^{\mu} = p_{k}^{\mu} + p_{i}^{\mu} - (1-x_{ik,a})p_{a}^{\mu} 
+        \end{equation}
+        such that
+        \begin{equation}\label{eqn:IF_mapping_conditions}
+            \begin{split}
+            &p_{i}^{\mu} + p_{k}^{\mu} - p_{a}^{\mu} = \tilde{p}_{k}^{\mu} - \tilde{p}_{ai}^{\mu} \, , \\
+            &\tilde{p}_{ai}^{2} = \tilde{p}_{k}^{2} = 0 \, .
+            \end{split}
+        \end{equation}
+        Note that the momentum $\tilde{p}_{ai}$ is parallel
+        to $p_{a}$, therefore the modified spectator momentum $\tilde{p}_{k}$
+        has to recoil the transverse momentum. This leads to $x_{ik,a}$
+        acting as a splitting variable.
+
+        With $x_{ik,a}$ taking the role of $\tilde{z}$, there is another
+        parameter, $u_{i}$, appearing in the splitting functions.
+        These are given as
+        \begin{equation}\label{eqn:x_ika}
+            x_{ik,a} = \dfrac{p_{k}p_{a}+p_{i}p_{a}-p_{i}p_{k}}{p_{i}p_{a} + p_{k}p_{a}} \, , \quad u_{i} = \dfrac{p_{i}p_{a}}{p_{i}p_{a} + p_{k}p_{a}} \, .
+        \end{equation}
+
+        The spin-averaged splitting functions are
+        \begin{align}\label{eqn:V_aik}
+            &\langle \boldsymbol{V}_{k}^{q_{a}g_{i}} \rangle = \dfrac{n_{s}(\tilde{q})}{n_{s}(q)} \; 8\pi \mu^{2\epsilon} \alpha_{s} C_{F} \left[\dfrac{2}{1-x_{ik,a} + u_{i}} + (1+x_{ik,a}) - \epsilon(1-x_{ik,a})\right] \, , \nonumber \\
+            &\langle \boldsymbol{V}_{k}^{g_{a}\bar{q}_{i}} \rangle = \dfrac{n_{s}(g)}{n_{s}(\tilde{q})} \; 8\pi \mu^{2\epsilon} \alpha_{s} T_{R} \left[1 - \dfrac{2x_{ik,a}(1-x_{ik,a})}{1-\epsilon}\right] \, , \nonumber \\
+            &\langle \boldsymbol{V}_{k}^{g_{a}g_{i}} \rangle = \dfrac{n_{s}(g)}{n_{s}(\tilde{g})} \; 16\pi \mu^{2\epsilon} \alpha_{s} C_{A} \left[\dfrac{1}{1-x_{ik,a} + u_{i}} + \dfrac{1-x_{ik,a}}{x_{ik,a}} - 1 + x_{ik,a}(1-x_{ik,a}) \right] \, , \nonumber \\
+            &\langle \boldsymbol{V}_{k}^{q_{a}q_{i}} \rangle = \dfrac{n_{s}(q)}{n_{s}(\tilde{g})} \; 8\pi \mu^{2\epsilon} \alpha_{s} C_{F} \left[(1-\epsilon)x_{ik,a} + 2\dfrac{1-x_{ik,a}}{x_{ik,a}} \right] \, ,
+        \end{align}
+        where $n_{s}(a)$ is the number of polarisations of particle a.
+        For fermions $n_{s}(q) = n_{s}(\bar{q}) = 2$
+        and for gluons $n_{s}(g) = d-2$. In the limit $\epsilon \rightarrow 0$
+        all ratios of $n_{s}$ cancel to give unity.
+
+        \subsubsection{Initial-initial dipoles}
+        In the case of two initial-state partons, $a$ and $b$, there
+        is an additional dipole contribution. The dipole factorisation
+        in this case is  given by
+        \begin{equation}\label{eqn:II_factorisation}
+            \begin{split}
+                {}_{n+1} \langle 1, \ldots, n+1; a,b| | 1, \ldots, n+1;a,b \rangle_{n+1} = \sum_{k \neq i} &\mathcal{D}_{k}^{ai}(p_{1}, \ldots, p_{n+1};p_{a},p_{b}) \\
+                + &\mathcal{D}^{ai,b}(p_{1}, \ldots, p_{n+1};p_{a},p_{b}) + \ldots \, ,
+            \end{split}
+        \end{equation}
+        where the new contribution is the dipole $\mathcal{D}^{ai,b}$ given by
+        \begin{equation}\label{eqn:D_aib}
+            \begin{split}
+                \mathcal{D}^{ai,b}(p_{1},\ldots,p_{n+1};p_{a},p_{b}) &= -\dfrac{1}{2p_{a}p_{i}} \dfrac{1}{x_{i,ab}} \\
+                &{}_{n}\langle \tilde{1}, \ldots, \widetilde{n+1}; \widetilde{ai}, b | \dfrac{\boldsymbol{T}_{b}\cdot\boldsymbol{T}_{ai}}{\boldsymbol{T}^{2}_{ai}}\boldsymbol{V}^{ai,b} | \tilde{1}, \ldots, \widetilde{n+1}; \widetilde{ai}, b \rangle_{n} \, .
+            \end{split}
+        \end{equation}
+        The initial-state parton $\widetilde{ai}$ is the emitter and the
+        other initial-state parton $b$ is the spectator. Notice that even the
+        momenta not involved in the dipole term have been modified with only
+        parton $b$ remaining unchanged. The momentum mapping for this case
+        has $\tilde{p}_{ai}$ parallel with $p_{a}$ and also
+        modifies the momenta for all other final-state momenta (even non-QCD particles) $k_{j}$:
+        \begin{align}\label{eqn:II_mapping}
+            \tilde{p}_{ai}^{\mu} &= x_{i,ab} \, p_{a}^{\mu} \, , \nonumber \\
+            \tilde{k}_{j}^{\mu} &= k_{j}^{\mu} - \dfrac{2k_{j}(K+\widetilde{K})}{(K+\widetilde{K})^{2}}(K+\widetilde{K})^{\mu} + \dfrac{2k_{j}K}{K^{2}}\widetilde{K}^{\mu} \, , \\
+            x_{i,ab} &= \dfrac{p_{a}p_{b} - p_{i}p_{a}-p_{i}p_{b}}{p_{a}p_{b}} \, . \nonumber
+        \end{align}
+        where $K$ and $\widetilde{K}$ are the total momenta
+        of the dipole before and after the mapping. They are given by
+        \begin{align}\label{eqn:II_mapping2}
+            \begin{split}
+            K^{\mu} &= p_{a}^{\mu} + p_{b}^{\mu} - p_{i}^{\mu} \, ,\\
+            \widetilde{K}^{\mu} &= \tilde{p}_{ai}^{\mu} + p_{b}^{\mu} \, . \\
+            \end{split}
+        \end{align}
+        This mapping conserves momentum and keeps mapped momenta on-shell
+        \begin{align}\label{eqn:II_mapping_conditions}
+            \begin{split}
+            &p_{a}^{\mu} + p_{b}^{\mu} - p_{i}^{\mu} - \sum_{j} k_{j}^{\mu} = \tilde{p}_{ai}^{\mu} + p_{b}^{\mu} - \sum_{j} \tilde{k}_{j}^{\mu} = 0\, , \\
+            &\tilde{p}_{ai}^{2} = \tilde{k}_{j}^{2} = 0 \, .
+            \end{split}
+        \end{align}
+        The spin-averaged splitting functions are proportional
+        to the Altarelli-Parisi splitting functions which we
+        quote here for completeness
+        \begin{align}\label{V_aib}
+            \begin{split}
+            &\langle \boldsymbol{V}^{q_{a}q_{i},b} \rangle = \dfrac{n_{s}(q)}{n_{s}(\tilde{g})} \; 8\pi \mu^{2\epsilon} \alpha_{s} C_{F} \left[\dfrac{1 + (1-x_{i,ab})^{2}}{x_{i,ab}}\right] \, , \\
+            &\langle \boldsymbol{V}^{q_{a}g_{i},b} \rangle = \dfrac{n_{s}(q)}{n_{s}(\tilde{q})} \; 8\pi \mu^{2\epsilon} \alpha_{s} C_{F} \left[\dfrac{1 + x_{i,ab}^{2}}{1 - x_{i,ab}}\right] \, , \\
+            &\langle \boldsymbol{V}^{g_{a}\bar{q}_{i},b} \rangle = \dfrac{n_{s}(g)}{n_{s}(\tilde{q})} \; 8\pi \mu^{2\epsilon} \alpha_{s} T_{R} \left[x_{i,ab}^{2} + (1-x_{i,ab})^{2}\right] \, , \\
+            &\langle \boldsymbol{V}^{g_{a}g_{i},b} \rangle = \dfrac{n_{s}(g)}{n_{s}(\tilde{g})} \; 16\pi \mu^{2\epsilon} \alpha_{s} C_{A} \left[\dfrac{x_{i,ab}}{1-x_{i,ab}} + \dfrac{1-x_{i,ab}}{x_{i,ab}} + x_{i,ab}(1-x_{i,ab})\right] \, .
+            \end{split}
+        \end{align}
+
+        \subsubsection{Final formula}
+        Now that we have described each of the dipole contributions,
+        we can collect them all into a master dipole factorisation
+        formula
+        \begin{equation}\label{eqn:dipole_factorisation_master}
+            \begin{split}
+            {}_{n+1} \langle 1, \ldots, n+1 | &| 1, \ldots, n+1 \rangle_{n+1} = \\
+            &\sum_{k \neq i, j} \mathcal{D}_{ij,k}(p_{1},\ldots,p_{n+1};p_{a},p_{b}) + \mathcal{D}_{ij}^{a}(p_{1},\ldots,p_{n+1};p_{a},p_{b}) \\
+            &\sum_{k \neq i} \mathcal{D}_{k}^{ai}(p_{1},\ldots,p_{n+1};p_{a},p_{b}) + \mathcal{D}^{ai,b}(p_{1},\ldots,p_{n+1};p_{a},p_{b}) + \ldots \, ,
+            \end{split}
+        \end{equation}
+        where it is understood that for a given partonic process,
+        the matrix element in the soft and collinear limits will be
+        reproduced by summing over all the possible dipoles that are
+        relevant.
+\subsection{Antenna functions}\label{sec:antenna_functions}
 \end{document}

### 2022-09-28 qcd.tex
wrote: CS massive dipoles

@@
         \begin{equation}\label{eqn:FF_factorisation}
-            {}_{n+1} \langle 1, \ldots, n+1 | | 1, \ldots, n+1 \rangle_{n+1} = \sum_{k\neq i,j}\mathcal{D}_{ij,k}(p_{1},\ldots,p_{n+1}) + \ldots \, ,
+            {}_{n+1} \langle 1, \ldots, n+1 | | 1, \ldots, n+1 \rangle_{n+1} = \sum_{i,j}\sum_{k\neq i,j}\mathcal{D}_{ij,k}(p_{1},\ldots,p_{n+1}) + \ldots \, ,
         \end{equation}
@@
             \begin{split}
-                {}_{n+1}\langle 1, \ldots, n+1; a, \ldots || 1, \ldots, n+1; a, \ldots \rangle_{n+1} = \sum_{k\neq i,j} &\mathcal{D}_{ij,k}(p_{1},\ldots,p_{n+1};p_{a},\ldots) \\
-                + &\mathcal{D}_{ij}^{a}(p_{1},\ldots,p_{n+1};p_{a},\ldots) + \ldots \, ,
+                {}_{n+1}\langle 1, \ldots, n+1; a, \ldots |&| 1, \ldots, n+1; a, \ldots \rangle_{n+1} = \\
+                &\sum_{i,j}\sum_{k\neq i,j} \mathcal{D}_{ij,k}(p_{1},\ldots,p_{n+1};p_{a},\ldots) \\
+                &\sum_{i,j}\sum_{a} \mathcal{D}_{ij}^{a}(p_{1},\ldots,p_{n+1};p_{a},\ldots) + \ldots \, ,
             \end{split}
@@
         \begin{equation}\label{eqn:IF_factorisation}
-            {}_{n+1} \langle 1, \ldots, n+1; a || 1, \ldots, n+1; a \rangle_{n+1} = \sum_{k \neq i}\mathcal{D}_{k}^{ai}(p_{1},\ldots,p_{n+1};p_{a}) + \ldots \, ,
+            {}_{n+1} \langle 1, \ldots, n+1; a || 1, \ldots, n+1; a \rangle_{n+1} = \sum_{a,i}\sum_{k \neq i}\mathcal{D}_{k}^{ai}(p_{1},\ldots,p_{n+1};p_{a}) + \ldots \, ,
         \end{equation}
@@
             \begin{split}
-                {}_{n+1} \langle 1, \ldots, n+1; a,b| | 1, \ldots, n+1;a,b \rangle_{n+1} = \sum_{k \neq i} &\mathcal{D}_{k}^{ai}(p_{1}, \ldots, p_{n+1};p_{a},p_{b}) \\
-                + &\mathcal{D}^{ai,b}(p_{1}, \ldots, p_{n+1};p_{a},p_{b}) + \ldots \, ,
+                {}_{n+1} \langle 1, \ldots, n+1; a,b|&| 1, \ldots, n+1;a,b \rangle_{n+1} = \\
+                &\sum_{a,i}\sum_{k \neq i} \mathcal{D}_{k}^{ai}(p_{1}, \ldots, p_{n+1};p_{a},p_{b}) \\
+                &\sum_{a,i}\sum_{b \neq a} \mathcal{D}^{ai,b}(p_{1}, \ldots, p_{n+1};p_{a},p_{b}) + \ldots \, ,
             \end{split}
@@
             \begin{split}
-            {}_{n+1} \langle 1, \ldots, n+1 | &| 1, \ldots, n+1 \rangle_{n+1} = \\
-            &\sum_{k \neq i, j} \mathcal{D}_{ij,k}(p_{1},\ldots,p_{n+1};p_{a},p_{b}) + \mathcal{D}_{ij}^{a}(p_{1},\ldots,p_{n+1};p_{a},p_{b}) \\
-            &\sum_{k \neq i} \mathcal{D}_{k}^{ai}(p_{1},\ldots,p_{n+1};p_{a},p_{b}) + \mathcal{D}^{ai,b}(p_{1},\ldots,p_{n+1};p_{a},p_{b}) + \ldots \, ,
+            {}_{n+1} \langle 1, &\ldots, n+1 || 1, \ldots, n+1 \rangle_{n+1} = \\
+            &\sum_{i,j}\sum_{k \neq i, j} \mathcal{D}_{ij,k}(p_{1},\ldots,p_{n+1};p_{a},p_{b}) + \sum_{i,j}\sum_{a} \mathcal{D}_{ij}^{a}(p_{1},\ldots,p_{n+1};p_{a},p_{b}) \\
+            &\sum_{a,i}\sum_{k \neq i} \mathcal{D}_{k}^{ai}(p_{1},\ldots,p_{n+1};p_{a},p_{b}) + \sum_{a,i}\sum_{b \neq a} \mathcal{D}^{ai,b}(p_{1},\ldots,p_{n+1};p_{a},p_{b}) + \ldots \, ,
             \end{split}
@@
         where it is understood that for a given partonic process,
-        the matrix element in the soft and collinear limits will be
+        the matrix element in the different soft and collinear limits will be
         reproduced by summing over all the possible dipoles that are
-        relevant.
+        relevant. The omitted  terms are non-divergent.
+
+        \subsubsection{Treatment of massive partons}
+        The dipoles described in this section so far have
+        treated all partons to be massless. Since quarks are massive
+        particles it is not possible in general to treat them
+        as massless. This is especially true for the top
+        quark which has a mass of approximately 175 GeV, the
+        most massive SM particle. In the case of partonic
+        processes involving top quarks, it is therefore
+        necessary to include mass effects into the dipole
+        functions.
+
+        The universality of infrared divergent structure
+        extends to massive quarks \cite{Catani:2000ef}.
+        Massive partons are treated in detail in Reference\cite{Catani:2002hc}
+        where the authors derive dipole
+        functions for massive partons. The expressions derived therein
+        are utilised in Chapter~\ref{chapter:fame3}, but will
+        not be written here as the philosophy of constructing
+        the dipole functions is similar to those already described.
+
+        The extension to massive quarks necessitates the straightforward
+        modification of the master dipole factorisation formula
+        (\ref{eqn:dipole_factorisation_master}) to replace
+        any massless dipole with its massive counterpart whenever
+        there is a massive parton in the dipole.
+
 \subsection{Antenna functions}\label{sec:antenna_functions}
+
 \end{document}

### 2022-09-29 qcd.tex
wrote: antenna functions

--- 2022-09-28 qcd.tex
@@
 
-\subsection{Factorisation of colour-ordered amplitudes}
+\subsection{Factorisation of colour-ordered amplitudes}\label{sec:CO_factorisation}
     The factorisation expressed in (\ref{eqn:soft_factorisation}),
@@
     \end{equation}
-    where the leptonic current is $V^{\mu}$ and the hadronic
+    where $V^{\mu}$ is the leptonic current and the hadronic
     current is
@@
     structure has been factored into a product of fundamental group
-    generators where the indices $a_{i} \in {1,\ldots, N_{c}^{2}-1}$
-    and $c_{i} \in {1,\ldots,N_{c}}$. This leaves the colour-ordered
+    generators where the indices $a_{i} \in \{1,\ldots, N_{c}^{2}-1\}$
+    and $c_{i} \in \{1,\ldots,N_{c}\}$. This leaves the colour-ordered
     partial amplitude $S_{\mu}(q_{1};1,\ldots,n;\bar{q}_{2})$ depending
-    only on kinematical variables, and where the gluons are emitted in an
-    ordered fashion from the quarks. The sum over $P(1,\ldots,n)$
-    represents a sum over all permutations of gluon emissions
-    which accounts for all Feynman diagrams and colour structures.
+    only on kinematical variables. In this parton amplitude, the gluons
+    are emitted in an ordered fashion from the quarks.
+    The sum over $P(1,\ldots,n)$ represents a sum over all permutations
+    of gluon emissions which accounts for all Feynman diagrams and colour structures.
 
@@
     For partons which are not colour connected (are not neighbouring
-    partons in the colour ordered amplitude), there will be no singular
+    partons in the colour-ordered amplitude), there will be no singular
     behaviour as $s_{ij} \rightarrow 0$.
@@
     It has been shown that one-loop colour-ordered amplitudes
-    also factorise in the soft and collinear limits. At the one-loop level,
-    there are new universal singular functions and the
-    factorisation formula are modified.
+    also factorise in the soft and collinear limits \cite{Bern:1994zx,Bern:1998sc,Kosower:1999xi}.
+    At the one-loop level, there are new universal singular
+    functions and the factorisation formulae are modified.
 
     In the soft limit, a one-loop colour-ordered amplitude
-    factorises as (cite here)
+    factorises as
     \begin{equation}\label{eqn:1L_soft_factorisation}
@@
     where $S_{ijk}^{(0)}$ is the eikonal factor (\ref{eqn:eikonal_factor}),
-    and $S_{ijk}^{(1)}(\epsilon)$ is the one-loop soft radiation function (cite here).
+    and $S_{ijk}^{(1)}(\epsilon)$ is the one-loop soft radiation function \cite{Bern:1999ry}.
 
@@
     (\ref{eqn:AP_kernels}) and $P_{ij}^{(1)}(z, \epsilon)$
-    are the one-loop splitting functions (cite here).
+    are the one-loop splitting functions \cite{Bern:1999ry}.
 
@@
     amplitude, plus a one-loop splitting function multiplied
-    by a tree-level amplitude as diagrammatically shown
-    in Fig ...
+    by a tree-level amplitude as illustrated in in Fig ...
 
@@
     at NLO. However, this is an active area of
-    research. See (cite black book) for a review of methods
-    that have been applied successfuly in NNLO QCD and beyond.
+    research. See Reference \cite{TorresBobadilla:2020ekr} for a review of methods
+    that have been applied to NNLO QCD.
 
@@
 
-\subsection{Catani-Seymour dipoles}\label{sec:CS_dipoles}
+\section{Catani-Seymour dipoles}\label{sec:CS_dipoles}
     Catani-Seymour dipoles introduced in Ref.~\cite{Catani:1996vz}
@@
 
-    \subsubsection{Final-final dipoles}
+    \subsection{Final-final dipoles}
         The utilisation of dipoles is encapsulated in the
@@
 
-        \subsubsection{Final-initial dipoles}
+    \subsection{Final-initial dipoles}
         For the case of final-state emitter and initial-state
@@
 
-        \subsubsection{Initial-final dipoles}
+    \subsection{Initial-final dipoles}
         For the case of initial-state emitter and final-state spectator,
@@
 
-        \subsubsection{Initial-initial dipoles}
+    \subsection{Initial-initial dipoles}
         In the case of two initial-state partons, $a$ and $b$, there
@@
 
-        \subsubsection{Final formula}
+    \subsection{Master factorisation formula}
         Now that we have described each of the dipole contributions,
@@
         where it is understood that for a given partonic process,
-        the matrix element in the soft and collinear limits will be
+        the matrix element in the different soft and collinear limits will be
         reproduced by summing over all the possible dipoles that are
-        relevant.
-\subsection{Antenna functions}\label{sec:antenna_functions}
+        relevant. The omitted  terms are non-divergent.
+
+    \subsection{Treatment of massive partons}
+        The dipoles described in this section so far have
+        treated all partons to be massless. Since quarks are massive
+        particles it is not possible in general to treat them
+        as massless. This is especially true for the top
+        quark which has a mass of approximately 175 GeV, the
+        most massive SM particle. In the case of partonic
+        processes involving top quarks, it is therefore
+        necessary to include mass effects into the dipole
+        functions.
+
+        The universality of infrared divergent structure
+        extends to massive quarks \cite{Catani:2000ef}.
+        Massive partons are treated in detail in Reference\cite{Catani:2002hc}
+        where the authors derive dipole
+        functions for massive partons. The expressions derived therein
+        are utilised in Chapter~\ref{chapter:fame3}, but will
+        not be written here as the philosophy of constructing
+        the dipole functions is similar to those already described.
+
+        The extension to massive quarks necessitates the straightforward
+        modification of the master dipole factorisation formula
+        (\ref{eqn:dipole_factorisation_master}) to replace
+        any massless dipole with its massive counterpart whenever
+        there is a massive parton in the dipole.
+
+\section{Antenna functions}\label{sec:antenna_functions}
+    Another set of functions that can be used to construct
+    subtractions terms in NLO \cite{Kosower:1997zr,Campbell:1998nn}
+    and NNLO QCD \cite{Gehrmann-DeRidder:2004ttg,Gehrmann-DeRidder:2005btv}
+    are the antenna functions. They are derived from physical
+    matrix elements and so by construction contain the correct
+    IR behaviour in the soft and collinear limits. More specifically,
+    they are derived from colour-ordered matrix elements so they
+    follow the colour-ordered factorisation properties outlined
+    in Section~\ref{sec:CO_factorisation}. In the factorisation
+    formulae, the colour-ordered matrix elements factorised
+    exactly into a colour-ordered matrix element with one parton
+    removed multiplied by a singular factor. Since this factorisation
+    is exact, and the IR behaviour is universal, it is possible
+    to extract these singular factors once and for all.
+
+    In contrast to the dipole functions, where there is an identified
+    emitter, antenna functions can have the unresolved parton
+    radiate from either of the two hard partons. In this sense,
+    an antenna is a linear combination of two dipoles.
+
+    The discussion in this section will be limited to the
+    situation of a massive colour-neutral boson decaying
+    to massless QCD partons, suitable for multijet production
+    in electron-positron annihilation. Additionally, we
+    assume that there will only ever be at most one unresolved
+    parton in the final-state. The antenna formalism has been
+    extended to more general cases than this \cite{Daleo:2006xa,Daleo:2009yj,Boughezal:2010mc,Gehrmann:2011wi,Gehrmann-DeRidder:2012too}.
+
+    Antenna functions are constructed from a ratio of
+    colour-ordered matrix elements.
+    For the case of one unresolved parton, the three-parton
+    tree-level antenna function is given by
+    \begin{equation}\label{eqn:X30}
+        X_{3}^{0}(i, j, k) = S_{ijk,IK} \dfrac{|M_{n+1}^{(0)}(i,j,k)|^{2}}{|M_{n}^{(0)}(I,K)|^{2}}
+    \end{equation}
+    where $|M_{n}^{(\ell)}|^{2}$ is the $n$-body colour-ordered matrix
+    element at loop-level $\ell$. $S_{ijk,IK}$ is a symmetry factor
+    accounting for identical particles in the final-state
+    and for the presence of multiple antennae in the two-
+    parton process. $I$ and $K$ are hard partons forming a colour
+    connected antenna that radiates particle $j$. The identities
+    of $i/I$ and $k/K$ defines the specific class of the
+    antenna function. There are three classes corresponding
+    to the three underlying two-parton processes: quark-antiquark,
+    quark-gluon and gluon-gluon. Therefore, there will
+    be an antenna for every radiative correction to these
+    two-parton processes. The physical matrix elements used
+    to the derive the antenna functions for each class are as follows:
+    \begin{itemize}
+        \item Quark-antiquark: $\gamma^{*} \rightarrow q\bar{q}$ + (partons),
+        the decay of a virtual photon into a quark-antiquark pair with QCD radiation from the quark pair \cite{Gehrmann-DeRidder:2004ttg}.
+        \item Quark-gluon: $\tilde{\chi} \rightarrow \tilde{g}g$ + (partons),
+        the decay of a heavy neutralino into a gluino and gluon with QCD radiation from the gluon-gluino pair \cite{Gehrmann-DeRidder:2005svg}.
+        \item Gluon-gluon: $H \rightarrow gg$ + (partons),
+        the decay of a Higgs boson into a pair of gluons with QCD from the gluon pair \cite{Gehrmann-DeRidder:2005alt}.
+    \end{itemize}
+    Along with the tree-level antenna, there are also the
+    one-loop antennae. For the three-parton case these are given by
+    \begin{equation}\label{eqn:X31}
+        X_{3}^{1}(i,j,k) = S_{ijk,IK} \dfrac{|M_{n+1}^{(1)}(i,j,k)|^{2}}{|M_{n}^{(0)}(I,K)|^{2}} - X_{3}^{0}(i,j,k) \dfrac{|M_{n}^{(1)}(I,K)|^{2}}{|M_{n}^{(0)}(I,K)|^{2}} \, ,
+    \end{equation}
+    where $X_{3}^{1}$ are defined such that they are proportional
+    to the one-loop singular functions. Since the one-loop
+    factorisation formulae (\ref{eqn:1L_soft_factorisation})
+    and (\ref{eqn:1L_collinear_factorisation}) are of the form
+    $|M^{(1)}_{n+1}|^{2}\rightarrow S^{(0)}|M^{(1)}_{n}|^{2} + S^{(1)}|M^{(0)}_{n}|^{2}$,
+    the term proportional to $X_{3}^{0}$ has to be removed
+    to fulfill this requirement.
+    The antenna functions are summarised in Table~\ref{table:antenna_functions}, where
+    antenna functions are categorised by their class and radiative
+    processes. A pictorial representation is also given in Figure ...
+
+    \begin{table}[ht]
+        \begin{tabular}{ cccc } 
+            \toprule
+            \multirow{2}[3]{*}{Class} & \multirow{2}[3]{*}{Radiation} & \multicolumn{2}{c}{Antenna functions} \\
+            \cmidrule(lr){3-4}
+             & & Tree-level & One-loop \\
+            \midrule
+            Quark-antiquark & $q \bar{q} \rightarrow q g \bar{q}$& $A_{3}^{0}$ & $A_{3}^{1}$, $\tilde{A}_{3}^{1}$, $\hat{A}_{3}^{1}$ \\
+            \midrule
+            \multirow{2}{*}{Quark-gluon} & $q g \rightarrow q g g$ & $D_{3}^{0}$ & $D_{3}^{1}$, $\hat{D}_{3}^{1}$ \\
+            \cmidrule(lr){2-4}
+             & $qg \rightarrow q Q \bar{Q}$ & $E_{3}^{0}$ & $E_{3}^{1}$, $\tilde{E}_{3}^{1}$, $\hat{E}_{3}^{1}$ \\
+            \midrule
+            \multirow{2}{*}{Gluon-gluon} & $g g \rightarrow g g g$ & $F_{3}^{0}$ & $F_{3}^{1}$, $\hat{F}_{3}^{1}$ \\
+            \cmidrule(lr){2-4}
+            & $g g \rightarrow g q \bar{q}$ & $G_{3}^{0}$ & $G_{3}^{1}$, $\tilde{G}_{3}^{1}$, $\hat{G}_{3}^{1}$ \\
+            \bottomrule
+        \end{tabular}
+        \caption{Three-parton antenna functions at tree-level, $X_{3}^{0}$,
+        and one-loop level, $X_{3}^{1}$. The different antenna functions at
+        one-loop level correspond to the leading colour ($X_{3}^{1}$),
+        sub-leading colour ($\tilde{X}_{3}^{1}$), and closed quark loop
+        ($\hat{X}_{3}^{1}$) contributions.}
+        \label{table:antenna_functions}
+    \end{table}
+
+    \begin{figure}
+        \includegraphics[width=\linewidth]{qcd/antenna.pdf}
+        \caption{Three-parton antenna functions with the grey
+        blob representing all diagrams that give rise to the given
+        external states. The hard radiators are depicted in red,
+        where as the unresolved parton is coloured in blue.}
+        \label{fig:antenna}
+    \end{figure}
+
+    Another requirement in constructing a subtraction term
+    is that the functions have to be analytically integrable
+    over the unresolved phase-space.
+    For one unresolved parton, this requirement is met with a
+    suitable $3 \rightarrow 2$ momentum mapping that allows
+    the $(n+1)$-body phase-space to factorise into a product of
+    the mapped $n$-body phase-space and the antenna phase-space.
+    The antenna phase-space is independent of the $n$-body phase-space
+    and depends only on the momenta of partons $i, j, k$ appearing
+    in the antenna.
+    In the case of antenna functions, the mapping of choice is
+    the Kosower mapping \cite{Kosower:2002su} which maps
+    $p_{i} + p_{j} + p_{k} \rightarrow p_{I} + p_{K}$. It reads
+    \begin{align}\label{eqn:kosower_mapping}
+        \begin{split}
+        p_{I} &= x p_{i} + r p_{j} + y p_{k} \, \\
+        p_{K} &= (1-x) p_{i} + (1-r) p_{j} + (1-y) p_{k} \, ,
+        \end{split}
+    \end{align}
+    where the mapping variables are given as
+    \begin{align}\label{eqn:kosower_variables}
+        \begin{split}
+        s_{ijk} &= s_{ij} + s_{ik} + s_{jk} \, , \\
+        r &= \dfrac{s_{jk}}{s_{ij}s_{jk}} \, , \\
+        \rho &= \sqrt{1 + 4r(1-r)\dfrac{s_{ij}s_{jk}}{s_{ijk}{s_{ik}}}} \, , \\
+        x &= \dfrac{(1+\rho)s_{ijk}-2r s_{jk}}{2(s_{ij}+s_{ik})} \, , \\
+        y &= \dfrac{(1-\rho)s_{ijk}-2r s_{ij}}{2(s_{jk}+s_{ik})} \, . \\
+        \end{split}
+    \end{align}
+    This momentum mapping maintains momentum conservation
+    and keeps the mapped partons on-shell
+    \begin{equation}\label{eqn:kosower_conditions}
+        \begin{split}
+        &p_{i} + p_{j} + p_{k} = p_{I} + p_{K} \, , \\
+        &p_{I}^{2} = p_{K}^{2} = 0 \, .
+        \end{split}
+    \end{equation}
+
 \end{document}

### 2022-10-04 qcd.tex
wrote: antenna function expressions (D, F)

@@
 
+    The antenna functions have some functions in common,
+    namely the singularity operators (cite here)
+    \begin{align}\label{eqn:singularity_operators}
+        \begin{split}
+        \mathbf{I}_{q\bar{q}}^{(1)}(\epsilon, s_{q\bar{q}}) &= -\dfrac{e^{\epsilon\gamma}}{2\Gamma(1-\epsilon)} \bigg[\dfrac{1}{\epsilon^{2}} + \dfrac{3}{2\epsilon}\bigg] \mathrm{Re}(-s_{q\bar{q}})^{-\epsilon} \, , \\
+        \mathbf{I}_{qg}^{(1)}(\epsilon, s_{qg}) &= -\dfrac{e^{\epsilon\gamma}}{2\Gamma(1-\epsilon)} \bigg[\dfrac{1}{\epsilon^{2}} + \dfrac{5}{3\epsilon}\bigg] \mathrm{Re}(-s_{qg})^{-\epsilon} \, , \\
+        \mathbf{I}_{gg}^{(1)}(\epsilon, s_{gg}) &= -\dfrac{e^{\epsilon\gamma}}{2\Gamma(1-\epsilon)} \bigg[\dfrac{1}{\epsilon^{2}} + \dfrac{11}{6\epsilon}\bigg] \mathrm{Re}(-s_{gg})^{-\epsilon} \, , \\
+        \mathbf{I}^{(1)}_{g\bar{q}}(\epsilon, s_{g\bar{q}}) &= \mathbf{I}_{qg}^{(1)}(\epsilon,s_{g\bar{q}}) \, ,
+        \end{split}
+    \end{align}
+    and dilogarithms which are embedded in the function
+    \begin{equation}\label{eqn:R}
+        \begin{split}
+        R(y, z) = &\log{(y)}\log{(z)} - \log{(y)}\log{(1-y)} - \log{(z)}\log{(1-z)} \\
+        &+ \dfrac{\pi^{2}}{6} - \mathrm{Li}_{2}(y) - \mathrm{Li}_{2}(z) \, .
+        \end{split}
+    \end{equation}
+    It will also be useful to introduce the variables $y_{ij} = s_{ij}/s_{ijk}$.
+
+    \subsection{Quark-gluon antenna functions}
+    \subsubsection{Three-parton tree-level antenna function}
+    \begin{equation}\label{eqn:D30}
+        \begin{split}
+        D_{3}^{0}(1_{q},3_{g},4_{g}) = \dfrac{1}{s_{134}^{2}}&\left(\dfrac{2s_{134}^{2}s_{14}}{s_{13}s_{34}} + \dfrac{2s_{134}^{2}s_{13}}{s_{14}s_{34}} + \dfrac{s_{14}s_{34}+s_{34}^{2}}{s_{13}}\right.
+\\
+        & \left. + \dfrac{s_{13}s_{34}+s_{34}^{2}}{s_{14}} + \dfrac{2s_{13}s_{14}}{s_{34}} + 5s_{134} + s_{34}\right) + \mathcal{O}(\epsilon) \, .
+        \end{split}
+    \end{equation}
+    \subsubsection{Three-parton one-loop antenna functions}
+    \begin{align}\label{eqn:D31}
+        \mathcal{P}oles\left(D_{3}^{1}(1_{q},3_{g},4_{g})\right) =& \; 2\left(\mathbf{I}_{qg}^{(1)}(\epsilon, s_{13}) + \mathbf{I}_{qg}^{(1)}(\epsilon, s_{14}) + \mathbf{I}_{gg}^{(1)}(\epsilon, s_{34}) \right.
+\nonumber \\
+        &\left. \quad -2\mathbf{I}_{qg}^{(1)}(\epsilon, s_{134}) \vphantom{\mathbf{I}_{gg}^{(1)}} \right) D_{3}^{0}(1_{q},3_{g},4_{g}) \, , \\
+        \mathcal{F}inite\left(D_{3}^{1}(1_{q},3_{g},4_{g})\right) =& \; -\left(R(y_{13},y_{34})+R(y_{14},y_{34})+R(y_{13},y_{14}) + \dfrac{5}{3}\log{y_{13}}\nonumber \right. \\
+        &\left. \vphantom{\dfrac{1}{1}} \quad + \dfrac{5}{3}\log{y_{14}} + \dfrac{11}{6}\log{y_{34}}\right) D_{3}^{0}(1_{q},3_{g},4_{g}) + \dfrac{1}{3s_{34}} \, .
+    \end{align}
+
+    \subsection{Gluon-gluon antenna functions}
+    \subsubsection{Three-parton tree-level antenna function}
+    \begin{align}\label{eqn:F30}
+        F_{3}^{0}(3_{g},4_{g},5_{g}) =& \; \dfrac{2}{s_{345}^{2}}\left(\dfrac{s_{345}^{2}s_{34}}{s_{35}s_{45}} + \dfrac{s_{345}^{2}s_{35}}{s_{34}s_{45}} + \dfrac{s_{345}^{2}s_{45}}{s_{34}s_{35}} \right.
+\nonumber \\
+        &\left.
+\vphantom{\dfrac{s_{345}^{2}s_{34}}{s_{35}s_{45}}} \quad \quad +\dfrac{s_{34}s_{35}}{s_{45}} + \dfrac{s_{34}s_{45}}{s_{35}} + \dfrac{s_{35}s_{45}}{s_{34}} + 4s_{345} + \mathcal{O}(\epsilon) \right) \, .
+    \end{align}
+    \subsubsection{Three-parton one-loop antenna functions}
+    \begin{align}\label{eqn:F31}
+        \mathcal{P}oles\left(F_{3}^{1}(3_{g},4_{g},5_{g})\right) =& \; 2\left(\mathbf{I}_{gg}^{(1)}(\epsilon,s_{34}) + \mathbf{I}_{gg}^{(1)}(\epsilon,s_{35}) + \mathbf{I}_{gg}^{(1)}(\epsilon,s_{45}) \right.
+\nonumber \\
+        &\left. \quad -2\mathbf{I}_{gg}^{(1)}(\epsilon,s_{345})
+        \right) F_{3}^{0}(3_{g},4_{g},5_{g}) \, , \\
+        \mathcal{F}inite\left(F_{3}^{1}(3_{g},4_{g},5_{g})\right) =& -\left(R(y_{34},y_{35}) + R(y_{35},y_{45}) + R(y_{34},y_{45}) \vphantom{\dfrac{1}{1}} \right. \nonumber \\
+        &\left. \quad + \, \dfrac{11}{6}\log{y_{34}} + \dfrac{11}{6}\log{y_{35}} + \dfrac{11}{6}\log{y_{45}} \right) F_{3}^{0}(3_{g},4_{g},5_{g}) \nonumber \\
+        & \quad + \dfrac{1}{3s_{34}} + \dfrac{1}{3s_{35}} + \dfrac{1}{3s_{45}} + \dfrac{1}{3s_{345}} \, .
+    \end{align}
 \end{document}

### 2022-10-05 qcd.tex
wrote: antenna expressions (A)

--- 2022-09-29 qcd.tex
@@
     antenna functions are categorised by their class and radiative
-    processes. A pictorial representation is also given in Figure ...
+    processes. A pictorial representation is also given in Figure~\ref{fig:antenna}.
 
@@
 
+    Below we list the antenna functions relevant
+    for multijet production in electron-positron
+    annihiliation with a single quark flavour,
+    namely the partonic channel $e^{-}e^{+} \rightarrow
+    d \bar{d}$ + $ng$.
+    For this we require the $A$, $D$, and $F$ antenna functions,
+    see Table~\ref{table:antenna_functions}.
+    The $i,j,k$ indices will be fixed such that $d = 1$, $\bar{d} = 2$,
+    and $g \in \{3, 4, 5\}$.
+
+    It will be useful to define the $\mathcal{P}oles$
+    and $\mathcal{F}inite$ operators to extract the
+    singular and finite contributions from an antenna
+    function. However, there are still finite terms
+    from the $\mathcal{P}oles$ operator coming from
+    the IR singularity operators (cite here). Therefore,
+    care has to be taken to extract all the finite
+    remainders from the antenna functions.
+
+    For the one-loop antenna functions, there are three
+    different functions corresponding to the: leading colour,
+    subleading colour, and closed quark look contributions.
+    It is sufficient to consider the leading colour
+    structure in this thesis as the remaining contributions
+    are subleading. They become important when
+    building subtraction terms to exactly cancel divergences
+    but we do not require this level of accuracy for
+    our applications.
+
+    \subsection{Quark-antiquark antenna functions}
+    The quark-antiquark antenna functions are derived by normalising
+    the colour-ordered QCD radiative corrections to $\gamma^{*} \rightarrow q \bar{q}$
+    at NNLO.
+
+    \subsubsection{Three-parton tree-level antenna function}
+    The tree-level three-parton antenna function is given by
+    \begin{equation}\label{eqn:A30}
+        A_{3}^{0}(1_{q}, 3_{g}, 2_{\bar{q}}) = \dfrac{1}{s_{123}}\left(\dfrac{s_{13}}{s_{23}} + \dfrac{s_{23}}{s_{13}} + 2\dfrac{s_{12}s_{123}}{s_{13}s_{23}}\right) + \mathcal{O}(\epsilon)
+    \end{equation}
+    where the IR structure is apparent. The first two terms
+    correspond to the gluon going collinear with the quark
+    and antiquark, where as the final term accounts for the soft
+    limit.
+
+    \subsubsection{Three-parton one-loop antenna function}
+    The leading colour one-loop antenna functions are
+    given by
+    \begin{align}\label{eqn:A31}
+        \mathcal{P}oles\left(A_{3}^{1}(1_{q}, 3_{g}, 2_{\bar{q}})\right) =& \; 2\left(\mathbf{I}_{qg}^{(1)}(\epsilon, s_{13}) + \mathbf{I}_{qg}^{(1)}(\epsilon, s_{23}) - \mathbf{I}_{q\bar{q}}^{(1)}(\epsilon,s_{123})\right)A_{3}^{0}(1_{q},3_{g},2_{\bar{q}}) \, , \\
+        \mathcal{F}inite\left(A_{3}^{1}(1_{q}, 3_{g}, 2_{\bar{q}})\right) =& -\left(R(y_{13}, y_{23}) + \dfrac{5}{3}\log{y_{13}}+\dfrac{5}{3}\log{y_{23}}\right)A_{3}^{0}(1_{q}, 3_{g}, 2_{\bar{q}}) \nonumber\\
+        &+\dfrac{1}{s_{123}} + \dfrac{s_{12}+s_{23}}{2s_{123}s_{13}} + \dfrac{s_{12}+s_{13}}{2s_{123}s_{23}} \nonumber \\
+        &-\dfrac{s_{13}}{2s_{123}(s_{12}+s_{13})}-\dfrac{s_{23}}{2s_{123}(s_{12}+s_{23})} \nonumber \\
+        &+\dfrac{\log{y_{13}}}{s_{123}}\left(2-\dfrac{1}{2}\dfrac{s_{13}s_{23}}{(s_{12}+s_{23})^{2}} + 2\dfrac{s_{13}-s_{23}}{s_{12}+s_{23}}\right) \nonumber  \\
+        &+\dfrac{\log{y_{23}}}{s_{123}}\left(2-\dfrac{1}{2}\dfrac{s_{13}s_{23}}{(s_{12}+s_{13})^{2}} + 2\dfrac{s_{23}-s_{13}}{s_{12}+s_{13}}\right) \, , 
+    \end{align}
+
 \end{document}

### 2022-10-05 qcd.tex

--- 2022-10-04 qcd.tex
@@
 
+    Below we list the antenna functions relevant
+    for multijet production in electron-positron
+    annihiliation with a single quark flavour,
+    namely the partonic channel $e^{-}e^{+} \rightarrow
+    d \bar{d}$ + $ng$.
+    For this we require the $A$, $D$, and $F$ antenna functions,
+    see Table~\ref{table:antenna_functions}.
+    The $i,j,k$ indices will be fixed such that $d = 1$, $\bar{d} = 2$,
+    and $g \in \{3, 4, 5\}$.
+
+    It will be useful to define the $\mathcal{P}oles$
+    and $\mathcal{F}inite$ operators to extract the
+    singular and finite contributions from an antenna
+    function. However, there are still finite terms
+    from the $\mathcal{P}oles$ operator coming from
+    the IR singularity operators (cite here). Therefore,
+    care has to be taken to extract all the finite
+    remainders from the antenna functions.
+
+    For the one-loop antenna functions, there are three
+    different functions corresponding to the: leading colour,
+    subleading colour, and closed quark look contributions.
+    It is sufficient to consider the leading colour
+    structure in this thesis as the remaining contributions
+    are subleading. They become important when
+    building subtraction terms to exactly cancel divergences
+    but we do not require this level of accuracy for
+    our applications.
+
     The antenna functions have some functions in common,
@@
     It will also be useful to introduce the variables $y_{ij} = s_{ij}/s_{ijk}$.
+
+    \subsection{Quark-antiquark antenna functions}
+    The quark-antiquark antenna functions are derived by normalising
+    the colour-ordered QCD radiative corrections to $\gamma^{*} \rightarrow q \bar{q}$
+    at NNLO.
+
+    \subsubsection{Three-parton tree-level antenna function}
+    The tree-level three-parton antenna function is given by
+    \begin{equation}\label{eqn:A30}
+        A_{3}^{0}(1_{q}, 3_{g}, 2_{\bar{q}}) = \dfrac{1}{s_{123}}\left(\dfrac{s_{13}}{s_{23}} + \dfrac{s_{23}}{s_{13}} + 2\dfrac{s_{12}s_{123}}{s_{13}s_{23}}\right) + \mathcal{O}(\epsilon)
+    \end{equation}
+    where the IR structure is apparent. The first two terms
+    correspond to the gluon going collinear with the quark
+    and antiquark, where as the final term accounts for the soft
+    limit.
+
+    \subsubsection{Three-parton one-loop antenna function}
+    The leading colour one-loop antenna functions are
+    given by
+    \begin{align}\label{eqn:A31}
+        \mathcal{P}oles\left(A_{3}^{1}(1_{q}, 3_{g}, 2_{\bar{q}})\right) =& \; 2\left(\mathbf{I}_{qg}^{(1)}(\epsilon, s_{13}) + \mathbf{I}_{qg}^{(1)}(\epsilon, s_{23}) - \mathbf{I}_{q\bar{q}}^{(1)}(\epsilon,s_{123})\right)A_{3}^{0}(1_{q},3_{g},2_{\bar{q}}) \, , \\
+        \mathcal{F}inite\left(A_{3}^{1}(1_{q}, 3_{g}, 2_{\bar{q}})\right) =& -\left(R(y_{13}, y_{23}) + \dfrac{5}{3}\log{y_{13}}+\dfrac{5}{3}\log{y_{23}}\right)A_{3}^{0}(1_{q}, 3_{g}, 2_{\bar{q}}) \nonumber\\
+        &+\dfrac{1}{s_{123}} + \dfrac{s_{12}+s_{23}}{2s_{123}s_{13}} + \dfrac{s_{12}+s_{13}}{2s_{123}s_{23}} \nonumber \\
+        &-\dfrac{s_{13}}{2s_{123}(s_{12}+s_{13})}-\dfrac{s_{23}}{2s_{123}(s_{12}+s_{23})} \nonumber \\
+        &+\dfrac{\log{y_{13}}}{s_{123}}\left(2-\dfrac{1}{2}\dfrac{s_{13}s_{23}}{(s_{12}+s_{23})^{2}} + 2\dfrac{s_{13}-s_{23}}{s_{12}+s_{23}}\right) \nonumber  \\
+        &+\dfrac{\log{y_{23}}}{s_{123}}\left(2-\dfrac{1}{2}\dfrac{s_{13}s_{23}}{(s_{12}+s_{13})^{2}} + 2\dfrac{s_{23}-s_{13}}{s_{12}+s_{13}}\right) \, , 
+    \end{align}
 
@@
     \end{align}
+
 \end{document}

### 2022-10-07 intro.tex
wrote: colour-ordered and one-loop factorisation
wrote: antenna functions
wrote: monte carlo

@@
 
-    \begin{figure}\label{fig:qcd_feynman_rules}
+    \begin{figure}
         \includegraphics[width=\linewidth]{introduction/qcd_feynman_rules.pdf}
         \caption{QCD Feynman rules.}
+        \label{fig:qcd_feynman_rules}
     \end{figure}
@@
     \end{equation}
-    where $P_{ba}(\dfrac{x}{z})$ are splitting functions
+    where $P_{ba}(\frac{x}{z})$ are splitting functions
     representing parton $b$ emitting a parton $a$ carrying
@@
     the renormalisation scale, $\mu_{R}$
-    (see \ref{chapter:qcd}).
-    {\color{red} Need to mention how these scales
-    are unphysical but there is a residual dependence
-    on them at fixed-order calculations. Scale
-    variations are a conventional way to determine
-    the effects of higher order terms.
-    Renormalisation will be discussed in Chapter 2
-    when divergences are introduced.}
+    (see Section~\ref{sec:renormalisation}). This will
+    be discussed in Section~\ref{sec:scale_variations}
+    in the context of theoretical uncertainties.
 
@@
     $\mathcal{M}$, is defined in the following expression
-    \begin{equation}\label{eqn:matrix_element}
+    \begin{equation}\label{eqn:S_matrix_element}
         \langle f | \mathcal{T} | i \rangle = (2\pi)^{4} \delta^{4}\left(p_{a} + p_{b} - \sum_{f=1}^{n} p_{f}^{\mu}\right) \mathcal{M} \, ,
@@
     cross-sections complete, we will move the discussion onto
-    Monte Carlo event generators. These event
-    generators are multipurpose tools that
-    incorporate many aspects of the theory discussed
-    so far (and to be discussed) for practical use.
-    Most importantly, they are
-    used to compare theoretical predictions to
-    experimental measurements.
+    more practical aspects of computations within the
+    QCD framework. More specifically, we will discuss
+    the divergent structure of matrix elements
+    and the machinery developed to tackle these
+    unphysical singularities.
 
-\section{Monte Carlo event generator overview}
-\subsection{Event unweighting}
-\begin{itemize}
-    \item MC generators are the main tools that
-    people use to compare with experimental data.
-    There are many groups that have their own tool.
-    \item The usage of MC tools came about because
-    the multi-dimensional phase-space integrals
-    are too complex to do analytically.
-    \item This doesn't mean that they only do
-    the integrals, the event generators basically
-    contain all the tools from phase-space generation
-    to matrix element evaluation, to parton showers
-    to hadronisation.
-    \item Have a look at Katherina thesis to see
-    how much I can skip because only really want to
-    introduce unweighting. Probably along the lines
-    of event weight is statistically sampled vs
-    what they measure in real life. Therefore we need
-    to unweight events.
-\end{itemize}
 \end{document}

### 2022-10-07 mc.tex
wrote: monte carlo

\documentclass[main.tex]{subfiles}
\begin{document}
\chapter{Monte Carlo Event Generators}\label{chap:MC}
    At a particle collider experiment,
    the collisions between incoming particles can lead
    to highly complex final-states with a large number
    of particles. In a hadron collider, such as the LHC,
    the particles that we observe are not the constituents
    of the proton, the quarks and gluons. Instead we observe
    hadrons which are the bound states of these partons
    due to the property of confinement at low energy scales.

    In order to describe the entire chain of events
    from the initial energetic collision to the production
    of the lower energy hadrons, it is customary to
    split up the process into stages characterised by
    their kinematic regime. The standard tool used
    to simulate these stages of an event are general-purpose
    Monte Carlo event generators (event genrators), see
    Reference \cite{Buckley:2011ms} for a review.
    They are indispensable tools that help
    bridge the gap between theoretical predictions
    and experimental measurements with nearly all
    analyses made at the ATLAS and CMS experiments
    involving some form of event generator usage.

    General-purpose event generators available
    such as Herwig \cite{Bahr:2008pv,Bellm:2015jjp},
    Pythia \cite{Sjostrand:2006za,Bierlich:2022pfr},
    and Sherpa \cite{Gleisberg:2008ta,Sherpa:2019gpd} all broadly share 
    the same features at a high-level overview.

    In the following we will motivate the necessity
    for Monte Carlo methods and give a brief overview
    of the main stages involved in simulating an event.
    We will give particular attention to the concept of event
    unweighting which is required to compare probabilistically
    generated events to experimental results.

    \section{Monte Carlo integration}\label{sec:MC_integration}
    As seen in Section~\ref{sec:xs}, the task of
    computing partonic cross-sections boils down
    to integrating the matrix elements over the
    final-state phase-space. Whilst it is possible to
    carry out these integrals analytically for the simplest
    cases, the integral quickly becomes intractable when the
    final-state multiplicity becomes large. For an $n$-body
    final-state the dimensionality of the integral scales as $d=3n-4$,
    where momentum conservation and on-shell conditions
    have already been applied to reduce the dimensionality.
    Therefore, instead of carrying out the integrals analytically,
    numerical methods become more attractive. The current
    prevailing method is Monte Carlo integration which
    has no scaling with the dimensionality of the integral,
    so can be used for arbitrarily high multiplicity final-states.
    The purpose of this chapter is to give a brief overview
    of event generators, a more indepth treatment of the
    specifics of Monte Carlo methods is given in Reference
    \cite{James:1980yn}.

    The basic idea of Monte Carlo integration is to approximate
    the integral of a function $f(\vec{x})$ by sampling it
    at $d$-dimensional parameter space points $\vec{x}$
    \begin{equation}\label{eqn:MC_integration}
        I = \int_{V} \mathrm{d}^{D}x \; f(\vec{x}) \approx \langle I(f) \rangle_{x} = V \dfrac{1}{N} \sum_{i=1}^{N} f(\vec{x}_{i}) = V \langle f \rangle_{x} \, ,
    \end{equation}
    where the points $\vec{x}$ are randomly (hence Monte Carlo),
    uniformly sampled in the volume $V$. $\langle f \rangle_{x}$
    denotes the average of $f$ over these uniformly sampled
    points. The law of large numbers ensures that the
    estimator $\langle I(f) \rangle_{x}$ will converge to
    the true value $I$ for $N \rightarrow \infty$. As a
    result of the central limit theorem, the function evaluations
    $f(\vec{x}_{i})$ will follow a Gaussian distribution with
    mean $\langle I(f) \rangle_{x}$,
    meaning an error esimate on the estimator is given by
    the standard error of the mean
    \begin{equation}\label{eqn:MC_error}
        \sigma_{I} = V \sqrt{\dfrac{\langle f^{2} \rangle_{x} - \langle f \rangle_{x}^{2}}{N}}
    \end{equation}
    which falls as $1/\sqrt{N}$, with no dependence
    on $d$.

    An algorithm that samples points $\vec{x}$ is called
    an integrator. To make use of the random numbers sampled,
    they need to be mapped to quantities that are useful in a simulation
    setting. The most commonly used mapping is to map
    the random numbers to four-momenta. An example of a
    simple integrator is the RAMBO algorithm \cite{Kleiss:1985gy,Platzer:2013esa}
    which samples four-momenta space uniformly. Since the integrator
    now effectively samples in momentum space, we refer to the points
    generated as phase-space points and so the terms integrator
    and phase-space generator are used interchangeably.

    Whilst Monte Carlo integration is independent of the
    dimensionality of the integral, meaning convergence is
    guaranteed for large enough $N$, there are variance
    reduction methods to more rapidly reduce the error
    on the estimator.

    One such method is importance sampling which
    aims to reduce the variance via a remapping of the
    uniformly sampled variables $\vec{x}$
    to a more suitable non-uniform distribution.
    In the context of particle physics,
    this corresponds to using a mapping of random numbers
    to four-momenta that captures the distribution of the
    matrix elements and their divergences, i.e. samples
    more frequently in regions where the matrix element
    is large in magnitude. Integrators inspired by the pole
    structure of multi-parton QCD processes have been
    studied in References \cite{Draggiotis:2000gm,vanHameren:2002tc}.

    Another common variance reduction technique is
    stratified sampling which divides the integration
    volume into sub-volumes, or bins. The overall integral
    and variances are the sums of the partial results
    in each bin. The overall variance is minimised when
    the the contributions from each bin are equal. This
    is achieved by sampling more from bins where the integral
    is rapidly fluctuating (and not necesarily large),
    and less from bins where the integral has less fluctuations.

    In practice, state-of-the-art event generators
    combine the concepts of importance sampling and stratified
    sampling with adaptive sampling \cite{Lepage:1977sw,Lepage:2020tgj}
    and multi-channelling methods \cite{Kleiss:1994qy,Ohl:1998jn}
    to achieve faster rates of convergence.

    \section{Event unweighting}\label{sec:unweighting}
    Once a suitable configuration of momenta
    has been found by the integrator, this point forms the basis of an
    event. The matrix element of interest can be evaluated at
    this phase-space point at the desired order in perturbation
    theory (see Section~\ref{sec:me_generators}). Additionally, there will be an associated Jacobian
    from the mapping from random variables to the four-momenta.
    This Jacobian is referred to as the phase-space weight, which together
    with the matrix element forms the event weight\footnote{For
    the more sophisticated sampling methods employed in event generators
    there will be additional weights involved.}
    \begin{equation}\label{eqn:event_weight}
        w_{i} = |\mathcal{M}_{i}|^{2} J_{i} \, .
    \end{equation}
    The event weight can be interpreted as the probability
    of this particular event occurring at an experiment \cite{Plehn:2009nd}.
    By generating a number of events in this manner,
    and histogramming them, a sample of weighted events is created.

    At experiments, measured events are either recorded
    or they are not, i.e. events come with unit-weight,
    not a variable weight as they are currently being
    generated with. In order to generate events that
    compare to reality, it has become customary to translate
    weighted events into unit-weight events.
    This process is known as unweighting and is typically
    carried out using a rejection sampling algorithm
    outlined in Algorithm~\ref{alg:unweighting}. The rejection
    sampling algorithm makes use of $w_{\mathrm{max}}$ which is the
    maximal event weight in the integration volume.

    \begin{algorithm}
        \caption{Rejection sampling for unweighting events}\label{alg:unweighting}
        \While{unweighting}{
            generate random phase-space point $\vec{x}$\;
            evaluate event weight $w \gets w(\vec{x})$\;
            generate uniform random number $r \gets$ Random(0, 1)\;
            \If{$w / w_{\mathrm{max}} > r$}{
                \KwRet $\vec{x}$ and $\tilde{w} \gets \max(1,\; w / w_{\mathrm{max}})$\;
            }
        }
    \end{algorithm}

    From this description, it is clear that the number
    of accepted events will generally be lower than
    the number of total events generated. This is encapsulated
    in the unweighting efficiency
    \begin{equation}\label{eqn:unweighting_efficiency}
        \varepsilon = \dfrac{N_{\mathrm{accepted}}}{N_{\mathrm{total}}} \, ,
    \end{equation}
    whose inverse is the average number of trial
    events required to obtain a single unweighted event.
    Unweighting efficiency for high-multiplicity processes
    are often well below 1\% \cite{Hoche:2019flt,Gao:2020zvv} and development
    of methods to improve the efficiency have been
    explored \cite{Jadach:1999sf,Jadach:2002kn}.

    An alternative to improving the unweighting efficiency
    itself is to accelerate the process of unweighting. This
    was explored in Reference \cite{Danziger:2021eeg} where a
    novel two-stage unweighting procedure employing a surrogate
    model of the exact event weight was used to decrease the amount of
    time spent unweighting events. This idea will be elaborated on
    in Chapter~\ref{chapter:fame3} where we build upon
    Reference \cite{Danziger:2021eeg} by increasing the fidelity of
    the surrogate model.

    \section{Anatomy of an event}\label{sec:anatomy_event}
    The simulation of an event can be split up into three
    main stages: the hard scattering process, followed by
    the parton shower, and finally hadronisation.
    This is motivated by the factorisation theorem
    which separates the high energy regimes from the low
    energy regimes.
    \subsection*{Hard scattering}\label{sec:me_generators}
    The first stage of an event simulation begins with
    the computation of the matrix element. The partonic
    channel of interest sets the multiplicity of the process,
    therefore informing the integrator on how to sample the phase-space
    accordingly.

    As seen in (\ref{eqn:matrix_element}),
    the exact matrix element is approximated by a perturbative
    series in a coupling constant. The calculations of
    tree-level (cite here) and one-loop (cite here)
    matrix elements are now largely fully automated
    in matrix element generators with a common generci
    interface, the Binoth Les Houches Accord (cite here).
    For more information on the machinery developed to
    compute nearly arbitrary matrix elements, see Reference (cite here).
    NNLO and N$^{3}$LO QCD corrections have been computed
    for select processes but have yet to be reach the
    level of automation of NLO.
    For a review of recent advancements see References (cite here).

    Matrix element calculations make up a large
    portion of the total event generation time (cite here).
    Reducing the time spent in this stage of the event
    generation therefore represents a sizeable increase
    in the number of events generated given a fixed computing
    budget. This thesis presents a novel technique to
    accelerate matrix element calculations by building
    accurate emulation models from modern machine learning
    techniques.

    \subsection*{Parton shower}\label{sec:parton_showers}
    Fixed-order calculations of matrix elements are limited
    to relatively few final-state particles due to the complexity
    of analytical expressions and are only accurate in the
    high energy limit. To evolve from the fixed-order
    calculation to the non-perturbative regime where hadrons
    are formed, parton showers are employed in event generators.

    Parton showers are implemented as algorithms that
    iteratively emit radiation (gluons and quarks in QCD
    parton showers) described by splitting functions.
    By repeating this procedure of emissions, parton
    showers produce a cascade of partons that naturally
    evolves the energy scale from the hard process to
    an infrared scale where non-perturbative effects
    begin to set in. In doing so parton showers account
    for potentially large logarithms that arise in
    these kinematic regimes at all orders in the coupling
    parameter, this is known as resummation.

    In event generators, fixed-order matrix elements
    and parton showers have to be utilised together to describe
    the many different observables at experiments. In order
    to combine the two stages of event generation in a consistent
    manner there are matching (cite here) and merging (cite here)
    schemes.

    \subsection*{Hadronisation}\label{sec:hadronisation}
    The formation of hadrons in the low energy regime
    of an event generator follow non-perturbative models.
    The two main classes of models used to simulate
    hadronisation are the cluster models (cite here)
    and string models (cite here).
    These models share a similar concept of grouping together
    quarks and antiquarks to ensure only hadrons remain
    in the final-state. For more discussion on hadronisation
    see References (cite here).

    \section{A note on theoretical uncertainties}\label{sec:scale_variations}
    The partonic cross-section, $\hat{\sigma}(\mu_{F}, \mu_{R})$ has
    a dependence on both the factorisation scale and the
    renormalisation scale. Although these scales are unphysical,
    and so any observable should not depend on them,
    there is a residual dependence on them in fixed-order
    calculations due to missing higher order terms.

    The most widely adopted approach to estimate the
    theoretical uncertainties associated with these
    missing higher order terms is to carry out a seven-point
    scale variation. Typically in a calculation we take
    $\mu = \mu_{F} = \mu_{R}$ where $\mu$ depends on the specifics
    of the process of interest. Taking the value of the
    partonic cross-section at this scale as the central value,
    the seven variations correspond to multiplying $\mu_{F}$
    and $\mu_{R}$ by the factors
    \begin{equation}\label{eqn:scale_variations}
        \dfrac{\mu_{i}}{\mu} = \left\{(1, 1), (2,2),(0.5,0.5),(2,1),(1,2),(1,0.5),(0.5,1)\right\} \, .
    \end{equation}
    The estimate of the uncertainty is then taken
    to be the half-width of the maximum interval.

\end{document}

### 2022-10-07 qcd.tex
wrote: monte carlo

--- 2022-10-04 qcd.tex
@@
 
+    Below we list the antenna functions relevant
+    for multijet production in electron-positron
+    annihiliation with a single quark flavour,
+    namely the partonic channel $e^{-}e^{+} \rightarrow
+    q \bar{q}$ + $ng$.
+    For this we require the $A$, $D$, and $F$ antenna functions,
+    see Table~\ref{table:antenna_functions}.
+    The $i,j,k$ indices will be fixed such that $q = 1$, $\bar{q} = 2$,
+    and $g \in \{3, 4, 5\}$ in view of using the
+    antenna functions in a calculation where the momenta
+    of partons will have fixed indices in a phase-space
+    point.
+
+    It will be useful to define the $\mathcal{P}oles$
+    and $\mathcal{F}inite$ operators to extract the
+    singular and finite contributions from an antenna
+    function. However, there are still finite terms
+    from the $\mathcal{P}oles$ operator coming from
+    the IR singularity operators (cite here). Therefore,
+    care has to be taken to extract all the finite
+    remainders from the antenna functions.
+
+    For the one-loop antenna functions, there are three
+    different functions corresponding to the: leading colour,
+    subleading colour, and closed quark look contributions.
+    It is sufficient to consider the leading colour
+    structure in this thesis as the remaining contributions
+    are subleading. Ignoring subleading terms in
+    colour is equivalent to taking the large $N_{c}$
+    limit. For QCD where $N_{c} = 3$, neglecting
+    subleading colour terms amounts to an
+    approximately 10\% effect.
+
     The antenna functions have some functions in common,
@@
 
+    \subsection{Quark-antiquark antenna functions}
+    The quark-antiquark antenna functions are derived by normalising
+    the colour-ordered QCD radiative corrections to $\gamma^{*} \rightarrow q \bar{q}$
+    at NNLO.
+    \subsubsection{Three-parton tree-level antenna function}
+    \begin{equation}\label{eqn:A30}
+        A_{3}^{0}(1_{q}, 3_{g}, 2_{\bar{q}}) = \dfrac{1}{s_{123}}\left(\dfrac{s_{13}}{s_{23}} + \dfrac{s_{23}}{s_{13}} + 2\dfrac{s_{12}s_{123}}{s_{13}s_{23}}\right) + \mathcal{O}(\epsilon)
+    \end{equation}
+
+    \subsubsection{Three-parton one-loop antenna function}
+    \begin{align}\label{eqn:A31}
+        \mathcal{P}oles\left(A_{3}^{1}(1_{q}, 3_{g}, 2_{\bar{q}})\right) =& \; 2\left(\mathbf{I}_{qg}^{(1)}(\epsilon, s_{13}) + \mathbf{I}_{qg}^{(1)}(\epsilon, s_{23}) \right. \nonumber \\
+        &\quad \left. -\mathbf{I}_{q\bar{q}}^{(1)}(\epsilon,s_{123})\right)A_{3}^{0}(1_{q},3_{g},2_{\bar{q}}) \, , \\
+        \mathcal{F}inite\left(A_{3}^{1}(1_{q}, 3_{g}, 2_{\bar{q}})\right) =& -\left(R(y_{13}, y_{23}) + \dfrac{5}{3}\log{y_{13}}+\dfrac{5}{3}\log{y_{23}}\right)A_{3}^{0}(1_{q}, 3_{g}, 2_{\bar{q}}) \nonumber\\
+        &\quad +\dfrac{1}{s_{123}} + \dfrac{s_{12}+s_{23}}{2s_{123}s_{13}} + \dfrac{s_{12}+s_{13}}{2s_{123}s_{23}} \nonumber \\
+        &\quad -\dfrac{s_{13}}{2s_{123}(s_{12}+s_{13})}-\dfrac{s_{23}}{2s_{123}(s_{12}+s_{23})} \nonumber \\
+        &\quad +\dfrac{\log{y_{13}}}{s_{123}}\left(2-\dfrac{1}{2}\dfrac{s_{13}s_{23}}{(s_{12}+s_{23})^{2}} + 2\dfrac{s_{13}-s_{23}}{s_{12}+s_{23}}\right) \nonumber  \\
+        &\quad +\dfrac{\log{y_{23}}}{s_{123}}\left(2-\dfrac{1}{2}\dfrac{s_{13}s_{23}}{(s_{12}+s_{13})^{2}} + 2\dfrac{s_{23}-s_{13}}{s_{12}+s_{13}}\right) \, .
+    \end{align}
+
     \subsection{Quark-gluon antenna functions}
+    The quark-gluon antenna functions are obtained by considering
+    the QCD real radiation corrections in the process $\tilde{\chi} \rightarrow \tilde{g} g$.
     \subsubsection{Three-parton tree-level antenna function}
@@
     \subsection{Gluon-gluon antenna functions}
+    The gluon-gluon antenna functions are obtained from the
+    QCD real radiation corrections in the process $H \rightarrow g g$.
     \subsubsection{Three-parton tree-level antenna function}
@@
     \end{align}
+    \subsection{Limiting behaviour of antenna functions}
+    Examining (\ref{eqn:A30}) more closely, it becomes clear
+    that the correct IR behaviour is exhibited by the antenna
+    function. The first two terms correspond to the gluon
+    going collinear with the quark and antiquark. The final
+    term encapsulates singularity in the limit of the gluon
+    going soft.
+
+    The other antenna functions have similar limiting behaviour,
+    collapsing to the single collinear and soft splitting functions
+    in the relevant regions of phase-space.
 \end{document}

### 2022-10-10 mc.tex
add references (MC)

@@
     explored \cite{Jadach:1999sf,Jadach:2002kn}.
+    These low efficiencies lead to most of the
+    yearly CPU budgets of LHC experiments to be spent on
+    unweighting events \cite{HSFPhysicsEventGeneratorWG:2020gxw}.
 
@@
     series in a coupling constant. The calculations of
-    tree-level (cite here) and one-loop (cite here)
+    tree-level \cite{Mangano:2002ea,Krauss:2001iv,Gleisberg:2008fv,Cafarella:2007pc,Alwall:2011uj,Kilian:2007gr}
+    and one-loop \cite{Berger:2008sj,Cullen:2011ac,Cullen:2014yla,Bevilacqua:2011xh,Hirschi:2011pa,Alwall:2014hca,Frederix:2018nkq,Campbell:1999ah,Campbell:2021vlt,Badger:2012pg,Cascioli:2011va,Buccioni:2019sur,Actis:2016mpe,Denner:2017wsf}
     matrix elements are now largely fully automated
-    in matrix element generators with a common generci
-    interface, the Binoth Les Houches Accord (cite here).
+    in NLO QCD and electroweak corrections.
+    Many of the modern matrix element generators are supplied
+    with a common generic interface to event generators,
+    the Binoth Les Houches Accord \cite{Binoth:2010xt,Alioli:2013nda}.
     For more information on the machinery developed to
-    compute nearly arbitrary matrix elements, see Reference (cite here).
+    compute nearly arbitrary matrix elements at tree-level and
+    one-loop level, see for instance Reference \cite{Campbell:2017hsr}.
     NNLO and N$^{3}$LO QCD corrections have been computed
@@
     level of automation of NLO.
-    For a review of recent advancements see References (cite here).
+    For a review of recent advancements see References \cite{Caola:2022ayt,Campbell:2022qmc}.
 
     Matrix element calculations make up a large
-    portion of the total event generation time (cite here).
+    portion of the total event generation time \cite{Bothmann:2022thx}.
     Reducing the time spent in this stage of the event
@@
     in the number of events generated given a fixed computing
-    budget. This thesis presents a novel technique to
+    budget \cite{HSFPhysicsEventGeneratorWG:2020gxw}.
+    This thesis presents a novel technique to
     accelerate matrix element calculations by building
@@
     these kinematic regimes at all orders in the coupling
-    parameter, this is known as resummation.
+    parameter, this is known as resummation \cite{Buckley:2011ms}.
 
@@
     to combine the two stages of event generation in a consistent
-    manner there are matching (cite here) and merging (cite here)
+    manner there are matching \cite{Frixione:2002ik,Frixione:2007vw,Jadach:2015mza}
+    and merging \cite{Catani:2001cc,Lonnblad:2011xx}
     schemes.
@@
     The two main classes of models used to simulate
-    hadronisation are the cluster models (cite here)
-    and string models (cite here).
+    hadronisation are the cluster models \cite{Webber:1983if,Winter:2003tt}
+    and string models \cite{Andersson:1983ia}.
     These models share a similar concept of grouping together
@@
     in the final-state. For more discussion on hadronisation
-    see References (cite here).
+    see for example Reference \cite{Webber:1999ui}.
 
@@
     \begin{equation}\label{eqn:scale_variations}
-        \dfrac{\mu_{i}}{\mu} = \left\{(1, 1), (2,2),(0.5,0.5),(2,1),(1,2),(1,0.5),(0.5,1)\right\} \, .
+        \dfrac{\mu_{i}}{\mu} = \left\{(1, 1), (2,2),(0.5,0.5),(2,1),(1,2),(1,0.5),(0.5,1)\right\} \, ,
     \end{equation}
+    and re-evaluating the
+    cross-section at these scales.
     The estimate of the uncertainty is then taken

### 2022-10-13 ml.tex
wrote: ml intro

@@
 \begin{document}
-\chapter{Machine learning}
+\chapter{Modern machine learning algorithms and their uses}
 \label{chapter:ml}
-\section{Introduction to machine learning}
-\subsection{Supervised and unsupervised learning}
+\section{Rise of machine learning in HEP}
+    In high energy physics where analyses
+    depend on making use of large computing clusters
+    and international grid efforts, advancements
+    in technology and more efficient methods are
+    required to keep up with the needs of the
+    experimental and theoretical community.
+    However, the computational resources needed
+    are at risk of outpacing the growth in these research
+    areas \cite{Bothmann:2022thx}. Therefore new techniques and
+    algorithms from the machine learning community
+    have been adopted to augment the ongoing efforts
+    in the high energy physics community. Machine
+    learning is a category of artificial intelligence
+    concerned with the design of algorithms that
+    automates learning from data.
+    By now machine learning is utilised throughout
+    every stage of particle physics analysis, from collection
+    of data from experiments all the way to novel
+    applications on the theory side.
+
+    Around the turn of the last decade the culmination
+    in advancement in hardware and techniques led to an
+    explosion in image classification research in the
+    machine learning community, specifically in the
+    utilisation of very large neural networks. The accuracy
+    of these deep neural networks greatly outperformed the
+    previous state-of-the-art \cite{NIPS2012_c399862d,https://doi.org/10.48550/arxiv.1409.0575} and
+    lead to the widespread use of neural networks for a
+    wide array of tasks \cite{Schmidhuber_2015}.
+    This led to the rise of programming
+    frameworks such as scikit-learn \cite{scikit-learn},
+    TensorFlow \cite{https://doi.org/10.48550/arxiv.1603.04467},
+    PyTorch \cite{https://doi.org/10.48550/arxiv.1912.01703},
+    and XGBoost \cite{Chen:2016:XST:2939672.2939785}
+    to name a few which implement neural networks, alongside
+    other mainstay algorithms such as decisions trees
+    and support vector machines.
+    These algorithms are now extremely commonplace
+    in many domains, including in particle physics.
+
+    Collisions at the LHC can produce hundreds of
+    particles with complex final-state configurations
+    leading to the design of the ATLAS and CMS experiments
+    to contain of the order 100 million detection elements
+    in an attempt to disentangle these events. With these
+    large arrays of detection elements there is an enormouse
+    amount of data harvested from each collision, terabytes
+    per second, which presents a challenge in the form of
+    data collection.
+    Already at this stage, boosted decision trees
+    have been used to enhance triggers \cite{CMS:2020cmk} to accept or
+    reject data from being saved to disk. At the LHCb
+    experiment 70\% of all data retained have had
+    some involvement with machine learning algorithms \cite{LHCb:2014set}.
+
+    On the theory side of things, there have been numerous
+    triumphs in the application of machine learning.
+    Perhaps the most well-known is the fitting
+    of PDFs performed by the NNPDF collaboration \cite{NNPDF:2021njg}.
+    This area is rapidly growing with many new
+    ideas being explored at each stage of the event
+    generation process.
+    A living review that is actively being updated
+    archiving advancements in the field is available
+    at \cite{Feickert:2021ajf}.
+    For more traditional reviews, see for instance
+    \cite{Guest:2018yhq,Radovic:2018dip,Butter:2022rso}.
+
+    In this chapter we give an overview of neural
+    networks: how they are constructed, how
+    they are trained, and how they are deployed.
+    We will then proceed to discuss their
+    applicability as surrogate models for matrix
+    elements.
+
 \section{Neural networks}
@@
 \subsection{Optimisation of weights}
-\subsection{ML applications in HEP}
+\section{Neural networks as matrix element surrogate models}
 \end{document}

### 2022-10-18 ml.tex
wrote: NN

@@
 \documentclass[main.tex]{subfiles}
+
 \begin{document}
@@
     of these deep neural networks greatly outperformed the
-    previous state-of-the-art \cite{NIPS2012_c399862d,https://doi.org/10.48550/arxiv.1409.0575} and
+    previous state-of-the-art \cite{NIPS2012_c399862d,russakovsky2015imagenet} and
     lead to the widespread use of neural networks for a
@@
     frameworks such as scikit-learn \cite{scikit-learn},
-    TensorFlow \cite{https://doi.org/10.48550/arxiv.1603.04467},
-    PyTorch \cite{https://doi.org/10.48550/arxiv.1912.01703},
+    TensorFlow \cite{tensorflow2015-whitepaper},
+    PyTorch \cite{paszke2019pytorch},
     and XGBoost \cite{Chen:2016:XST:2939672.2939785}
@@
 \section{Neural networks}
-\subsection{Perceptron}
-\subsection{Densely connected neural networks}
-\subsection{Optimisation of weights}
-\section{Neural networks as matrix element surrogate models}
+    Neural networks are a machine learning algorithm
+    inspired vaguely by the structure of the human brain.
+    That is to say they are an interconnected network
+    of neurons used for analysing data. Here we will
+    introduce the densely-connected neural network
+    which will be the main workhorse algortihm used
+    in this thesis, and frame the discussion around a
+    regression task. Namely, how to use sampled training data
+    from a target function $g(\mathbf{x})$ to create a surrogate
+    model with the neural network algorithm.
+
+\subsection{Model of a neuron}
+    In this context a neuron is modelled by
+    \begin{equation}\label{eqn:neuron}
+        y = \phi(z) = \phi(\mathbf{w}^{T}\mathbf{x} + b)
+    \end{equation}
+    where the model inputs $\mathbf{x}$ are multiplied
+    by the model weights $\mathbf{w}$ before being summed
+    with a bias term $b$. Collectively t he model weights
+    and biases are the model parameters $\theta$.
+    This combination of terms is then modified
+    by an activation function $\phi$ which is carefully
+    chosen to perform a non-linear transformation. A neuron
+    is illustrated in Figure~\ref{fig:neuron}.
+
+    \begin{figure}
+        \ctikzfig{img/ml/neuron}
+        \caption{A schematic diagram of a neuron. The model inputs
+        $\mathbf{x}$ and weights $\mathbf{w}$ are vectors, where as
+        the bias $b$ is a scalar. The weighted inputs are summed with
+        the bias before passing through an activation function
+        to give the scalar output $y$.}
+        \label{fig:neuron}
+    \end{figure}
+
+    Some of the more well-known activation functions are
+    the sigmoid function, hyperbolic tangent, rectified linear unit (ReLU).
+    Less familiar activation functions are the class of
+    sigmoid linear units (SiLU) of which the swish function
+    is a specific case. The functional forms of these
+    functions are summarised in Table~\ref{table:activation_functions}.
+    There is no single activation function that is the best
+    for any given task, it is common practice to choose activation
+    functions on a case-by-case basis.
+
+    \begin{table}
+        \centering
+        \begin{tabular}{lll}
+            \toprule
+            Activation functions & Functional form & Gradient \\
+            \midrule
+            Linear & $f(x) = x$ & $f'(x) = 1$ \\
+            \midrule
+            Sigmoid & $f(x) = \dfrac{1}{1+e^{-x}}$ & $f'(x) = f(x)(1-f(x))$ \\
+            \midrule
+            Tanh & $f(x) = \tanh(x)$ & $f'(x) = 1 - f(x)^{2}$ \\
+            \midrule
+            ReLU & $f(x) = \begin{cases} 0 \, , \quad \mathrm{for} \quad x \leq 0 \\ x \, , \quad \mathrm{for} \quad x > 0 \end{cases}$ & $f'(x) = \begin{cases} 0 \, , \quad \mathrm{for} \quad x \leq 0 \\ 1 \, , \quad \mathrm{for} \quad x > 0 \end{cases}$ \tablefootnote{The gradient of ReLU is not defined at $x=0$ but for a numerical implementation, defining it to be 0 at this point is sufficient.} \\
+            \midrule
+            swish & $f(x) = \dfrac{x}{1+e^{-x}}$ & $f'(x) = \dfrac{1}{x}f(x) \left[(1+x) - f(x)\right]$ \\
+            \bottomrule
+        \end{tabular}
+        \caption{Summary of activation functions and their gradients.}
+        \label{table:activation_functions}
+    \end{table}
+
+\subsection{Densely-connected neural networks}
+    In order to build a neural network, the neurons
+    have to be connected in some fashion. Perhaps
+    the most straightforward method is to create
+    layers of neurons, and then connect every neuron
+    is a layer with every neuron in neighbouring layers.
+    This is illustrated in Figure~\ref{fig:nn_example}.
+    Such a configuration is called a fully-connected,
+    or densely-connected neural network. The constituents
+    of this network structure are the input layer where
+    model inputs enter, hidden layers where the bulk of
+    model parameters live, and the
+    output layer which outputs the model prediction(s).
+    The training data only provides concrete desired
+    outputs for the overall model, and not for the layers
+    before the output layer. This is why mediating layers
+    between inputs and outputs are so-called hidden.
+    The neural network is free to change the parameters
+    in these layers to best approximate the target function
+    $g$.
+    Output layers are functionally the same as hidden
+    layers, except that their outputs are taken as the
+    model prediction and so can be compared with the truth
+    value.
+
+    \begin{figure}
+        \includegraphics[width=\linewidth]{ml/nn_example.pdf}
+        \caption{An example of a densely-connected
+        neural network with two hidden layers.
+        Each node on this image represents
+        a neuron and the connections between then
+        represents the inputs/outputs to neighbouring neurons.}
+        \label{fig:nn_example}
+    \end{figure}
+
+    The depth of a neural network
+    is generally denoted by the number of hidden layers,
+    or alternatively, the number of activation functions
+    between the input and output layers. On the other hand,
+    the width of a neural network is generally referring
+    to the number of nodes in the hidden layers.
+
+    The output of the densely-connected neural network can
+    be formulated by repeatedly applying (\ref{eqn:neuron})
+    to give
+    \begin{equation}\label{eqn:nn_output}
+        f(\mathbf{x}; \theta) = \boldsymbol{\phi}^{(n)}(\ldots \boldsymbol{\phi}^{(2)}(\; \boldsymbol{\phi}^{(1)}(\mathbf{x}; \theta^{(1)})\; ; \theta^{(2)}) \; ; \theta^{(n)}) \, ,
+    \end{equation}
+    where $\boldsymbol{\phi}^{(n)}$ and $\theta^{(n)}$
+    denotes the outputs and parameters of layer $n$,
+    where the output layer is included in this notation.
+    By chaining together activation functions, and especially
+    non-linear activation functions, neural networks
+    become good function approximators
+    \begin{equation}\label{eqn:nn_approx}
+        f(\mathbf{x}; \theta) \approx g(\mathbf{x}) \, .
+    \end{equation}
+    The universal approximation theorem states that neural networks are
+    able to represent any continuous function $g(\mathbf{x})$
+    with one hidden layer and a finite number of neurons \cite{HORNIK1991251}.
+    However it is very difficult to achieve this due to
+    practical constraints such as limited data and network size.
+    It is much more common to link
+    together a larger number of hidden layers, leading
+    to deep neural networks since better learning algorithms
+    have been found for this case \cite{lu2017expressive}.
+
+\subsection{Loss functions and optimisation of parameters}
+    To summarise so far, neural networks are a vast network
+    of connected nodes with a large number of parameters that
+    can be tuned for the problem at hand.
+    The task of optimising the parameters is the main challenge
+    in training a neural network and is a key area of research.
+
+    In order to quantify the performance of a network for
+    a given task it is useful to define a loss function.
+    For regression tasks, such as approximating a function,
+    a commonly used loss function is the mean squared error
+    \begin{equation}\label{eqn:MSE}
+        L_{\mathrm{MSE}} = \dfrac{1}{N} \sum_{i=0}^{N} (g(\mathbf{x}_{i}) - y(\mathbf{x}_{i}; \theta))^{2} \, ,
+    \end{equation}
+    where $g(\mathbf{x}_{i})$ and $f(\mathbf{x}_{i}; \theta)$
+    are samples from the target distribution, and model predictions
+    for inputs $\mathbf{x}_{i}$ belonging to a data set with $N$
+    samples. The loss function
+    encodes the discrepency between the truth value
+    and the model prediction, therefore, the task of finding
+    an acceptable set of parameters can be reframed as an optimisation
+    of the loss function. It should be noted that minimisation of
+    the loss function is simply a proxy for maximising neural network
+    predictive power. Due to the training and testing datasets being finite,
+    the generalisation of the neural network is not guaranteed for the
+    true underlying distribution from which the datasets are sampled
+    from. This problem is generally referred to as overfitting.
+    We will refer back to this problem in the context of
+    fitting matrix elements in Section~\ref{sec:NN_ME}.
+
+    Since neural networks are generally aimed at
+    learning non-linear functions, the loss surface
+    corresponding to (\ref{eqn:MSE}) is
+    a function of many variables, and is highly
+    non-convex with many local minima.
+    The methods of choice for traversing these loss surfaces
+    are all iterative gradient-based methods with modifications
+    to improve convergence. Broadly speaking, the algorithms
+    iteratively update the neural network parameters based on the
+    gradient of the loss with respect to these parameters
+    with some step size (or learning rate) $\eta$. That is
+    \begin{equation}\label{eqn:gradient_descent}
+        \theta = \theta - \eta \cdot \nabla_{\theta} L(\mathbf{x}_{\mathrm{batch}}; \theta) \, ,
+    \end{equation}
+    where $\nabla_{\theta}L(\mathbf{x}_{\mathrm{batch}})$ is the
+    gradient of the loss with respect to the parameters of the
+    model averaged. The initial state of the parameters
+    are usually distributed according to normal
+    or uniform distributions \cite{pmlr-v9-glorot10a,he_initialiser}.
+    This variation of gradient descent is referred to
+    as mini-batch gradient descent because the gradient
+    $\nabla_{\theta}L$ is averaged over a mini-batch of
+    samples $\mathbf{x}_{\mathrm{batch}}$. Mini-batch gradient
+    descent strikes a balance between stochastic gradient
+    descent (estimating gradient with one sample at a time),
+    and batch gradient descent (calculating gradient with
+    respect to entire training set) by having an efficient
+    estimation of the gradient that is accurate enough
+    for practical applications. In modern machine learning
+    libraries, the updating of weights in (\ref{eqn:gradient_descent})
+    is carried out via matrix multiplications which are
+    highly efficient on graphics-processing units (GPUs),
+    and the gradients are computed numerically with the aid
+    of automatic differentiation tools. For these reasons, deep
+    neural networks have become much more widespread.
+
+    The basic gradient descent algorithm described above
+    is the basis upon which many of the most widely
+    adopted optimisers \cite{kiefer1952stochastic,hintonrmsprop,Kingma:2014vow}
+    are based on. For an overview of these optimisers, see for instance
+    Reference \cite{Ruder:2016lil}.
+
+    \subsection{Optimisation of hyperparameters}
+    Hyperparameters are parameters that are
+    not explicitly parameters of the model, instead they control
+    the speed and quality of the training process.
+    The learning rate in {\ref{eqn:gradient_descent}} is
+    an important hyperparameter whose initial value has
+    to be chosen carefully to observe a good rate of convergence.
+    Some other examples would be
+    the number of hidden layers, the choice of activation
+    function, or the mini-batch size.
+
+    The choice of hyperparameters is an important
+    factor during the training process as a suboptimal set
+    of hyperparameters can perform significantly worse
+    than a more carefully chosen set. The process of tuning
+    hyperparameters is often computationally expensive for two
+    reason. Firstly the parameter space to choose from is vast,
+    and the curse of dimensionality kicks in rapidly.
+    Secondly there is no exact a priori way to evaluate the
+    final performance of the neural network given a set of
+    hyperparameters meaning training has to take place for
+    some steps before the final model performance can be estimated.
+
+    Some of the simplest methods of hyperparameter optimisation
+    are grid search and random search. Grid search is simply
+    an exhaustive scan of parameters systematically sampled
+    in a grid-like fashion, where as random search replaces
+    this systematic sampling with random sampling. Whilst
+    these methods are quick to implement, the number of trials
+    to land on a good set of hyperparameters could potentially
+    be very large. Methods which aim to reduce the number
+    of trials have been proposed in the literature, see for example
+    Reference \cite{Yu2020HyperParameterOA} for a review.
+
+    Bayesian optimisation methods have been explored
+    to carry out hyperparameter optimisation \cite{bergstra2013making,bergstra2015hyperopt}.
+    The method  begins with the construction of a probabilistic surrogate
+    model for the performance of the neural network. This
+    surrogate model collects evidence through successive trainings
+    of the neural network. By selecting promising hyperparameter
+    sets based on the updated model, the exploration of
+    hyperparameter space is much more efficient.
+
+    Another group of methods are based on the principle
+    that the first few epochs of training are indicative
+    of the final model performance. This in combination
+    with a fixed time budget gives rise to multi-armed
+    bandit strategies \cite{jamieson2016non,li2017hyperband}
+    that allocate more budget to better
+    performing models whilst simultaneously dropping poorly
+    performing models. By iterating this procedure only the 
+    best performing models remain. Recently, there have
+    been proposals of combining Bayesian optimsation and
+    the bandit-based approaches \cite{falkner2018bohb}.
+
+    In practice the area of hyperparameter optimisation
+    is still extremely computationally expensive even with
+    the more advanced methods. Given a sensible set of hyperparameters,
+    one has to balance the additional time spent tuning with the
+    relatively incremental improvements in performance. For
+    novel machine learning experiments in high energy physics,
+    the hyperparameter tuning is not of paramount importance,
+    and generally more focus is spent on the model building itself.
+
+\section{Neural networks as matrix element surrogate models}\label{sec:NN_ME}
+    Motivated by the ability of neural networks to
+    model non-linear functions, as well as the rise
+    in GPU usage in high energy physics, it seems
+    natural to leverage neural networks to build surrogate
+    models for matrix elements. There are multiple advantages
+    to this: the time taken to evaluate matrix elements remains
+    a bottleneck in event generation, and so building a
+    fast surrogate model would accelerate this process;
+    once a fast and accurate surrogate model has been
+    built it can supplement the traditional matrix
+    element generator to increase statistical power of the sample
+    by predicting on many more phase-space points. This point
+    requires the accumulated error in the surrogate model
+    predictions to be lower than the statistical error in the
+    integration of matrix elements.
+
+    The task of building a surrogate
+    model for matrix elements is in principle straightforward.
+    The dataset usually consists of phase-space kinematics
+    in the form of four-momenta and the targets are the
+    corresponding matrix elements for the process of interest,
+    at the desired order in perturbation theory. As already
+    discussed in Sections~\ref{sec:MC_integration} and \ref{sec:me_generators},
+    phase-space samplers and matrix element generators are
+    now widely available so the acquisition of data is only
+    limited by time. Even then, this is a bottleneck only
+    for the most expensive of processes as generation
+    of data is trivial to parallelise. In the ideal case, the
+    surrogate model would be faster than the matrix
+    element generators, and sufficiently accurate. Neural
+    network predictions are fast by virtue of them being
+    simple matrix multiplications, and because they are predicted
+    in batches vectorisation is automatic. Therefore the difficulty
+    in building a good surrogate model is in making it accurate.
+
+    The complication in accurately emulating matrix elements
+    on a per-point basis arises due to the infrared divergent behaviour
+    as discussed in Section~\ref{sec:IR_divergences}
+    (UV divergences have been removed via renormalisation).
+    Outside of these soft and collinear limits, the matrix element
+    is a well-behaved function that varies smoothly with phase-space.
+    However, once these infrared regions are approached, small changes in the
+    kinematics can lead to very large changes in the matrix
+    element. This rapid response in the target
+    function makes it difficult to accurately model the
+    non-divergent regions simultaneously with the divergent
+    regions due to the disparity in scales. This problem
+    becomes more apparent for higher multiplicity final-states
+    as there are more regions in phase-space for which
+    the matrix element diverges.
+
+    Related to the problem of modelling matrix elements, in recent
+    years there has been work in applying machine learning
+    methods to learn: event weights \cite{Danziger:2021eeg},
+    cross-sections \cite{Otten:2018kum,Buckley:2020bxg},
+    analytic expressions of squared amplitudes \cite{Alnuqaydan:2022ncd},
+    contour deformations for multi-loop integrals \cite{Winterhalder:2021ngy},
+    and simplifying polylogarithms \cite{Dersy:2022bym}.
+
+    Emulation of matrix elements directly have been
+    carried out in recent years in various projects.
+    Bishara and Montull showed \cite{Bishara:2019iwh} that it was possible to
+    build a boosted decision tree model to reproduce
+    the matrix elements of the loop induced process $gg \rightarrow ZZ$ with good accuracy --
+    below 0.1\% errors for fully differential distributions.
+    However, for this $2 \rightarrow 2$ process it was already demonstrated
+    that training a single model on the entire phase-space sampled
+    was suboptimal compared to training a number of regressors
+    on subdomains of the phase-space.
+
+    This trend was observed in a future work by Aylett-Bullock
+    and Badger \cite{Badger:2020uow} where the authors emulated tree-level
+    matrix elements and NLO QCD K-factors for $e^{+}e^{-} \rightarrow q \bar{q}$ +
+    up to 3 gluons. With a higher final-state multiplicity, controlling
+    the infrared divergences became a crucial challenge.
+    Their approach was to first partition phase-space
+    into divergent and non-divergent regions, then subsequently
+    partition the divergent regions further according to partition functions
+    based on FKS subtraction.
+    In each partition the authors trained a separate neural network,
+    meaning that their model was an ensemble of neural networks.
+    The performance of this method was shown to be an improvement
+    over a single neural network trained on the unpartitioned phase-space,
+    with good agreement in histogrammed distributions. However,
+    the per-point agreement was lacking.
+
+    This FKS partitioning method was applied in Reference \cite{Aylett-Bullock:2021hmo}
+    for diphoton plus jet production, where the authors presented a
+    use case of the emulator with a novel interface to the event
+    generator Sherpa. This work was revisited in \cite{Butter:2022rso},
+    where particular attention was paid to quantifying the uncertainty associated
+    with the neural network prediction. This uncertainty was
+    modelled with a Bayesian neural network which enabled
+    the boosted training on regions of phase-space that were
+    lacking accuracy. Additionally, the authors showed
+    that accuracy was improved with a more careful
+    preprocessing of the training data compared to
+    the previous treatment.
+    However, both of these works showed that increasing
+    multiplicity from $2 \rightarrow 3$ to $2 \rightarrow 4$
+    represented a large decrease in accuracy.
+
+    To summarise, the current state of matrix element
+    emulation has to trade-off higher multiplicity processes
+    with accuracy. The research in this thesis attempts to
+    tackle both of these problems by exploiting the factorisation
+    property of matrix elements (Sections \ref{sec:me_factorisation},
+    \ref{sec:CO_factorisation}, and \ref{sec:OL_factorisation}),
+    to build into the emulator the universal singular functions
+    that describe the matrix elements in the soft and collinear limits
+    (Sections \ref{sec:CS_dipoles} and \ref{sec:antenna_functions}).
+    Since this method relies on the factorisation of matrix
+    elements, we collectively refer to these family of models
+    as factorisation-aware models.
+    In Chapter~\ref{chapter:fame1}, we show that by using
+    Catani-Seymour dipoles as a set of basis functions, tree-level
+    matrix elements for $e^{+}e^{-} \rightarrow q \bar{q}$ + up to 3 gluons
+    can be emulated accurately to well below a 1\% accuracy,
+    even for the highest multiplicity. Extending this study
+    to the one-loop level is done in Chapter~\ref{chapter:fame2}
+    where antenna functions replace Catani-Seymour dipoles as
+    the basis functions.
+
+    The question of uncertainty of the neural network
+    prediction is important in quantifying the theoretical
+    uncertainty. In Reference \cite{Danziger:2021eeg},
+    it was shown that with a surrogate model of the event
+    weight it was possible to use a double unweighting
+    algorithm to increase the rate of event unweighting.
+    In Chapter~\ref{chapter:fame3} we examine replacing
+    the original emulator in said work with a factorisation-aware
+    model to study the potential gain factors with a more
+    accuracte model. A consequence of this two-stage unweighting
+    algorithm is that the exact event weight has to be evaluated
+    at some point, meaning the uncertainty of the surrogate
+    model is practically irrelevant.
 \end{document}

### 2022-11-02 qcd.tex
update: figures

@@
     \begin{figure}
-        \includegraphics[scale=1.0]{qcd/bubble_diagram.pdf}
+        \includegraphics[scale=1.2]{qcd/bubble_diagram.pdf}
         \caption{Bubble diagram}
@@
     \begin{figure}
-        \includegraphics[scale=1.0]{qcd/vertex_correction.pdf}
+        \includegraphics[scale=1.2]{qcd/vertex_correction.pdf}
         \caption{Vertex correction}
@@
     \begin{figure}
-        \includegraphics[width=\linewidth]{qcd/CS_dipoles.pdf}
+        \centering
+        \begin{subfigure}{0.49\linewidth}
+            \includegraphics[width=\linewidth]{qcd/Dijk.pdf}
+            \caption{FF dipole}
+            \vspace{20pt}
+        \end{subfigure}
+        \begin{subfigure}{0.49\linewidth}
+            \includegraphics[width=\linewidth]{qcd/Daij.pdf}
+            \caption{FI dipole}
+            \vspace{20pt}
+        \end{subfigure}
+        \begin{subfigure}{0.49\linewidth}
+            \includegraphics[width=\linewidth]{qcd/Daik.pdf}
+            \caption{IF dipole}
+        \end{subfigure}
+        \begin{subfigure}{0.49\linewidth}
+            \includegraphics[width=\linewidth]{qcd/Daib.pdf}
+            \caption{II dipole}
+        \end{subfigure}
+        \vspace{30pt}
         \caption{Schematic diagrams of the four classes of Catani-Seymour
-        dipoles, $\mathcal{D}$. Each dipole consists of a composite particle
+            dipoles, $\mathcal{D}$. The dipoles are named according to
+            whether the emitter and spectator are in the initial or final-state.
+            lEach dipole consists of a composite particle (denoted by tilde)
             that decays into two partons, and a spectator that recoils
-        to conserve momentum. $\boldsymbol{V}$ represents the splitting
+            to conserve momentum. The grey blob represents the hard scattering process,
+            with incoming and outgoing lines representing initial and final-state
+            partons, respectively. represents the splitting
             function within the dipole function which contains the divergent
-        behaviour. The solid black blob represents the hard scattering process,
-        with incoming and outgoing lines representing initial and final-state
-        partons, respectively.}
+            behaviour.}
         \label{fig:CS_dipoles}
@@
     antenna functions are categorised by their class and radiative
-    processes. A pictorial representation is also given in Figure~\ref{fig:antenna}.
+    processes.
 
@@
 
-    \begin{figure}
-        \includegraphics[width=\linewidth]{qcd/antenna.pdf}
-        \caption{Three-parton antenna functions with the grey
-        blob representing all diagrams that give rise to the given
-        external states. The hard radiators are depicted in red,
-        where as the unresolved parton is coloured in blue.}
-        \label{fig:antenna}
-    \end{figure}
-
     Another requirement in constructing a subtraction term
@@
 
+    \begin{figure}
+        \centering
+        \begin{subfigure}{0.4\linewidth}
+            \includegraphics[width=\linewidth]{qcd/A30.pdf}
+            \caption{Quark-antiquark antenna, $A_{3}$}
+            \vspace{20pt}
+        \end{subfigure}
+        \begin{subfigure}{0.4\linewidth}
+            \includegraphics[width=\linewidth]{qcd/D30.pdf}
+            \caption{Quark-gluon antenna, $D_{3}$}
+            \vspace{20pt}
+        \end{subfigure}
+        \begin{subfigure}{0.4\linewidth}
+            \includegraphics[width=\linewidth]{qcd/F30.pdf}
+            \caption{Gluon-gluon antenna, $F_{3}$}
+        \end{subfigure}
+        \vspace{10pt}
+        \caption{Three-parton antenna functions where the loop order
+            is implicitly illustrated in the grey circle,
+            and the grey ellipse represents all diagrams that give
+            rise to the given external states. The hard radiators are
+            depicted in red, where as the unresolved parton is coloured
+            in blue. External fermion lines have been drawn without
+            an arrow to depict the equivalence of quark and antiquark
+            antenna functions.}
+        \label{fig:antenna}
+    \end{figure}
+
     Below we list the antenna functions relevant
@@
     For this we require the $A$, $D$, and $F$ antenna functions,
-    see Table~\ref{table:antenna_functions}.
+    which are illustrated in Figure~\ref{fig:antenna}.
     The $i,j,k$ indices will be fixed such that $q = 1$, $\bar{q} = 2$,

### 2022-11-02 fame_appendex1.tex
add paper1

\documentclass[main.tex]{subfiles}
\begin{document}
\chapter{Appendices for Chapter~\ref{chapter:fame1}}
\label{appendix:fame1}

\section{Azimuthal angle $\phi_{ij}$ calculation}\label{appendix:fame1_phi}
To calculate the azimuthal angle $\phi_{ij}$ for a pair of particle momenta $p_i$ 
and $p_j$ we first consider the plane perpendicular to the momentum 
\begin{eqnarray}\label{eq:pij}
    \vec{p}_{ij} = \vec{p}_{i} + \vec{p}_{j} \, .
\end{eqnarray}
We project the unit vector in the $z$ direction and the momentum of particle $i$ onto 
this plane\footnote{using particle $j$ instead results in a shift of $\phi_{ij}$ by 
$\pi$ which makes no difference for $\sin2\phi$ or $\cos2\phi$.}: 
\begin{eqnarray}\label{eq:north_east}
    \vec{r}_{z} &=& \vec{e}_{z} - \left(\dfrac{\vec{p}_{ij} \cdot \vec{e}_{z}}{\vec{p}_{ij}^{2}}\right) \vec{p}_{ij} \, , \\
    \vec{r}_{i} &=& \vec{p}_{i} - \left(\dfrac{\vec{p}_{ij} \cdot \vec{p}_{i}}{\vec{p}_{ij}^{2}}\right) \vec{p}_{ij} \,.
\end{eqnarray}
The angle $\phi_{ij}$ is the angle between these two projected vectors.
\begin{eqnarray}\label{eq:sin_cos}
    \sin{\phi_{ij}} = 
    \hat r_{ij}\cdot\left( \hat{r}_{i} \times \hat{r}_{z} \right)\, , \quad \cos{\phi_{ij}} = \hat{r}_{i} \cdot \hat{r}_{z} \, ,
\end{eqnarray}
where we have normalised all vectors to be unit vectors:
\begin{equation}
\hat r_z = \frac{\vec{r}_z}{|r_z|}\;, \qquad \hat r_i = \frac{\vec{r}_i}{|r_i|}\;, \qquad \hat r_{ij} = \frac{\vec{p}_{ij}}{|p_{ij}|}\;.
\end{equation}

\section{Jensen's Inequality}\label{appendix:fame1_jensen}
Jensen's inequality states that for concave functions
\begin{eqnarray}
    f(\mathbb{E}[Y]) \geq \mathbb{E}[f(Y)] \, ,
\end{eqnarray}
where $\mathbb{E}$ is the expectation value, the function $f$ in our case is $\mathrm{arcsinh}$ that behaves similarly to the natural logarithm due to the scale of the problem, and $Y$ is a random variable representing our target distribution.
This inequality can be rewritten as
\begin{eqnarray}
    f(\mathbb{E}[Y]) - \mathbb{E}[f(Y)] \geq 0 \, ,
\end{eqnarray}
which is known as Jensen's gap.
With a concave function, the mean of the transformed target distribution will always be underestimating the actual mean, i.e.
\begin{eqnarray}
    \mathbb{E}[Y] \geq \sinh{(\mathbb{E}[\mathrm{arcsinh}{(y)}])} \, ,
\end{eqnarray}
where the LHS is the actual expectation value (cross-section) of the random variable $Y$ and the RHS is the shifted expectation value that the neural network learns.
In our scenario the neural network reduces the variance of the residual distribution, meaning the gap in reality is small but there will always be an offset in the mean value learnt.

\section{Phase-space trajectories}\label{appendix:fame1_trajectory_generation}
We can see the RAMBO algorithm as a map from the unit hypercube in some high dimension into the $n$-particle phase-space. A flat distribution in the hypercube maps to a flat distribution in the multi-particle phase-space. 
To generate our phase-space trajectories we pick two points at random in the unit hypercube and map the line between them using the RAMBO mapping. As a result, each point on the resulting trajectory has equal probability density in phase-space. 
Any other phase-space generator that smoothly maps the unit hypercube to a multi-particle phase-space could replace RAMBO in this procedure and could lead to trajectories with very different characteristics. One could for example imagine a 
sophisticated algorithm that only maps points close to the boundary of the hypercube to soft or collinear configurations. Using such an algorithm would have trajectories that avoid configurations with many particles more collinear or soft than 
the end-points of the trajectories. We find that with RAMBO the trajectories tend not to avoid difficult phase-space configurations and are therefore a good test of the extrapolation properties of our method.

Figure~\ref{fig:detectorpath} shows the trajectory we chose in Section~\ref{sec:trajectories}.

\begin{figure}
    \centering
    \includegraphics[width=0.45\textwidth]{fame1/yphi.png}
    \includegraphics[width=0.45\textwidth]{fame1/E.png}
    \caption{Left: rapidity and azimuthal angle trajectories for the five final state particles in the trajectories used in Section~\ref{sec:trajectories}.
    Right: Evolution of the same particle energies as a function of the position along the segment between the two random points in the unit hypercube.
    }
    \label{fig:detectorpath}
\end{figure}

\end{document}

### 2022-11-02 main.tex
wrote about running coupling
wrote: matrix element factorisation
wrote: monte carlo
add paper1

@@
 	submitted elsewhere for any degree or qualification.
+
+	Research presented in this thesis are based on joint work
+	published in the articles:
+	\begin{itemize}
+		\item Chapter \ref{chapter:fame1} is based on \cite{Maitre:2021uaa}: Daniel Ma\^{i}tre and Henry Truong, \textit{A factorisation-aware Matrix element emulator}, Journal of High Energy Physics, 2021 (11).
+066.
+		\item Chapter \ref{chapter:fame2} is based on ...
+		\item Chapter \ref{chapter:fame3} is based on ...
+	\end{itemize}
+
+	Other research projects carried out during my studies but not included
+	in this thesis:
+	\begin{itemize}
+		\item \cite{doi:10.1098/rsos.210506}: J.
+Aylett-Bullock, C.
+Cuesta-Lazaro, A.
+Quera-Bofarull, M.
+Icaza-Lizaola, A.
+Sedgewick, H.
+Truong, A.
+Curran, E.
+Elliott, T.
+Caulfield, K.
+Fong, I.
+Vernon, J.
+Williams, R.
+Bower and F.
+Krauss, \textit{June: open-source individual-based epidemiology simulation}, (2021), R.
+Soc.
+open sci.8210506210506.
+		\item \cite{doi:10.1098/rsta.2022.0039}: I.
+Vernon, J.
+Owen, J.
+Aylett-Bullock, C.
+Cuesta-Lazaro, J.
+Frawley, A.
+Quera-Bofarull, A.
+Sedgewick, D.
+Shi, H.
+Truong, M.
+Turner, J.
+Walker, T.
+Caulfield, K.
+Fong and F.
+Krauss, \textit{Bayesian emulation and history matching of JUNE}, (2022), Phil.
+Trans.
+R.
+Soc.
+A.3802022003920220039.
+	\end{itemize}
+
 %
@@
 \subfile{qcd}
+\subfile{MC}
 \subfile{ml}
@@
 \subfile{paper3}
+\subfile{conclusion}
 
 \appendix
-%\subfile{appendix1}
+\subfile{appendix1}
+\subfile{fame_appendex1}
 %\subfile{appendix2}

### 2022-11-02 paper1.tex
initial draft of chapters
add paper1

--- 2022-08-30 ml.tex
@@
 \begin{document}
-\chapter{Machine learning}
-\label{chapter:ml}
-\section{Introduction to machine learning}
-\subsection{Supervised and unsupervised learning}
-\section{Neural networks}
-\subsection{Perceptron}
-\subsection{Densely connected neural networks}
-\subsection{Optimisation of weights}
-\subsection{ML applications in HEP}
+\chapter{Emulation of tree-level $e^{+}e^{-}$ matrix elements}
+\label{chapter:fame1}
+
+\section{Introduction}
+\label{sec:intro}
+Neural networks (NN) first appeared in the field of particle physics as a tool for event discrimination in the analysis of collider data. 
+Since then neural networks and other machine learning techniques have proved useful in many other areas of the field. 
+On the theory prediction side they have been used 
+to improve the efficiency of Monte Carlo sampling \cite{Bendavid:2017zhk,Klimek:2018mza,Bothmann:2020ywa,Verheyen:2020bjw,Chen_2021}, 
+to accelerate the simulation of radiation within a jet~\cite{Carrazza:2019cnt,Bothmann:2018trh,Dohi:2020eda}, 
+to streamline the processes of generation and unweighting of simulated event samples 
+\cite{Gao:2020zvv,Otten:2018kum,Hashemi:2019fkn,DiSipio:2019imz,Butter:2019cae,Bishara:2019iwh,Backes:2020vka,Butter:2020qhk,Alanazi:2020klf,Nachman:2020fff}.
+Closer to the experimental measurements 
+they have also been used to emulate detector simulation \cite{Paganini:2017dwg,SHiP:2019gcl,Derkach:2019qfk,Alanazi:2020jod},
+ they can be used to perform unfolding \cite{Andreassen:2019cjw} or correcting for detector effects \cite{Bellagente:2019uyp},
+and perform pileup subtraction \cite{Andreassen:2019cjw}.
+
+The capacity of neural networks to approximate intricate functions have already been 
+used to provide fast calculations of production cross-sections~\cite{otten2019deepxs,Buckley:2020bxg}.
+In this article we investigate whether NNs can approximate production cross-sections more differentially by replacing computationally expensive 
+matrix element calculations. 
+The challenge with this endeavour is that matrix elements are plagued with numerous divergences that arise from infrared divergences.
+In previous works \cite{Badger:2020uow,Aylett-Bullock:2021hmo} a combination of individual neural networks were used 
+to approximate matrix elements. In order to deal with the complex structure of the matrix elements 
+the authors of these studies divided the phase-space into sectors according to the infrared singularities and trained networks on these sectors,
+thereby limiting the complexity of the fit by isolating a single divergence per sector. All sectors were then combined to make the final prediction.
+While the authors of the initial study for electron-positron annihilation showed good agreement between the total cross-section and histogrammed distributions both at LO and NLO,
+we note that the accuracy of the interpolation at the level of individual points was a lot worse than when averaged in the 
+histograms. In addition, the performance of the extrapolation outside of its training phase-space 
+(i.e. more singular configurations than those considered to fit the model) is problematic.
+
+In this article we present a different approach to the emulation of matrix elements that incorporates the factorisation 
+properties of the matrix elements in the interpolation model and therefore is able to safely 
+extrapolate the matrix element in regions more singular than that covered by the data used for the training of the emulator. We find
+that our interpolator also displays a much improved pointwise accuracy.
+
+The article is organised as follows. 
+Section~\ref{sec:fitting} introduces our factorisation-aware 
+deep neural network model, Section~\ref{sec:results} showcases our results with a comparison with the 
+model from Ref.~\cite{Badger:2020uow} in Section~\ref{sec:n3jet_comparison}, and an analysis of the performance of our model in Section~\ref{sec:mainresult}. 
+To show that the neural network is both interpolating and extrapolating well we show its 
+behaviour on random trajectories in phase-space in Section~\ref{sec:trajectories}. 
+Our conclusions are presented in Section~\ref{sec:conclusion} 
+and some discussion of specific details are collected in the appendices so as not to distract from the main discussion.
+
+We provide code to reproduce the methodology detailed in this article \cite{fame_repo}.
+
+\section{Fitting framework}\label{sec:fitting}
+For this work we consider the $e^+e^-\rightarrow Z / \gamma^{*} \rightarrow q\bar q +n g$ matrix elements for $n$ up to and including 3, 
+which corresponds to events with up to 5 jets. We denote the number of jets in the final state as $n_{j}$.
+
+We formulate the problem of emulating matrix elements as a supervised regression task with a set of phase-space points' kinematic information as input and the 
+values of the matrix element for each of these phase-space points as the targets. Section~\ref{sec:data} describes how these matrix elements were obtained.
+
+A neural network can be seen as a function $f(\vec{x}; \theta) = \vec{y}$, where $f: \mathbb{R}^{d} \rightarrow \mathbb{R}$ maps a 
+$d$-dimensional vector $\vec{x}$ of inputs onto the vector of outputs $\vec{y}$, and where $\theta$ are the parameters of the neural network which we aim to optimise 
+such that the outputs $\vec{y}$ of the neural network match the target as well as possible. 
+The simplest implementation of an emulator would be to take the input as the kinematic information of the phase-space point and the output to be the full matrix element. 
+Our approach modifies both the input of the NN and its target, as described in the following sections.
+
+\subsection{Infrared divergences and dipole factorisation formula}
+It is well known that in soft and collinear limits the matrix element in $n+1$-body phase-space factorises into a singular factor and a reduced matrix element in $n$-body phase-space \cite{Altarelli:1977zs,BASSETTO1983201}.
+This factorisation was used by Catani and Seymour \cite{Catani:1996vz} to construct subtraction terms for the real radiation part of a NLO calculation.
+They introduced a factorisation formula with universal dipoles that smoothly interpolates 
+between the soft and collinear limits to capture the singular structure in these regions of phase-space.
+The dipole factorisation formula can be written schematically as
+\begin{eqnarray}\label{eq:factorisation}
+    |\mathcal{M}_{n+1}|^{2} \rightarrow |\mathcal{M}_{n}|^{2} \otimes \mathbf{V}_{ij,k} \, ,
+\end{eqnarray}
+where $\mathbf{V}_{ij,k}$ is a process independent, singular factor. It depends on the momenta and quantum numbers (colour and spin) of partons $i,j,k$, where $i$ is the emitter parton, 
+$j$ is the emitted parton, and $k$ is the spectator parton.
+For singly unresolved limits, this factorisation isolates all the divergent behaviour in $\mathbf{V}_{ij,k}$ and the factor $|\mathcal{M}_{n}|^{2}$ is free of divergences, which makes 
+it more amenable to emulation through a neural network.
+The dipole factorisation formula forms the basis of our fitting ansatz which we present in detail in Section~\ref{sec:ansatz}.
+
+\subsection{Fitting coefficients of Catani-Seymour dipoles}\label{sec:ansatz}
+Instead of using a neural network to fit the matrix element directly, we use the dipole factorisation formula to build an ansatz of the colour and helicity summed $n+1$-body matrix element,
+\begin{eqnarray}\label{eq:ansatz}
+    \ME{n+1} = \sum_{\{ijk\}} C_{ijk} D_{ij,k} \, ,
+\end{eqnarray}
+where $D_{ij,k} = \langle V_{ij,k} \rangle / s_{ij}$ are the spin-averaged Catani-Seymour dipoles divided by the corresponding Mandelstam invariant and $C_{ijk}$ are the coefficients we train the neural network to fit.
+$C_{ijk}$ can be interpreted as the reduced matrix element in $n$-body phase-space. 
+Since the input for the $C_{ijk}$ function is the full $n+1$ phase-phase information, the neural network will also model the phase-space mappings usually introduced in the factorisation formula.
+A schematic diagram illustrating our ansatz is given in Figure~\ref{fig:nn_architecture}.
+The sum over ${\{ijk\}}$ denotes the sum over relevant permutations of the external outgoing legs.
+More detail on this is given in Section~\ref{sec:architecture}.
+The representation (\ref{eq:ansatz}) is not unique but 
+through appropriate training, the neural network takes advantage of the right ingredients to model 
+the divergent soft and collinear behaviour of the matrix elements.
+
+This form of the ansatz allows the neural network to avoid fitting a rapidly varying function over the phase-space, leaving the Catani-Seymour dipoles to reproduce the correct singular behaviour, meaning a single neural network can interpolate a now relatively smooth function over the phase-space.
+
+\begin{figure}
+    \includegraphics[width=\textwidth]{fame1/nn_architecture.pdf}
+    \caption{Schematic diagram of our neural network architecture.
+    We have a densely-connected neural network with inputs phase-space points, $p$, and recoil factors, $y_{ij,k}$, propagated through hidden layers to the output layer which outputs $C_{ijk}$.
+    These coefficients are combined with their corresponding spin-averaged dipoles as in (\ref{eq:ansatz}) to produce an approximation of the matrix element.
+    Diagram of neural network generated with the aid of \cite{LeNail2019}.}
+    \label{fig:nn_architecture}
+\end{figure}
+
+\subsection{Data generation}\label{sec:data}
+For all multiplicities, phase-space is sampled uniformly using the RAMBO algorithm \cite{Kleiss:1985gy} with a centre-of-mass energy $\sqrt{s_{\mathrm{com}}} = 1000$ GeV.
+Phase-space points are subsequently clustered using {\FastJet} \cite{Cacciari:2011ma,noel_dawe_2021_4446849} with the $e^{+}e^{-}$ $k_{t}$ algorithm \cite{Catani:1991hj}.
+Global phase-space cuts are applied according to the criterion $y_{\mathrm{cut}} \leq y_{ij}$ where $y_{ij}$ are the Mandelstam invariants normalised by $s_{\mathrm{com}}$.
+Jets are clustered exclusively where $d_{\mathrm{cut}}$ was supplied to {\FastJet}. We took $d_{\mathrm{cut}} = \mathrm{max}(2 \times y_{\mathrm{cut}}, 0.01 \times s_{\mathrm{com}})$.
+We explore three different values of the global phase-space cut parameter, $y_{\mathrm{cut}} = [0.01, 0.001, 0.0001]$, to demonstrate the ability of the factorisation-aware neural network to effectively interpolate in more and more singular regions of phase-space.
+
+The generated phase-space points are fed to the {\NJet} package \cite{Badger:2012pg} to calculate colour and helicity summed tree-level matrix elements. 
+All external legs have been considered to be massless. 
+The strong coupling constant has been set to $\alpha_{s} = 0.118$, and the electromagnetic coupling constant has been set to $\alpha_{e} = 1.0 / 132.5070$.
+The mass of the $Z$-boson is taken to be $m_{Z} = 91.188$ GeV.
+These parameters and those not listed, are consistent with those in the Standard Model mode of \aNLO \cite{Alwall_2014}.
+
+The phase-space points generated form the basis of our inputs to the neural network with the matrix elements as our fitting targets.
+As with most machine learning applications, we need to demonstrate that our neural network emulator has managed to generalise outside of the training dataset.
+We do this by firstly testing on an independent testing dataset that is never exposed to the network during training, and secondly by predicting on random trajectories in phase-space.
+Generation of phase-space trajectories is described in Appendix~\ref{appendix:fame1_trajectory_generation}.
+We believe that accurate predictions on random phase-space trajectories demonstrates the ability of the neural network to extrapolate to never before seen data 
+that is of a different nature to both the training and testing datasets.
+
+\subsection{Neural network emulator}\label{sec:network}
+We construct our emulator using a densely-connected neural network built using the {\Keras} API \cite{chollet2015keras} with the {\TensorFlow} back-end \cite{tensorflow2015-whitepaper} with GPU support.
+A simple model architecture such as a densely-connected neural network allows for quicker training and inference compared to more complicated setups.
+\subsubsection{Inputs and outputs}\label{sec:input_outputs}
+\paragraph*{Inputs to neural network}
+As mentioned in Section~\ref{sec:data}, phase-space points form the basis of the inputs to our neural network.
+We input the 4-momenta of each outgoing parton as an $(n_{j} \times 4)$ array with each component of its 4-momenta standardised to zero mean and unit variance across the training dataset.
+Although it would be feasible to omit the energy component of the momenta or leave one outgoing parton out of the training due to our datasets being generated with a fixed centre-of-mass energy, 
+we find that keeping all momenta information to improve the network's performance.
+With the acceleration of training of neural networks on GPUs, the slowdown in keeping all outgoing momenta components is negligible.
+
+Along with the 4-momenta, we also include the recoil factors
+\begin{eqnarray}\label{eq:recoil_factor}
+    y_{ij,k} = \dfrac{p_{i}p_{j}}{p_{i}p_{j} + p_{j}p_{k} + p_{i}p_{k}} \, ,
+\end{eqnarray}
+as input for relevant permutations of $\{ijk\}$.
+We take the natural logarithm of $y_{ij,k}$ before standardising them to a zero mean and unit variance across the training dataset.
+We find that the addition of these recoil factors significantly improves performance of our neural network emulator during training and testing.
+This can be attributed to the fact that the coefficients $C_{ijk}$ rely heavily on the mapped momenta in $n$-body phase-space, where $y_{ij,k}$ is usually required when performing the momentum mapping, see e.g.
+\cite{Catani:1996vz}.
+
+Following our ansatz (\ref{eq:ansatz}), we must provide spin-averaged Catani-Seymour dipoles to the neural network.
+These dipoles are computed for all phase-space points before training begins, but they are not passed directly to the neural network as input features.
+Instead, they are only included in our custom loss function which will be explained in the next subsection.
+
+\paragraph*{Accounting for spin-correlation in $g \rightarrow gg$}
+In addition to the Catani-Seymour dipoles, we include other functions to account for the spin-correlations of $g \rightarrow gg$ and $g \rightarrow q\bar q$ splittings which are present in the factorisation formula but averaged out in the spin-averaged dipoles.
+This effect becomes relevant when there are two or more gluons in the final state.
+We seek to capture this behaviour by introducing a pair of terms of the form 
+\begin{equation}
+S_{ij} \sin\left(2 \phi_{ij}\right) + C_{ij} \cos\left(2\phi_{ij}\right)
+\end{equation}
+in the fitting ansatz for each gluon pair.
+The coefficients $S_{ij}$ and $C_{ij}$ are fitted by the neural network along with the dipole coefficients.
+The angle $\phi_{ij}$ is the azimuthal angle of the decay particles in the plane perpendicular to the parent particle momentum.
+The procedure to obtain this angle is described in Appendix~\ref{appendix:fame1_phi}.
+
+\paragraph*{Outputs of neural network}
+Denoting the raw output of our neural networks as $c_{ijk}$, they are transformed to $C_{ijk}$ according to
+\begin{eqnarray}\label{eq:coef_output}
+    C_{ijk} = S_{\mathrm{coef}} \times \sinh{(c_{ijk})} \,
+\end{eqnarray}
+where $S_{\mathrm{coef}} = S_{\mathrm{pred}} / S_{\mathrm{dipole}}$.
+$S_{\mathrm{pred}}$ is the prediction scale, taken to be the minimum of the matrix elements in the training set and $S_{\mathrm{dipole}}$ is the dipole scale, taken to be the mean of all dipoles in the training set.
+These scaled coefficients are then multiplied with their corresponding Catani-Seymour dipole $D_{ij,k}$, and then summed to produce an estimation of the matrix element.
+The targets are the matrix elements corresponding to the phase-space point inputs in the training dataset. We transform the targets according to
+\begin{eqnarray}\label{eq:arcsinh}
+    y = \mathrm{arcsinh}{\left(\dfrac{\ME{n+1}}{S_{\mathrm{pred}}}\right)} \, ,
+\end{eqnarray}
+to reduce the orders of magnitude the matrix elements span.
+This performs a similar transformation to taking the natural logarithm except that it remains a valid transformation for negative arguments.
+We require this transformation to allow negative values as arguments because the coefficients $C_{ijk}$ will not be restricted to only positive values, meaning that during training there is a possibility for the outputs of the network to go negative.
+By reducing the span of the targets, the neural network is able to more effectively pick out patterns across the entire training dataset rather than a smaller region, helping it to generalise.
+This technique is employed in other studies, see for example \cite{Coccaro_2020,Bury_2021}.
+We would like to stress that we do not expect, and have not observed, negative predictions for the matrix element from the neural network, as this would be unphysical.
+$y$ is finally standardised to a zero mean and unit variance.
+
+\paragraph*{Training, validation, and testing datasets}
+In Ref. \cite{Badger:2020uow}, the authors used 500k training samples to train their models.
+For a fair comparison of our respective methods, we follow their methodology closely by constructing our models to be as close as possible to theirs and generate training and testing data by using code from their project repository \cite{n3jet_repo}\footnote{for our main results we generate all data using our own code}.
+Our model architectures will not be identical due to the difference in our methods, details of our model architecture are given in Section~\ref{sec:architecture}.
+Results of this comparison are presented in Section~\ref{sec:n3jet_comparison}.
+
+While training on 500k samples gives acceptable performance for the total cross-section, per-point accuracy is lacking.
+We find that increasing the size of the training dataset drastically improves the per-point prediction accuracy.
+Neural networks have been shown to scale well with large datasets \cite{hestness2017deep} and given that they only need to be trained one time, it is useful to provide neural networks which have been pre-trained with maximum accuracy in mind.
+In addition to improving the per-point accuracy, we aim to overcome the problem of extrapolating to more singular regions.
+It is well known that neural networks in general do not extrapolate well \cite{xu2021neural,kaplan2020scaling}, but with our factorisation-aware model we show that letting the models learn about the infrared structure of QCD alleviates this problem.
+
+To demonstrate the scaling performance of our model, we present our main results with models trained on more training samples where details of data generation are given in Section~\ref{sec:data}.
+For each multiplicity and global phase-space cut we generate a dataset consisting of 60 million phase-space points and their corresponding recoil factors, dipoles, and matrix elements.
+We then split this dataset into training, validation and testing datasets in a ratio of 4:1:1, meaning we have 40 million training samples, 10 million validation, and 10 million testing samples.
+The validation dataset is used to monitor model performance after each epoch of training, whereas the testing dataset is used as an out-of-sample check of the model's performance after training is complete.
+
+\subsubsection{Architecture}\label{sec:architecture}
+We have a fixed base to our neural network architecture that is used for all processes considered in this work, with variations in the number of nodes in the input and output layers due to the change in number of outgoing partons.
+Our base neural network consists of one input layer, eight hidden layers consisting of (64, 128, 256, 512, 768, 386, 128, 64) nodes, and one output layer.
+
+The number of nodes in the input and output layers scales with the number of jets in the final state.
+The input layer has $(n_{j} \times 4) + n_{\mathrm{rel}}$ nodes, and the output layer has $n_{\mathrm{rel}} + (2 \times n_{\phi})$ nodes, where $n_{\mathrm{rel}}$ is the number of relevant permutations and $n_{\phi}$ denotes the number of $\phi_{ij}$ angles.
+Relevant permutations, ${\{ijk\}}$, are the set of permutations for which their corresponding dipole could have the possibility to have a meaningful contribution to the matrix element for a given process.
+They are a subset of all the possible permutations for a given multiplicity, $P(n_{j}, 3)$.
+Relevant permutations exclude any permutation where a quark or anti-quark are emitted (i.e. $j = q$ or $\bar{q}$), as low energy quarks do not give rise to singularities in our processes of interest.
+Furthermore, we can remove degenerate permutations where swapping $i=g$ and $j=g$ has no effect as they have identical Catani-Seymour dipoles, e.g.
+$D_{34, 1} = D_{43, 1}$, so we only keep $D_{34, 1}$.
+The omission of these redundant permutations speeds up training of the neural networks as we have fewer inputs, fewer dot products to compute in the loss function, as well as speeding up inference due to there being fewer dipoles to compute.
+For reference, we list the number of input and output nodes for each multiplicity we consider in Table \ref{table:nodes}.
+
+\begin{table}
+    \centering
+    \begin{tabular}{|c|c|c|}
+        \hline
+        Number of final state jets & Input nodes & Output nodes \\
+        \hline
+        3 & $(3 \times 4) + 2  = 14$ & $2 + (2 \times 0)  = 2$   \\
+        4 & $(4 \times 4) + 10 = 26$ & $10 + (2 \times 1) = 12$ \\
+        5 & $(5 \times 4) + 27 = 47$ & $27 + (2 \times 3) = 33$ \\
+        \hline
+    \end{tabular}
+    \caption{List of the number of input and output nodes for every process we consider.}
+    \label{table:nodes}
+\end{table}
+
+The neural network weights are initialised according to the `Glorot uniform' distribution as described in \cite{pmlr-v9-glorot10a}.
+We use the $\tanh$ activation function for all nodes in the hidden layers, and have a linear activation function for nodes in the output layer.
+Initial learning rate is set to 0.001 and training mini-batch size is set to 4096.
+During training, we reduce the learning rate by a factor of $0.7$ whenever there is no improvement in validation loss for 20 epochs.
+We use the {\Keras} callback {\ReduceLROnPlateau} to achieve this.
+Model training is terminated after the validation loss does not improve after 40 epochs of training using the {\EarlyStopping} callback in {\Keras}.
+We find that reducing the learning rate during training helps the model to converge to more optimal parameter sets.
+Since there are periods during training where the validation loss stagnates, reducing the learning rate helps to reach minima which otherwise wouldn't be accessible due to a too large learning rate.
+There is a possibility to reduce the learning rate too rapidly causing the model to have suboptimal optimisation, but this is countered by the high patience we set for the {\ReduceLROnPlateau} callback.
+
+These choices of hyperparameters were not results of extensive scanning of parameter space and were chosen heuristically.
+Hyperparameters can be tuned more optimally using more sophisticated methods, but they all rely on training a large number of models which is computationally expensive and time-consuming.
+
+\paragraph*{Custom loss function}
+To assess the model's performance when training we need to compare the network predictions with the targets. Our metric for the model's regression performance is the mean squared error (MSE)
+\begin{eqnarray}\label{eq:mse}
+    L_{\mathrm{MSE}} = \dfrac{1}{N} \sum_{i=1}^{N} \left(y_{i} - p(\vec{x}_{i}; \theta)\right)^{2} \, ,
+\end{eqnarray}
+where $y_{i}$ is the target for the $i$-th sample out of $N$ samples, and $p(\vec{x}_{i}; \theta)$ is the corresponding prediction
+obtained by combining the dipole factors and the azimuthal dependency terms multiplied by their NN-learned coefficients. 
+In order to correct for the scale difference, we need to apply the transformation (\ref{eq:arcsinh}) on the neural network prediction first, with the same $S_{\mathrm{pred}}$ that was used for scaling the targets.
+We choose the mean squared error to measure the network's performance because it is sensitive to outliers in the target distribution.
+This is useful because even though we have taken measures to reduce the span of the targets, the distribution still contains a tail towards larger values which correspond to soft and collinear configurations.
+These points have large contributions to the cross-section when integrating over phase-space, meaning it is important that we accurately predict these points.
+It is also convenient that the mean squared error tends to learn the mean of the target distribution which, in our case, corresponds to the cross-section.
+
+In addition to the MSE, we introduce a regularisation term to penalise non-sparse representations of the matrix element.
+We know that in soft and collinear limits there will be dominant dipoles that have large contributions to the matrix element and there will be other dipoles with minimal contribution.
+We try to suppress the coefficients associated with these minimally contributing dipoles with the penalty term
+\begin{eqnarray}\label{eq:penalty}
+    L_{\mathrm{pen}} = J \, \sum_{i} \dfrac{D_{i}^{-2}}{\sum_{j}D_{j}^{-2}} |C_{i}D_{i}| \,
+\end{eqnarray}
+where the sum over $i$ replaces the sum over $\{ijk\}$ for brevity, and $\sum_{j} D_{j}^{-2}$ is the sum over all dipoles for a phase-space point, acting as a normalisation factor.
+$J$ is a tunable parameter that scales the importance of $L_{\mathrm{pen}}$ versus $L_{\mathrm{MSE}}$.
+We found that models perform the best when $L_{\mathrm{pen}} < L_{\mathrm{MSE}}$, so $J$ is tuned accordingly.
+It is possible for the product $C_{i}D_{i}$ to be large due to $D_{i}$ alone, so to penalise this we regularise the product rather than just the coefficient, since it is the product that contributes to the matrix element.
+
+The form of $L_{\mathrm{pen}}$ is reminiscent of the usual L1 regularisation which promotes sparse models.
+Regularisation is usually included to prevent overfitting by making the model make decisions on the most important features, reducing other features to zero.
+In our case we would like the neural network to learn about the universal factorisation property in QCD by making it choose a minimal amount of dipoles to represent the matrix element in singular regions.
+In addition to preventing overfitting, (\ref{eq:penalty}) helps the neural network to extrapolate to more soft and collinear regions as it has learnt to choose which dipoles are relevant in specific configurations.
+For non-singular configurations, the neural network is free to interpolate as there is not a clear set of dipoles that dominate, meaning $L_{\mathrm{pen}}$ is small.
+
+Combining the regression loss term and the regularisation term, our expression for the total loss is
+\begin{eqnarray}\label{eq:loss}
+    L = L_{\mathrm{MSE}} + L_{\mathrm{pen}} \, ,
+\end{eqnarray}
+which is minimised through mini-batch gradient descent \cite{bottou2018optimization} with the Adam optimiser \cite{Kingma:2014vow} to find optimal parameters $\theta$ for the neural network.
+
+\paragraph*{Ensemble of models}
+Due to the random initialisation of weights in the neural network, and the fact that the optimisation procedure is carried out on mini-batches of the full training dataset, every neural network trained will be similar but non-identical, even with identical model architecture.
+This is partly because the loss surface is unlikely to be a completely smooth surface with a single global minimum, instead it is likely to contain multiple local minima, meaning it would be optimistic to believe that a single neural network is able to find the most optimal set of parameters.
+Given that we don't expect a single network to perform optimally, we train a number of models and aggregate their predictions to create an ensemble of models.
+Ensembling models is a well-known technique within machine learning, for a review see \cite{ZHOU2002239}.
+
+Each model in the ensemble is initialised with different weights according to the `Glorot uniform' distribution, and trained on the same but randomly shuffled dataset, resulting in models that have been exposed to different distributions of the training data.
+After sufficient training, each model will have a distinct set of parameters that have similar predictive power.
+The prediction for the matrix element is then the mean of the outputs of all models in the ensemble\footnote{if not explicitly stated, all references to neural network predictions from henceforth refers to the prediction from the ensemble of networks}.
+Taking the mean will give a more accurate and robust prediction as averaging over the different models will reduce variance due to over/underfitting in the training phase.
+We choose to have 20 models in our ensembles because we begin to see diminishing returns in the accuracy of per-point predictions after this.
+Another advantage of training an ensemble of models is that we have a measure of uncertainty, due to the neural networks, on the model predictions by calculating the standard error of the mean.
+That is we take the standard deviation of predictions across the ensemble and divide by the square root of number of models in the ensemble. This would not be possible with just a single model.
+
+By choosing to build an ensemble, there is a performance impact because we have to spend more resources on training, and inference is slower due to having to predict on all models in the ensemble.
+However, we believe that having a more robust prediction with a measure of uncertainty outweigh these negatives.
+Additionally, each model only needs to be trained once, and slowdown during inference is alleviated with GPUs.
+
+\section{Results}\label{sec:results}
+In this section we present results for our matrix element emulator for $e^{+}e^{-}$ annihilation into up to 5-jets.
+We first compare results obtained with our method to the tree-level results of Ref.~\cite{Badger:2020uow}, then proceed to exhibit results from larger NNs
+that have been exposed to larger training datasets to demonstrate the full scaling performance of our model.
+We then further demonstrate our method's capability to generalise to unseen regions of phase-space by assessing the prediction accuracy on random phase-space trajectories that venture well outside of the phase-space region used for the training.
+
+\subsection{Comparison with previous work}\label{sec:n3jet_comparison}
+We compare our method to methods for tree-level matrix elements emulation from Ref.~\cite{Badger:2020uow}.
+The authors presented two methods: `single' and `ensemble' models. A `single' model indicates that there is 
+one neural network trained across the entire phase-space, and an `ensemble' model indicates that there is a group of neural networks trained together
+with weighting functions that focus individual networks on a specific divergent region of phase-space.
+In order to conduct a fair comparison, we have made attempts to follow their training methodology closely by 
+constructing the neural networks at the centre of our model with a similar structure (e.g. same hidden layer structure, same activation functions, 
+same random initial weight distribution), training methodology (same {\EarlyStopping} criteria, same initial learning rate), and have generated training and testing datasets using code 
+available on the authors' project repository with the relevant cuts.
+
+We compare the distribution of errors for the matrix elements predictions on 3 million phase-space points.
+The distribution of errors is crucial because it informs us of the performance of our emulators at the matrix element level.
+It is well known that a neural network optimised with the mean squared error loss function has tendencies to learn the mean of the target distribution \cite{Nachman_2020},
+meaning the quality of the cross-section prediction can potentially belie the point-to-point accuracy of emulators.
+In Figure \ref{fig:comparison}, we plot the distribution of errors for matrix element predictions where our method is labelled `Dipole NN'.
+We compare against the `single' and `ensemble' methods by training and testing using the corresponding datasets.
+For more detailed information on the differences please refer to \cite{Badger:2020uow}.
+Note that the height of the peak for the dipole histograms are not illustrated in the figure as it would not fit on the current axes but that is not important for this comparison.
+We can see that the prediction-to-truth ratio distribution for our method is much narrower and consistently peaked around the ideal accuracy, indicating our model performs better on a per-point basis for all multiplicities.
+Even with this reduced NN size we can see that incorporating the known
+divergent structure explicitly in the model gives better results, as it
+uses the NN representation to learn a function that is more
+suitably approximated by a NN.
+For example, even though the three jet matrix element has a fairly trivial analytical structure, a standard
+fitting approach using a NN typically struggles to reproduce
+divergences. In our approach the NN only needs to emulate a non-singular
+modulation on top of the main divergent behaviour and is therefore more
+suited to the task.
+
+\begin{figure}
+    \centering
+    \includegraphics[width=\textwidth]{fame1/n3jet_comparison_inset.pdf}
+    \caption{Error distribution compared to Figure 3 in Ref.~\cite{Badger:2020uow}, where data to reproduce the histograms were provided by the authors.
+    We plot the log ratio of the matrix element as predicted by the neural network ensemble and the value from {\NJet} on the main axes for comparison.
+The blue and orange dipole histograms representing our method
+    are cut off at the top on the main axes, but the most important feature is the narrowness of the peak centred around the ideal value.
+    The insets show the detailed distribution of our result on a linear scale.}
+    \label{fig:comparison}
+\end{figure}
+
+\subsection{Main results}\label{sec:mainresult}
+Here we present our main results which are obtained using the larger NNs described in Section~\ref{sec:architecture} along with larger training datasets described in Section~\ref{sec:data}.
+In Figure~\ref{fig:main_error_distribution}, we show the error distributions on 10 million matrix element predictions for each multiplicity and global phase-space cut.
+
+\begin{figure}
+    \centering
+    \includegraphics[width=\textwidth]{fame1/main_error_distribution.pdf}
+    \caption{Error distributions for all three multiplicities (rows) and global phase-space cuts.
+    Left: absolute percentage difference between NN prediction and {\NJet}.
+    Right: ratio of matrix elements from NN and {\NJet}.
+    Axis scales have been fixed for each column of subplots for ease of comparison between multiplicities.}
+    \label{fig:main_error_distribution}
+\end{figure}
+
+Predicting on a large number of phase-space points allows us to explore singular regions with higher statistics. 
+The ratios (right) clearly highlight the symmetry of the errors with Gaussian-like distributions tightly centred around 0.
+We have also included the absolute percentage difference distribution for more easily interpretable errors where the bulk of predictions are below the 0.1\% error level.
+With increasing multiplicity, the fitting gets more challenging due to the rise in the number of singular regions in phase-space and the dimensionality of the phase-space, which can be seen in the decrease in accuracy as we increase multiplicity.
+Although the errors do increase, practically all matrix element predictions are below the 1\% error level even for the most challenging scenario.
+Relaxing the global phase-space cut for the training and testing set is also expected to decrease model performance as allowing more singular regions of phase-space stretches the span of the target distribution making it difficult to fit with a single neural network.
+Our method manages to retain good performance while global phase-space cuts have been relaxed by a factor of 100 with only a small decrease in accuracy as illustrated in the left column of Figure~\ref{fig:main_error_distribution}.
+This is because most of the span is accounted for by the dipole factors, while the coefficients themselves vary less.
+In the 3-jet case there is negligible difference between the different phase-space cuts while the 4 and 5 jet cases see less than a factor of 10 difference in the peaks of the absolute percentage difference distributions going from $y_{\mathrm{cut}} = 0.01$ to $y_{\mathrm{cut}} = 0.0001$.
+
+By increasing the size of the training datasets we aim to expose the neural network to more samples of the phase-space, thereby increasing accuracy on predictions on as much of the phase-space as possible.
+Along with increasing the number of training samples, the neural network architecture has been expanded to include more hidden nodes and hidden layers.
+The extra hidden nodes and layers introduces more parameters into the model allowing the neural network to utilise the additional data. 
+We found that to get good performance, we had to balance the size of the training dataset used and the size of the network.
+i.e.
+a small network is not expected to capture all the variations in a large dataset as easily as a larger network would, due to the smaller number of parameters available to the network.
+Of course, we also had to consider more physical constraints such as the time spent on training the neural networks which limits both the size of training datasets and architecture.
+In Figure~\ref{fig:error_500k_vs_40m}, we show the improvements in accuracy of our main NNs compared to the smaller NNs from Section~\ref{sec:n3jet_comparison}.
+Although the training and testing data for the smaller NNs are not identical due to differences in code used to generate the sets\footnote{although RAMBO is used for phase-space generation in both works, the selection criteria is different (JADE algorithm vs $k_{t}$ algorithm)}, for this comparison it suffices to show that the larger NNs are orders of magnitude more accurate than the smaller NNs.
+
+\begin{figure}
+    \centering
+    \includegraphics[width=\textwidth]{fame1/error_distribution_500k_vs_40m.pdf}
+    \caption{Comparison of error distributions of 10 million matrix element predictions between NNs used in Section~\ref{sec:n3jet_comparison} labelled as `Small' and our larger NNs labelled as `Large'.
+    Note that `small' and `large' NNs were trained on 500k and 40m training samples, respectively.
+    The testing data is identical to those shown in Figure~\ref{fig:main_error_distribution}, for the relevant global phase-space cuts.}
+    \label{fig:error_500k_vs_40m}
+\end{figure}
+
+Improvements in per-point accuracy translate to improved total cross-section predictions.
+In Figure~\ref{fig:cross_section_diffs} we show the percentage differences of the NN cross-section predicted compared to those from {\NJet}.
+There is a similar trend of errors increasing with increasing multiplicity and more inclusive phase-space cuts.
+All total cross-section predictions are well below 0.1\% error.
+There is a small systematic offset of the neural network cross-section compared to the {\NJet} cross-section that becomes apparent under closer inspection.
+This was discussed in Ref.~\cite{Badger:2020uow} and we provide an additional explanation in Appendix~\ref{appendix:fame1_jensen}.
+
+We also show in Figure~\ref{fig:cross_section_diffs} the estimated statistical Monte Carlo integration relative error for comparison.
+The fact that the accumulated error on the matrix element is much lower than the MC error on the cross-section opens up the possibility 
+to use the knowledge the network has gathered on the matrix element to augment the dataset to reduce the statistical error. Using such an 
+augmentation technique would introduce a new systematic error on the prediction related to the accumulated network interpolation/extrapolation error, 
+which would have to be balanced with the reduction in the MC integration error. Figure~\ref{fig:cross_section_diffs} suggests that the dataset could be 
+augmented in such a way by a large factor before reaching a minimal overall uncertainty. This opportunity might not seem very useful for this particular example of leading 
+order matrix elements where evaluations are relatively cheap computationally, but 
+if a similar degree of accuracy in the emulation can be obtained for higher order matrix elements, this procedure could reduce the resource cost of matrix element calculation
+significantly. We defer the study of this augmentation method to future work.
+
+\begin{figure}
+    \centering
+    \includegraphics[width=\textwidth]{fame1/main_cross_section_diffs_mc_error.pdf}
+    \caption{Absolute percentage differences between NN and {\NJet} cross-section predictions in solid lines.
+    Dashed lines represent the Monte Carlo error expressed as a percentage error relative to the {\NJet} cross-section calculated with the corresponding number of phase-space points.
+    Cross-sections have been calculated at intervals of 100k points up to the full 10 million phase-space points in the test set.
+    Axis scales have been fixed to highlight the differences between multiplicities.}
+    \label{fig:cross_section_diffs}
+\end{figure}
+
+\begin{figure}
+    \centering
+    \includegraphics[width=\textwidth]{fame1/errors_relaxed_cuts.pdf}
+    \caption{Left: error distribution on the $y_{\mathrm{cut}} = 0.01$ testing dataset as predicted by $y_{\mathrm{cut}} = 0.01$ and $y_{\mathrm{cut}} = 0.0001$ models.
+    Right: error distribution on the $y_{\mathrm{cut}} = 0.001$ testing dataset as predicted by $y_{\mathrm{cut}} = 0.001$ and $y_{\mathrm{cut}} = 0.0001$ models.
+    We use a logarithmic vertical axis to highlight the right-hand tail of the error distributions.}
+    \label{fig:errors_relaxed_cuts}
+\end{figure}
+Since we retain good performance by relaxing the global phase-space cut, 
+we carry out a simple test of generalisability by using the 5-jet $y_{\mathrm{cut}} = 0.0001$ model to infer on the two datasets with harsher cuts.
+This is shown in Figure~\ref{fig:errors_relaxed_cuts} where we see that accuracy is comparable to the reference (blue) in both cases.
+In the case of $y_{\mathrm{cut}} = 0.01$ (left), the model trained with more of the phase-space reduces errors in the right-hand tail of the distribution. 
+This proves that enlarging the training phase-space can be done without having a large detrimental effect on the overall accuracy, and can significantly reduce the number of
+large prediction errors.
+
+\subsection{Random trajectories}\label{sec:trajectories}
+As another test of our emulator we assess the accuracy of our predictions on random phase-space trajectories.
+These random phase-space trajectories are generated by connecting two random points in phase-space continuously without excluding any region of phase-space.
+This presents an interesting and challenging test of the interpolation and extrapolation abilities of the NNs as some parts of the trajectories may lie outside of the phase-space region of the training datasets.
+We show the results for 5-jet trajectories as this is the highest multiplicity we considered, predictions on lower multiplicities are better-behaved.
+We investigated 50 different random trajectories\footnote{the corresponding plots are available in a document attached to this submission}. For the discussion in this article 
+we chose one that contains many interesting features, namely the matrix elements span many orders of magnitudes and there are distinct peaks in the trajectory.
+In Figure~\ref{fig:random_trajectory} we show the predictions by the three 5-jet models trained on data with different global phase-space cuts for this trajectory.
+Left column shows the actual matrix element prediction and right column shows the ratios of the prediction to {\NJet}.
+We analysed the predictions for the 50 random trajectories and measured the fraction of their length where the accuracy falls within given intervals.
+Table~\ref{table:trajectory_predictions} shows the result for the regions of phase-space where training data was available, and those falling beyond the data available to the model.
+
+NN predictions are depicted as coloured scatter plots where the colour indicates the value of the minimum $s_{ij}$ between any pair of final-state particles at that phase-space point.
+To more easily visualise the extrapolation performance of the NNs we highlight the regions where the minimum $s_{ij}$ goes below the global phase-space cut applied to the training set the models were trained on,
+for each cut made. The regions of the plots where this occurs have been coloured in red.  
+With these trajectories being completely randomly selected in phase-space there is a possibility for there to be doubly singular points or worse.
+To check for this we used {\FastJet} to cluster the phase-space points in the same way we did for data generation, see Section~\ref{sec:data}.
+The pink regions indicate points which have two separate single unresolved limits, we label this configuration `Single+single'.
+The purple regions indicate points which have a double unresolved limit (i.e. three particles in one jet), we label this configuration `Double'.
+The blue regions indicate points which have both a double unresolved limit and a separate single unresolved limit, we label this configuration `Double+single'.
+We do not include the quark-anti-quark invariant in defining these regions as there is no associated infrared divergence.
+The pink, purple, and blue bands indicates regions of points which would have been discarded for our training and testing datasets.
+
+Accuracy is high when the minimum $s_{ij}$ of the trajectory is not below any $y_{\mathrm{cut}}$, i.e. when the NN prediction curve is blue.
+This is demonstrating that the NNs are interpolating well.
+Performance generally declines in the red, pink, purple, and blue regions, which is not unexpected as the NNs are extrapolating.
+Given that this trajectory has regions which go more collinear than any points the networks have been exposed to before, we would expect the networks which
+have been trained with the smallest $y_{\mathrm{cut}}$ parameter to perform best.
+We see that this is the case as accuracy is acceptable in the $y_{\mathrm{cut}} = 0.0001$ models, including in the regions where the NN is extrapolating.
+
+\begin{figure}
+    \centering
+    \includegraphics[width=\textwidth]{fame1/trajectory_1271_double.pdf}
+    \caption{Left: matrix element prediction of random phase-space trajectory.
+    Right: ratio of NN and {\NJet}.
+    NN predictions are coloured by minimum $s_{ij}$ pair.
+    Red bands indicate min($s_{ij}$) is smaller than $y_{\mathrm{cut}}$.
+    Pink bands indicate where there are two separate single unresolved limits.
+    Purple bands indicate double unresolved limits, i.e. three particles in one jet.
+    Blue bands indicate where there is both a double and a single unresolved limit.
+    The pink, purple, and blue bands represent regions of phase-space which would have been excluded by {\FastJet}.}
+    \label{fig:random_trajectory}
+\end{figure}
+
+\begin{table}
+    \centering
+    \begin{tabular}{|c|c|c|c|c|c|}
+        \hline
+        Region & $y_{\mathrm{cut}}$ & Fraction of pts & Pts outside $0.1\%$ & Pts outside $1\%$ & Pts outside $5\%$ \\
+        \hline
+        \multirow{3}{*}{White} & 0.01   & 36.5\% & 1.9\% & 0.0\% & 0.0\% \\
+        \cline{2-6}
+                               & 0.001  & 74.8\% & 0.98\% & 0.0\% & 0.0\% \\
+        \cline{2-6}
+                               & 0.0001 & 78.4\% & 1.9\% & 0.0\%  & 0.0\% \\
+        \hline
+        \hline
+        \multirow{3}{*}{Pink}  & 0.01   & 27.6\% & 58.2\% & 25.2\% & 12.9\% \\
+        \cline{2-6}
+                               & 0.001  & 13.5\% & 31.7\% & 6.5\% & 0.53\% \\
+        \cline{2-6}
+                               & 0.0001 & 13.5\% & 38.9\% & 9.3\% & 2.8\% \\
+        \hline
+        \hline
+        \multirow{3}{*}{Purple}& 0.01   & 7.2\% & 69.1\% & 32.1\% & 20.3\% \\
+        \cline{2-6}
+                               & 0.001  & 3.6\% & 31.3\% & 7.9\% & 4.5\% \\
+        \cline{2-6}
+                               & 0.0001 & 3.6\% & 30.5\% & 1.6\% & 0.0\% \\
+        \hline
+        \hline
+        \multirow{3}{*}{Blue}  & 0.01   & 1.5\% & 79.3\% & 41.6\% & 0.77\% \\
+        \cline{2-6}
+                               & 0.001  & 0.4\% & 29.2\% & 0.0\% & 0.0\% \\
+        \cline{2-6}
+                               & 0.0001 & 0.4\% & 76.4\% & 0.0\% & 0.0\% \\
+        \hline
+        \hline
+        \multirow{3}{*}{Red}   & 0.01   & 52.5\% & 76.2\% & 38.6\% & 20.5\% \\
+        \cline{2-6}
+                               & 0.001  & 7.0\%  & 69.9\% & 25.2\% & 9.5\% \\
+        \cline{2-6}
+                               & 0.0001 & 1.1\%  & 90.8\% & 33.8\% & 1.5\% \\
+        \hline
+    \end{tabular}
+    \caption{The performance of trajectory predictions separated for white, pink, purple, blue, and red regions.
+    We present the percentage of points that lie outside 0.1\%/1.0\%/5.0\% errors.
+    Fraction of points indicates the percentage of points that lie in the region of interest, out of all phase-space points from the 50 trajectories we examine.
+    }
+    \label{table:trajectory_predictions}
+\end{table}
+
+In summary, we have shown that the neural networks show acceptable performance on random phase-space trajectories which are of different nature to the datasets used to train and test the networks.
+Given that the general performance of the NNs of all three phase-space cuts are similar, it would make sense to use the models trained with the most inclusive phase-space cuts as it has been exposed
+to more of the complete phase-space.
+
+\section{Conclusion}\label{sec:conclusion}
+In this article we presented a new strategy to emulate matrix elements using a neural network. By leveraging the knowledge of 
+the factorisation properties of the matrix elements our model is able to extrapolate well outside of its training range. 
+We showed that using this method we obtain significantly improved per-point accuracy than obtained in previous works. 
+We also showed that the per-point accuracy of the model is not significantly affected by the generation cut for the training, which means that
+it would be possible to train our emulator on very inclusive cuts, allowing them to be applied in a multitude of settings.
+
+The accuracy of the emulation could allow users to augment the training dataset to reduce the MC error of cross-sections or distributions while using 
+fewer computing resources compared to the original calculation. We leave the investigation of this aspect to further work.
+
+Our method was demonstrated in this article using a tree-level process, but it could be generalised to higher order matrix elements by adapting the set 
+of ingredients made available to the network for the interpolation.
+Specifically, we can include extra terms into our ansatz to help capture additional divergences at higher orders.
+We look forward to applying this method to such cases.
+
 \end{document}

### 2022-11-02 intro.tex
proof read chapter 1

@@
     where subscripts denote the charges of the gauge groups. The first gauge
-    group with colour charge $C$ describes the interactions of the
+    group with colour charge $c$ describes the interactions of the
     strong force within the theory of quantum chromodynamics (QCD), which we
-    will elaborate more on in the next section. The second
+    will elaborate more on in Section~\ref{sec:qcd}. The second
     and third gauge groups represent the electroweak sector of the SM with
@@
     \subsection{The QCD Lagrangian}\label{sec:L_qcd}
-    Quantum chromodynamics (QCD) is the sector of the SM that
+    Quantum chromodynamics is the sector of the SM that
     describes the strong interaction. 
     QCD is a non-Abelian gauge theory with gauge group
-    $\mathrm{SU}(3)_{c}$ where the charge is named colour.
+    $\mathrm{SU}(N_{c} = 3)_{c}$ where the charge is named colour.
     The gauge and fermion part of the QCD Lagrangian is
@@
         \begin{split}
-            T^{a}_{ij}T^{a}_{jk} &= \delta{ik}C_{F} \, , \quad C_{F} = \dfrac{N_{c}^2-1}{2N_{c}} \, , \\
+            T^{a}_{ij}T^{a}_{jk} &= \delta_{ik}C_{F} \, , \quad C_{F} = \dfrac{N_{c}^2-1}{2N_{c}} \, , \\
             f^{abc}f^{abd} &= \delta^{cd}C_{A} \, , \quad C_{A} = N_{c} \, ,
@@
     is due to the process of renormalisation which will be discussed
-    in section ...
-    The implications of this is that at collider
-    experiments where collisions are extremely high energy,
+    in Section~\ref{sec:renormalisation}.
+    The implication of this is that at collider
+    experiments where collisions occur at extremely high energy,
     the strong coupling constant becomes small, a
@@
     we can examine the terms to write down how the
-    gauge bosons interact with the fermions. A convenient
-    way to do this is with Feynman rules, which can be
+    gauge bosons interact with the fermions, and especially
+    important for QCD, the gauge boson self interactions. A convenient
+    way to do this is by deriving the Feynman rules of the theory, which can be
     pieced together to form Feynman diagrams. We will see
-    that Feynman diagrams are a simple, but powerful, tool
-    to systematically build all the possible ways a process
-    can occur. This will be discussed further in the
-    scattering amplitude section.
+    that Feynman diagrams are a simple but powerful tool
+    to systematically build up all the possible ways a process
+    can occur.
 
@@
     anti-quark, the second term is a triple gluon
-    self-interaction and the third term is a four gluon
+    self-interaction, and the third term is a four gluon
     self-interaction. These mixing terms can be recast
-    as Feynman rules as interaction vertices.
+    into Feynman rules representing interaction vertices.
 
@@
         \includegraphics[width=\linewidth]{introduction/qcd_feynman_rules.pdf}
-        \caption{QCD Feynman rules.}
+        \caption{QCD Feynman rules with the Minkowski metric signature $g^{\mu\nu} = \mathrm{diag}(1, -1, -1, -1)$.}
         \label{fig:qcd_feynman_rules}
@@
 
-    In Figure~\ref{fig:qcd_feynman_rules} we have
-    collected the QCD Feynman rules.
-    Note that we have only written down the vertices,
-    propagators, and external lines for the physical gluons and quarks,
+    From these Lagrangians, we can read off the QCD Feynman rules
+    which we have collected in Figure~\ref{fig:qcd_feynman_rules}.
+    Note that we have only written down the vertices and propagators
+    for the physical gluons and quarks,
     ghosts also have associated Feynman rules but we do
@@
     quark-antiquark-gluon and triple gluon vertices
-    are associated with one power of $\alpha_{s}$,
+    are associated with one power of $\alpha_{s}$ (due to squaring $g_{s}$),
     and the four gluon vertex is associated with
@@
 \section{Factorisation theorem}\label{sec:factorisation}
-    At the LHC, collisions occur between two protons,
+    At the LHC collisions occur between two protons,
     which have constituent quarks and gluons, collectively
@@
     the factorisation theorem which enables the
-    separation of soft and hard energy scales.
-    Within the proton, there are composite quarks
+    separation of low and high energy scales (so-called soft and hard).
+    Within the proton, there are quarks
     and gluons that are constantly absorbed and
     emitted on a timescale inversely proportional
-    to the mass of the proton (or hadron in general).
+    to the mass of the proton.
     The timescale of this fluctuation is much longer
@@
     For a review of PDFs and how they are determined
-    see Ref. \cite{Ethier:2020way,Jimenez-Delgado:2013sma}.
+    see Refs. \cite{Ethier:2020way,Jimenez-Delgado:2013sma}.
     In practice, these PDF fits are accessed through the
@@
     representing parton $b$ emitting a parton $a$ carrying
-    a momentum fraction $z$. These splitting functions are calculable
+    a momentum fraction $x$. These splitting functions are calculable
     as a power series in $\alpha_{s}$ where they have been computed
@@
     \end{equation}
-
     $\hat{\sigma}_{ab \rightarrow n}(Q, \mu_{F}, \mu_{R})$
@@
     where the scattering occurs at energy scale $Q$,
-    which is often taken to the $\hat{s}$.
+    which is often taken to be $\hat{s}$.
     Notice that the partonic cross-section depends on
@@
     experiment is much greater than the mass
-    of the colliding particles.
+    of the colliding particles, and so they can
+    be neglected.
 
@@
     phase-space which contains all the possible
-    configurations of the $n$-particle final state.
+    configurations of the $n$-particle final-state.
     Absorbing the momentum conserving $\delta$-function
@@
     where the second $\delta$-function restricts
-    the final state particles to be on mass-shell
+    the final-state particles to be on mass-shell
     (on-shell). The step function selects only
-    the positive energy solution for this constraint.
+    the positive energy solution.
     This gives an intuitive picture of the partonic
@@
 
-    From (\ref{eqn:S_prob}), it becomes clear
+    From the definition of the transition probability
+    (\ref{eqn:S_prob}), it becomes clear
     that the matrix element is the object which
@@
     are exactly the sum over all Feynman diagrams
-    \footnote{technically the interference of
-    Feynman diagrams as we are working on the level
-    of $|\mathcal{M}|^{2}$}
+    \footnote{Since we are working on the level
+    of $|\mathcal{M}|^{2}$, it would technically
+    be the sum of all interference of Feynman diagrams.}
     for a particular process. However, there are
@@
     draw as there is no constraints on how the
-    vertices and propagators can be connected.
+    vertices and propagators can be connected, as
+    long as the diagram contains the desired initial
+    and final-states.
     Fortunately because $\alpha_{s}$
@@
     because the expansion $\alpha_{s}$
-    is small, each additional term contribute less.
+    is small, each additional term should contribute
+    less and less.
     This means that the first terms dominate the expansion,
@@
     reduces to the computation of the matrix elements
-    up to a fixed-order in the coupling expansion,
+    up to a fixed-order in the coupling parameter,
     which is then integrated over the valid phase-space of

### 2022-11-02 main.tex
proof read chapter 1

@@
 
-	Research presented in this thesis are based on joint work
+	Research presented in this thesis is based on joint work
 	published in the articles:
@@
 
-	Other research projects carried out during my studies but not included
+	Other research projects published during my studies but not included
 	in this thesis:

### 2022-11-04 mc.tex
first proofread

@@
 \chapter{Monte Carlo Event Generators}\label{chap:MC}
-    At a particle collider experiment,
+    At particle collider experiments,
     the collisions between incoming particles can lead
@@
     to simulate these stages of an event are general-purpose
-    Monte Carlo event generators (event genrators), see
-    Reference \cite{Buckley:2011ms} for a review.
+    Monte Carlo event generators (event generators), see
+    for example Reference \cite{Buckley:2011ms} for a review.
     They are indispensable tools that help
@@
     analyses made at the ATLAS and CMS experiments
-    involving some form of event generator usage.
-
+    involve some form of event generator usage.
     General-purpose event generators available
@@
     and Sherpa \cite{Gleisberg:2008ta,Sherpa:2019gpd} all broadly share 
-    the same features at a high-level overview.
+    the same features at a high-level and it is common
+    to utilise many different event generators
+    in one analysis to get an estimate of the uncertainty
+    due to the different approaches from each collaboration.
 
-    In the following we will motivate the necessity
-    for Monte Carlo methods and give a brief overview
+    In the following we will motivate the usage of
+    Monte Carlo methods and give a brief overview
     of the main stages involved in simulating an event.
@@
     of event generators, a more indepth treatment of the
-    specifics of Monte Carlo methods is given in Reference
+    specifics of Monte Carlo methods can be found in, for example, Reference
     \cite{James:1980yn}.
@@
     in each bin. The overall variance is minimised when
-    the the contributions from each bin are equal. This
+    the contributions from each bin are equal. This
     is achieved by sampling more from bins where the integral
@@
     \begin{equation}\label{eqn:event_weight}
-        w_{i} = |\mathcal{M}_{i}|^{2} J_{i} \, .
+        w_{i} = |\mathcal{M}|_{i}^{2} J_{i} \, .
     \end{equation}
@@
     or they are not, i.e. events come with unit-weight,
-    not a variable weight as they are currently being
-    generated with. In order to generate events that
+    not a variable weight in weighted events. In order to generate events that
     compare to reality, it has become customary to translate
@@
     These low efficiencies lead to most of the
-    yearly CPU budgets of LHC experiments to be spent on
+    yearly event generator CPU budgets of LHC experiments to be spent on
     unweighting events \cite{HSFPhysicsEventGeneratorWG:2020gxw}.
@@
     An alternative to improving the unweighting efficiency
-    itself is to accelerate the process of unweighting. This
+    itself is to accelerate the process of unweighting itself. This
     was explored in Reference \cite{Danziger:2021eeg} where a
@@
     energy regimes.
+
     \subsection*{Hard scattering}\label{sec:me_generators}
@@
     the exact matrix element is approximated by a perturbative
-    series in a coupling constant. The calculations of
+    series in a coupling constant. By now it is standard to
+    have at least the LO and NLO contributions in this expansion.
+    The calculations of
     tree-level \cite{Mangano:2002ea,Krauss:2001iv,Gleisberg:2008fv,Cafarella:2007pc,Alwall:2011uj,Kilian:2007gr}
@@
     to relatively few final-state particles due to the complexity
-    of analytical expressions and are only accurate in the
-    high energy limit. To evolve from the fixed-order
+    of analytical expressions and their phase-space integrals.
+    They are also only accurate in the high energy limit.
+    To evolve from the fixed-order
     calculation to the non-perturbative regime where hadrons
@@
     partonic cross-section at this scale as the central value,
-    the seven variations correspond to multiplying $\mu_{F}$
-    and $\mu_{R}$ by the factors 
+    the seven variations correspond to multiplying ($\mu_{F}, \, \mu_{R}$) by the factors 
     \begin{equation}\label{eqn:scale_variations}
-        \dfrac{\mu_{i}}{\mu} = \left\{(1, 1), (2,2),(0.5,0.5),(2,1),(1,2),(1,0.5),(0.5,1)\right\} \, ,
+        S \in \left\{(1, 1), (\sfrac{1}{2}, \sfrac{1}{2}), (2, 2), (\sfrac{1}{2}, 1), (1, \sfrac{1}{2}), (2, 1), (1, 2) \right\} \, ,
     \end{equation}
@@
     The estimate of the uncertainty is then taken
-    to be the half-width of the maximum interval.
+    to be half the width of the maximum interval.
 

### 2022-11-04 ml.tex
first proofread

@@
     depend on making use of large computing clusters
-    and international grid efforts, advancements
+    and international grid efforts \cite{WLCG}, advancements
     in technology and more efficient methods are
@@
     are at risk of outpacing the growth in these research
-    areas \cite{Bothmann:2022thx}. Therefore new techniques and
+    areas \cite{Bothmann:2022thx}, especially with
+    HL-LHC over the horizon. Therefore new techniques and
     algorithms from the machine learning community
@@
     to name a few which implement neural networks, alongside
-    other mainstay algorithms such as decisions trees
-    and support vector machines.
+    other mainstay algorithms such as decisions trees.
     These algorithms are now extremely commonplace
@@
     in an attempt to disentangle these events. With these
-    large arrays of detection elements there is an enormouse
+    large arrays of detection elements there is an enormous
     amount of data harvested from each collision, terabytes
@@
 
-    On the theory side of things, there have been numerous
-    triumphs in the application of machine learning.
-    Perhaps the most well-known is the fitting
-    of PDFs performed by the NNPDF collaboration \cite{NNPDF:2021njg}.
-    This area is rapidly growing with many new
-    ideas being explored at each stage of the event
+    On the theory side of things, machine learning
+    has been applied to many stages of the event
     generation process.
+    To name a few examples of active research areas:
+    phase-space sampling, matrix element modelling, hadronisation modelling,
+    and anomaly searches for new physics.
+    The research carried out in these areas have shown
+    promising results but a challenge that remains
+    is to interface these novel machine learning methods
+    to existing event generators.
     A living review that is actively being updated
@@
     networks: how they are constructed, how
-    they are trained, and how they are deployed.
+    they are trained, and how their parameters are optimised.
     We will then proceed to discuss their
-    applicability as surrogate models for matrix
-    elements.
+    application to modelling matrix
+    elements with a review of the current
+    state-of-the-art methods.
 
@@
     by the model weights $\mathbf{w}$ before being summed
-    with a bias term $b$. Collectively t he model weights
+    with a bias term $b$. Collectively the model weights
     and biases are the model parameters $\theta$.
@@
     the sigmoid function, hyperbolic tangent, rectified linear unit (ReLU).
-    Less familiar activation functions are the class of
-    sigmoid linear units (SiLU) of which the swish function
-    is a specific case. The functional forms of these
+    A less familiar class of activation functions are the
+    sigmoid linear units (SiLU) of which swish is an example,
+    which have been shown to outperform ReLU is some cases
+    \cite{DBLP:journals/corr/abs-1710-05941}. The functional forms of these
     functions are summarised in Table~\ref{table:activation_functions}.
     There is no single activation function that is the best
-    for any given task, it is common practice to choose activation
+    for any given task, and it is common practice to choose activation
     functions on a case-by-case basis.
@@
             \toprule
-            Activation functions & Functional form & Gradient \\
+            Activation & Functional form & Gradient \\
             \midrule
@@
         \end{tabular}
-        \caption{Summary of activation functions and their gradients.}
+        \caption{Functional forms of common activation functions and their gradients.}
         \label{table:activation_functions}
@@
     layers of neurons, and then connect every neuron
-    is a layer with every neuron in neighbouring layers.
+    in a layer with every neuron in neighbouring layers.
     This is illustrated in Figure~\ref{fig:nn_example}.
@@
     in these layers to best approximate the target function
-    $g$.
+    $g(\mathbf{x})$.
     Output layers are functionally the same as hidden
@@
     or alternatively, the number of activation functions
-    between the input and output layers. On the other hand,
+    between the input and output layers. Similarly,
     the width of a neural network is generally referring
@@
 
-    The output of the densely-connected neural network can
+    The output of a densely-connected neural network can
     be formulated by repeatedly applying (\ref{eqn:neuron})
@@
     corresponding to (\ref{eqn:MSE}) is
-    a function of many variables, and is highly
+    a function of many variables, and is likely highly
     non-convex with many local minima.
@@
     \end{equation}
-    where $\nabla_{\theta}L(\mathbf{x}_{\mathrm{batch}})$ is the
+    where $\nabla_{\theta}L(\mathbf{x}_{\mathrm{batch}}; \theta)$ is the
     gradient of the loss with respect to the parameters of the
-    model averaged. The initial state of the parameters
+    model, averaged over the batch of inputs $\mathbf{x}_{\mathrm{batch}}$.
+    The initial state of the parameters
     are usually distributed according to normal
@@
     of automatic differentiation tools. For these reasons, deep
-    neural networks have become much more widespread.
+    neural networks have become much more widespread due
+    to the ease of constructing and training them, as well
+    as their good predictive performance for general problems.
 
@@
     \subsection{Optimisation of hyperparameters}
+    Another set of parameters that need to be optimised
+    are the model hyperparameters.
     Hyperparameters are parameters that are
@@
     the speed and quality of the training process.
-    The learning rate in {\ref{eqn:gradient_descent}} is
+    For instance, the learning rate in ({\ref{eqn:gradient_descent}}) is
     an important hyperparameter whose initial value has
@@
     hyperparameters is often computationally expensive for two
-    reason. Firstly the parameter space to choose from is vast,
-    and the curse of dimensionality kicks in rapidly.
+    reason. Firstly the dimensionality of parameter space to choose from is high,
+    and so the curse of dimensionality kicks in rapidly.
     Secondly there is no exact a priori way to evaluate the
@@
     Reference \cite{Yu2020HyperParameterOA} for a review.
+    Below we discuss two paradigms of hyperparameter optimisation
+    that aim to reduce number of trials before landing
+    on a set of decent hyperparameters.
 
@@
     predictions to be lower than the statistical error in the
-    integration of matrix elements.
+    integration of matrix elements, which is a key motivation
+    for building an accurate emulator.
 
@@
     simple matrix multiplications, and because they are predicted
-    in batches vectorisation is automatic. Therefore the difficulty
-    in building a good surrogate model is in making it accurate.
+    in batches, vectorisation is automatic. Therefore the difficulty
+    in building a good surrogate model is in controlling
+    the accuracy.
 
@@
     on a per-point basis arises due to the infrared divergent behaviour
-    as discussed in Section~\ref{sec:IR_divergences}
+    as discussed in Section~\ref{sec:IR_divergences} which occurs
+    in corners of the phase-space.
     (UV divergences have been removed via renormalisation).
@@
     for diphoton plus jet production, where the authors presented a
-    use case of the emulator with a novel interface to the event
-    generator Sherpa. This work was revisited in \cite{Butter:2022rso},
+    use case of the emulator with a novel interface to the Sherpa event generator.
+    This work was revisited in \cite{Badger:2022hwf},
     where particular attention was paid to quantifying the uncertainty associated
@@
     the boosted training on regions of phase-space that were
-    lacking accuracy. Additionally, the authors showed
+    lacking in accuracy. Additionally, the authors showed
     that accuracy was improved with a more careful
@@
     matrix elements for $e^{+}e^{-} \rightarrow q \bar{q}$ + up to 3 gluons
-    can be emulated accurately to well below a 1\% accuracy,
+    can be emulated accurately to well below 1\% accuracy,
     even for the highest multiplicity. Extending this study
@@
     The question of uncertainty of the neural network
-    prediction is important in quantifying the theoretical
-    uncertainty. In Reference \cite{Danziger:2021eeg},
-    it was shown that with a surrogate model of the event
+    prediction is important as it contributes to the total
+    theoretical uncertainty. An alternative to quantifying
+    the uncertainty of the neural network prediction, is to
+    bypass the usage of the neural network prediction itself
+    in production.
+    Instead it can be used as a proxy for the exact matrix
+    element in event unweighting. This was explored in
+    Reference \cite{Danziger:2021eeg},
+    where it was shown that with a surrogate model of the event
     weight it was possible to use a double unweighting
@@
     model to study the potential gain factors with a more
-    accuracte model. A consequence of this two-stage unweighting
+    accuracte model. As previously mentioned, a consequence of
+    this two-stage unweighting
     algorithm is that the exact event weight has to be evaluated

### 2022-11-04 paper1.tex
first proofread

@@
 The capacity of neural networks to approximate intricate functions have already been 
-used to provide fast calculations of production cross-sections~\cite{otten2019deepxs,Buckley:2020bxg}.
+used to provide fast calculations of production cross-sections~\cite{Otten:2018kum,Buckley:2020bxg}.
 In this article we investigate whether NNs can approximate production cross-sections more differentially by replacing computationally expensive 

### 2022-11-04 qcd.tex
first proofread

@@
     This provides a bridge to compare
-    theoretical predictions and experiment.
-    From (\ref{eqn:dsigma}), the partonic cross-section
+    theoretical predictions and experimental measurements.
+    The relevant quantity calculated in perturbation theory
+    is the partonic cross-section which
     is the matrix element integrated over the relevant
@@
     calculated order-by-order in the strong coupling
-    $\alpha_{s}$, (\ref{eqn:matrix_element}),
+    $\alpha_{s}$ (c.f. \ref{eqn:matrix_element}),
     where each order corresponds to a set of Feynman diagrams
@@
     some solutions that have been widely adopted to circumvent
-    these issues to provide real-world applicable predictions.
-    This chapter forms the theoretical foundations on which
-    the work of this thesis is built upon.
+    these issues in order to provide real-world applicable predictions.
+    This chapter forms the theoretical foundations upon which
+    this thesis is built on.
 
@@
 \section{Divergent structures}\label{sec:divergences}
-    We have seen that matrix element computations
-    result in evaluating Feynman diagrams. From
-    the QCD Feynman rules (ref here), it becomes
-    clear that integrating over all momentum modes
-    will give rise to singularities. The divergences
+    The computation of matrix elements essentially reduces down
+    to evaluating Feynman diagrams which are analytical expressions
+    built up from the Feynman rules of a theory (c.f. QCD Feynman rules
+    \ref{fig:qcd_feynman_rules}). In evaluating certain
+    topologies of Feynman diagrams, integral expressions
+    containing unconstrained momenta will give rise to
+    singularities. The divergences
     associated with high energy modes are ultraviolet
-    (UV) divergences. On the opposite end of the energy
-    spectrum, low energy modes also give rise to 
+    (UV) divergences and on the opposite end of the energy
+    spectrum, low energy modes give rise to 
     infrared (IR) divergences.
@@
     During intermediate steps of calculations,
-    such as trying to compute loop diagrams seen in
-    the Figure~\ref{fig:bubble_diagram},
+    such as the compututation of loop diagrams seen in
+    Figure~\ref{fig:bubble_diagram},
     we have to evaluate integrals of the form
@@
     capture the divergence as the loop momenta
-    $k \rightarrow \infty$. It is clear that the
+    $\ell \rightarrow \infty$. It is clear that the
     integral diverges in this high energy limit,
@@
         \includegraphics[scale=1.2]{qcd/bubble_diagram.pdf}
-        \caption{Bubble diagram}
+        \caption{Massless bubble diagram which is an example of where unconstrained
+        loop momenta can lead to UV divergences.}
         \label{fig:bubble_diagram}
@@
     \begin{equation}\label{eqn:dim_reg}
-        % \int \dfrac{\mathrm{d}^{d}\ell}{(2\pi)^{d}}\dfrac{1}{\ell^{2}(\ell+p)^{2}} = \dfrac{\mathrm{i}(-p^{2})^{-\epsilon}}{(4\pi)^{2-\epsilon}}\dfrac{\Gamma(1-\epsilon)^{2}}{\Gamma(2-2\epsilon)}\Gamma(\epsilon) \, ,
-        I_{\mathrm{UV}} = \int \dfrac{\mathrm{d}^{d}\ell}{(2\pi)^{d}}\dfrac{1}{\ell^{2}(\ell+p)^{2}} = \dfrac{\mathrm{i}}{(4\pi)^{2}}\left(\dfrac{-p^{2}}{4\pi}\right)^{-\epsilon}\dfrac{\Gamma(1-\epsilon)^{2}}{\Gamma(2-2\epsilon)}\Gamma(\epsilon) \, ,
+        I_{\mathrm{UV}} = \int \dfrac{\mathrm{d}^{d}\ell}{(2\pi)^{d}}\dfrac{1}{\ell^{2}(\ell+p)^{2}} = \dfrac{\mathrm{i}}{(4\pi)^{2}}\left(\dfrac{-p^{2}}{4\pi}\right)^{-\epsilon}\dfrac{\Gamma(1-\epsilon)^{2}}{\Gamma(2-2\epsilon)} \, \Gamma(\epsilon) \, ,
     \end{equation}
@@
     With the UV divergence regularised by dimensional
-    regularisation, it is now possible to construct the forms
-    of the counterterms order-by-order in $\alpha_{s}$ to
-    explicitly cancel them.
+    regularisation, it is now possible to construct the
+    counterterms order-by-order in $\alpha_{s}$ to
+    explicitly cancel the divergences.
 
@@
     of the Lagrangian has to change accordingly to retain a
-    dimensionless action. We can account for this by
-    introducing an arbitrary scale $\mu_{R}$, the renormalisation
+    dimensionless action in natural units. We can account for this by
+    introducing an arbitrary energy scale $\mu_{R}$, the renormalisation
     scale, into the gauge coupling
@@
     \end{equation}
-    where it is understood the original gauge coupling is
+    where it is understood that the original gauge coupling is
     dimensionless. The order of $\mu_{R}$ is determined by examining
@@
     on the coupling constant $\alpha_{s}$, a point we will return
-    to in \ref{sec:alpha_running}.
+    to in Section~\ref{sec:alpha_running}.
     % In a calculation,
@@
     The inclusion of counterterms into the theory can be
-    thought of as renormalisations of the the bare fields
-    and couplings in (\ref{eqn:L_QCD}) to restore predictive
+    thought of as renormalisations of the bare fields
+    and couplings in the QCD Lagrangian (\ref{eqn:L_QCD}) to restore predictive
     power to the theory. The renormalised fields and couplings
-    can be given as (see e.g. Ref.~\cite{mangano1999introduction})
+    can be given as (see for instance Ref.~\cite{mangano1999introduction})
     \begin{align}\label{eqn:renormlised_fields}
@@
     \begin{equation}
-        \mathcal{L}_{\mathrm{renorm}} = \mathcal{L}_{\mathrm{bare}} + \mathcal{L}_{\mathrm{c.t.}} \, ,
+        \mathcal{L}_{R} = \mathcal{L}_{\mathrm{bare}} + \mathcal{L}_{\mathrm{c.t.}} \, ,
     \end{equation}
@@
     (MS) scheme subtracts only the epsilon pole appearing
-    in loop integrals (\ref{eqn:gamma_laurent}). However, the most
+    in loop integrals. However, the most
     commonly used regularisation scheme is the modified minimal
@@
     with this residual dependence is by carrying out
-    a scale variation. This involves evaluating the
-    relevant quantities at the nominal renormalisation
-    scale and again with the scale varied by a factor of
-    two above and below.
+    a scale variation (see Section~\ref{sec:sec:scale_variations}).
 
 \subsection{Running of the coupling constant}\label{sec:alpha_running}
-    The bare coupling should not depend on the actual
-    value of the renormalisation scale,
+    The renormalisation scale introduced during regularisation
+    is a mathematical artifact and should not impact any measurable
+    quantities. Therefore, the bare coupling should not depend on
+    the renormalisation scale,
     \begin{equation}\label{eqn:d_g_bare_d_mu}
-        \dfrac{\mathrm{d}g_{\mathrm{bare}}}{\mathrm{d}\mu_{R}} = 0 \, ,
+        \dfrac{\mathrm{d}g_{s,\mathrm{bare}}}{\mathrm{d}\mu_{R}} = 0 \, ,
     \end{equation}
     or said another way, the renormalisation process
-    is independent of the renormalisation scale. The
-    consequence of this is that $\alpha_{s}$ has to
-    depends on the renormalisation scale.
-
-    The behaviour of $\alpha_{s}$
+    is independent of the actual value of the renormalisation scale.
+    The consequence of this is that the renormalised coupling, $\alpha_{s}(\mu_{R})$, has to
+    depend on the renormalisation scale instead.
+
+    The renormalisation scale dependence of $\alpha_{s}$
     is governed by the Callan-Symanzik \cite{Callan:1970yg,Symanzik:1970rt}
@@
     \begin{equation}\label{eqn:beta_fn}
-        \mu^{2} \dfrac{\partial \alpha_{s}(\mu^{2})}{\partial \mu^{2}} = \beta(\alpha_{s}) \, ,
+        \mu_{R}^{2} \dfrac{\partial \alpha_{s}(\mu_{R}^{2})}{\partial \mu_{R}^{2}} = \beta(\alpha_{s}) \, ,
     \end{equation}
@@
     \begin{equation}\label{eqn:1l_alpha}
-        \alpha_{s}(\mu^{2}) = \dfrac{1}{\frac{\beta_{0}}{4\pi}\log\left(\frac{\mu^{2}}{\Lambda_{\mathrm{QCD}}^{2}}\right)} \, ,
+        \alpha_{s}(\mu_{R}^{2}) = \dfrac{1}{\frac{\beta_{0}}{4\pi}\log\left(\frac{\mu_{R}^{2}}{\Lambda_{\mathrm{QCD}}^{2}}\right)} \, ,
     \end{equation}
@@
     In QCD, $C_{A}=3$ and $T_{R}=\frac{1}{2}$, meaning for
-    $n_{f} < \frac{33}{2}$ the sign of $\beta_{0}$ is positive,
-    which is what we have observed in nature with six
-    flavours of quarks. This behaviour of positive
-    $\beta_{n}$ is also the case in all coefficients
-    computed to date, meaning that $\alpha_{s}$
+    $n_{f} < \frac{33}{2}$ the sign of $\beta_{0}$ is positive.
+    In nature we have observed six flavours of quarks, so
+    $\beta_{0} > 0$. In fact all $\beta$-coefficients
+    computed to date are positive, meaning that $\alpha_{s}$
     decreases with increasing energy due to the minus sign
@@
     asymptotic freedom \cite{Gross:1973id,Politzer:1973fx},
-    and is justification for treating QCD perturbatively.
+    and is justification for treating QCD perturbatively
+    when the energy scale is high, such as at collider experiments.
     The running of the coupling constant has been
@@
     and the behaviour of hadrons cannot be predicted
-    from perturbation theory.
+    from perturbation theory. Instead phenomenological
+    models have to be used to describe these processes.
 
@@
     are two cases in which these divergences arise:
-    in loop integrals (virtual) and when an emission
-    of an extra particle (c.f. splitting functions
-    (\ref{eqn:AP_kernels})) has vanishing energy (soft),
-    or becomes parallel to an external colourful line (collinear).
+    \begin{itemize}
+        \item virtual divergences in loop integrals,
+        \item real emission divergences when an emission
+    of an extra particle has vanishing energy (soft),
+    or becomes parallel to an external leg (collinear).
+    \end{itemize}
 
@@
     \begin{equation}\label{eqn:virtual_IR_integral}
-        I_{v} = \int \dfrac{\mathrm{d}^{4}\ell}{(2\pi)^{4}} \dfrac{1}{\ell^{2}(\ell+p_{1})^{2}(\ell-p_{2})^{2}} \, ,
+        I_{\mathrm{V}} = \int \dfrac{\mathrm{d}^{4}\ell}{(2\pi)^{4}} \dfrac{1}{\ell^{2}(\ell+p_{1})^{2}(\ell-p_{2})^{2}} \, ,
     \end{equation}
@@
     \begin{equation}\label{eqn:virtual_IR_DR}
-        I_{v} = \dfrac{\mathrm{i}}{(4\pi)^{2}Q^{2}}\left(\dfrac{-Q^{2}}{4\pi}\right)^{-\epsilon}\dfrac{\Gamma(1+\epsilon)\Gamma(1-\epsilon)^{2}}{\Gamma(1-2\epsilon)}\left[\dfrac{1}{\epsilon^{2}}\right] \, ,
+        I_{\mathrm{V}} = \dfrac{\mathrm{i}}{(4\pi)^{2}Q^{2}}\left(\dfrac{-Q^{2}}{4\pi}\right)^{-\epsilon}\dfrac{\Gamma(1+\epsilon)\Gamma(1-\epsilon)^{2}}{\Gamma(1-2\epsilon)}\left[\dfrac{1}{\epsilon^{2}}\right] \, ,
     \end{equation}
@@
         \includegraphics[scale=1.2]{qcd/vertex_correction.pdf}
-        \caption{Vertex correction}
+        \caption{One-loop vertex correction for the gluon-quark-antiquark vertex which diverges
+        for $\ell \rightarrow 0$.}
         \label{fig:vertex_correction}
@@
     the phase-space integrals of the matrix elements are rarely carried
-    out analytically due to the high dimensionality of the integral.
-    Instead they are done numerically with the use
-    of Monte Carlo methods, which requires the integral to be in
+    out analytically due to the high dimensionality of the integral,
+    rendering them intractable.
+    Instead they are done numerically through the use
+    of Monte Carlo methods (see Chapter~\ref{chap:MC}),
+    which requires the integral to be in
     integer dimensions. Therefore, numerical techniques
@@
     by a matrix element with the unresolved parton
-    being absorbed by an emitting leg (reduced matrix
-    element). This factorisation, however, is not exact
+    being absorbed by an emitting leg -- we call this the reduced matrix
+    element. This factorisation, however, is not exact
     in QCD where there are spin and colour correlations.
@@
     functions only depends on three partons, and not all
-    of the $n+1$ external states.
+    of the external states.
 
@@
     where terms less singular than $1 / \lambda^{2}$ have been
-    neglected. The reduced matrix element is obtained by removing
-    the soft gluon. The indices $i$, $j$, and $k$ label partons involved
+    neglected. The reduced matrix element represented by the
+    colour + helicity vector inner product is obtained by removing
+    the soft gluon from the $(n+1)$-body matrix element.
+    The indices $i$, $j$, and $k$ label partons involved
     in the factorsation process: $i$ is the emitter parton,
@@
     renormalisation scale, and $\bm{T_{i}}$ is the colour-charge
-    operator. It acts on the colour vectors to give factors of
-    $C_{A}$ and $C_{F}$, hence this factorisation is not exact.
+    operator. It acts on the colour vectors in the reduced matrix element
+    to give factors of $C_{A}$ and $C_{F}$, hence this factorisation is not exact.
 
@@
         \begin{split}
-        p_{i}^{\mu} &= z p^{\mu} + k_{\perp}^{\mu} - \dfrac{k_{\perp}^{2}}{z} \dfrac{n^{\mu}}{2 p n} \, \\
-        p_{j}^{\mu} &= (1-z) p^{\mu} - k_{\perp}^{\mu} - \dfrac{k_{\perp}^{2}}{1-z}\dfrac{n^{\mu}}{2 p n} \, , 
+        p_{i}^{\mu} &= z p^{\mu} + k_{\perp}^{\mu} - \dfrac{k_{\perp}^{2}}{z} \dfrac{n^{\mu}}{2 p \cdot n} \, \\
+        p_{j}^{\mu} &= (1-z) p^{\mu} - k_{\perp}^{\mu} - \dfrac{k_{\perp}^{2}}{1-z}\dfrac{n^{\mu}}{2 p \cdot n} \, , 
         \end{split}
@@
     the transverse direction perpendicular to the
-    collinear direciton ($p k_{\perp} = 0$), and $n^{\mu}$ is an auxiliary
-    vector satisfying the conditions $n^{2} = 0$ and $k_{\perp} n = 0$.
+    collinear direciton ($p \cdot k_{\perp} = 0$), and $n^{\mu}$ is an auxiliary
+    vector satisfying the conditions $n^{2} = 0$ and $k_{\perp} \cdot n = 0$.
     $z$ is the fraction of momenta carried away from
@@
     $ij \rightarrow i + j$ and the transverse momentum $k_{\perp}$.
-    Each splitting function is matrix acting on the spin indices of $ij$.
-    Due to these spin correlations the matrix
+    Each splitting function is a matrix acting on the spin indices of $ij$.
+    Due to these spin correlations the reduced matrix
     element does not factorise from the splitting functions exactly.
@@
 
-    In general, it is possible for the case in which a final-state
-    parton $i$ becomes collinear with an initial-state parton $a$.
+    In general, it is possible for a final-state
+    parton $i$ to become collinear with an initial-state parton $a$.
     This is described by the splitting process $a \rightarrow ai + i$.
@@
         \begin{split}
-        p_{i}^{\mu} &= (1-x)p_{a}^{\mu} + k_{\perp}^{\mu} - \dfrac{k_{\perp}^{2}}{1-x}\dfrac{n^{\mu}}{2p_{a} n} \, , \\
+        p_{i}^{\mu} &= (1-x)p_{a}^{\mu} + k_{\perp}^{\mu} - \dfrac{k_{\perp}^{2}}{1-x}\dfrac{n^{\mu}}{2p_{a} \cdot n} \, , \\
         p_{a}^{\mu} &= x p_{a}^{\mu} \, ,
@@
 
-    Similar to (\ref{eqn:soft_factorisation}), (\ref{eqn:F_collinear_factorisation})
+    Similar to (\ref{eqn:soft_factorisation}) where the factorisation was
+    only true in the strict soft limit, (\ref{eqn:F_collinear_factorisation})
     and (\ref{eqn:I_collinear_factorisation}) are only
@@
 \subsection{Factorisation of colour-ordered amplitudes}\label{sec:CO_factorisation}
-    The factorisation expressed in (\ref{eqn:soft_factorisation}),
+    The factorisation formulae expressed in (\ref{eqn:soft_factorisation}),
     (\ref{eqn:F_collinear_factorisation}), and (\ref{eqn:I_collinear_factorisation})
     were not exact because of colour and spin correlations.
-
     It can be shown that the factorisation of matrix
@@
 
-    For the scope of this thesis it is sufficient
+    In general any QCD amplitude can be
+    colour decomposed, that is the colour structure is
+    separated from the kinematics. Consider the process $e^{+}e^{-} \rightarrow q\bar{q} + n \; g$
+    \footnote{For the scope of this thesis it is sufficient
     to consider colour decomposition of electron-positron
-    annihilation. In general any QCD amplitude can be
-    colour decomposed. Consider the process $e^{+}e^{-} \rightarrow q\bar{q} + n g$,
+    annihilation into a single quark pair plus gluons,
+    and not the more general case of multiple quark
+    flavours in the final-state.},
     the amplitude can be written as a product of hadronic and
-    leptonic currents \cite{Campbell:1998nn} (on which this discussion is based on)
+    leptonic currents \cite{Campbell:1998nn}
     \begin{equation}\label{eqn:epem_amplitude}
@@
     partial amplitude $S_{\mu}(q_{1};1,\ldots,n;\bar{q}_{2})$ depending
-    only on kinematical variables. In this parton amplitude, the gluons
-    are emitted in an ordered fashion from the quarks.
-    The sum over $P(1,\ldots,n)$ represents a sum over all permutations
+    only on kinematical variables. In this partial amplitude, the gluons
+    are emitted in an ordered fashion from the quarks, meaning the quarks
+    have a fixed position in the partial amplitude.
+    The sum over $P(1,\ldots,n)$ represents the sum over all permutations
     of gluon emissions which accounts for all Feynman diagrams and colour structures.
@@
     \begin{equation}\label{eqn:epem_me2}
-        \left|\hat{\mathcal{S}}_{\mu}^{n+2}V^{\mu}\right|^{2} = e^{2}\left(\dfrac{g_{s}^{2}N}{2}\right)^{n}\left(\dfrac{N_{c}^{2}-1}{N_{c}}\right) \sum_{P(1,\ldots,n)} \left(\left|S_{\mu}(q_{1};1,\ldots,n;\bar{q}_{2})V^{\mu}\right|^{2} + \mathcal{O}\left(\dfrac{1}{N_{c}^{2}}\right)\right) \, ,
-    \end{equation}
-    where the subleading colour terms have been omitted.
+        \left|\hat{\mathcal{S}}_{\mu}^{n+2}V^{\mu}\right|^{2} = e^{2}\left(\dfrac{g_{s}^{2}N_{c}}{2}\right)^{n}\left(\dfrac{N_{c}^{2}-1}{N_{c}}\right) \sum_{P(1,\ldots,n)} \left(\left|S_{\mu}(q_{1};1,\ldots,n;\bar{q}_{2})V^{\mu}\right|^{2} + \mathcal{O}\left(\dfrac{1}{N_{c}^{2}}\right)\right) \, ,
+    \end{equation}
+    where the subleading colour terms proportional to $1/N_{c}^{2}$ have been omitted.
+    The term proportional to $1/N_{c}$ is the leading colour term
+    and is the dominant term in the colour expansion.
 
@@
     partial amplitudes, it is now possible to factorise (\ref{eqn:epem_me2})
-    exactly in the soft and collinear limits. In the limit where gluon
-    $j$ is soft, we have
+    exactly in the soft and collinear limits. Consider the limit where a final-state
+    gluon $j$ is soft, we have
     \begin{equation}\label{eqn:CO_soft_factorisation}
@@
     is the well-known eikonal factor. We see that the colour-ordered
-    amplitude on the RHS of (\ref{eqn:eikonal_factor}) has gluon $j$
+    amplitude on the RHS of (\ref{eqn:CO_soft_factorisation}) has gluon $j$
     removed but the ordering of all hard partons remain unchanged.
@@
     amplitude, plus a one-loop splitting function multiplied
-    by a tree-level amplitude as illustrated in in Fig ...
+    by a tree-level amplitude.
 
@@
     isolate these singularities such that it is possible
-    to numerically achieve predictions of physical
-    quantities that are finite. To tackle this task, three
+    to make finite predictions of physical quantities numerically.
+    To tackle this problem, three
     main mathods have been proposed: phase-space slicing \cite{Fabricius:1981sx,Kramer:1986mc,Giele:1991vf},
@@
     carry out a numerical calculation, it is therefore necessary
-    to regulate these divergences. In dimensional regularisation these
+    to regulate these divergences to make them explicit.
+    Using dimensional regularisation these
     divergences are mapped to poles in $\epsilon$.
@@
     divergences in (\ref{eqn:NLO_cs}) can be cancelled upon the insertion
-    of a counterterm evaluated in $\Phi_{n+1}$ and an integrated counterterm
-    evaluated in $\Phi_{n}$ such that the condition
+    of a counterterm evaluated in $\Phi_{n+1}$, $\mathcal{C}_{n+1}$,
+    and an integrated counterterm evaluated in $\Phi_{n}$, $\mathcal{I}_{n}$,
+    such that the condition
     \begin{equation}\label{eqn:subtraction_term_condition}
@@
     \end{equation}
-    holds. Here we have made use of the fact that
+    holds. Here phase-space factorisation is utilised:
     $\Phi_{n+1} \rightarrow \Phi_{n}\Phi_{1}$ when using
@@
     \begin{equation}\label{eqn:integrated_counterterm}
-        \mathcal{I}_{n} = \int \Phi_{1} \; \mathcal{C}_{n+1}
+        \mathcal{I}_{n} = \int \Phi_{1} \; \mathcal{C}_{n+1} \, ,
     \end{equation}
@@
     Additionally, since the pole structure in $\mathcal{V}_{n}$ is identical
-    to $\mathcal{R}_{n+1}$, $\mathcal{V}_{n} + \mathcal{I}_{n}$ will be finite
+    to $\mathcal{R}_{n+1}$ but with an opposite sign, $\mathcal{V}_{n} + \mathcal{I}_{n}$ will be finite
     by construction.
-    Inserting this into (\ref{eqn:NLO_cs}) we get
+    Inserting the subtraction terms into (\ref{eqn:NLO_cs}) we get
     \begin{equation}\label{eqn:NLO_subtraction}
@@
     In the next sections we will describe in detail two sets of functions
-    that are used to build subtraction terms: Catani-Seymour dipoles,
+    that are used to build subtraction terms: Catani-Seymour dipoles, and
     antenna functions. However, we will not construct the counterterms explicitly.
@@
     in the soft and collinear limits as these are universal and can
-    be applied to any process. These functions will become instrumental
+    be applied to any process, and not the subtraction terms themselves.
+    These functions will become instrumental
     during our construction of matrix element emulators in Chapters
@@
     Since the dipoles only depend on three partons, to map out
-    all the singular limits, it will be required to construct all
-    permutations of dipole functions.
+    all the singular limits in a process, it is necessary to construct all
+    the permutations of dipole functions.
 
@@
     In the following we will describe in detail the FF dipoles
+    and the associated matrix element factorisation formula,
     but only give a brief description of the remaining dipoles as
-    the structure of the factorisation formula is essentially identical.
+    the structure of the terms and factorisation formulae
+    are essentially identical.
 
@@
             dipoles, $\mathcal{D}$. The dipoles are named according to
-            whether the emitter and spectator are in the initial or final-state.
-            lEach dipole consists of a composite particle (denoted by tilde)
+            whether the emitter and spectator are in the initial (upper indices) or final-state (lower indices).
+            Each dipole consists of a composite particle (denoted by tilde)
             that decays into two partons, and a spectator that recoils
@@
             with incoming and outgoing lines representing initial and final-state
-            partons, respectively. represents the splitting
+            partons, respectively. The black circle represents the splitting
             function within the dipole function which contains the divergent
@@
         \end{equation}
-        where $\boldsymbol{T}_{i}$ are colour charges and $\boldsymbol{V}_{ij,k}$
+        where $\boldsymbol{T}_{i}$ are the colour-charge operators and $\boldsymbol{V}_{ij,k}$
         is a matrix in the helicity space of the emitter embedding the
         IR divergent behaviour.
-        The sum can be understood as summing over all possible
+        The sum in (\ref{eqn:FF_factorisation}) can be understood as summing over all possible
         3 leg permutations to capture all the soft and collinear limits.
@@
         Equations (\ref{eqn:FF_mapping}) and (\ref{eqn:y_ijk})
-        are a $2\rightarrow 3$ momenta mapping that maps
+        are a $3 \rightarrow 2$ momenta mapping that maps
         $p_{i} + p_{j} + p_{k} \rightarrow \tilde{p}_{ij} + \tilde{p}_{k}$.
@@
         \begin{equation}\label{eqn:zi}
-            \tilde{z}_{i} = \dfrac{p_{i}p_{k}}{p_{i}p_{k} + p_{j}p_{k}} \, , \quad \tilde{z}_{j} = \dfrac{p_{j}p_{k}}{p_{i}p_{k} + p_{j}p_{k}} = 1 - \tilde{z}_{i} \, ,
-        \end{equation}
-        which are analogous to the $z$ from Altarelli-Parisi splitting functions.
-        $\boldsymbol{V}_{ij,k}$ acts on the spin indices of the emitter and emitted
-        parton, but is independent of the type of the spectator. Making the dependence
-        on partons $i$ and $j$ explicit we list all the kernels here
+            \tilde{z}_{i} = \dfrac{p_{i}p_{k}}{p_{i}p_{k} + p_{j}p_{k}} \, , \quad \tilde{z}_{j} = \dfrac{p_{j}p_{k}}{p_{i}p_{k} + p_{j}p_{k}} = 1 - \tilde{z}_{i} \, .
+        \end{equation}
+        They are analogous to the $z$ from Altarelli-Parisi splitting functions.
+        $\boldsymbol{V}_{ij,k}$ acts on the spin indices of the composite particle $\widetilde{ij}$
+        , but is independent of the type of the spectator. Making the spin-dependence
+        on parton $\widetilde{ij}$ explicit ($s$ and $s'$ for $\widetilde{ij}=\,$fermion, and $\mu$ and $\nu$ for $\widetilde{ij}=\,$gluon)
+        we list all the kernels here
         \begin{align}\label{eqn:FF_V_ijk}
-            &\langle s | \boldsymbol{V}_{q_{i}g_{j},k} | s' \rangle = 8\pi \mu^{2\epsilon} \alpha_{s} C_{F} \left[\dfrac{2}{1-\tilde{z}_{i}(1-y_{ij,k})} - (1+\tilde{z}_{i}) - \epsilon(1-\tilde{z}_{i})\right] \delta_{ss'} \, , \nonumber \\
+            &\langle s | \boldsymbol{V}_{q_{i}g_{j},k} | s' \rangle = 8\pi \mu^{2\epsilon} \alpha_{s} C_{F} \left[\dfrac{2}{1-\tilde{z}_{i}(1-y_{ij,k})} - (1+\tilde{z}_{i}) - \epsilon \tilde{z}_{j} \right] \delta_{ss'} \, , \nonumber \\
             &\langle \mu | \boldsymbol{V}_{q_{i},\bar{q}_{j},k} | \nu \rangle = 8\pi \mu^{2\epsilon} \alpha_{s} T_{R} \left[-g^{\mu\nu} - \dfrac{2}{p_{i}p_{j}}(\tilde{z}_{i}p_{i}^{\mu}-\tilde{z}_{j}p_{j}^{\mu})(\tilde{z}_{i}p_{i}^{\nu}-\tilde{z}_{j}p_{j}^{\nu})\right] \, , \nonumber \\
@@
 
-        A key feature of these kernels and are that they smoothly interpolate
+        A key feature of these kernels are that they smoothly interpolate
         between the soft and collinear limits, meaning they do not double count
@@
         \begin{align}\label{eqn:FF_avg_V_ijk}
-            &\langle \boldsymbol{V}_{q_{i}g_{j},k} \rangle = 8\pi \mu^{2\epsilon} \alpha_{s} C_{F} \left[\dfrac{2}{1-\tilde{z}_{i}(1-y_{ij,k})} - (1+\tilde{z}_{i}) - \epsilon(1-\tilde{z}_{i})\right] \, , \nonumber \\
+            &\langle \boldsymbol{V}_{q_{i}g_{j},k} \rangle = 8\pi \mu^{2\epsilon} \alpha_{s} C_{F} \left[\dfrac{2}{1-\tilde{z}_{i}(1-y_{ij,k})} - (1+\tilde{z}_{i}) - \epsilon \tilde{z}_{j} \right] \, , \nonumber \\
             &\langle \boldsymbol{V}_{q_{i}\bar{q}_{j},k} \rangle = \, 8\pi \mu^{2\epsilon} \alpha_{s} T_{R} \left[1-\dfrac{2\tilde{z}_{i}\tilde{z}_{j}}{1-\epsilon}\right] \, , \nonumber \\
@@
         \end{align}
-        These spin-averaged functions will form the set of basis functions
-        that map out all the IR singular limits in matrix elements
-        in Chapter~\ref{chapter:fame1}.
 
@@
         \end{equation}
-        where the new dipole contribution appears on the second line. $a$ represents an initial-state parton.
+        where the new dipole contribution appears on the second line.
+        The initial-state parton is represented as $a$.
         The dipole term $\mathcal{D}_{ij}^{a}$ is given by
@@
         \begin{align}\label{eqn:x_ija}
-            &x_{ij,a} = \dfrac{p_{i}p_{a} + p_{j}p_{a} - p_{i}p_{j}}{(p_{i}p_{a} + p_{j}p_{a}} \, , \\
-            &\tilde{z}_{i} = \dfrac{p_{i}p_{a}}{p_{i}p_{a} + p_{j}p_{a}} \, , \quad \tilde{z}_{j} = \dfrac{p_{j}p_{k}}{p_{i}p_{k} + p_{j}p_{k}} = 1 - \tilde{z}_{1} \, .
+            &x_{ij,a} = \dfrac{p_{i}p_{a} + p_{j}p_{a} - p_{i}p_{j}}{(p_{i}p_{a} + p_{j}p_{a})} \, , \\
+            &\tilde{z}_{i} = \dfrac{p_{i}p_{a}}{p_{i}p_{a} + p_{j}p_{a}} \, , \quad \tilde{z}_{j} = \dfrac{p_{j}p_{a}}{p_{i}p_{a} + p_{j}p_{a}} = 1 - \tilde{z}_{i} \, .
         \end{align}
@@
         \begin{align}\label{eqn:FI_V_aij}
-            &\langle \boldsymbol{V}_{q_{i}g_{j},a} \rangle = 8\pi \mu^{2\epsilon}\alpha_{s} C_{F} \left[\dfrac{2}{1 - \tilde{z}_{i} + (1-x_{ij,a})} - (1+\tilde{z}_{i}) - \epsilon(1-\tilde{z}_{i})\right] \, , \nonumber \\
+            &\langle \boldsymbol{V}_{q_{i}g_{j},a} \rangle = 8\pi \mu^{2\epsilon}\alpha_{s} C_{F} \left[\dfrac{2}{1 - \tilde{z}_{i} + (1-x_{ij,a})} - (1+\tilde{z}_{i}) - \epsilon \tilde{z}_{j} \right] \, , \nonumber \\
             &\langle \boldsymbol{V}_{q_{i}\bar{q}_{j},a} \rangle = 8\pi \mu^{2\epsilon}\alpha_{s} T_{R} \, \left[1-\dfrac{2\tilde{z}_{i}\tilde{z}_{j}}{1-\epsilon}\right] \, , \nonumber \\
@@
         \begin{align}\label{eqn:V_aik}
-            &\langle \boldsymbol{V}_{k}^{q_{a}g_{i}} \rangle = \dfrac{n_{s}(\tilde{q})}{n_{s}(q)} \; 8\pi \mu^{2\epsilon} \alpha_{s} C_{F} \left[\dfrac{2}{1-x_{ik,a} + u_{i}} + (1+x_{ik,a}) - \epsilon(1-x_{ik,a})\right] \, , \nonumber \\
+            &\langle \boldsymbol{V}_{k}^{q_{a}g_{i}} \rangle = \dfrac{n_{s}(q)}{n_{s}(\tilde{q})} \; 8\pi \mu^{2\epsilon} \alpha_{s} C_{F} \left[\dfrac{2}{1-x_{ik,a} + u_{i}} + (1+x_{ik,a}) - \epsilon(1-x_{ik,a})\right] \, , \nonumber \\
             &\langle \boldsymbol{V}_{k}^{g_{a}\bar{q}_{i}} \rangle = \dfrac{n_{s}(g)}{n_{s}(\tilde{q})} \; 8\pi \mu^{2\epsilon} \alpha_{s} T_{R} \left[1 - \dfrac{2x_{ik,a}(1-x_{ik,a})}{1-\epsilon}\right] \, , \nonumber \\
@@
         \end{align}
-        where $n_{s}(a)$ is the number of polarisations of particle a.
+        where $n_{s}(a)$ is the number of polarisations of particle $a$.
         For fermions $n_{s}(q) = n_{s}(\bar{q}) = 2$
@@
             \tilde{p}_{ai}^{\mu} &= x_{i,ab} \, p_{a}^{\mu} \, , \nonumber \\
-            \tilde{k}_{j}^{\mu} &= k_{j}^{\mu} - \dfrac{2k_{j}(K+\widetilde{K})}{(K+\widetilde{K})^{2}}(K+\widetilde{K})^{\mu} + \dfrac{2k_{j}K}{K^{2}}\widetilde{K}^{\mu} \, , \\
+            \tilde{k}_{j}^{\mu} &= k_{j}^{\mu} - \dfrac{2k_{j} \cdot (K+\widetilde{K})}{(K+\widetilde{K})^{2}}(K+\widetilde{K})^{\mu} + \dfrac{2k_{j} \cdot K}{K^{2}}\widetilde{K}^{\mu} \, , \\
             x_{i,ab} &= \dfrac{p_{a}p_{b} - p_{i}p_{a}-p_{i}p_{b}}{p_{a}p_{b}} \, . \nonumber
@@
         reproduced by summing over all the possible dipoles that are
-        relevant. The omitted  terms are non-divergent.
+        relevant and hence will form the set of basis functions for
+        approximating the matrix elements in Chapters \ref{chapter:fame1}
+        and \ref{chapter:fame3}. The omitted terms are non-divergent
+        in all of phase-space.
 
@@
         extends to massive quarks \cite{Catani:2000ef}.
-        Massive partons are treated in detail in Reference\cite{Catani:2002hc}
-        where the authors derive dipole
-        functions for massive partons. The expressions derived therein
+        In Reference \cite{Catani:2002hc} the authors derive dipole
+        functions for massive partons which
         are utilised in Chapter~\ref{chapter:fame3}, but will
         not be written here as the philosophy of constructing
-        the dipole functions is similar to those already described.
+        the massive dipole functions is similar to massless ones.
 
@@
     emitter, antenna functions can have the unresolved parton
-    radiate from either of the two hard partons. In this sense,
-    an antenna is a linear combination of two dipoles.
+    radiate from either of the two hard partons in the antenna.
+    In this sense, an antenna is a linear combination of two dipoles.
 
@@
     accounting for identical particles in the final-state
-    and for the presence of multiple antennae in the two-
+    and for the presence of multiple antennae in the two
     parton process. $I$ and $K$ are hard partons forming a colour
@@
         \item Gluon-gluon: $H \rightarrow gg$ + (partons),
-        the decay of a Higgs boson into a pair of gluons with QCD from the gluon pair \cite{Gehrmann-DeRidder:2005alt}.
+        the decay of a Higgs boson into a pair of gluons with QCD radiation from the gluon pair \cite{Gehrmann-DeRidder:2005alt}.
     \end{itemize}
@@
     to fulfill this requirement.
-    The antenna functions are summarised in Table~\ref{table:antenna_functions}, where
-    antenna functions are categorised by their class and radiative
+    The antenna functions are summarised in Table~\ref{table:antenna_functions},
+    where they are categorised by their class and radiative
     processes.
@@
         one-loop level correspond to the leading colour ($X_{3}^{1}$),
-        sub-leading colour ($\tilde{X}_{3}^{1}$), and closed quark loop
+        subleading colour ($\tilde{X}_{3}^{1}$), and closed quark loop
         ($\hat{X}_{3}^{1}$) contributions.}
@@
         \caption{Three-parton antenna functions where the loop order
-            is implicitly illustrated in the grey circle,
+            is implicitly illustrated by the grey circle,
             and the grey ellipse represents all diagrams that give
@@
     namely the partonic channel $e^{-}e^{+} \rightarrow
-    q \bar{q}$ + $ng$.
-    For this we require the $A$, $D$, and $F$ antenna functions,
+    q \bar{q}$ + $n \, g$. The emulation of the NLO QCD K-factors
+    of this partonic process is examined in Chapter~\ref{chap:fame2}.
+    For this process we require the $A$, $D$, and $F$ antenna functions,
     which are illustrated in Figure~\ref{fig:antenna}.
@@
     and $g \in \{3, 4, 5\}$ in view of using the
-    antenna functions in a calculation where the momenta
+    antenna functions where the momenta
     of partons will have fixed indices in a phase-space
-    point.
+    point represented as an array.
 
@@
     function. However, there are still finite terms
-    from the $\mathcal{P}oles$ operator coming from
-    the IR singularity operators (cite here). Therefore,
+    from $\mathcal{P}oles(X)$ coming from the $\epsilon$-expansion
+    of the IR singularity operators given below. Therefore,
     care has to be taken to extract all the finite
-    remainders from the antenna functions.
+    remainders from the antenna functions when trying
+    to use them as basis functions for emulating the
+    finite part of the one-loop matrix element.
 
@@
     The antenna functions have some functions in common,
-    namely the singularity operators (cite here)
+    namely the singularity operators
     \begin{align}\label{eqn:singularity_operators}
@@
     in the relevant regions of phase-space.
+
+    Utilising this property of the antenna functions, and the fact
+    that they naturally interpolate between the soft and collinear
+    limits due to being derived from physical matrix elements,
+    they can be used as a set of basis functions for approximating
+    matrix elements. This will be explored in Chapter~\ref{chapter:fame2}.
 \end{document}

### 2022-11-04 paper1.tex
fix typos

@@
 The mass of the $Z$-boson is taken to be $m_{Z} = 91.188$ GeV.
-These parameters and those not listed, are consistent with those in the Standard Model mode of \aNLO \cite{Alwall_2014}.
+These parameters and those not listed, are consistent with those in the Standard Model mode of \aNLO \cite{Alwall:2014hca}.
 
@@
     \centering
+    \small
     \begin{tabular}{|c|c|c|c|c|c|}
         \hline
-        Region & $y_{\mathrm{cut}}$ & Fraction of pts & Pts outside $0.1\%$ & Pts outside $1\%$ & Pts outside $5\%$ \\
+        Region & $y_{\mathrm{cut}}$ & Frac. of pts & Pts outside $0.1\%$ & Pts outside $1\%$ & Pts outside $5\%$ \\
         \hline

### 2022-11-08 intro.tex
rewrite intro

@@
 \label{chapter:intro}
-The standard model of particle physics (SM) is our current best working
-theory that describes the phenomenon we observe, for  example,
- at collider experiments.
-The large hadron collider (LHC) is a high energy proton-proton
-collider experiment that allows us to study the fundamental particles
-laid out in the standard model.
 
-The interaction of these fundamental particles are described
-by scattering amplitudes. The standard model allows us to
-calculate these amplitudes, or matrix elements, from first
-principles. A key challenge in the past few decades has been
-to compute these matrix elements for increasing number of
-final state particles and for higher orders in perturbative
-expansions in fixed order calculations. Although there
-has been tremendous progress in this sector a complication
-that naturally arises is that the evaluation of these matrix
-elements becomes increasingly time-consuming due to their
-algebraic complexity.
+The Large Hadron Collider (LHC) at CERN has been colliding
+particles since 2009 with each successive run increasing the centre-of-mass
+energy of the collision (7-13.6 TeV). With the experiment operating
+over the course of decades, there is a vast quantity of
+experimental data collected that has to be compared with
+theoretical predictions. Currently, the theory predictions
+are made with the Standard Model (SM) of
+particle physics which describes the fundamental particles
+of nature and their interactions.
 
-The topic of this thesis is to apply modern machine learning
-techniques to reduce the time spent evaluating these
-complex matrix elements. Considering the physical properties
-of these matrix elements allows us to implement physics
-knowledge into these machine learning algorithms giving
-us higher levels of performance than would otherwise be
-achievable.
+The standard paradigm for comparing
+theory and experiment is to simulate the particle collisions
+from the initial collision all the way through
+to the detection of the hundreds of particles produced in the collision.
+General purpose Monte Carlo event generators are the de-facto
+tool designed to generate these simulated events.
+With the large number of scattering events collected at
+the LHC already, and with even more expected at the High Luminosity LHC,
+the rate of event generation is at risk of
+being outpaced by experimental data collection.
 
-The structure of this thesis is as follows: in Chapter~\ref{chapter:intro}
-I will recap the necessary concepts from the standard
-model, in particular, on quantum chromodynamics (QCD), the sector
+The process of generating simulated events can be broadly
+split up into three segments:
+the highly energetic hard scattering process, 
+the subsequent cascading of particles (parton shower),
+and the formation of bound states detected (hadronisation).
+While hadronisation is based on phenomological models,
+the hard scattering and parton shower are derived
+from the SM and are theoretically well-motivated.
+
+The hard scattering process can be formulated as a probability
+for particles to collide, represented by matrix elements.
+These matrix elements are formally calculated order-by-order
+in perturbation theory with each order leading
+to increasingly difficult computations. In practice,
+these matrix elements are provided in libraries interfaced
+to the event generators and evaluation is largely automated for many
+of the most important processes. However, the evaluation time
+for these complex expressions becomes problematic
+when generating the large samples required to compare with
+experiments, especially when a significant portion of time
+spent simulating an event is spent on the matrix element
+evaluation.
+
+The focus of this thesis will be on applying modern
+machine learning methods to accelerate the evaluation
+of matrix elements, hence speeding up the event
+generation process in general. Machine learning algorithms
+have become ubiquitous in many fields and their applicability
+to a wide range of problems have made them a popular choice
+in the high energy physics community as well. By combining the
+well understood behaviour of matrix elements in specific
+kinematic regions with powerful machine learning
+algorithms, the construction of physics-inspired machine
+learning models unlocks higher levels of performance
+than would otherwise be achievable.
+
+The structure of this thesis is as follows: in this chapter
+I will recap the necessary concepts from the SM,
+in particular quantum chromodynamics (QCD), the sector
 which governs the behaviour of quarks and gluons. Furthermore,
-the relationship between theory and experiment will be elaborated on.
+the relationship between theory and experiment will be elaborated on
+with the introduction of cross-sections and their relation
+to matrix elements.
 
-In Chapter~\ref{chapter:qcd} I will discuss the challenges that arise
-in fixed order perturbative calculations and the current methods
-that have been adopted in the community to deal with them. These
-methods also form the basis of our physics knowledge which we
+In Chapter~\ref{chapter:qcd}, I will discuss the challenges that arise
+in fixed-order perturbative calculations and the current methods
+that have been adopted in the community to deal with them. Some of
+these methods also form the theoretical foundation which we
 embed into the machine learning algorithm. This procedure is 
@@
 A similar philosophy is applied in Chapter~\ref{chapter:fame2}
-where next-to-leading order QCD k-factors are emulated.
+where next-to-leading order QCD k-factors are emulated for
+the same processes.
 These two chapters discuss in detail the procedure for
-electron-positron annihilation, however, the most relevant
-experiment currently is the LHC which is colliding protons.
-The extension to hadron-hadron colliders is detailed in
-Chapter~\ref{chapter:fame3} where we explore using the
+constructing emulators for electron-positron annihilation,
+however, the more relevant processes for the LHC are in
+proton-proton initiated collisions. The extension
+to this scenario is detailed in Chapter~\ref{chapter:fame3}
+where we also explore using the
 emulator in a novel implementation to replace traditional
 matrix element providers in the event generator SHERPA.
-
-Finally, I will conclude the thesis in Chapter~\ref{chapter:conclusion}.
+Finally, I will summarise and conclude the thesis in
+Chapter~\ref{chapter:conclusion}.
 
@@
     and the discovery of the Higgs boson in 2012 by the ATLAS \cite{ATLAS:2012yve}
-    and CMS experiments \cite{CMS:2012qbp} at the Large Hadron Collider
-    (LHC) which was theorised decades prior.
+    and CMS experiments \cite{CMS:2012qbp} at the LHC
+    which was theorised decades prior.
 
@@
     of this is that predictions made in QCD can be expressed
-    in the form of perturbative expansions in $\alpha_{s}$.
+    in the form of perturbative expansions in $\alpha_{s}$ (see Section~\ref{sec:xs}).
 

### 2022-11-08 qcd.tex
fix typos
rewrite intro

@@
     with this residual dependence is by carrying out
-    a scale variation (see Section~\ref{sec:sec:scale_variations}).
+    a scale variation (see Section~\ref{sec:scale_variations}).
 
@@
     q \bar{q}$ + $n \, g$. The emulation of the NLO QCD K-factors
-    of this partonic process is examined in Chapter~\ref{chap:fame2}.
+    of this partonic process is examined in Chapter~\ref{chapter:fame2}.
     For this process we require the $A$, $D$, and $F$ antenna functions,

### 2022-11-15 mc.tex
chapter 2.2, 2.3, 2.4 and 3 changes

@@
     of particles. In a hadron collider, such as the LHC,
-    the particles that we observe are not the constituents
+    the particles we observe are not the constituents
     of the proton, the quarks and gluons. Instead we observe
@@
 
-    In order to describe the entire chain of events
+    In order to describe the entire process
     from the initial energetic collision to the production
@@
     split up the process into stages characterised by
-    their kinematic regime. The standard tool used
+    their kinematic hardness (c.f. Section~\ref{sec:factorisation}).
+    The standard tools used
     to simulate these stages of an event are general-purpose
@@
     for example Reference \cite{Buckley:2011ms} for a review.
-    They are indispensable tools that help
+    These are indispensable tools that help
     bridge the gap between theoretical predictions
@@
     analyses made at the ATLAS and CMS experiments
-    involve some form of event generator usage.
-    General-purpose event generators available
+    involving some form of event generator usage.
+    General-purpose event generators
     such as Herwig \cite{Bahr:2008pv,Bellm:2015jjp},
@@
     and Sherpa \cite{Gleisberg:2008ta,Sherpa:2019gpd} all broadly share 
-    the same features at a high-level and it is common
-    to utilise many different event generators
-    in one analysis to get an estimate of the uncertainty
-    due to the different approaches from each collaboration.
+    the same features at a high-level, but each framework
+    carries subtle differences. Hence, it is common
+    to use many different event generators
+    in one analysis to obtain an estimate of the uncertainty
+    due to the different approaches in each framework.
 
@@
     has no scaling with the dimensionality of the integral,
-    so can be used for arbitrarily high multiplicity final-states.
+    and so can be used for arbitrarily high multiplicity final-states
+    with no performance penalty compared to lower multiplicities.
     The purpose of this chapter is to give a brief overview
-    of event generators, a more indepth treatment of the
+    of Monte Carlo integration in the context of event
+    generation, a more in-depth treatment of the
     specifics of Monte Carlo methods can be found in, for example, Reference
@@
     the integral of a function $f(\vec{x})$ by sampling it
-    at $d$-dimensional parameter space points $\vec{x}$
+    in $d$-dimensional parameter space points $\vec{x}$
     \begin{equation}\label{eqn:MC_integration}
-        I = \int_{V} \mathrm{d}^{D}x \; f(\vec{x}) \approx \langle I(f) \rangle_{x} = V \dfrac{1}{N} \sum_{i=1}^{N} f(\vec{x}_{i}) = V \langle f \rangle_{x} \, ,
+        I = \int_{V} \mathrm{d}^{d}x \; f(\vec{x}) \approx \langle I(f) \rangle_{x} = V \dfrac{1}{N} \sum_{i=1}^{N} f(\vec{x}_{i}) = V \langle f \rangle_{x} \, ,
     \end{equation}
-    where the points $\vec{x}$ are randomly (hence Monte Carlo),
+    where the points $\vec{x}$ are randomly,
     uniformly sampled in the volume $V$. $\langle f \rangle_{x}$
@@
     mean $\langle I(f) \rangle_{x}$,
-    meaning an error esimate on the estimator is given by
+    meaning an error estimate on the estimator is given by
     the standard error of the mean
@@
     is achieved by sampling more from bins where the integral
-    is rapidly fluctuating (and not necesarily large),
+    is rapidly fluctuating (and not necessarily large),
     and less from bins where the integral has less fluctuations.
@@
     from the mapping from random variables to the four-momenta.
-    This Jacobian is referred to as the phase-space weight, which together
+    This Jacobian, along with the PDF weights and flux factors,
+    are referred to as the phase-space weight, $J_{i}$. Taking
+    this phase-space weight together
     with the matrix element forms the event weight\footnote{For
@@
     The event weight can be interpreted as the probability
-    of this particular event occurring at an experiment \cite{Plehn:2009nd}.
+    of the event occurring at an experiment \cite{Plehn:2009nd}.
     By generating a number of events in this manner,
@@
 
-    At experiments, measured events are either recorded
-    or they are not, i.e. events come with unit-weight,
-    not a variable weight in weighted events. In order to generate events that
-    compare to reality, it has become customary to translate
+    At experiments, collision events are either observed or unobserved,
+    meaning each observed configuration contributes equally to
+    the measured cross-sections, and so are equally weighted.
+    Therefore, in order to generate events that
+    compare to experiments, it is necessary to translate
     weighted events into unit-weight events.
     This process is known as unweighting and is typically
-    carried out using a rejection sampling algorithm
+    carried out using a rejection sampling algorithm, as
     outlined in Algorithm~\ref{alg:unweighting}. The rejection
@@
             \If{$w / w_{\mathrm{max}} > r$}{
-                \KwRet $\vec{x}$ and $\tilde{w} \gets \max(1,\; w / w_{\mathrm{max}})$\;
+                \KwRet $\vec{x}$ and $\tilde{w} \gets 1$\;
+                % $\tilde{w} \gets \max(1,\; w / w_{\mathrm{max}})$\;
             }
@@
     events required to obtain a single unweighted event.
-    Unweighting efficiency for high-multiplicity processes
-    are often well below 1\% \cite{Hoche:2019flt,Gao:2020zvv} and development
-    of methods to improve the efficiency have been
-    explored \cite{Jadach:1999sf,Jadach:2002kn}.
+    Unweighting efficiencies for high-multiplicity processes
+    are often well below 1\% \cite{Hoche:2019flt,Gao:2020zvv} and the development
+    of methods to improve the efficiency are an active area of research
+    \cite{Jadach:1999sf,Jadach:2002kn}.
     These low efficiencies lead to most of the
@@
     itself is to accelerate the process of unweighting itself. This
-    was explored in Reference \cite{Danziger:2021eeg} where a
+    was explored in Reference \cite{Danziger:2021eeg} wherein a
     novel two-stage unweighting procedure employing a surrogate
@@
     the parton shower, and finally hadronisation.
-    This is motivated by the factorisation theorem
+    This is motivated by the factorisation theorem (Section~\ref{sec:factorisation})
     which separates the high energy regimes from the low
@@
 
-    As seen in (\ref{eqn:matrix_element}),
+    As seen in Eq. (\ref{eqn:matrix_element}),
     the exact matrix element is approximated by a perturbative
@@
     For more information on the machinery developed to
-    compute nearly arbitrary matrix elements at tree-level and
-    one-loop level, see for instance Reference \cite{Campbell:2017hsr}.
+    compute matrix elements at tree-level and
+    one-loop level required for SM predictions, see for instance Reference \cite{Campbell:2017hsr}.
     NNLO and N$^{3}$LO QCD corrections have been computed
@@
 
-    Matrix element calculations make up a large
+    Matrix element calculations account for a large
     portion of the total event generation time \cite{Bothmann:2022thx}.
@@
     of analytical expressions and their phase-space integrals.
-    They are also only accurate in the high energy limit.
+    They are also only accurate at high energy scales.
     To evolve from the fixed-order
@@
     an infrared scale where non-perturbative effects
-    begin to set in. In doing so parton showers account
+    begin to set in. In doing so, parton showers account
     for potentially large logarithms that arise in
@@
     \subsection*{Hadronisation}\label{sec:hadronisation}
-    The formation of hadrons in the low energy regime
-    of an event generator follow non-perturbative models.
+    The simulation of the formation of hadrons in the low-energy regime
+    of an event generator follow phenomenological, non-perturbative models.
+    These hadronisation models require the combination
+    of quarks and antiquarks to form colourless hadrons.
     The two main classes of models used to simulate
@@
     and string models \cite{Andersson:1983ia}.
-    These models share a similar concept of grouping together
-    quarks and antiquarks to ensure only hadrons remain
-    in the final-state. For more discussion on hadronisation
+    For more discussion on hadronisation
     see for example Reference \cite{Webber:1999ui}.
@@
     \section{A note on theoretical uncertainties}\label{sec:scale_variations}
-    The partonic cross-section, $\hat{\sigma}(\mu_{F}, \mu_{R})$ has
+    The partonic cross-section, $\hat{\sigma}(\mu_{F}, \mu_{R})$, has
     a dependence on both the factorisation scale and the
@@
     missing higher order terms is to carry out a seven-point
-    scale variation. Typically in a calculation we take
-    $\mu = \mu_{F} = \mu_{R}$ where $\mu$ depends on the specifics
+    scale variation. Typically in a calculation,
+    $\mu_{0} = \mu_{F} = \mu_{R}$ where $\mu_{0}$ depends on the specifics
     of the process of interest. Taking the value of the
@@
     \end{equation}
-    and re-evaluating the
-    cross-section at these scales.
-    The estimate of the uncertainty is then taken
-    to be half the width of the maximum interval.
+    and re-evaluating the cross-section at these scales.
+    The deviations from the central cross-section value
+    forms an estimate of the uncertainty in the perturbatively
+    computed cross-section, when compared to the true cross-section.
+    This can be written as
+    \begin{equation}
+        \hat{\sigma}(\mu_{F}, \mu_{R}) = \hat{\sigma}(\mu_{0}, \mu_{0}) \pm \delta \, ,
+    \end{equation}
+    where $\delta$ is determined from the scale variations.
+    However, there is no single, well agreed upon definition
+    of $\delta$ \cite{Cacciari:2011ze}.
+    The reasoning behind this approach is that $\delta$
+    is approximately the same order as the true missing
+    higher order terms.
 

### 2022-11-15 qcd.tex
chapter 2.2, 2.3, 2.4 and 3 changes

@@
     \begin{equation}\label{eqn:soft_factorisation}
-        |\mathcal{M}_{n+1}|^{2} \rightarrow -\dfrac{1}{\lambda^{2}}8\pi \mu^{2\epsilon}\alpha_{s} \sum_{i}\dfrac{1}{p_{i}q}\sum_{k \neq i} \dfrac{p_{k}p_{i}}{(p_{i} + p_{k})q} \; {}_{n}\langle 1, \ldots, n+1 | \bm{T}_{k} \cdot \bm{T}_{i} | 1, \ldots, n+1 \rangle_{n} \, ,
+        \begin{split}
+        |\mathcal{M}_{n+1}|^{2} \rightarrow &-\dfrac{1}{\lambda^{2}}8\pi \mu^{2\epsilon}\alpha_{s} \sum_{i}\dfrac{1}{p_{i}q} \sum_{k \neq i} \dfrac{p_{k}p_{i}}{(p_{i} + p_{k})q} \\
+        &{}_{n}\langle 1, \ldots, j-1, j+1, \ldots, n+1 | \bm{T}_{k} \cdot \bm{T}_{i} | 1, \ldots, j-1, j+1, \ldots, n+1 \rangle_{n} \, ,
+        \end{split}
     \end{equation}
@@
     \begin{equation}\label{eqn:F_collinear_factorisation}
-        |\mathcal{M}_{n+1}|^{2} \rightarrow \dfrac{1}{p_{i}p_{j}}4\pi \mu^{2\epsilon}\alpha_{s} \; {}_{n} \langle 1, \ldots, n+1 | \hat{P}_{ij}(z, k_{\perp}) | 1, \ldots, n+1 \rangle_{n} \, ,
+        |\mathcal{M}_{n+1}|^{2} \rightarrow \dfrac{1}{p_{i}p_{j}}4\pi \mu^{2\epsilon}\alpha_{s} \; {}_{n} \langle 1, \ldots, ij, \ldots, n+1 | \hat{P}_{ij}(z, k_{\perp}) | 1, \ldots, ij, \ldots, n+1 \rangle_{n} \, ,
     \end{equation}
@@
     $e$ and the strong gauge coupling $g_{s}$ appearing. The colour
-    structure has been factored into a product of fundamental group
+    structure has been factorised into a product of fundamental group
     generators where the indices $a_{i} \in \{1,\ldots, N_{c}^{2}-1\}$
@@
     where the subleading colour terms proportional to $1/N_{c}^{2}$ have been omitted.
-    The term proportional to $1/N_{c}$ is the leading colour term
+    The first term in the sum is the leading colour term
     and is the dominant term in the colour expansion.
@@
     To tackle this problem, three
-    main mathods have been proposed: phase-space slicing \cite{Fabricius:1981sx,Kramer:1986mc,Giele:1991vf},
+    main methods have been proposed: phase-space slicing \cite{Fabricius:1981sx,Kramer:1986mc,Giele:1991vf},
     sector decomposition \cite{Binoth:2000ps,Binoth:2003ak}, and subtraction \cite{Ellis:1980nc}.
@@
     $\Phi_{n+1} \rightarrow \Phi_{n}\Phi_{1}$ when using
-    an appropriate $3 \rightarrow 2$ momentum mapping such that
+    an appropriate $3 \rightarrow 2$ momentum mapping \cite{Catani:1996vz,Kosower:2002su} such that
     \begin{equation}\label{eqn:integrated_counterterm}
@@
     Catani-Seymour dipoles introduced in Ref.~\cite{Catani:1996vz}
-    are process independent functions that reproduce the IR
+    are process-independent functions that reproduce the IR
     singular behaviour of matrix elements. They depend on the
@@
     the unresolved emitted parton, and $k$ is a spectator parton.
-    Since the dipoles only depend on three partons, to map out
-    all the singular limits in a process, it is necessary to construct all
-    the permutations of dipole functions.
+    In order to map out all the singular limits in a process, it is
+    necessary to construct all permutations of the dipole functions
+    since the dipoles only depends on three partons.
 
@@
     initial-final (IF) dipoles and initial-initial (II) dipoles where
-    the nomenclature is emitter-spectator dipole.
-    For electron-positron annihilations only FF dipoles are required,
-    but for any hadronic collision the inclusion of the remaining
-    dipoles are required to capture all IR singular behaviour.
+    the nomenclature refers to the emitter-spectator dipole.
+    For electron-positron annihilations only FF dipoles are required
+    since electrons (positrons) do not carry any colour charge
+    and so there is no initial-state radiation.
+    However, for any hadronic collision the inclusion of the remaining
+    dipoles are required to capture all IR-singular behaviour arising
+    from the initial-state emissions.
 
@@
     the structure of the terms and factorisation formulae
-    are essentially identical.
+    generalise analogously.
 
@@
             to conserve momentum. The grey blob represents the hard scattering process,
-            with incoming and outgoing lines representing initial and final-state
+            with incoming and outgoing lines representing initial- and final-state
             partons, respectively. The black circle represents the splitting
@@
         \begin{equation}\label{eqn:FF_factorisation}
-            {}_{n+1} \langle 1, \ldots, n+1 | | 1, \ldots, n+1 \rangle_{n+1} = \sum_{i,j}\sum_{k\neq i,j}\mathcal{D}_{ij,k}(p_{1},\ldots,p_{n+1}) + \ldots \, ,
+            {}_{n+1} \langle 1, \ldots, n+1 | 1, \ldots, n+1 \rangle_{n+1} = \sum_{i,j}\sum_{k\neq i,j}\mathcal{D}_{ij,k}(p_{1},\ldots,p_{n+1}) + \ldots \, ,
         \end{equation}
@@
         $k$ with a parton $\tilde{k}$. $\tilde{k}$ has all the
-        same quantum numbers as $k$, where as the
+        same quantum numbers as $k$, whereas the
         partonic nature of $\widetilde{ij}$ depends on
@@
         They are analogous to the $z$ from Altarelli-Parisi splitting functions.
-        $\boldsymbol{V}_{ij,k}$ acts on the spin indices of the composite particle $\widetilde{ij}$
-        , but is independent of the type of the spectator. Making the spin-dependence
+        $\boldsymbol{V}_{ij,k}$ acts on the spin indices of the composite particle $\widetilde{ij}$,
+        but is independent of the type of the spectator. Making the spin-dependence
         on parton $\widetilde{ij}$ explicit ($s$ and $s'$ for $\widetilde{ij}=\,$fermion, and $\mu$ and $\nu$ for $\widetilde{ij}=\,$gluon)
@@
         In this thesis we will be focussing on the emulation of
-        colour and spin-averaged matrix elements. Therefore the spin-indices
+        colour- and spin-averaged matrix elements. Therefore the spin-indices
         are not explicitly available. It is possible to average over the spin indices in (\ref{eqn:FF_V_ijk})
@@
             \begin{split}
-                {}_{n+1}\langle 1, \ldots, n+1; a, \ldots |&| 1, \ldots, n+1; a, \ldots \rangle_{n+1} = \\
+                {}_{n+1}\langle 1, \ldots, n+1; a, \ldots |& 1, \ldots, n+1; a, \ldots \rangle_{n+1} = \\
                 &\sum_{i,j}\sum_{k\neq i,j} \mathcal{D}_{ij,k}(p_{1},\ldots,p_{n+1};p_{a},\ldots) \\
@@
         where the new dipole contribution appears on the second line.
-        The initial-state parton is represented as $a$.
+        The initial-state parton is denoted by $a$.
         The dipole term $\mathcal{D}_{ij}^{a}$ is given by
@@
         The recoil parameter $x_{ij,a}$ and splitting variables $\tilde{z}$ are given by
-        \begin{align}\label{eqn:x_ija}
+        \begin{equation}\label{eqn:x_ija}
+        \begin{split}
                 &x_{ij,a} = \dfrac{p_{i}p_{a} + p_{j}p_{a} - p_{i}p_{j}}{(p_{i}p_{a} + p_{j}p_{a})} \, , \\
                 &\tilde{z}_{i} = \dfrac{p_{i}p_{a}}{p_{i}p_{a} + p_{j}p_{a}} \, , \quad \tilde{z}_{j} = \dfrac{p_{j}p_{a}}{p_{i}p_{a} + p_{j}p_{a}} = 1 - \tilde{z}_{i} \, .
-        \end{align}
+            \end{split}
+        \end{equation}
         The spin-averaged splitting functions are given as
@@
         \begin{equation}\label{eqn:IF_factorisation}
-            {}_{n+1} \langle 1, \ldots, n+1; a || 1, \ldots, n+1; a \rangle_{n+1} = \sum_{a,i}\sum_{k \neq i}\mathcal{D}_{k}^{ai}(p_{1},\ldots,p_{n+1};p_{a}) + \ldots \, ,
+            {}_{n+1} \langle 1, \ldots, n+1; a | 1, \ldots, n+1; a \rangle_{n+1} = \sum_{a,i}\sum_{k \neq i}\mathcal{D}_{k}^{ai}(p_{1},\ldots,p_{n+1};p_{a}) + \ldots \, ,
         \end{equation}
@@
         where $K$ and $\widetilde{K}$ are the total momenta
-        of the dipole before and after the mapping. They are given by
+        of the dipole before and after the mapping, respectively. They are given by
         \begin{align}\label{eqn:II_mapping2}
@@
             \begin{split}
-            {}_{n+1} \langle 1, &\ldots, n+1 || 1, \ldots, n+1 \rangle_{n+1} = \\
+            {}_{n+1} \langle 1, &\ldots, n+1 | 1, \ldots, n+1 \rangle_{n+1} = \\
             &\sum_{i,j}\sum_{k \neq i, j} \mathcal{D}_{ij,k}(p_{1},\ldots,p_{n+1};p_{a},p_{b}) + \sum_{i,j}\sum_{a} \mathcal{D}_{ij}^{a}(p_{1},\ldots,p_{n+1};p_{a},p_{b}) \\
@@
         the matrix element in the different soft and collinear limits will be
-        reproduced by summing over all the possible dipoles that are
-        relevant and hence will form the set of basis functions for
+        reproduced by summing over all the relevant dipoles. Meaning that
+        the dipole functions can be used as a set of basis functions for
         approximating the matrix elements in Chapters \ref{chapter:fame1}
@@
         as massless. This is especially true for the top
-        quark which has a mass of approximately 175 GeV, the
+        quark which has a mass of 172.69 GeV (cite here, pdg), the
         most massive SM particle. In the case of partonic
@@
         functions for massive partons which
-        are utilised in Chapter~\ref{chapter:fame3}, but will
-        not be written here as the philosophy of constructing
+        are used in Chapter~\ref{chapter:fame3}, but will
+        not be written here as the procedure of constructing
         the massive dipole functions is similar to massless ones.

### 2022-11-15 qcd.tex
changes for 2.1, 2.5 and typos

@@
     built up from the Feynman rules of a theory (c.f. QCD Feynman rules
-    \ref{fig:qcd_feynman_rules}). In evaluating certain
+    Figure \ref{fig:qcd_feynman_rules}). In evaluating certain
     topologies of Feynman diagrams, integral expressions
@@
     associated with high energy modes are ultraviolet
-    (UV) divergences and on the opposite end of the energy
-    spectrum, low energy modes give rise to 
-    infrared (IR) divergences.
+    (UV) divergences, and on the opposite end of the energy
+    spectrum, there are infrared (IR) divergences associated
+    with low energy modes.
 
     The de-facto methods to alleviate these
-    divergences are through regularisation
+    divergences are through renormalisation
     for UV divergences, and through subtraction
@@
     During intermediate steps of calculations,
-    such as the compututation of loop diagrams seen in
+    such as the computation of loop diagrams seen in
     Figure~\ref{fig:bubble_diagram},
@@
     \end{equation}
-    where a cutoff scale $\Lambda$ has been introduced to
+    where a cut-off scale $\Lambda$ has been introduced to
     capture the divergence as the loop momenta
@@
     hence the name ultraviolet divergence.
-    This UV divergence cannot influence the
-    experimentally observable quantities at experiments,
-    which are finite, because the final observable
-    should not depend on the intermediate steps taken
-    to get there.
-    Infinities associated with UV divergences
-    are removed via the process of renormalisation
-    which introduces counterterms in the Lagrangian
-    to exactly cancel these divergences.
+    These divergences can be systematically
+    removed by replacing the bare masses
+    and fields of the theory with their measured
+    physical values, a process named renormalisation.
+    In practice, this amounts to introducing
+    counterterms that exactly cancel the corresponding
+    UV divergences at the same order in perturbation
+    theory.
 
@@
     out in $d=4$ space-time dimensions, would be finite
-    if $d < 4$.
+    if we move away from $d=4$ dimensions
+    \footnote{In dimensional regularisation it is possible to regulate
+    the integral regardless of if $d > 4$ or $d < 4$, it would amount
+    to a sign difference.}.
 
@@
     The inclusion of counterterms into the theory can be
-    thought of as renormalisations of the bare fields
+    thought of as redefinitions of the bare fields
     and couplings in the QCD Lagrangian (\ref{eqn:L_QCD}) to restore predictive
@@
     models have to be used to describe these processes.
+    This will be briefly touched on in Section~\ref{sec:anatomy_event}.
 
@@
     where $Q = (p_{1} + p_{2})$. There is a double $\epsilon$ pole manifest,
-    corresponding to the soft and collinear divergence.
+    corresponding to the associated soft and collinear divergences in real emission corrections.
 
@@
     in the hard scattering.
-    When either of the partons beecome soft, $E_{i,j}\rightarrow 0$,
+    When either of the partons become soft, $E_{i,j}\rightarrow 0$,
     or when they go collinear to each other, $\theta_{ij} \rightarrow 0$,
     the matrix element will diverge if $s_{ij}$ appears
-    in the demoninator. These scenarios correspond to an external
+    in the denominator. These scenarios correspond to an external
     particle becoming unresolved.
@@
     The indices $i$, $j$, and $k$ label partons involved
-    in the factorsation process: $i$ is the emitter parton,
+    in the factorisation process: $i$ is the emitter parton,
     $j$ is the emitted parton, and $k$ is a parton accounting
@@
     the transverse direction perpendicular to the
-    collinear direciton ($p \cdot k_{\perp} = 0$), and $n^{\mu}$ is an auxiliary
+    collinear direction ($p \cdot k_{\perp} = 0$), and $n^{\mu}$ is an auxiliary
     vector satisfying the conditions $n^{2} = 0$ and $k_{\perp} \cdot n = 0$.
@@
     partial amplitude $S_{\mu}(q_{1};1,\ldots,n;\bar{q}_{2})$ depending
-    only on kinematical variables. In this partial amplitude, the gluons
+    only on kinematic variables. In this partial amplitude, the gluons
     are emitted in an ordered fashion from the quarks, meaning the quarks
@@
     \end{equation}
-    where the subleading colour terms proportional to $1/N_{c}^{2}$ have been omitted.
-    The first term in the sum is the leading colour term
+    where the sub-leading colour terms proportional to $1/N_{c}^{2}$ have been omitted.
+    The left-most term in the sum is the leading colour term
     and is the dominant term in the colour expansion.
@@
     In both these formulae the structure is of the form:
-    tree-level splitting function multipled by a one-loop
+    tree-level splitting function multiplied by a one-loop
     amplitude, plus a one-loop splitting function multiplied
@@
     that is universally used, the general form of a subtraction term
-    must fulfill the requirements of replicating the matrix element
+    must fulfil the requirements of replicating the matrix element
     behaviour in all IR singular limits, and be analytically integrable
@@
     At the next order in perturbation theory, we have the NLO
-    cross-section which recieves contributions from the real and virtual
+    cross-section which receives contributions from the real and virtual
     corrections
@@
     The counterterm $\mathcal{C}_{n+1}$ has been kept general
-    but specific exmples of automated subtractions schemes include
+    but specific examples of automated subtractions schemes include
     Catani-Seymour (CS) \cite{Catani:1996vz,Catani:2002hc},
@@
         behaviour seen in (\ref{eqn:soft_factorisation}) and (\ref{eqn:F_collinear_factorisation}),
-        repsectively. In particular, $\boldsymbol{V}_{ij,k}$ becomes
+        respectively. In particular, $\boldsymbol{V}_{ij,k}$ becomes
         proportional to the eikonal factor and Altarelli-Parisi splitting
@@
         as massless. This is especially true for the top
-        quark which has a mass of 172.69 GeV (cite here, pdg), the
+        quark which has a mass of 172.69 GeV \cite{Workman:2022ynf}, the
         most massive SM particle. In the case of partonic
@@
     Another set of functions that can be used to construct
-    subtractions terms in NLO \cite{Kosower:1997zr,Campbell:1998nn}
+    subtraction terms in NLO \cite{Kosower:1997zr,Campbell:1998nn}
     and NNLO QCD \cite{Gehrmann-DeRidder:2004ttg,Gehrmann-DeRidder:2005btv}
@@
     in Section~\ref{sec:CO_factorisation}. In the factorisation
-    formulae, the colour-ordered matrix elements factorised
+    formulae, the colour-ordered matrix elements factorises
     exactly into a colour-ordered matrix element with one parton
@@
     situation of a massive colour-neutral boson decaying
-    to massless QCD partons, suitable for multijet production
+    to massless QCD partons, suitable for multi-jet production
     in electron-positron annihilation. Additionally, we
@@
     \end{itemize}
-    Along with the tree-level antenna, there are also the
+    Along with the tree-level antennae, there are also the
     one-loop antennae. For the three-parton case these are given by
@@
     the term proportional to $X_{3}^{0}$ has to be removed
-    to fulfill this requirement.
+    to fulfil this requirement.
     The antenna functions are summarised in Table~\ref{table:antenna_functions},
@@
         one-loop level correspond to the leading colour ($X_{3}^{1}$),
-        subleading colour ($\tilde{X}_{3}^{1}$), and closed quark loop
+        sub-leading colour ($\tilde{X}_{3}^{1}$), and closed quark loop
         ($\hat{X}_{3}^{1}$) contributions.}
@@
     Below we list the antenna functions relevant
-    for multijet production in electron-positron
-    annihiliation with a single quark flavour,
+    for multi-jet production in electron-positron
+    annihilation with a single quark flavour,
     namely the partonic channel $e^{-}e^{+} \rightarrow
@@
     different functions corresponding to the: leading colour,
-    subleading colour, and closed quark look contributions.
+    sub-leading colour, and closed quark look contributions.
     It is sufficient to consider the leading colour
     structure in this thesis as the remaining contributions
-    are subleading. Ignoring subleading terms in
+    are sub-leading. Ignoring sub-leading terms in
     colour is equivalent to taking the large $N_{c}$
     limit. For QCD where $N_{c} = 3$, neglecting
-    subleading colour terms amounts to an
+    sub-leading colour terms amounts to an
     approximately 10\% effect.
@@
     function. The first two terms correspond to the gluon
-    going collinear with the quark and antiquark. The final
+    going collinear with the quark and antiquark, respectively. The final
     term encapsulates singularity in the limit of the gluon

### 2022-11-16 ml.tex
chapter 4 changes

@@
 \section{Rise of machine learning in HEP}
-    In high energy physics where analyses
+    In high energy physics, where analyses
     depend on making use of large computing clusters
@@
     required to keep up with the needs of the
-    experimental and theoretical community.
+    experimental and theoretical communities.
     However, the computational resources needed
@@
     areas \cite{Bothmann:2022thx}, especially with
-    HL-LHC over the horizon. Therefore new techniques and
-    algorithms from the machine learning community
+    HL-LHC projected to begin operating at the end
+    of this decade \cite{ZurbanoFernandez:2020cco}.
+    Therefore new techniques and
+    algorithms from machine learning 
     have been adopted to augment the ongoing efforts
-    in the high energy physics community. Machine
+    in high energy physics. Machine
     learning is a category of artificial intelligence
@@
     automates learning from data.
-    By now machine learning is utilised throughout
-    every stage of particle physics analysis, from collection
+    By now, machine learning is ubiquitous in
+    particle physics analysis, from collection
     of data from experiments all the way to novel
@@
     Around the turn of the last decade the culmination
-    in advancement in hardware and techniques led to an
-    explosion in image classification research in the
+    of advancement in hardware and machine learning training algorithms
+    led to an explosion in research in the
     machine learning community, specifically in the
@@
     previous state-of-the-art \cite{NIPS2012_c399862d,russakovsky2015imagenet} and
-    lead to the widespread use of neural networks for a
-    wide array of tasks \cite{Schmidhuber_2015}.
+    lead to the widespread use of neural networks across
+    a range of tasks \cite{Schmidhuber_2015}.
     This led to the rise of programming
@@
     Collisions at the LHC can produce hundreds of
-    particles with complex final-state configurations
-    leading to the design of the ATLAS and CMS experiments
+    particles with complex final-state configurations.
+    This leads to the design of the ATLAS and CMS experiments
     to contain of the order 100 million detection elements
-    in an attempt to disentangle these events. With these
-    large arrays of detection elements there is an enormous
-    amount of data harvested from each collision, terabytes
-    per second, which presents a challenge in the form of
-    data collection.
-    Already at this stage, boosted decision trees
-    have been used to enhance triggers \cite{CMS:2020cmk} to accept or
-    reject data from being saved to disk. At the LHCb
-    experiment 70\% of all data retained have had
-    some involvement with machine learning algorithms \cite{LHCb:2014set}.
+    in an attempt to disentangle these events. With these large
+    arrays, terabytes of data are harvested per second from
+    each collision. This presents a practical challenge for data
+    collection where machine learning has could provide a solution.
+    Indeed, boosted decision trees have been use
+    to enhance triggers \cite{CMS:2020cmk} to accept or
+    reject data before being saved to disk.
+    % At the LHCb
+    % experiment, 70\% of all data retained have had
+    % some contact with machine learning algorithms \cite{LHCb:2014set}.
 
-    On the theory side of things, machine learning
-    has been applied to many stages of the event
+    In view of the upgrades at the LHC and the upcoming
+    HL-LHC providing unprecedented integrated luminosity,
+    it is important that the generation of simulated samples
+    speeds up. This prompted the explosion of research
+    in machine learning applications on the theory side.
+    Studies have been carried out on many stages of the event
     generation process.
-    To name a few examples of active research areas:
-    phase-space sampling, matrix element modelling, hadronisation modelling,
-    and anomaly searches for new physics.
+    To name just a few examples of active research areas:
+    phase-space sampling \cite{Bendavid:2017zhk,Klimek:2018mza,Bothmann:2020ywa,Verheyen:2020bjw,Gao:2020vdv},
+    matrix element modelling \cite{Bishara:2019iwh,Badger:2020uow,Aylett-Bullock:2021hmo,Maitre:2021uaa,Badger:2022hwf},
+    hadronisation modelling \cite{Biro:2021zgm,Ilten:2022jfm,Ghosh:2022zdz},
+    and end-to-end event generation \cite{Gao:2020zvv,DiSipio:2019imz,Butter:2019cae,Butter:2021csz,ArjonaMartinez:2019ahl}.
+    For a more complete overview of active research,
+    a living review that is archiving advancements in the
+    field is available at \cite{Feickert:2021ajf}.
+    For more traditional reviews, see for instance
+    \cite{Guest:2018yhq,Radovic:2018dip,Butter:2022rso}.
+
     The research carried out in these areas have shown
@@
     to existing event generators.
-    A living review that is actively being updated
-    archiving advancements in the field is available
-    at \cite{Feickert:2021ajf}.
-    For more traditional reviews, see for instance
-    \cite{Guest:2018yhq,Radovic:2018dip,Butter:2022rso}.
+    There is scepticism regarding the use of machine learning
+    algorithms in production due to their lack of interpretability \cite{Grojean:2022mef},
+    as well as their use as black-boxes which precludes predictability \cite{Schwartz:2022njo}.
+    Along similar lines, the quantification of the uncertainties
+    associated with these algorithms are not well defined \cite{Chen:2022pzc}.
 
@@
     That is to say they are an interconnected network
-    of neurons used for analysing data. Here we will
+    of neurons used for analysing data. Here, we will
     introduce the densely-connected neural network
-    which will be the main workhorse algortihm used
+    which will be the main workhorse algorithm used
     in this thesis, and frame the discussion around a
@@
     from a target function $g(\mathbf{x})$ to create a surrogate
-    model with the neural network algorithm.
+    model with a neural network algorithm.
+
+    In a machine learning context, the fitting of a
+    model to training data is termed training the model.
+    Once the model is fitted, it is customary to use
+    an unbiased dataset that is distinct from the training set
+    to evaluate the final model performance. This dataset is
+    the test set.
 
 \subsection{Model of a neuron}
-    In this context a neuron is modelled by
+    In this context, a neuron is described by
     \begin{equation}\label{eqn:neuron}
-        y = \phi(z) = \phi(\mathbf{w}^{T}\mathbf{x} + b)
+        y = \phi(\mathbf{w}^{T}\mathbf{x} + b)
     \end{equation}
-    where the model inputs $\mathbf{x}$ are multiplied
-    by the model weights $\mathbf{w}$ before being summed
-    with a bias term $b$. Collectively the model weights
-    and biases are the model parameters $\theta$.
+    where the model inputs, $\mathbf{x}$, are multiplied
+    by the model weights, $\mathbf{w}$, before being summed
+    with a bias term, $b$. Collectively, the model weights
+    and biases are the model parameters, $\theta$.
     This combination of terms is then modified
-    by an activation function $\phi$ which is carefully
+    by an activation function, $\phi$, which is carefully
     chosen to perform a non-linear transformation. A neuron
@@
     Some of the more well-known activation functions are
-    the sigmoid function, hyperbolic tangent, rectified linear unit (ReLU).
+    the sigmoid function, hyperbolic tangent, rectified linear unit (ReLU) \cite{Nair2010RectifiedLU}.
     A less familiar class of activation functions are the
-    sigmoid linear units (SiLU) of which swish is an example,
-    which have been shown to outperform ReLU is some cases
-    \cite{DBLP:journals/corr/abs-1710-05941}. The functional forms of these
+    sigmoid linear units (SiLU) \cite{Hendrycks2016BridgingNA},
+    of which swish \cite{DBLP:journals/corr/abs-1710-05941} is an example.
+    The functional forms of these
     functions are summarised in Table~\ref{table:activation_functions}.
@@
     or densely-connected neural network. The constituents
-    of this network structure are the input layer where
-    model inputs enter, hidden layers where the bulk of
-    model parameters live, and the
+    of this network structure are: the input layer where
+    model inputs enter; hidden layers, where the bulk of
+    model parameters live; and the
     output layer which outputs the model prediction(s).
-    The training data only provides concrete desired
-    outputs for the overall model, and not for the layers
-    before the output layer. This is why mediating layers
-    between inputs and outputs are so-called hidden.
+    The training data only provides concrete, desired
+    outputs for the overall model, and does not specify
+    anything about the layers preceding, hence the term hidden.
     The neural network is free to change the parameters
-    in these layers to best approximate the target function
+    in these layers to best approximate the target function,
     $g(\mathbf{x})$.
@@
     The universal approximation theorem states that neural networks are
-    able to represent any continuous function $g(\mathbf{x})$
+    able to represent any continuous function, $g(\mathbf{x})$,
     with one hidden layer and a finite number of neurons \cite{HORNIK1991251}.
-    However it is very difficult to achieve this due to
+    However, it is very difficult to achieve this due to
     practical constraints such as limited data and network size.
@@
     are samples from the target distribution, and model predictions
-    for inputs $\mathbf{x}_{i}$ belonging to a data set with $N$
+    for inputs, $\mathbf{x}_{i}$, belonging to a data set with $N$
     samples. The loss function
-    encodes the discrepency between the truth value
-    and the model prediction, therefore, the task of finding
+    encodes the discrepancy between the truth value
+    and the model prediction. Therefore, the task of finding
     an acceptable set of parameters can be reframed as an optimisation
@@
     \begin{equation}\label{eqn:gradient_descent}
-        \theta = \theta - \eta \cdot \nabla_{\theta} L(\mathbf{x}_{\mathrm{batch}}; \theta) \, ,
+        \theta \gets \theta - \eta \cdot \nabla_{\theta} L(\mathbf{x}_{\mathrm{batch}}; \theta) \, ,
     \end{equation}
@@
     hyperparameters is often computationally expensive for two
-    reason. Firstly the dimensionality of parameter space to choose from is high,
+    reasons. Firstly, the dimensionality of parameter space to choose from is large,
     and so the curse of dimensionality kicks in rapidly.
-    Secondly there is no exact a priori way to evaluate the
+    Secondly, there is no exact a priori way to evaluate the
     final performance of the neural network given a set of
-    hyperparameters meaning training has to take place for
+    hyperparameters, meaning training has to take place for
     some steps before the final model performance can be estimated.
@@
     an exhaustive scan of parameters systematically sampled
-    in a grid-like fashion, where as random search replaces
+    in a grid-like fashion, whereas random search replaces
     this systematic sampling with random sampling. Whilst
@@
     Below we discuss two paradigms of hyperparameter optimisation
-    that aim to reduce number of trials before landing
-    on a set of decent hyperparameters.
+    that aim to reduce number of trials when doing hyperparameter searches.
 
@@
     to carry out hyperparameter optimisation \cite{bergstra2013making,bergstra2015hyperopt}.
-    The method  begins with the construction of a probabilistic surrogate
-    model for the performance of the neural network. This
-    surrogate model collects evidence through successive trainings
-    of the neural network. By selecting promising hyperparameter
-    sets based on the updated model, the exploration of
+    The method begins with the construction of a 
+    Bayesian probabilistic model. This model is a mapping
+    from hyperparameter sets to the performance
+    of a neural network, as evaluated by a loss function. 
+    This Bayesian model collects evidence
+    through successive trainings of the neural network,
+    and so iteratively becomes more informed on the hyperparameter
+    space. By selecting promising hyperparameter
+    sets based on the successively updated model, the exploration of
     hyperparameter space is much more efficient.
@@
     that the first few epochs of training are indicative
-    of the final model performance. This in combination
+    of the final model performance \cite{pmlr-v28-bubeck13}. This in combination
     with a fixed time budget gives rise to multi-armed
@@
     best performing models remain. Recently, there have
-    been proposals of combining Bayesian optimsation and
+    been proposals of combining Bayesian optimisation and
     the bandit-based approaches \cite{falkner2018bohb}.
@@
     one has to balance the additional time spent tuning with the
-    relatively incremental improvements in performance. For
-    novel machine learning experiments in high energy physics,
-    the hyperparameter tuning is not of paramount importance,
-    and generally more focus is spent on the model building itself.
+    relatively incremental improvements in performance. For this reason,
+    the hyperparameter tuning in novel machine learning experiments
+    in high energy physics is not of paramount importance.
+    Instead more focus is spent on the model building itself, as
+    more sophisticated models represent a more substantial increase
+    in performance.
 
@@
     a bottleneck in event generation, and so building a
-    fast surrogate model would accelerate this process;
-    once a fast and accurate surrogate model has been
-    built it can supplement the traditional matrix
+    fast surrogate model would accelerate this process.
+    Once a fast and accurate surrogate model has been
+    built, it can supplement the traditional matrix
     element generator to increase statistical power of the sample
-    by predicting on many more phase-space points. This point
+    by predicting on many more phase-space points. This
     requires the accumulated error in the surrogate model
@@
     integration of matrix elements, which is a key motivation
-    for building an accurate emulator.
+    for building an accurate surrogate model.
 
@@
     The complication in accurately emulating matrix elements
-    on a per-point basis arises due to the infrared divergent behaviour
-    as discussed in Section~\ref{sec:IR_divergences} which occurs
-    in corners of the phase-space.
+    on a per-point basis arises due to the infrared divergent behaviour,
+    as discussed in Section~\ref{sec:IR_divergences}, which occurs
+    in corners of the phase-space
     (UV divergences have been removed via renormalisation).
@@
     non-divergent regions simultaneously with the divergent
-    regions due to the disparity in scales. This problem
+    regions, due to the disparity in scales. This problem
     becomes more apparent for higher multiplicity final-states
@@
     into divergent and non-divergent regions, then subsequently
-    partition the divergent regions further according to partition functions
+    weight the divergent regions further according to partition functions
     based on FKS subtraction.
-    In each partition the authors trained a separate neural network,
+    In each weighted partition, the authors trained a separate neural network,
     meaning that their model was an ensemble of neural networks.
@@
     even for the highest multiplicity. Extending this study
-    to the one-loop level is done in Chapter~\ref{chapter:fame2}
+    to the one-loop level is done in Chapter~\ref{chapter:fame2},
     where antenna functions replace Catani-Seymour dipoles as
@@
     theoretical uncertainty. An alternative to quantifying
-    the uncertainty of the neural network prediction, is to
+    the uncertainty of the neural network prediction is to
     bypass the usage of the neural network prediction itself
@@
     model to study the potential gain factors with a more
-    accuracte model. As previously mentioned, a consequence of
+    accurate model. As previously mentioned, a consequence of
     this two-stage unweighting

### 2022-11-17 intro.tex
changes to chapter 1

@@
 The Large Hadron Collider (LHC) at CERN has been colliding
-particles since 2009 with each successive run increasing the centre-of-mass
-energy of the collision (7-13.6 TeV). With the experiment operating
-over the course of decades, there is a vast quantity of
+particles since 2009, with each successive run increasing the centre-of-mass
+energy of the collision (7-13.6 TeV). As the experiment has
+operated for over a decade, there is a vast quantity of
 experimental data collected that has to be compared with
@@
 to the detection of the hundreds of particles produced in the collision.
-General purpose Monte Carlo event generators are the de-facto
+General purpose Monte Carlo event generators are the de facto
 tool designed to generate these simulated events.
 With the large number of scattering events collected at
-the LHC already, and with even more expected at the High Luminosity LHC,
+the LHC already, and with even more expected at the High Luminosity LHC (HL-LHC),
 the rate of event generation is at risk of
@@
 the subsequent cascading of particles (parton shower),
-and the formation of bound states detected (hadronisation).
-While hadronisation is based on phenomological models,
+and the formation of bound states detected (hadronisation),
+which makes their way to the detectors.
+While hadronisation is based on phenomenological models,
 the hard scattering and parton shower are derived
@@
 These matrix elements are formally calculated order-by-order
-in perturbation theory with each order leading
-to increasingly difficult computations. In practice,
+in perturbation theory, with each order leading
+to increasingly difficult computations at each order. In practice,
 these matrix elements are provided in libraries interfaced
@@
 when generating the large samples required to compare with
-experiments, especially when a significant portion of time
-spent simulating an event is spent on the matrix element
-evaluation.
+experiments. Especially when a significant portion of time
+spent simulating an event is spent on evaluating matrix
+elements.
 
@@
 machine learning methods to accelerate the evaluation
-of matrix elements, hence speeding up the event
-generation process in general. Machine learning algorithms
-have become ubiquitous in many fields and their applicability
+of matrix elements in order to speed up the event
+generation process. Machine learning algorithms
+have become ubiquitous in many fields, and their applicability
 to a wide range of problems have made them a popular choice
@@
 which governs the behaviour of quarks and gluons. Furthermore,
-the relationship between theory and experiment will be elaborated on
-with the introduction of cross-sections and their relation
+I will elaborate on the relationship between theory and experiment
+with the introduction of cross-sections, and their relation
 to matrix elements.
-
 In Chapter~\ref{chapter:qcd}, I will discuss the challenges that arise
 in fixed-order perturbative calculations and the current methods
-that have been adopted in the community to deal with them. Some of
-these methods also form the theoretical foundation which we
-embed into the machine learning algorithm. This procedure is 
-described in detail in Chapter~\ref{chapter:fame1} for electron-positron
-annihilation for tree-level matrix element emulation.
+that have been adopted in the community to deal with them.
+I will briefly introduce Monte Carlo event generators in
+Chapter~\ref{chapter:MC}, where the theory discussed in the first
+two chapters is applied in practice. In Chapter~\ref{chapter:ml},
+I discuss how bottlenecks in traditional event generator techniques
+motivates the usage of machine learning based approaches, with
+particular attention placed on using neural networks as emulators
+for matrix elements.
+This construction of this neural network emulator is
+described in detail in Chapter~\ref{chapter:fame1} for tree-level
+electron-positron annihilation matrix element emulation.
 A similar philosophy is applied in Chapter~\ref{chapter:fame2}
-where next-to-leading order QCD k-factors are emulated for
+where next-to-leading order QCD K-factors are emulated for
 the same processes.
@@
 to this scenario is detailed in Chapter~\ref{chapter:fame3}
-where we also explore using the
-emulator in a novel implementation to replace traditional
-matrix element providers in the event generator SHERPA.
+where I also explore using the
+emulator in a novel implementation to supplement traditional
+matrix element providers in the event generator {\Sherpa}.
 Finally, I will summarise and conclude the thesis in
@@
 \section{The Standard Model of particle physics}
-    The Standard Model of particle physics (SM) is our current
-    best working theory to describe all known elementary particles
+    The Standard Model of particle physics is our current
+    best working theory to describe all known elementary particles,
     as well as three of the four fundamental forces. Developed
     predominantly in the latter half of the 20th century, it is
-    one of the most well tested theories we have in science today.
+    one of the most well tested theories that we have in science today.
     Some highlights include the highly precise predictions of the
-    anomalous magnetic moment of the electron, agreeing with experimental
+    anomalous magnetic moment of the electron, which agrees with experimental
     measurements to more than 10 significant figures \cite{Aoyama:2017uqe},
     and the discovery of the Higgs boson in 2012 by the ATLAS \cite{ATLAS:2012yve}
-    and CMS experiments \cite{CMS:2012qbp} at the LHC
+    and CMS experiments \cite{CMS:2012qbp} at the LHC,
     which was theorised decades prior.
@@
     The SM is a gauge quantum field theory (QFT) where particles
-    are described as excitations in quantum fields. It can be specified
-    by the gauge group
+    are described as excitations of quantum fields. The symmetry group
+    of the SM is
     \begin{equation}\label{eqn:SM_gauge}
@@
     group with colour charge $c$ describes the interactions of the
-    strong force within the theory of quantum chromodynamics (QCD), which we
+    strong force within the theory of quantum chromodynamics, which we
     will elaborate more on in Section~\ref{sec:qcd}. The second
-    and third gauge groups represent the electroweak sector of the SM with
-    charges $L$ for left and $Y$ for hypercharge. Under electroweak
+    and third gauge groups represent the electroweak sector of the SM, where
+    $L$ denotes left-chiral fields that carry weak isospin, whilst $Y$ denotes
+    the weak hypercharge. Under electroweak
     spontaneous symmetry breaking (EWSB), this product becomes
@@
     \end{equation}
-    giving rise to the electromagnetic and weak forces we observe.
+    giving rise to the electromagnetic and weak forces that we observe.
     The effect of gravity is considered to be negligible on the scales
-    considered in the SM and so it is not described.
+    considered in the SM, and so it is not described.
     Each of the three fundamental forces described by the SM is mediated
-    by the exchange of a gauge boson. The gauge bosons mediating the strong
-    and EM force are the massless gluon $g$ and massless photon $\gamma$,
+    by the exchange of a gauge boson. The massless gauge bosons mediating the strong
+    and EM force are the gluon $g$ and photon $\gamma$,
     respectively. For the weak force the gauge bosons are the $W^{\pm}$
-    and $Z^{0}$ bosons which attain a mass through the Higgs mechanism \cite{Englert:1964et,Higgs:1964pj,Guralnik:1964eu}
-    during EWSB, which elucidates the Higgs boson $H$.
+    and $Z^{0}$ bosons, which attain a mass through the Higgs mechanism \cite{Englert:1964et,Higgs:1964pj,Guralnik:1964eu}
+    during EWSB, elucidating the Higgs boson $H$.
 
-    The matter content of the SM consists of fermions which can be split
+    The matter content of the SM consists of fermions, which can be split
     into quarks and leptons. Quarks are massive and experience the strong,
@@
     neutrino $\nu_{\tau}$ and their antiparticles) which only experience the
-    weak force. Neutrinos are massless in the SM but this is not a requirement
-    of the model. In fact they have been observed to have mass \cite{Super-Kamiokande:1998kpq,SNO:2002tuh}.
+    weak force. Neutrinos are massless in the SM, however, they have been
+    observed to have mass \cite{Super-Kamiokande:1998kpq,SNO:2002tuh}. This
+    prompts physics beyond the Standard Model (BSM) to describe these observed masses.
 
-    The particle content of the SM is summarised in Figure \ref{fig:SM_particles}
-    showing the quarks, leptons and bosons along with their masses, charges
+    The particle content of the SM is summarised in Figure \ref{fig:SM_particles},
+    which shows the quarks, leptons and bosons along with their masses, charges
     and spin.
@@
         \includegraphics[width=\linewidth]{introduction/sm_particles.pdf}
-        \caption{Particle content of the SM split into quarks, leptons,
+        \caption{Particle content of the SM, split into quarks, leptons,
         gauge bosons and scalar bosons. The columns for the fermions depict
-        the three different generations. Masses, electric charge, and spin
+        the three different generations. The masses, electric charge, and spin
         are given for each particle. The yellow contours indicate the coupling
         of bosons to fermions, illustrating the forces experienced by the
-        fermions inside the contour. Figure reference \cite{SM_figure}.}
+        fermions inside the contour. Figure reproduced from \cite{SM_figure}.}
         \label{fig:SM_particles}
     \end{figure}
-    The interactions of these fields are governed by the SM Lagrangian
-    \footnote{technically Lagrangian density but we use the terms
-    Lagrangian and Lagrangian density interchangeably}
+    The interactions of these fields are governed by the SM Lagrangian\footnote{Technically Lagrangian density but we use the terms
+    Lagrangian and Lagrangian density interchangeably.}
     which can be written as
@@
     All terms in the SM Lagrangian are invariant under local transformations
-    of the gauge group (\ref{eqn:SM_gauge}).
+    of the gauge group Eq. (\ref{eqn:SM_gauge}).
 
     In the following section we will focus on the gauge, fermion, gauge-fixing
-    and ghost Lagrangian terms in the framework of QCD as that will be the
-    most relevant sector for this thesis. The remaining terms are discussed at
+    and ghost Lagrangian terms in the framework of QCD, which is the most
+    relevant sector for this thesis. The remaining terms are discussed at
     length in standard reference texts \cite{Peskin:1995ev,Schwartz:2014sze,Romao:2012pq}.
@@
     \end{equation}
-    where twice repeated indices are summed over.
-    The fields $\psi_{i}^{f}$ ($\bar{\psi}_{i}^{f}$) are the fermions (antifermions),
-    representing quarks (antiquarks) with flavours $f$: 
+    where repeated indices are summed over.
+    The fields $\psi_{i}^{f}$ are the fermions field operators,
+    representing quarks and antiquarks with flavours $f$: 
     up $u$, down $d$, strange $s$, charm $c$, top $t$, and bottom $b$,
-    with masses $m_{f}$. They transform under the fundamental
+    with masses $m_{f}$. The field operators transform under the fundamental
     representation with indices $i, j \in \{1, 2, 3\}$, named colour indices.
@@
     Gauge fields transform under the adjoint representation, 
-    which is an $N_{c}^{2}-1$ dimensional representation, meaning
+    which is an $N_{c}^{2}-1$ dimensional representation such that
     the adjoint indices $a,b,c \in \{1,...,8\}$. $T^{a}_{ij}$ are the
-    group generators in a fundamental representation. In SU(3) it is
+    group generators in the fundamental representation. In SU(3) it is
     common to write the group generators as
@@
     where $\lambda^{a}_{ij}$ are the Gell-Mann matrices \cite{Gell-Mann:1962yej}.
-    The generators of the group have to obey the Lie algebra
+    The generators of the group obey the Lie algebra
     \begin{equation}\label{eqn:lie_algebra}
@@
     where $f^{abc}$ are the structure constants of SU(3).
-    The generators are normalised by convention to be
+    By convention, the generators are normalised to be
     \begin{equation}\label{eqn:generator_normalisation}
-        \mathrm{Tr}(T^{a}T^{b}) = \delta^{ab}T_{R} \, , \quad T_{R} = \dfrac{1}{2} \, .
+        \mathrm{Tr}(T^{a}T^{b}) = \delta^{ab}T_{R} \, , \quad \mathrm{where} \quad T_{R} = \dfrac{1}{2} \, .
     \end{equation}
@@
     where $C_{F}=4/3$ and $C_{A}=3$ for QCD.
-    They are collectively referred to as colour factors.
+    These are collectively referred to as colour factors.
 
@@
     \end{equation}
-    The strong coupling constant, is in fact not constant, but
-    is dependent on the energy scale of the process. This
-    is due to the process of renormalisation which will be discussed
+    The strong coupling constant is in fact not constant, and
+    depends on the energy scale of the process. This
+    is due to the process of renormalisation, which will be discussed
     in Section~\ref{sec:renormalisation}.
@@
     in the form of perturbative expansions in $\alpha_{s}$ (see Section~\ref{sec:xs}).
+    This is normally precluded as $\alpha_{s} > 1$ at low
+    energies.
 
     To remove unphysical degrees of freedom from the
-    theory we need to gauge-fix the theory
+    theory we need to fix the gauge
     and add a ghost Lagrangian. The gauge-fixing term
-    in the $R_{\xi}$ gauges is written as
+    in the $R_{\xi}$ gauge is written as
     \begin{equation}\label{eqn:L_GF}
@@
     \end{equation}
-    where $c^{a}$ ($\bar{c}^{a}$) are Faddeev-Popov ghosts (anti-ghosts).
-    They are unphysical external states but must be
-    included in internal lines for a Lorentz invariant
-    perturbative gauge theory.
+    where $c^{a}$ ($\bar{c}^{a}$) are Faddeev-Popov ghosts (antighosts).
+    The ghosts are unphysical external states but must be
+    included in internal lines for a gauge invariant theory.
 
@@
     Now that we have written down the QCD Lagrangian,
-    we can examine the terms to write down how the
-    gauge bosons interact with the fermions, and especially
+    we can examine the terms that govern the interactions between
+    bosons and fermions, and especially
     important for QCD, the gauge boson self interactions. A convenient
@@
     that Feynman diagrams are a simple but powerful tool
-    to systematically build up all the possible ways a process
+    to systematically build up all the possible ways in which a process
     can occur.
 
-    Expanding (\ref{eqn:L_QCD}) and extracting the terms
+    Expanding Eq. (\ref{eqn:L_QCD}) and extracting the terms
     that mix the gauge and fermions fields, we get an interaction Lagrangian
@@
     which we can interpret as follows: the first term
-    is an interaction between a gluon, a quark, and an
-    anti-quark, the second term is a triple gluon
+    is an interaction between a gluon and two fermion field operators,
+    the second term is a three gluon
     self-interaction, and the third term is a four gluon
@@
     we can read off from the kinetic Lagrangian where
-    we collect terms from (\ref{eqn:L_QCD}) and (\ref{eqn:L_GF}), 
+    we collect terms from Eqns. (\ref{eqn:L_QCD}) and (\ref{eqn:L_GF}), 
     \begin{equation}\label{L_kin}
@@
         \includegraphics[width=\linewidth]{introduction/qcd_feynman_rules.pdf}
-        \caption{QCD Feynman rules with the Minkowski metric signature $g^{\mu\nu} = \mathrm{diag}(1, -1, -1, -1)$.}
+        \caption{QCD Feynman rules with the Minkowski metric $g^{\mu\nu} = \mathrm{diag}(1, -1, -1, -1)$.}
         \label{fig:qcd_feynman_rules}
@@
     Note that we have only written down the vertices and propagators
-    for the physical gluons and quarks,
-    ghosts also have associated Feynman rules but we do
-    not collect those here. A full list of Feynman rules
-    in the SM can be found at Ref. \cite{Romao:2012pq}.
+    for the physical gluons and quarks.
+    Ghosts also have associated Feynman rules but we do
+    not specify those here. A full list of Feynman rules
+    in the SM can be found in Ref. \cite{Romao:2012pq}.
 
     From these Feynman rules, we identify that each
-    quark-antiquark-gluon and triple gluon vertices
+    fermion-gluon and three gluon vertices
     are associated with one power of $\alpha_{s}$ (due to squaring $g_{s}$),
@@
     us to systematically build up Feynman diagrams
-    that correspond to a fixed order in $\alpha_{s}$.
+    that correspond to a fixed-order in $\alpha_{s}$.
     This point will be elaborated on in Section
@@
     named partons.
-    The application of QCD to describe phenomenon
+    The application of QCD to describe phenomena
     in these proton-proton collisions rests on the use of
-    the factorisation theorem which enables the
+    the factorisation theorem, which enables the
     separation of low and high energy scales (so-called soft and hard).
@@
     from another proton. This is the scenario
-    that occurs at collisions at the LHC.
+    that occurs in collisions at the LHC.
     During the collision, or the hard scattering,
@@
     This heuristic argument motivates the form of
-    the factorisation equation, where the production of
-    particles from a hadronic collision, the hadronic
-    cross-section, can be written as
+    the factorisation equation, where the cross-section,
+    which is proportional to the production rate of
+    particles from a hadronic collision, can be written as
     \begin{equation}\label{eqn:hadronic_cs}
@@
     This factorisation is not exact as indicated by the
-    correction term inversely proportional to characteristic
-    hard scale $Q$. However, for high energy collisions, where
+    correction term, which is inversely proportional to the characteristic
+    hard scale $Q$. The other scale involved, $\Lambda_{\mathrm{QCD}}$, the QCD scale,
+    is the scale at which $\alpha_{s}$ becomes
+    large enough that perturbation theory breaks down.
+    However, for high energy collisions, where
     $Q \gg \Lambda_{\mathrm{QCD}}$, this term and higher order
@@
 
-    $f_{a/h}(x, \mu)$ are parton distribution functions
-    (PDFs) which depend on the momenta fractions $x$ parton $a$
-    has with respect to its parent hadron $h$, at a scale $\mu$,
-    usually taken to be the factorisation scale $\mu_{F}$,
-    the interface between soft and hard physics.
+    The parton distribution functions (PDFs), $f_{a/h}(x, \mu)$,
+    depend on the momenta fraction $x$ carried by parton $a$,
+    with respect to its parent hadron $h$, at a scale $\mu$,
+    usually taken to be the factorisation scale $\mu_{F}$.
+    The factorisation scale is the interface between soft and hard physics.
     The interpretation of PDFs at leading order are
-    as probability distributions. $f_{a/h}(x, \mu)$ is
+    as probability distributions: $f_{a/h}(x, \mu)$ is
     the probability of finding parton $a$ within $h$
@@
     PDFs are non-perturbative objects that encapsulate
-    the soft effects of the scattering which occur below energy
+    the soft effects of the scattering occurring below energy
     $\mu_{F}$. They are non-perturbative because
     they are determined by fitting to experimental data,
-    instead of calculated in perturbation theory.
+    instead of being calculated in perturbation theory.
 
@@
     through the use of the Dokshitzer-Gribov-Lipatov-Altarelli-Parisi (DGLAP)
-    equations \cite{Dokshitzer:1977sg,Gribov:1972ri,Lipatov:1974qm,Altarelli:1977zs}.
+    equations \cite{Dokshitzer:1977sg,Gribov:1972ri,Lipatov:1974qm,Altarelli:1977zs},
     \begin{equation}\label{eqn:DGLAP}
@@
     where $P_{ba}(\frac{x}{z})$ are splitting functions
-    representing parton $b$ emitting a parton $a$ carrying
+    representing parton $b$ emitting a parton $a$, that carries
     a momentum fraction $x$. These splitting functions are calculable
@@
     up to three-loops \cite{Moch:2004pa,Vogt:2004mw}.
-    At leading order \cite{Altarelli:1977zs} they are
-    \begin{align}\label{eqn:AP_kernels}
-        P^{(0)}_{qq}(z) &= C_{F}\left[\dfrac{1+z^{2}}{(1-z)_{+}} + \dfrac{3}{2}\delta(1-z)\right] \\
-        P^{(0)}_{qg}(z) &= T_{R}\left[z^{2} + (1-z)^{2}\right] \\
-        P^{(0)}_{gq}(z) &= C_{F}\left[\dfrac{1+(1-z)^{2}}{z^{2}}\right] \\
-        P^{(0)}_{gg}(z) &= 2C_{A}\left[\dfrac{z}{(1-z)_{+}}+\dfrac{1-z}{z}+z(1-z)\right] + \left(\dfrac{11C_{A}-4T_{R}n_{f}}{6}\right)\delta(1-z)
-    \end{align}
-    where the `$+$'-prescription has been used to regulate divergences
-    at $z = 1$.
+    At leading order they are \cite{Altarelli:1977zs}
+    \begin{equation}\label{eqn:AP_kernels}
+        \begin{split}
+            P^{(0)}_{qq}(z) &= C_{F}\left[\dfrac{1+z^{2}}{(1-z)_{+}} + \dfrac{3}{2}\delta(1-z)\right] \, , \\
+            P^{(0)}_{qg}(z) &= T_{R}\left[z^{2} + (1-z)^{2}\right] \, ,\\
+            P^{(0)}_{gq}(z) &= C_{F}\left[\dfrac{1+(1-z)^{2}}{z^{2}}\right] \, ,\\
+            P^{(0)}_{gg}(z) &= 2C_{A}\left[\dfrac{z}{(1-z)_{+}}+\dfrac{1-z}{z}+z(1-z)\right] + \left(\dfrac{11C_{A}-4T_{R}n_{f}}{6}\right)\delta(1-z) \, ,
+        \end{split}
+    \end{equation}
+    where the divergences at $z=1$ have been regulated
+    with the `$+$'-prescription, which is defined as
     \begin{equation}\label{eqn:plus_prescription}
-        \int_{0}^{1} \mathrm{d}z \; [g(z)]_{+}f(z) = \int_{0}^{1}\mathrm{d}z \; g(z)(f(z)-f(1)) \, ,
+        \int_{0}^{1} \mathrm{d}z \; [g(z)]_{+}f(z) = \int_{0}^{1}\mathrm{d}z \; g(z)(f(z)-f(1)) \, .
     \end{equation}
-    where the function $f(z)$ cancels the divergence if it is
-    sufficiently smooth at $z=1$.
+    With this prescription, the divergence at $g(z=1)$ is cancelled,
+    given that the function $f(z)$ is sufficiently smooth at $z=1$.
     At leading order, there is a physical interpretation
     of the splitting kernels as the probability of finding a parton
-    $a$ in parton $b$ carrying a fraction $x$ of the parent parton.
+    $a$ in parton $b$ carrying a momentum fraction $x$ of the parent parton.
 
@@
     \end{equation}
-    $\hat{\sigma}_{ab \rightarrow n}(Q, \mu_{F}, \mu_{R})$
-    is the partonic cross-section which descibes the interaction of
+    The partonic cross-section, $\hat{\sigma}_{ab \rightarrow n}(Q, \mu_{F}, \mu_{R})$,
+    describes the interaction of
     partons $a$ and $b$ scattering into $n$ particles,
@@
     Notice that the partonic cross-section depends on
-    both the factorisation scale, $\mu_{F}$ and
+    both the factorisation scale, $\mu_{F}$, and
     the renormalisation scale, $\mu_{R}$
-    (see Section~\ref{sec:renormalisation}). This will
+    (see Section~\ref{sec:renormalisation}). This dependence will
     be discussed in Section~\ref{sec:scale_variations}
@@
     where the electron-positron annihilates to form a
-    photon ($Z$ boson) which fluctuates into a quark-antiquark
+    photon or $Z$ boson which fluctuates into a quark-antiquark
     pair. For a sufficiently high energy collision,
@@
     particles, there is no requirement for PDFs,
-    therefore (\ref{eqn:hadronic_cs}) simplifies
+    therefore Eq. (\ref{eqn:hadronic_cs}) simplifies
     to just the partonic cross-section, which
-    will be computed in perturbation theory.
+    can be computed in perturbation theory.
 
-    The details of how $\hat{\sigma}_{ab \rightarrow n}$
-    is computed is discussed in the next section.
+    The details of how partonic cross-sections, $\hat{\sigma}_{ab \rightarrow n}$,
+    are computed is discussed in the next section.
 
 \section{Scattering amplitudes and cross-sections}\label{sec:xs}
-    As mentioned in the previous section, at
-    collider experiments we typically collide
-    two beams, consisting of bundles of energetic
-    particles, and analyse the resulting products
-    of these collisions. The predictions we make
+    At collider experiments we typically collide
+    two beams consisting of bundles of energetic
+    particles and analyse the resulting products
+    of these collisions. The predictions that we make
     from our theoretical model are the partonic
@@
     as the identity matrix describes the free theory.
-    By imposing a momentum conserving constraint
+    By imposing a momentum conservation constraint
     on the $S$-matrix, the matrix element,
@@
     particles in the initial-state, and $p_{f}$ to be the
-    momenta of the $n$-particle final states.
-    The transition probability of this scattering process
+    momenta of the $n$ final-state particles.
+    The probability of this process occurring
     is the modulus squared of the scattering amplitude
@@
     where $|\langle f | \mathcal{M} | i \rangle|^{2}  \equiv |\mathcal{M}|^{2}$
-    is the matrix element squared \footnote{from henceforth we refer to the matrix
-    element squared as matrix element}.
+    is the matrix element squared\footnote{Henceforth we refer to the matrix
+    element squared as matrix element, unless stated otherwise.}.
 
@@
     the cross-section, $\sigma$, as introduced in
-    (\ref{eqn:hadronic_cs}). The cross-section is a property
+    Eq. (\ref{eqn:hadronic_cs}). The cross-section is a property
     of the particles being scattered, and is independent
@@
     quantities, the partonic cross-section directly
-    relates the matrix elements which we calculate in our QFT,
+    relates the matrix elements which we calculate in our QFT
     to measurable observables at experiments.
@@
     will be produced with a range of, for example,
-    energies and momenta so the cross-section would
+    energies and momenta, so the cross-section would
     be vanishing for an exact momentum configuration.
@@
     \begin{equation}\label{eqn:dsigma}
-        \mathrm{d}\hat{\sigma} = \dfrac{1}{\mathcal{F}}|\mathcal{M}_{2 \rightarrow n}|^{2} \mathrm{d}\Phi_{n} \, .
+        \mathrm{d}\hat{\sigma} = \dfrac{1}{\mathcal{F}}|\mathcal{M}_{ab \rightarrow n}|^{2} \mathrm{d}\Phi_{n} \, ,
     \end{equation}
-    $\mathcal{F}$, the flux factor is
+    where the flux factor, $\mathcal{F}$, is given by
     \begin{equation}\label{eqn:flux}
-        \mathcal{F} = 4\sqrt{(p_{a}p_{b})^{2} - m_{a}^{2}m_{b}^{2}} \, ,
+        \mathcal{F} = 4\sqrt{(p_{a}p_{b})^{2} - m_{a}^{2}m_{b}^{2}} \, .
     \end{equation}
-    which in the massless limit reduces to
+    In the massless limit the flux factor reduces reduces to
     \begin{equation}\label{eqn:massless_flux}
@@
     experiment is much greater than the mass
-    of the colliding particles, and so they can
-    be neglected.
+    of the colliding particles, and so we can
+    neglect the masses.
 
-    $\mathrm{d}\Phi_{n}$ is the Lorentz-invariant
-    phase-space which contains all the possible
+    The Lorentz-invariant phase-space, $\mathrm{d}\Phi_{n}$,
+    contains all the possible
     configurations of the $n$-particle final-state.
     Absorbing the momentum conserving $\delta$-function
-    from (\ref{eqn:matrix_element}),
+    from Eq. (\ref{eqn:S_matrix_element}),
     we can write it as
@@
     the final-state particles to be on mass-shell
-    (on-shell). The step function selects only
+    (on-shell), and the step function selects only
     the positive energy solution.
@@
     From the definition of the transition probability
-    (\ref{eqn:S_prob}), it becomes clear
+    Eq. (\ref{eqn:S_prob}), it becomes clear
     that the matrix element is the object which
@@
     Feynman diagrams. In this picture, matrix elements
-    are exactly the sum over all Feynman diagrams
-    \footnote{Since we are working on the level
-    of $|\mathcal{M}|^{2}$, it would technically
+    are exactly the sum over all Feynman diagrams\footnote{Since we are working on the level
+    of $|\mathcal{M}|^{2}$, this would
     be the sum of all interference of Feynman diagrams.}
-    for a particular process. However, there are
-    an infinite number of Feynman diagrams you can
-    draw as there is no constraints on how the
-    vertices and propagators can be connected, as
-    long as the diagram contains the desired initial
-    and final-states.
-    Fortunately because $\alpha_{s}$
+    for a particular process. However, for a given
+    process the exact matrix element is a sum over
+    infinitely many Feynman diagrams.
+    Fortunately, because $\alpha_{s}$
     is small in high energy collisions, we can expand
@@
     where $m$ corresponds to the powers of $\alpha_{s}$
-    in the simplest Feynman diagrams you can draw
+    in the simplest Feynman diagrams
     for the process of interest.
@@
     N$^{\mathrm{k}}$LO where each additional power of $\alpha_{s}$
-    increases complixity of the computations, however,
+    increases complexity of the computations, however,
     because the expansion $\alpha_{s}$
@@
     less and less.
-    This means that the first terms dominate the expansion,
-    and it is sufficient to terminate the series after
+    In principle, this means that the first terms dominate the expansion,
+    and it should be sufficient to terminate the series after
     a few terms to reach an acceptable level of accuracy.
-    A further discussion of this point will be given
-    in Chapter~\ref{chapter:qcd} where we break down
-    the computation of each term in (\ref{eqn:matrix_element}).
 
@@
     collisions the partonic cross-section also needs to
-    be convoluted with the PDFs to obtain the hadronic
+    be convolved with the PDFs to obtain the hadronic
     cross-sections.

### 2022-11-17 qcd.tex
changes to chapter 2

@@
     calculated order-by-order in the strong coupling
-    $\alpha_{s}$ (c.f. \ref{eqn:matrix_element}),
+    $\alpha_{s}$ (c.f. Eq. \ref{eqn:matrix_element}),
     where each order corresponds to a set of Feynman diagrams
@@
     these issues in order to provide real-world applicable predictions.
-    This chapter forms the theoretical foundations upon which
-    this thesis is built on.
-
     We will consider the divergent nature of matrix
@@
     to tame these singularities.
+    This chapter forms the theoretical foundations upon which
+    this thesis is built on.
 
 \section{Divergent structures}\label{sec:divergences}
-    The computation of matrix elements essentially reduces down
+    The computation of matrix elements essentially reduces
     to evaluating Feynman diagrams which are analytical expressions
@@
 
-    The de-facto methods to alleviate these
+    The de facto methods to alleviate these
     divergences are through renormalisation
@@
     is based on the observation that the integral
-    (\ref{eqn:loop_integral}), which is carried
+    in Eq. (\ref{eqn:loop_integral}), which is carried
     out in $d=4$ space-time dimensions, would be finite
-    if we move away from $d=4$ dimensions
-    \footnote{In dimensional regularisation it is possible to regulate
+    if we move away from $d=4$ dimensions\footnote{In dimensional regularisation it is possible to regulate
     the integral regardless of if $d > 4$ or $d < 4$, it would amount
@@
     In DR we define $d = 4 - 2\epsilon$ such that
-    after Feynman parametrisation (\ref{eqn:loop_integral})
+    after Feynman parametrisation Eq. (\ref{eqn:loop_integral})
     becomes
@@
     the dimensions of the gauge and fermion fields
-    in the Lagrangian (\ref{eqn:L_QCD}).
+    in the Lagrangian in Eq. (\ref{eqn:L_QCD}).
     We see that this introduces a scale dependence
@@
     thought of as redefinitions of the bare fields
-    and couplings in the QCD Lagrangian (\ref{eqn:L_QCD}) to restore predictive
+    and couplings in the QCD Lagrangian to restore predictive
     power to the theory. The renormalised fields and couplings
@@
     such that we can set $Z_{n} = 1 + \delta_{n}$ for $n \in \{1, 2, 3\}$.
-    In this way we can write the Lagrangian as
+    In this way we can write the renormalised Lagrangian as
     \begin{equation}
-        \mathcal{L}_{R} = \mathcal{L}_{\mathrm{bare}} + \mathcal{L}_{\mathrm{c.t.}} \, ,
+        \mathcal{L}_{\mathrm{renorm}} = \mathcal{L}_{\mathrm{bare}} + \mathcal{L}_{\mathrm{c.t.}} \, ,
     \end{equation}
@@
     subtraction ($\overline{\mathrm{MS}}$) scheme which subtracts the
-    epsilon pole plus a universal constant appearing in all
+    epsilon pole along with a universal constant appearing in all
     loop integrals.
@@
     there would be no dependence on the renormalisation
-    scale as it is a remnant of the regularisation prescription
+    scale as it is a remnant of the regularisation prescription.
     However, because in perturbative QCD observables
@@
     the missing higher order terms. The conventional way
-    that the community estimates the uncertainty associated
-    with this residual dependence is by carrying out
+    that theoretical uncertainties associated
+    with this residual dependence are determined is by carrying out
     a scale variation (see Section~\ref{sec:scale_variations}).
@@
     The renormalisation scale dependence of $\alpha_{s}$
-    is governed by the Callan-Symanzik \cite{Callan:1970yg,Symanzik:1970rt}
-    $\beta$-function
+    is governed by the Callan-Symanzik 
+    $\beta$-function \cite{Callan:1970yg,Symanzik:1970rt}
     \begin{equation}\label{eqn:beta_fn}
@@
     up to $\beta_{4}$ \cite{Baikov:2016tgj,Luthe:2017ttg}.
-    The solution to (\ref{eqn:beta_fn}) to first order
+    The solution to Eq. (\ref{eqn:beta_fn}) to first order
     is
@@
     \end{equation}
-    where $\Lambda_{\mathrm{QCD}}$ is the QCD scale,
-    the scale at which $\alpha_{s}$ becomes
-    large enough that perturbation theory breaks down,
-    and $\beta_{0} = (11C_{A} - 4T_{R}n_{f})/3$.
+    where $\beta_{0} = (11C_{A} - 4T_{R}n_{f})/3$.
     In QCD, $C_{A}=3$ and $T_{R}=\frac{1}{2}$, meaning for
@@
     decreases with increasing energy due to the minus sign
-    in (\ref{eqn:beta_series}). This property is known as
+    in Eq. (\ref{eqn:beta_series}). This property is known as
     asymptotic freedom \cite{Gross:1973id,Politzer:1973fx},
@@
         as determined by experiments, with QCD theory prediction in black.
-        Figure reference \cite{Workman:2022ynf}.}
+        Figure reproduced from \cite{Workman:2022ynf}.}
         \label{fig:alpha_s_running}
@@
     modes. Since these divergences emerge from low energy
-    behaviour they are called infrared (IR) divergences. There
+    behaviour they are called infrared divergences. There
     are two cases in which these divergences arise:
@@
         \item virtual divergences in loop integrals,
-        \item real emission divergences when an emission
+        \item real-emission divergences when an emission
     of an extra particle has vanishing energy (soft),
@@
     which is encountered when calculating the one-loop
-    virtual correction to the gluon-quark-antiquark vertex
+    virtual correction to the gluon-fermion vertex
     as shown in Figure~\ref{fig:vertex_correction}.
@@
     These situations correspond to the gluon in the loop propagator
-    going soft, or collinear to the external quark or antiquark.
+    going soft, or collinear to the external quarks, respectively.
     These divergences are regulated in DR to give
@@
     \end{equation}
-    where $Q = (p_{1} + p_{2})$. There is a double $\epsilon$ pole manifest,
+    where $Q = (p_{1} + p_{2})$. In this expression, there is a double $\epsilon$ pole manifest,
     corresponding to the associated soft and collinear divergences in real emission corrections.
@@
 
-    Real IR divergences are manifest when integrating
+    Real IR divergences arise when integrating
     matrix elements over the phase-space of external state momenta.
@@
     \end{equation}
-    where $p_{i}$ and $p_{j}$ are the 4-momenta of partons $i$ and $j$
+    where $p_{i}$ and $p_{j}$ are the 4-momenta of two partons $i$ and $j$
     in the hard scattering.
@@
     it is not physically possible to observe arbitrarily soft
-    partons. It is also not possible to distinguish two partons
-    at arbitrarily small angles from one parton with the combined
+    particles. It is also not possible to distinguish two particles
+    at arbitrarily small angles from one particle with the combined
     momenta. These limitations along with the KLN theorem
@@
     with fixed-order perturbation theory must be insensitive to the
-    emissions of soft or collinear partons. Any observable obeying
+    emissions of soft or collinear particles. Any observable obeying
     this criteria is an infrared and collinear (IRC) safe observable.
@@
     Instead they are done numerically through the use
-    of Monte Carlo methods (see Chapter~\ref{chap:MC}),
+    of Monte Carlo methods (see Chapter~\ref{chapter:MC}),
     which requires the integral to be in
@@
     of Ref.~\cite{Catani:1996vz}, namely that the colour and
-    helicity summed $n$-body matrix element written as
+    helicity summed $n$-body matrix element can be written as
     \begin{equation}\label{eqn:CS_matrix_element}
@@
     The limit of the soft gluon with momentum
-    $p_{j}$ can be parametrised by
+    $p_{j}$ can be parametrised as
     \begin{equation}\label{eqn:soft_parametrisation}
@@
     colour + helicity vector inner product is obtained by removing
-    the soft gluon from the $(n+1)$-body matrix element.
+    the soft gluon, $j$, from the $(n+1)$-body matrix element.
     The indices $i$, $j$, and $k$ label partons involved
@@
     $j$ is the emitted parton, and $k$ is a parton accounting
-    for the colour correlations. The scale $\mu$ can be identified as the
-    renormalisation scale, and $\bm{T_{i}}$ is the colour-charge
+    for the colour correlations.
+    The scale $\mu$ can be identified as the
+    renormalisation scale, and $\bm{T}_{i}$ is the colour-charge
     operator. It acts on the colour vectors in the reduced matrix element
@@
 
-    The matrix elements in (\ref{eqn:soft_factorisation}) are
+    The matrix elements in Eq. (\ref{eqn:soft_factorisation}) are
     unambiguously defined only when momentum conservation is
@@
     This reduced matrix element is obtained by replacing the partons
-    $i$ and $j$ with a single parton $ij$ which carries the
+    $i$ and $j$ with a single parton $ij$, as explicitly shown in the reduced
+    matrix element. This composite parton carries the
     momentum $p^{\mu}$ and suitable quantum numbers depending
@@
     element does not factorise from the splitting functions exactly.
-
     $\hat{P}_{ij}$ become the more recognisable Altarelli-Parisi
-    splitting functions (\ref{eqn:AP_kernels}) once spin-averaged
+    splitting functions in Eq. (\ref{eqn:AP_kernels}) once spin-averaged
     and the $\epsilon \rightarrow 0$ limit is taken.
@@
 
-    The analogous expression of (\ref{eqn:F_collinear_factorisation})
+    The analogous expression of Eq. (\ref{eqn:F_collinear_factorisation})
     for initial-state splitting is
@@
     where the replacement of partons $a$ and $i$ have been made
-    explicit by the presence of the parton $ai$ in the colour + helicity vector.
+    explicit by the presence of the parton $ai$ in the colour + helicity vectors.
     The specific type of parton $ai$ depends upon the types of $a$ and $i$.
 
-    Similar to (\ref{eqn:soft_factorisation}) where the factorisation was
-    only true in the strict soft limit, (\ref{eqn:F_collinear_factorisation})
+    Similar to Eq. (\ref{eqn:soft_factorisation}) where the factorisation was
+    only true in the strict soft limit, Eqns. (\ref{eqn:F_collinear_factorisation})
     and (\ref{eqn:I_collinear_factorisation}) are only
     true in the strict collinear limit. Away from these
-    limits care has to be taken to conserve momenta
+    limits, care has to be taken to conserve momenta
     via the use of momenta mappings.
@@
 \subsection{Factorisation of colour-ordered amplitudes}\label{sec:CO_factorisation}
-    The factorisation formulae expressed in (\ref{eqn:soft_factorisation}),
+    The factorisation formulae expressed in Eqns. (\ref{eqn:soft_factorisation}),
     (\ref{eqn:F_collinear_factorisation}), and (\ref{eqn:I_collinear_factorisation})
@@
 
-    In general any QCD amplitude can be
+    In general, any QCD amplitude can be
     colour decomposed, that is the colour structure is
-    separated from the kinematics. Consider the process $e^{+}e^{-} \rightarrow q\bar{q} + n \; g$
-    \footnote{For the scope of this thesis it is sufficient
+    separated from the kinematics. Consider the process $e^{+}e^{-} \rightarrow q\bar{q} + n \; g$\footnote{For the scope of this thesis it is sufficient
     to consider colour decomposition of electron-positron
@@
 
-    Upon squaring the amplitude (\ref{eqn:epem_amplitude}), we get
+    Upon squaring the amplitude in Eq. (\ref{eqn:epem_amplitude}), we get
     \begin{equation}\label{eqn:epem_me2}
@@
     With the amplitude written in terms of the colour-ordered
-    partial amplitudes, it is now possible to factorise (\ref{eqn:epem_me2})
+    partial amplitudes, it is now possible to factorise Eq. (\ref{eqn:epem_me2})
     exactly in the soft and collinear limits. Consider the limit where a final-state
@@
     is the well-known eikonal factor. We see that the colour-ordered
-    amplitude on the RHS of (\ref{eqn:CO_soft_factorisation}) has gluon $j$
+    amplitude on the RHS of Eq. (\ref{eqn:CO_soft_factorisation}) has gluon $j$
     removed but the ordering of all hard partons remain unchanged.
@@
     where $P_{ij}(z)$ are the spin-averaged Altarelli-Parisi splitting
-    functions (\ref{eqn:AP_kernels}) with the colour factors removed.
+    functions with the colour factors removed.
     In this expression it is understood that the indices $i$, $j$, and $k$
-    have to be self-consistent for the different partonic splittings.
+    have to be self-consistent for the different types of partonic splittings.
     For partons which are not colour connected (are not neighbouring
@@
 
-    In (\ref{eqn:CO_soft_factorisation}) and (\ref{eqn:CO_collinear_factorisation})
+    In Eqns. (\ref{eqn:CO_soft_factorisation}) and (\ref{eqn:CO_collinear_factorisation}),
     the colour-ordered amplitude factorises exactly into a
@@
     \end{equation}
-    where $S_{ijk}^{(0)}$ is the eikonal factor (\ref{eqn:eikonal_factor}),
+    where $S_{ijk}^{(0)}$ is the eikonal factor in Eq. (\ref{eqn:eikonal_factor}),
     and $S_{ijk}^{(1)}(\epsilon)$ is the one-loop soft radiation function \cite{Bern:1999ry}.
@@
     where $P_{ij}^{(0)}(z)$ are the tree-level splitting functions
-    (\ref{eqn:AP_kernels}) and $P_{ij}^{(1)}(z, \epsilon)$
+    in Eq. (\ref{eqn:AP_kernels}) and $P_{ij}^{(1)}(z, \epsilon)$
     are the one-loop splitting functions \cite{Bern:1999ry}.
@@
     at the one-loop level for squared matrix elements. These
-    antenna functions are then applied in the context of NLO
-    k-factor emulation in Chapter~\ref{chapter:fame2}.
+    antenna functions are then applied in the context of NLO QCD
+    K-factor emulation for electron-positron annihilation in Chapter~\ref{chapter:fame2}.
 
@@
     isolate these singularities such that it is possible
-    to make finite predictions of physical quantities numerically.
+    to make finite predictions of physical quantities.
     To tackle this problem, three
@@
     The motivation behind the subtraction method is that the
-    divergences in (\ref{eqn:NLO_cs}) can be cancelled upon the insertion
+    divergences in Eq. (\ref{eqn:NLO_cs}) can be cancelled upon the insertion
     of a counterterm evaluated in $\Phi_{n+1}$, $\mathcal{C}_{n+1}$,
@@
     by construction.
-    Inserting the subtraction terms into (\ref{eqn:NLO_cs}) we get
+    Inserting the subtraction terms into Eq. (\ref{eqn:NLO_cs}) we get
     \begin{equation}\label{eqn:NLO_subtraction}
@@
     The counterterm $\mathcal{C}_{n+1}$ has been kept general
-    but specific examples of automated subtractions schemes include
+    but specific examples of subtractions schemes include
     Catani-Seymour (CS) \cite{Catani:1996vz,Catani:2002hc},
@@
     that are used to build subtraction terms: Catani-Seymour dipoles, and
-    antenna functions. However, we will not construct the counterterms explicitly.
-    In other words, we are interested in the approximations of the matrix elements
+    antenna functions. We are interested in the approximations of the matrix elements
     in the soft and collinear limits as these are universal and can
@@
     Namely, there are final-final (FF) dipoles, final-initial dipoles (FI),
-    initial-final (IF) dipoles and initial-initial (II) dipoles where
+    initial-final (IF) dipoles, and initial-initial (II) dipoles where
     the nomenclature refers to the emitter-spectator dipole.
     For electron-positron annihilations only FF dipoles are required
-    since electrons (positrons) do not carry any colour charge
+    since electrons and positrons do not carry any colour charge
     and so there is no initial-state radiation.
@@
         IR divergent behaviour.
-        The sum in (\ref{eqn:FF_factorisation}) can be understood as summing over all possible
-        3 leg permutations to capture all the soft and collinear limits.
-        The reduced matrix element on the RHS of (\ref{eqn:D_ijk})
+        The sum in Eq. (\ref{eqn:FF_factorisation}) can be understood as summing over all possible
+        three leg permutations to capture all the soft and collinear limits.
+        The reduced matrix element on the RHS of Eq. (\ref{eqn:D_ijk})
         is obtained by replacing the partons $i$ and $j$ with
         a single parton $\widetilde{ij}$, and replacing parton
-        $k$ with a parton $\tilde{k}$. $\tilde{k}$ has all the
+        $k$ with a parton $\tilde{k}$. The parton $\tilde{k}$ has all the
         same quantum numbers as $k$, whereas the
@@
         \end{equation}
-        Equations (\ref{eqn:FF_mapping}) and (\ref{eqn:y_ijk})
+        Eqns. (\ref{eqn:FF_mapping}) and (\ref{eqn:y_ijk})
         are a $3 \rightarrow 2$ momenta mapping that maps
@@
         \end{equation}
-        They are analogous to the $z$ from Altarelli-Parisi splitting functions.
+        These splitting variables are analogous to the $z$ from Altarelli-Parisi splitting functions.
         $\boldsymbol{V}_{ij,k}$ acts on the spin indices of the composite particle $\widetilde{ij}$,
@@
         $\mathcal{D}_{ij,k}$ correctly reproduces the matrix element factorisation
-        behaviour seen in (\ref{eqn:soft_factorisation}) and (\ref{eqn:F_collinear_factorisation}),
+        behaviour seen in Eqns. (\ref{eqn:soft_factorisation}) and (\ref{eqn:F_collinear_factorisation}),
         respectively. In particular, $\boldsymbol{V}_{ij,k}$ becomes
@@
         colour- and spin-averaged matrix elements. Therefore the spin-indices
-        are not explicitly available. It is possible to average over the spin indices in (\ref{eqn:FF_V_ijk})
+        are not explicitly available. It is possible to average over the spin
+        indices in Eq. (\ref{eqn:FF_V_ijk})
         to obtain the spin-averaged splitting functions
@@
         \end{equation}
-        such that the following momentum conservation and on-shell conditions are met
+        such that momentum conservation and on-shell conditions are met:
         \begin{equation}\label{eqn:FI_mapping_conditions}
@@
         \begin{equation}\label{eqn:IF_mapping}
-            \tilde{p}_{ai}^{\mu} = x_{ik,a} \, p_{a}^{\mu} \, , \quad \tilde{p}_{k}^{\mu} = p_{k}^{\mu} + p_{i}^{\mu} - (1-x_{ik,a})p_{a}^{\mu} 
+            \tilde{p}_{ai}^{\mu} = x_{ik,a} \, p_{a}^{\mu} \, , \quad \tilde{p}_{k}^{\mu} = p_{k}^{\mu} + p_{i}^{\mu} - (1-x_{ik,a})p_{a}^{\mu} \, ,
         \end{equation}
@@
         acting as a splitting variable.
-
         With $x_{ik,a}$ taking the role of $\tilde{z}$, there is another
         parameter, $u_{i}$, appearing in the splitting functions.
-        These are given as
+        These variables are given as
         \begin{equation}\label{eqn:x_ika}
@@
         In the case of two initial-state partons, $a$ and $b$, there
-        is an additional dipole contribution. The dipole factorisation
-        in this case is  given by
+        is an additional dipole contribution compared to the initial-final case.
+        The dipole factorisation is given by
         \begin{equation}\label{eqn:II_factorisation}
             \begin{split}
-                {}_{n+1} \langle 1, \ldots, n+1; a,b|&| 1, \ldots, n+1;a,b \rangle_{n+1} = \\
+                {}_{n+1} \langle 1, \ldots, n+1; a,b &| 1, \ldots, n+1;a,b \rangle_{n+1} = \\
                 &\sum_{a,i}\sum_{k \neq i} \mathcal{D}_{k}^{ai}(p_{1}, \ldots, p_{n+1};p_{a},p_{b}) \\
@@
         The initial-state parton $\widetilde{ai}$ is the emitter and the
-        other initial-state parton $b$ is the spectator. Notice that even the
+        other initial-state parton $b$ is the spectator. Notice that the
         momenta not involved in the dipole term have been modified with only
@@
     \subsection{Master factorisation formula}
-        Now that we have described each of the dipole contributions,
-        we can collect them all into a master dipole factorisation
-        formula
+        The combination of all types of dipole contributions
+        can be combined into a master dipole factorisation formula
         \begin{equation}\label{eqn:dipole_factorisation_master}
@@
         the dipole functions can be used as a set of basis functions for
-        approximating the matrix elements in Chapters \ref{chapter:fame1}
+        approximating the matrix elements for arbitrary SM processes.
+        This will be put into practice in Chapters \ref{chapter:fame1}
         and \ref{chapter:fame3}. The omitted terms are non-divergent
@@
         not be written here as the procedure of constructing
-        the massive dipole functions is similar to massless ones.
+        the massive dipole functions is similar to their massless counterparts.
 
@@
         modification of the master dipole factorisation formula
-        (\ref{eqn:dipole_factorisation_master}) to replace
+        in Eq. (\ref{eqn:dipole_factorisation_master}) to replace
         any massless dipole with its massive counterpart whenever
@@
     where $X_{3}^{1}$ are defined such that they are proportional
-    to the one-loop singular functions. Since the one-loop
-    factorisation formulae (\ref{eqn:1L_soft_factorisation})
-    and (\ref{eqn:1L_collinear_factorisation}) are of the form
+    to the one-loop singular functions mentioned in Section~\ref{sec:OL_factorisation}.
+    Since the one-loop factorisation formulae, Eqns. (\ref{eqn:1L_soft_factorisation})
+    and (\ref{eqn:1L_collinear_factorisation}), are of the form
     $|M^{(1)}_{n+1}|^{2}\rightarrow S^{(0)}|M^{(1)}_{n}|^{2} + S^{(1)}|M^{(0)}_{n}|^{2}$,
-    the term proportional to $X_{3}^{0}$ has to be removed
-    to fulfil this requirement.
+    the term proportional to $S^(0)$, that is $X_{3}^{0}$, has to be removed
+    in Eq. (\ref{eqn:X31}) to fulfil this requirement.
     The antenna functions are summarised in Table~\ref{table:antenna_functions},
@@
         \begin{split}
-        p_{I} &= x p_{i} + r p_{j} + y p_{k} \, \\
+        p_{I} &= x p_{i} + r p_{j} + y p_{k} \, , \\
         p_{K} &= (1-x) p_{i} + (1-r) p_{j} + (1-y) p_{k} \, ,
@@
 
-    Below we list the antenna functions relevant
+    In the following, we list the antenna functions relevant
     for multi-jet production in electron-positron
@@
     namely the partonic channel $e^{-}e^{+} \rightarrow
-    q \bar{q}$ + $n \, g$. The emulation of the NLO QCD K-factors
-    of this partonic process is examined in Chapter~\ref{chapter:fame2}.
+    q \bar{q}$ + $n \, g$.
     For this process we require the $A$, $D$, and $F$ antenna functions,
@@
     The $i,j,k$ indices will be fixed such that $q = 1$, $\bar{q} = 2$,
-    and $g \in \{3, 4, 5\}$ in view of using the
-    antenna functions where the momenta
-    of partons will have fixed indices in a phase-space
-    point represented as an array.
+    and $g \in \{3, 4, 5\}$, in view of using the
+    antenna functions with the momenta
+    represented as an array with partons having fixed indices.
 
@@
     singular and finite contributions from an antenna
-    function. However, there are still finite terms
+    function. However, there are remaining finite terms
     from $\mathcal{P}oles(X)$ coming from the $\epsilon$-expansion
@@
     care has to be taken to extract all the finite
-    remainders from the antenna functions when trying
+    contributions from the antenna functions when trying
     to use them as basis functions for emulating the
@@
     For the one-loop antenna functions, there are three
-    different functions corresponding to the: leading colour,
+    different functions corresponding to the leading colour,
     sub-leading colour, and closed quark look contributions.
@@
     \subsection{Limiting behaviour of antenna functions}
-    Examining (\ref{eqn:A30}) more closely, it becomes clear
+    Examining Eq. (\ref{eqn:A30}) more closely, it becomes clear
     that the correct IR behaviour is exhibited by the antenna
@@
     going collinear with the quark and antiquark, respectively. The final
-    term encapsulates singularity in the limit of the gluon
+    term encapsulates the singularity in the limit of the gluon
     going soft.
-
     The other antenna functions have similar limiting behaviour,
@@
     they can be used as a set of basis functions for approximating
-    matrix elements. This will be explored in Chapter~\ref{chapter:fame2}.
+    matrix elements. The emulation of the NLO QCD K-factors
+    for electron-positron annihilation is explored in Chapter~\ref{chapter:fame2}.
 \end{document}

### 2022-11-17 mc.tex
changes to chapter 3

@@
 \begin{document}
-\chapter{Monte Carlo Event Generators}\label{chap:MC}
+\chapter{Monte Carlo Event Generators}\label{chapter:MC}
     At particle collider experiments,
@@
     to simulate these stages of an event are general-purpose
-    Monte Carlo event generators (event generators), see
-    for example Reference \cite{Buckley:2011ms} for a review.
+    Monte Carlo event generators, see
+    Ref. \cite{Buckley:2011ms} for a review.
     These are indispensable tools that help
@@
     General-purpose event generators
-    such as Herwig \cite{Bahr:2008pv,Bellm:2015jjp},
-    Pythia \cite{Sjostrand:2006za,Bierlich:2022pfr},
-    and Sherpa \cite{Gleisberg:2008ta,Sherpa:2019gpd} all broadly share 
+    such as {\Herwig} \cite{Bahr:2008pv,Bellm:2015jjp},
+    {\Pythia} \cite{Sjostrand:2006za,Bierlich:2022pfr},
+    and {\Sherpa} \cite{Gleisberg:2008ta,Sherpa:2019gpd} all broadly share 
     the same features at a high-level, but each framework
@@
     generation, a more in-depth treatment of the
-    specifics of Monte Carlo methods can be found in, for example, Reference
+    specifics of Monte Carlo methods can be found in Ref.
     \cite{James:1980yn}.
@@
     The basic idea of Monte Carlo integration is to approximate
-    the integral of a function $f(\vec{x})$ by sampling it
-    in $d$-dimensional parameter space points $\vec{x}$
+    the integral of a function $f(\mathbf{x})$ by sampling it
+    in $d$-dimensional parameter space points $\mathbf{x}$
     \begin{equation}\label{eqn:MC_integration}
-        I = \int_{V} \mathrm{d}^{d}x \; f(\vec{x}) \approx \langle I(f) \rangle_{x} = V \dfrac{1}{N} \sum_{i=1}^{N} f(\vec{x}_{i}) = V \langle f \rangle_{x} \, ,
+        I = \int_{V} \mathrm{d}^{d}x \; f(\mathbf{x}) \approx \langle I(f) \rangle_{x} = V \dfrac{1}{N} \sum_{i=1}^{N} f(\mathbf{x}_{i}) = V \langle f \rangle_{x} \, ,
     \end{equation}
-    where the points $\vec{x}$ are randomly,
+    where the points $\mathbf{x}$ are randomly,
     uniformly sampled in the volume $V$. $\langle f \rangle_{x}$
@@
     result of the central limit theorem, the function evaluations
-    $f(\vec{x}_{i})$ will follow a Gaussian distribution with
+    $f(\mathbf{x}_{i})$ will follow a Gaussian distribution with
     mean $\langle I(f) \rangle_{x}$,
@@
 
-    An algorithm that samples points $\vec{x}$ is called
+    An algorithm that samples points $\mathbf{x}$ is called
     an integrator. To make use of the random numbers sampled,
@@
     aims to reduce the variance via a remapping of the
-    uniformly sampled variables $\vec{x}$
+    uniformly sampled variables $\mathbf{x}$
     to a more suitable non-uniform distribution.
@@
     structure of multi-parton QCD processes have been
-    studied in References \cite{Draggiotis:2000gm,vanHameren:2002tc}.
+    studied in Refs. \cite{Draggiotis:2000gm,vanHameren:2002tc}.
 
@@
     volume into sub-volumes, or bins. The overall integral
-    and variances are the sums of the partial results
+    and variance are the sums of the partial results
     in each bin. The overall variance is minimised when
@@
     is rapidly fluctuating (and not necessarily large),
-    and less from bins where the integral has less fluctuations.
+    and less from bins where the integral has fewer fluctuations.
 
-    In practice, state-of-the-art event generators
+    In practice, state of the art event generators
     combine the concepts of importance sampling and stratified
@@
     Once a suitable configuration of momenta
-    has been found by the integrator, this point forms the basis of an
+    has been found by the integrator, this phase-space point forms the basis of an
     event. The matrix element of interest can be evaluated at
     this phase-space point at the desired order in perturbation
-    theory (see Section~\ref{sec:me_generators}). Additionally, there will be an associated Jacobian
+    theory (see Section~\ref{sec:me_generators}).
+    Additionally, there will be an associated Jacobian
     from the mapping from random variables to the four-momenta.
@@
         \While{unweighting}{
-            generate random phase-space point $\vec{x}$\;
-            evaluate event weight $w \gets w(\vec{x})$\;
+            generate random phase-space point $\mathbf{x}$\;
+            evaluate event weight $w \gets w(\mathbf{x})$\;
             generate uniform random number $r \gets$ Random(0, 1)\;
             \If{$w / w_{\mathrm{max}} > r$}{
-                \KwRet $\vec{x}$ and $\tilde{w} \gets 1$\;
+                \KwRet $\mathbf{x}$ and $\tilde{w} \gets 1$\;
                 % $\tilde{w} \gets \max(1,\; w / w_{\mathrm{max}})$\;
@@
     itself is to accelerate the process of unweighting itself. This
-    was explored in Reference \cite{Danziger:2021eeg} wherein a
+    was explored in Ref. \cite{Danziger:2021eeg} wherein a
     novel two-stage unweighting procedure employing a surrogate
@@
     in Chapter~\ref{chapter:fame3} where we build upon
-    Reference \cite{Danziger:2021eeg} by increasing the fidelity of
+    Ref. \cite{Danziger:2021eeg} by increasing the fidelity of
     the surrogate model.
@@
     the parton shower, and finally hadronisation.
-    This is motivated by the factorisation theorem (Section~\ref{sec:factorisation})
+    This is motivated by the factorisation theorem (c.f. Section~\ref{sec:factorisation})
     which separates the high energy regimes from the low
@@
     the exact matrix element is approximated by a perturbative
-    series in a coupling constant. By now it is standard to
+    series in a coupling parameter. By now it is standard to
     have at least the LO and NLO contributions in this expansion.
@@
     compute matrix elements at tree-level and
-    one-loop level required for SM predictions, see for instance Reference \cite{Campbell:2017hsr}.
+    one-loop level required for SM predictions, see Ref. \cite{Campbell:2017hsr}.
     NNLO and N$^{3}$LO QCD corrections have been computed
@@
     level of automation of NLO.
-    For a review of recent advancements see References \cite{Caola:2022ayt,Campbell:2022qmc}.
+    For a review of recent advancements see Refs. \cite{Caola:2022ayt,Campbell:2022qmc}.
 
@@
     The simulation of the formation of hadrons in the low-energy regime
-    of an event generator follow phenomenological, non-perturbative models.
+    of an event generator follows phenomenological, non-perturbative models.
     These hadronisation models require the combination
@@
     For more discussion on hadronisation
-    see for example Reference \cite{Webber:1999ui}.
+    see Ref. \cite{Webber:1999ui}.
 
@@
     theoretical uncertainties associated with these
-    missing higher order terms is to carry out a seven-point
+    missing higher order terms is to carry out a seven point
     scale variation. Typically in a calculation,
@@
     of $\delta$ \cite{Cacciari:2011ze}.
-    The reasoning behind this approach is that $\delta$
+    The rationale behind taking this approach is that $\delta$
     is approximately the same order as the true missing

### 2022-11-17 ml.tex
changes to chapter 4

@@
     particle physics analysis, from collection
-    of data from experiments all the way to novel
+    of data from experiments, all the way to novel
     applications on the theory side.
@@
     Around the turn of the last decade the culmination
-    of advancement in hardware and machine learning training algorithms
+    of advancement in hardware and training algorithms
     led to an explosion in research in the
@@
     of these deep neural networks greatly outperformed the
-    previous state-of-the-art \cite{NIPS2012_c399862d,russakovsky2015imagenet} and
+    previous state of the art \cite{NIPS2012_c399862d,russakovsky2015imagenet} and
     lead to the widespread use of neural networks across
@@
     in an attempt to disentangle these events. With these large
-    arrays, terabytes of data are harvested per second from
-    each collision. This presents a practical challenge for data
+    arrays, terabytes of data are harvested from
+    every collision. This presents a practical challenge for data
     collection where machine learning has could provide a solution.
@@
     it is important that the generation of simulated samples
-    speeds up. This prompted the explosion of research
+    is accelerated. This prompted the explosion of research
     in machine learning applications on the theory side.
@@
     a living review that is archiving advancements in the
-    field is available at \cite{Feickert:2021ajf}.
-    For more traditional reviews, see for instance
+    field is available at Ref. \cite{Feickert:2021ajf}.
+    For more traditional reviews, see Refs.
     \cite{Guest:2018yhq,Radovic:2018dip,Butter:2022rso}.
@@
     is to interface these novel machine learning methods
-    to existing event generators.
+    to existing event generators for real-world use.
     There is scepticism regarding the use of machine learning
@@
     elements with a review of the current
-    state-of-the-art methods.
+    state of the art methods.
 
@@
     Once the model is fitted, it is customary to use
-    an unbiased dataset that is distinct from the training set
-    to evaluate the final model performance. This dataset is
-    the test set.
+    an unbiased dataset that is distinct from the training set, the test set,
+    to evaluate the final model performance.
 
@@
     \begin{equation}\label{eqn:neuron}
-        y = \phi(\mathbf{w}^{T}\mathbf{x} + b)
+        y = \phi(\mathbf{x}; \theta) = \phi(\mathbf{w}^{T}\mathbf{x} + b)
     \end{equation}
@@
         \caption{A schematic diagram of a neuron. The model inputs
-        $\mathbf{x}$ and weights $\mathbf{w}$ are vectors, where as
+        $\mathbf{x}$ and weights $\mathbf{w}$ are vectors, whereas
         the bias $b$ is a scalar. The weighted inputs are summed with
@@
     Some of the more well-known activation functions are
-    the sigmoid function, hyperbolic tangent, rectified linear unit (ReLU) \cite{Nair2010RectifiedLU}.
+    the sigmoid function, hyperbolic tangent, and the rectified linear unit (ReLU) \cite{Nair2010RectifiedLU}.
     A less familiar class of activation functions are the
@@
     The functional forms of these
-    functions are summarised in Table~\ref{table:activation_functions}.
+    functions and their gradients are summarised in Table~\ref{table:activation_functions}.
     There is no single activation function that is the best
@@
     The neural network is free to change the parameters
-    in these layers to best approximate the target function,
+    in these hidden layers to best approximate the target function
     $g(\mathbf{x})$.
@@
     The output of a densely-connected neural network can
-    be formulated by repeatedly applying (\ref{eqn:neuron})
+    be formulated by repeatedly applying Eq. (\ref{eqn:neuron})
     to give
@@
     learning non-linear functions, the loss surface
-    corresponding to (\ref{eqn:MSE}) is
+    corresponding to Eq. (\ref{eqn:MSE}) is
     a function of many variables, and is likely highly
@@
     for practical applications. In modern machine learning
-    libraries, the updating of weights in (\ref{eqn:gradient_descent})
+    libraries, the updating of weights in Eq. (\ref{eqn:gradient_descent})
     is carried out via matrix multiplications which are
@@
     are based on. For an overview of these optimisers, see for instance
-    Reference \cite{Ruder:2016lil}.
+    Ref. \cite{Ruder:2016lil}.
 
@@
     the speed and quality of the training process.
-    For instance, the learning rate in ({\ref{eqn:gradient_descent}}) is
+    For instance, the learning rate in Eq. ({\ref{eqn:gradient_descent}}) is
     an important hyperparameter whose initial value has
@@
     Secondly, there is no exact a priori way to evaluate the
-    final performance of the neural network given a set of
+    final performance of a neural network given a set of
     hyperparameters, meaning training has to take place for
@@
     be very large. Methods which aim to reduce the number
-    of trials have been proposed in the literature, see for example
-    Reference \cite{Yu2020HyperParameterOA} for a review.
+    of trials have been proposed in the literature, see Ref.
+    \cite{Yu2020HyperParameterOA} for a review.
     Below we discuss two paradigms of hyperparameter optimisation
@@
 
-    In practice the area of hyperparameter optimisation
+    In practice, the area of hyperparameter optimisation
     is still extremely computationally expensive even with
-    the more advanced methods. Given a sensible set of hyperparameters,
+    the more advanced methods outlined above. Given a sensible set of hyperparameters,
     one has to balance the additional time spent tuning with the
@@
     Once a fast and accurate surrogate model has been
-    built, it can supplement the traditional matrix
-    element generator to increase statistical power of the sample
+    built, it can supplement traditional matrix
+    element generators to increase statistical power of the sample
     by predicting on many more phase-space points. This
@@
     simple matrix multiplications, and because they are predicted
-    in batches, vectorisation is automatic. Therefore the difficulty
+    in batches, vectorisation is automatic\footnote{Event generators currently generate events one at a time, however, neural network predictions are still highly performant for single predictions.}.
+    Therefore the difficulty
     in building a good surrogate model is in controlling
@@
 
-    This FKS partitioning method was applied in Reference \cite{Aylett-Bullock:2021hmo}
+    This FKS partitioning method was applied in Ref. \cite{Aylett-Bullock:2021hmo}
     for diphoton plus jet production, where the authors presented a
-    use case of the emulator with a novel interface to the Sherpa event generator.
-    This work was revisited in \cite{Badger:2022hwf},
+    use case of the emulator with a novel interface to the {\Sherpa} event generator.
+    This work was revisited in Ref. \cite{Badger:2022hwf},
     where particular attention was paid to quantifying the uncertainty associated
@@
     matrix elements for $e^{+}e^{-} \rightarrow q \bar{q}$ + up to 3 gluons
-    can be emulated accurately to well below 1\% accuracy,
+    can be emulated accurately to well below 1\% per-point accuracy,
     even for the highest multiplicity. Extending this study
@@
     where antenna functions replace Catani-Seymour dipoles as
-    the basis functions.
+    the basis functions. Accuracy was observed to be of the
+    order 1\% for the 5 parton case, with good scaling
+    to higher multiplicities.
 
     The question of uncertainty of the neural network
-    prediction is important as it contributes to the total
+    prediction is important as it directly contributes to the total
     theoretical uncertainty. An alternative to quantifying
@@
     element in event unweighting. This was explored in
-    Reference \cite{Danziger:2021eeg},
+    Ref. \cite{Danziger:2021eeg},
     where it was shown that with a surrogate model of the event

### 2022-11-17 paper1.tex
update chapter 5

@@
 \label{sec:intro}
-Neural networks (NN) first appeared in the field of particle physics as a tool for event discrimination in the analysis of collider data. 
-Since then neural networks and other machine learning techniques have proved useful in many other areas of the field. 
-On the theory prediction side they have been used 
-to improve the efficiency of Monte Carlo sampling \cite{Bendavid:2017zhk,Klimek:2018mza,Bothmann:2020ywa,Verheyen:2020bjw,Chen_2021}, 
-to accelerate the simulation of radiation within a jet~\cite{Carrazza:2019cnt,Bothmann:2018trh,Dohi:2020eda}, 
-to streamline the processes of generation and unweighting of simulated event samples 
-\cite{Gao:2020zvv,Otten:2018kum,Hashemi:2019fkn,DiSipio:2019imz,Butter:2019cae,Bishara:2019iwh,Backes:2020vka,Butter:2020qhk,Alanazi:2020klf,Nachman:2020fff}.
-Closer to the experimental measurements 
-they have also been used to emulate detector simulation \cite{Paganini:2017dwg,SHiP:2019gcl,Derkach:2019qfk,Alanazi:2020jod},
- they can be used to perform unfolding \cite{Andreassen:2019cjw} or correcting for detector effects \cite{Bellagente:2019uyp},
-and perform pileup subtraction \cite{Andreassen:2019cjw}.
-
 The capacity of neural networks to approximate intricate functions have already been 
 used to provide fast calculations of production cross-sections~\cite{Otten:2018kum,Buckley:2020bxg}.
-In this article we investigate whether NNs can approximate production cross-sections more differentially by replacing computationally expensive 
+In this chapter we investigate whether NNs can approximate production cross-sections more differentially by replacing computationally expensive 
 matrix element calculations. 
@@
 
-In this article we present a different approach to the emulation of matrix elements that incorporates the factorisation 
+In this chapter we present a different approach to the emulation of matrix elements that incorporates the factorisation 
 properties of the matrix elements in the interpolation model and therefore is able to safely 
@@
 
-The article is organised as follows. 
+This chapter is organised as follows. 
 Section~\ref{sec:fitting} introduces our factorisation-aware 
@@
 
-We provide code to reproduce the methodology detailed in this article \cite{fame_repo}.
+Computer code to reproduce the methodology detailed in this chapter is provided at \cite{fame_repo}.
 
@@
 \subsection{Infrared divergences and dipole factorisation formula}
-It is well known that in soft and collinear limits the matrix element in $n+1$-body phase-space factorises into a singular factor and a reduced matrix element in $n$-body phase-space \cite{Altarelli:1977zs,BASSETTO1983201}.
-This factorisation was used by Catani and Seymour \cite{Catani:1996vz} to construct subtraction terms for the real radiation part of a NLO calculation.
+It is well known that in soft and collinear limits the matrix element in $(n+1)$-body phase-space factorises into a singular factor and a reduced matrix element in $n$-body phase-space \cite{Altarelli:1977zs,BASSETTO1983201} (Section~\ref{sec:me_factorisation}).
+This factorisation was used by Catani and Seymour \cite{Catani:1996vz} to construct subtraction terms for the real radiation part of a NLO calculation (Section~\ref{sec:CS_dipoles}).
 They introduced a factorisation formula with universal dipoles that smoothly interpolates 
@@
 In Ref. \cite{Badger:2020uow}, the authors used 500k training samples to train their models.
-For a fair comparison of our respective methods, we follow their methodology closely by constructing our models to be as close as possible to theirs and generate training and testing data by using code from their project repository \cite{n3jet_repo}\footnote{for our main results we generate all data using our own code}.
+For a fair comparison of our respective methods, we follow their methodology closely by constructing our models to be as close as possible to theirs and generate training and testing data by using code from their project repository \cite{n3jet_repo}\footnote{For our main results we generate all data using our own code}.
 Our model architectures will not be identical due to the difference in our methods, details of our model architecture are given in Section~\ref{sec:architecture}.
@@
 After sufficient training, each model will have a distinct set of parameters that have similar predictive power.
-The prediction for the matrix element is then the mean of the outputs of all models in the ensemble\footnote{if not explicitly stated, all references to neural network predictions from henceforth refers to the prediction from the ensemble of networks}.
+The prediction for the matrix element is then the mean of the outputs of all models in the ensemble\footnote{If not explicitly stated, all references to neural network predictions henceforth refers to the prediction from the ensemble of networks}.
 Taking the mean will give a more accurate and robust prediction as averaging over the different models will reduce variance due to over/underfitting in the training phase.
@@
 In Figure~\ref{fig:error_500k_vs_40m}, we show the improvements in accuracy of our main NNs compared to the smaller NNs from Section~\ref{sec:n3jet_comparison}.
-Although the training and testing data for the smaller NNs are not identical due to differences in code used to generate the sets\footnote{although RAMBO is used for phase-space generation in both works, the selection criteria is different (JADE algorithm vs $k_{t}$ algorithm)}, for this comparison it suffices to show that the larger NNs are orders of magnitude more accurate than the smaller NNs.
+Although the training and testing data for the smaller NNs are not identical due to differences in code used to generate the sets\footnote{Although RAMBO is used for phase-space generation in both works, the selection criteria is different (JADE algorithm vs $k_{t}$ algorithm).}, for this comparison it suffices to show that the larger NNs are orders of magnitude more accurate than the smaller NNs.
 
@@
 We show the results for 5-jet trajectories as this is the highest multiplicity we considered, predictions on lower multiplicities are better-behaved.
-We investigated 50 different random trajectories\footnote{the corresponding plots are available in a document attached to this submission}. For the discussion in this article 
+We investigated 50 different random trajectories. For the discussion in this section
 we chose one that contains many interesting features, namely the matrix elements span many orders of magnitudes and there are distinct peaks in the trajectory.
@@
 \section{Conclusion}\label{sec:conclusion}
-In this article we presented a new strategy to emulate matrix elements using a neural network. By leveraging the knowledge of 
+In this chapter we presented a new strategy to emulate matrix elements using a neural network. By leveraging the knowledge of 
 the factorisation properties of the matrix elements our model is able to extrapolate well outside of its training range. 
@@
 
-Our method was demonstrated in this article using a tree-level process, but it could be generalised to higher order matrix elements by adapting the set 
-of ingredients made available to the network for the interpolation.
-Specifically, we can include extra terms into our ansatz to help capture additional divergences at higher orders.
-We look forward to applying this method to such cases.
+Our method was demonstrated in this chapter using a tree-level process, but it could be generalised to higher order matrix elements by adapting the set 
+of ingredients made available to the network for the interpolation. This is explicitly shown in the next chapter for the same process at NLO QCD.
 

### 2022-11-17 intro.tex
changes to chapter 1

@@
 when generating the large samples required to compare with
-experiments. Especially when a significant portion of time
+experiments. Especially when a significant portion of the time
 spent simulating an event is spent on evaluating matrix
@@
     by the exchange of a gauge boson. The massless gauge bosons mediating the strong
-    and EM force are the gluon $g$ and photon $\gamma$,
+    and EM forces are the gluon $g$ and photon $\gamma$,
     respectively. For the weak force the gauge bosons are the $W^{\pm}$
@@
     into quarks and leptons. Quarks are massive and experience the strong,
-    weak, and EM force. Leptons are defined by their lack of colour charge,
+    weak, and EM forces. Leptons are defined by their lack of colour charge,
     meaning they do not experience the strong force. Leptons can be separated
@@
     $\mathcal{L}_{\mathrm{Higgs}}$ describes the Higgs field,
-    $\mathcal{L}_{\mathrm{Yukawa}}$ describes the interaction between the Higgs
+    $\mathcal{L}_{\mathrm{Yukawa}}$ describes the interactions between the Higgs
     field and fermions,
@@
     \end{equation}
-    where $\xi = 1$ corresponds to the Feynman-'t-Hooft gauge.
+    where $\xi = 1$ corresponds to the Feynman-'t Hooft gauge.
     The ghost Lagrangian is most commonly written
@@
     where $c^{a}$ ($\bar{c}^{a}$) are Faddeev-Popov ghosts (antighosts).
-    The ghosts are unphysical external states but must be
-    included in internal lines for a gauge invariant theory.
+    The ghosts are unphysical particles and so cannot appear as
+    external states but must be included in internal lines for
+    a gauge invariant theory.
 
@@
     Expanding Eq. (\ref{eqn:L_QCD}) and extracting the terms
-    that mix the gauge and fermions fields, we get an interaction Lagrangian
+    that mix the gauge and fermion fields, we get an interaction Lagrangian
     \begin{equation}\label{L_int}
@@
     From these Feynman rules, we identify that each
-    fermion-gluon and three gluon vertices
-    are associated with one power of $\alpha_{s}$ (due to squaring $g_{s}$),
+    fermion-gluon and three gluon vertex
+    is associated with one power of $\alpha_{s}$ (due to squaring $g_{s}$),
     and the four gluon vertex is associated with
@@
 
-    The momenta fractions, $x$, also link the
+    The momentum fractions, $x$, also link the
     squared centre-of-mass energies of the hadronic collision,
@@
 
-    The discussion so far has been focused hadron-hadron
-    initiated scattering, however, a similar argument
+    The discussion so far has been focused on hadron-hadron
+    initiated scattering, however a similar argument
     can be made about electron-positron annihilation,
     where the electron-positron annihilates to form a
-    photon or $Z$ boson which fluctuates into a quark-antiquark
+    photon or $Z$ boson which decays into a quark-antiquark
     pair. For a sufficiently high energy collision,
@@
     At collider experiments we typically collide
-    two beams consisting of bundles of energetic
+    two beams consisting of bunches of energetic
     particles and analyse the resulting products
@@
     a collider experiment, we look at the specific
-    case of two particles colliding into $n$ particles.
+    case of two particles colliding and producing $n$ particles.
     Scattering amplitudes are used to mathematically
@@
     the scattering amplitude can be written as
+    the overlap between the states
     \begin{equation}\label{eqn:scattering}
@@
     about how the initial-state will evolve
-    into the final-state over time. By writing
+    over time. By writing
     the $S$-matrix in this way, all the interactions
@@
     \begin{equation}\label{eqn:S_matrix_element}
-        \langle f | \mathcal{T} | i \rangle = (2\pi)^{4} \delta^{4}\left(p_{a} + p_{b} - \sum_{f=1}^{n} p_{f}^{\mu}\right) \mathcal{M} \, ,
+        \langle f | \mathcal{T} | i \rangle = (2\pi)^{4} \delta^{4}\left(p_{a} + p_{b} - \sum_{f=1}^{n} p_{f}\right) \mathcal{M} \, ,
     \end{equation}
@@
     particles in the initial-state, and $p_{f}$ to be the
-    momenta of the $n$ final-state particles.
+    momenta of the $n$ particles in the final-state.
     The probability of this process occurring
-    is the modulus squared of the scattering amplitude
+    is the modulus squared of the scattering amplitude\footnote{Neglecting the case of scattering with no interactions taking place.}
     \begin{equation}\label{eqn:S_prob}
@@
     of the particles being scattered, and is independent
-    of the way the experiment is carried out.
+    of the way the experiment is carried out (disregarding energy scale of experiment).
     Since the hadronic cross-section factorises
     the soft energy effects into the non-perturbative PDFs,
-    which are determined once and for all
+    which are determined once, for all
     processes, the object of interest now becomes the
@@
     \begin{equation}\label{eqn:dlips_4d}
-        \mathrm{d}\Phi_{n} = (2\pi)^{4}\delta^{4}\left(p_{a} + p_{b} - \sum_{f=1}^{n} p_{f}^{\mu}\right) \prod_{f=1}^{n} \dfrac{\mathrm{d}^{4}p_{f}}{(2\pi)^{3}}\delta(p_{f}^{2} - m_{f}^{2})\theta(E_{f}) \, ,
+        \mathrm{d}\Phi_{n} = (2\pi)^{4}\delta^{4}\left(p_{a} + p_{b} - \sum_{f=1}^{n} p_{f}\right) \prod_{f=1}^{n} \dfrac{\mathrm{d}^{4}p_{f}}{(2\pi)^{3}}\delta(p_{f}^{2} - m_{f}^{2})\theta(E_{f}) \, ,
     \end{equation}
@@
     of $|\mathcal{M}|^{2}$, this would
-    be the sum of all interference of Feynman diagrams.}
+    be the sum of all interference terms.}
     for a particular process. However, for a given
@@
     increases complexity of the computations, however,
-    because the expansion $\alpha_{s}$
-    is small, each additional term should contribute
-    less and less.
+    because $\alpha_{s}$ is small, each additional term
+    should contribute less and less.
     In principle, this means that the first terms dominate the expansion,

### 2022-12-28 appendix1.tex
for submission

@@
 \label{appendix:catani_seymour}
+    This appendix serves as supplementary material for Chapter~\ref{chapter:qcd},
+    specifically, for the tree-level matrix element factorisation in Section~\ref{sec:me_factorisation}
+    and Catani-Seymour dipoles in Section~\ref{sec:CS_dipoles}.
     \section{Notation}
@@
     The indices inside the vector denote the spin and colour indices,
-    while the subscript denotes the multiplicity of the vector.
+    while the subscript denotes the multiplicity of the matrix element.
     The matrix element squared, summed over final-state colours and spins,
@@
     Namely $n_{c}(q) = n_{c}(\bar{q}) = N_{c}$ and $n_{c}(g) = N_{c}^{2}-1$.
+
+    The colour-charge operators $\bm{T}_{i}$ act on colour space
+    to give the colour-correlated matrix element
+    \begin{equation}
+        \label{eqn:charge_operator}
+        \begin{split}
+            {}_{n} \langle 1, \ldots, n ; a, b | \bm{T}_{I} \cdot \bm{T}_{J} &| 1, \ldots, n ; a, b \rangle_{n} = \\
+            & \quad \dfrac{1}{\sqrt{n_{c}(a)n_{c}(b)}} \left[ \vphantom{\dfrac{1}{1}} \mathcal{M}_{n}^{c_{1},\ldots,c_{I},\ldots,c_{J},\ldots,c_{n}}(p_{1},\ldots,p_{n};p_{a},p_{b}) \right]^{*} \\
+            &\times \mathcal{T}^{e}_{c_{I}d_{I}}\mathcal{T}^{e}_{c_{J}d_{J}} \times \left[ \vphantom{\dfrac{1}{1}} \mathcal{M}^{d_{1},\ldots,d_{I},\ldots,d_{J},\ldots,d_{n}}(p_{1},\ldots,p_{n};p_{a},p_{b})\right] \, ,
+        \end{split}
+    \end{equation}
+    where $I$ and $J$ denotes initial- or final-state partons.
+    For a parton $I$, the matrices $\mathcal{T}^{e}_{c_{I},d_{I}}$ are defined as
+    \begin{equation}
+        \mathcal{T}^{e}_{c_{I}d_{I}} =
+        \begin{cases}
+            -i f_{cde} \hspace{-6pt} \quad \text{if} \quad I = \text{gluon} \, , \\
+            \hphantom{-} T^{e}_{cd} \quad \text{if} \quad I = \text{final-state quark or initial-state antiquark} \, , \\
+            -T^{e}_{cd} \quad \text{if} \quad I = \text{final-state antiquark or initial-state quark} \, .
+        \end{cases}
+    \end{equation}
+    Each colour vector $|1, \ldots, n; a, b \rangle$ is a colourless
+    state, meaning colour conservation can be written as
+    \begin{equation}
+        \sum_{I} \bm{T}_{I} |1, \ldots, n; a, b \rangle = 0 \, .
+    \end{equation}
+    Additionally, the colour-charge operators have the following properties
+    \begin{align}
+        \bm{T}_{I} \cdot \bm{T}_{J} &= \bm{T}_{J} \cdot \bm{T}_{I} \quad \text{for} \quad I \neq J \, , \nonumber \\
+        \bm{T}^{2}_{q} &= \bm{T}^{2}_{\bar{q}} = C_{F} \, ,\\
+        \bm{T}^{2}_{g} &= C_{A} \, . \nonumber
+    \end{align}
 \end{document}

### 2022-12-28 conclusion.tex
added intro to SM
for submission

--- 2022-08-30 qcd.tex
@@
 \begin{document}
-\chapter{Quantum chromodynamics}
-\label{chapter:qcd}
-\section{Fixed order perturbation theory}
-\subsection{NLO QCD}
-\section{Divergent structures}
-\section{Infrared factorisation}
-\subsection{Collinear limits}
-\subsection{Soft limits}
-\section{Subtraction}
-\subsection{Catani-Seymour dipoles}
-\subsection{Antenna subtraction}
+\chapter{Conclusion and outlook}
+\label{chapter:conclusion}
+In this thesis we have presented a physics-inspired
+neural network model to emulate matrix elements. The
+application of using a neural network to fit functions
+is a common scenario, however, the unique singularity
+structure of physical matrix elements precludes a naive
+approach if accuracy is desired. Indeed, the per-point
+accuracy of matrix elements is generally desired, especially
+in the infrared regions of phase-space where the emergence of
+large weights can contribute significantly to the total cross-section.
+Fortunately, matrix elements factorise in these infrared
+regions of phase-space, with well-understood explicit functional forms
+describing the single soft and collinear limits.
+By incorporating these functions into the neural network model via an
+ansatz composed of a linear combination of these singular
+functions, it is possible to transform the problem of fitting
+a potentially rapidly changing function over phase-space
+to the fitting of more well-behaved coefficients. We
+dubbed this family of models factorisation-aware
+as the neural network learns to choose the relevant
+functional behaviours in specific soft and collinear limits
+where the fitted coefficients are constrained. Outside
+of these limits, we make use of the well documented fitting capabilities
+of neural networks to form an interpolation.
+
+We first demonstrated the use of the factorisation-aware
+model in Chapter~\ref{chapter:fame1}
+where we utilised the Catani-Seymour dipole functions as
+the singular functions in the ansatz to model tree-level
+electron-positron annihilation matrix elements. It was
+shown that by building these functions into the emulator,
+there was drastic improvements in per-point accuracy compared to
+existing methods, with accuracy well below 1\% for all processes studied.
+Furthermore, the inclusion of the singular
+functions describing infrared behaviour of the matrix
+elements allowed well-behaved extrapolation of the model
+into untrained regions of phase-space which were potentially
+more singular than those seen previously during training.
+
+The success of this approach prompted the study of applying
+the factorisation-aware model to emulate the more expensive
+one-loop matrix elements for the same family of processes,
+as seen in Chapter~\ref{chapter:fame2}.
+It was found that by adapting the
+Catani-Seymour dipole functions to the more relevant antenna functions
+containing one-loop singular behaviour, the
+accurate emulation of NLO QCD k-factors was achieved.
+In this study we also demonstrated that it was possible
+for the emulator to learn the renormalisation scale dependence
+of the one-loop matrix elements, allowing it to be used to
+carry out scale variations.
+We showed that $e^{+}e^{-} \rightarrow q \bar{q} ggg$ NLO k-factors
+could be predicted with accuracy at the 1\% level, representing a speed up in
+evaluation of over 3 orders of magnitude compared to {\MadLoop} on
+a single CPU core.
+
+The study of electron-positron annihilation matrix elements
+whilst important for planned future lepton-lepton colliders,
+is not critical for making predictions for the LHC. For that,
+hadron-hadron initiated matrix elements are much more relevant.
+In Chapter~\ref{chapter:fame3}, we extended the factorisation-aware
+model to hadronic collisions by incorporating the full set
+of massless and massive dipole functions. The inclusion of these
+functions into the emulation model allowed us to model $Z+\{4,5\}$ jets
+and $t\bar{t}+\{3,4\}$ jets matrix elements at leading order, which
+represent computationally expensive and phenomenologically relevant
+processes. We found
+that the accuracy of the emulator, while lower than in the cleaner
+$e^{+}e^{-}$ environment, was not surprising given that accuracy
+scales inversely with multiplicity and these were the highest
+multiplicity processes considered in this thesis.
+However, the well-behaved
+predictions in the infrared limits and the low population tails
+makes it suitable for use as a surrogate model for event
+weights in the context of event unweighting, where stability
+of predictions is desirable.
+
+We showed that using the emulator in a novel two-stage
+unweighting procedure, where small mismatch from the full event
+weight is corrected for, could reduce the time spent in unweighting
+a sample of events. We saw that the unweighting of events
+could be accelerated by well over 2 orders of
+magnitude for the most computationally expensive processes,
+$Z+5$ jets and $t\bar{t}+4$ jets.
+
+The factorisation-aware model presented in this thesis
+has been demonstrated to be a powerful tool to emulate
+matrix elements for a variety of processes, at
+tree-level and one-loop level, owing to the
+flexibility of the modelling philosophy.
+The path forward is clear: to reliably accelerate the generation of
+simulated unit-weight events required for LHC experiments, the
+modelling procedure has to be extended to arbitrary SM
+processes, both for colour-summed, and colour-sampled matrix
+elements, with an easy-to-use interface to general
+purpose Monte Carlo event generators.
+We have taken the first steps towards this by proving the viability
+of the model in a production setting by deploying it
+within the {\Sherpa} framework to successfully accelerate the
+process of unweighting events.
+
 \end{document}

### 2022-12-28 intro.tex
for submission

@@
 particles since 2009, with each successive run increasing the centre-of-mass
-energy of the collision (7-13.6 TeV). As the experiment has
+energy of the collision, from 7 TeV in Run 1 to 13.6 TeV
+at the time of writing. As the experiment has
 operated for over a decade, there is a vast quantity of
@@
 the LHC already, and with even more expected at the High Luminosity LHC (HL-LHC),
-the rate of event generation is at risk of
-being outpaced by experimental data collection.
+the ability to generate the large simulated event samples
+required for theory predictions within the available computing
+budget presents a very real challenge.
 
@@
 the hard scattering and parton shower are derived
-from the SM and are theoretically well-motivated.
+from the SM and are perturbatively defined.
 
-The hard scattering process can be formulated as a probability
+The hard scattering cross-section can be formulated as a probability
 for particles to collide, represented by matrix elements.
@@
 algorithms, the construction of physics-inspired machine
-learning models unlocks higher levels of performance
+learning models unlocks higher levels of accuracy
 than would otherwise be achievable.
@@
 A similar philosophy is applied in Chapter~\ref{chapter:fame2}
-where next-to-leading order QCD K-factors are emulated for
+where next-to-leading order QCD k-factors are emulated for
 the same processes.
@@
 where I also explore using the
-emulator in a novel implementation to supplement traditional
-matrix element providers in the event generator {\Sherpa}.
+emulator in a novel implementation to accelerate event unweighting
+in the event generator {\Sherpa}.
 Finally, I will summarise and conclude the thesis in
@@
     giving rise to the electromagnetic and weak forces that we observe.
-    The effect of gravity is considered to be negligible on the scales
-    considered in the SM, and so it is not described.
+    The explanation of gravity in a QFT is an open problem and cannot currently
+    be included in the SM as any Lagrangian including gravity cannot be non-renormalised.
+    Fortunately, the effect of gravity is considered
+    to be negligible on the scales considered in the SM, and so it is ignored.
     Each of the three fundamental forces described by the SM is mediated
@@
     in the form of perturbative expansions in $\alpha_{s}$ (see Section~\ref{sec:xs}).
-    This is normally precluded as $\alpha_{s} > 1$ at low
-    energies.
 
@@
     where $\xi = 1$ corresponds to the Feynman-'t Hooft gauge.
+    Ghosts and antighosts are anti-commuting fields
+    introduced for each gauge field as a way to conveniently compute
+    determinants occurring in the gauge fixing procedure.
     The ghost Lagrangian is most commonly written
@@
     The ghosts are unphysical particles and so cannot appear as
-    external states but must be included in internal lines for
-    a gauge invariant theory.
+    external states. However, they must be included in internal lines for
+    a gauge invariant theory as the ghost loops cancel the unphysical
+    degrees of freedom in gluon loops.
 
@@
     we can read off from the kinetic Lagrangian where
-    we collect terms from Eqns. (\ref{eqn:L_QCD}) and (\ref{eqn:L_GF}), 
+    we collect terms from Eqs. (\ref{eqn:L_QCD}) and (\ref{eqn:L_GF}), 
     \begin{equation}\label{L_kin}
@@
     separation of low and high energy scales (so-called soft and hard).
-    Within the proton, there are quarks
+    Within the proton, there are quarks-antiquark pairs
     and gluons that are constantly absorbed and
@@
         \begin{pmatrix}
-            f_{q/h}(x,\mu_{F}) \\
-            f_{g/h}(x,\mu_{F})
+            f_{q/h}(z,\mu_{F}) \\
+            f_{g/h}(z,\mu_{F})
         \end{pmatrix} \, ,
@@
     theory.
-    Since electrons and positrons are not composite
-    particles, there is no requirement for PDFs,
-    therefore Eq. (\ref{eqn:hadronic_cs}) simplifies
-    to just the partonic cross-section, which
-    can be computed in perturbation theory.
+    Although electrons and positrons are fundamental
+    particles and not composite particles, the 
+    cross-sections in electron-positron collisions
+    have contributions from the initial-state radiation,
+    which also reduces the energy of the hard scattering.
+    One method to capture these effects is through
+    the use of process-independent structure functions \cite{Kuraev:1985hb}
+    which are analogous to PDFs. Therefore, the discussion
+    presented in this section applies to electron-positron
+    cross-sections as well.
 
-    The details of how partonic cross-sections, $\hat{\sigma}_{ab \rightarrow n}$,
+    We have seen that cross-section computations involve
+    the convolution of PDFs with partonic cross-sections,
+    $\hat{\sigma}_{ab \rightarrow n}$, which are
+    calculated in perturbation theory.
+    The details of how partonic cross-sections
     are computed is discussed in the next section.
@@
 
-    Since detectors at collider experiments
-    have limited resolution, single scattering
-    events are not measured. Instead we consider
+    An observable that can be measured at experiments is
     the cross-section, $\sigma$, as introduced in
@@
     Since the hadronic cross-section factorises
-    the soft energy effects into the non-perturbative PDFs,
+    the low energy effects into the non-perturbative PDFs,
     which are determined once, for all
@@
     in quantities such as energies and angles.
-    This is because the final-state particles
-    will be produced with a range of, for example,
-    energies and momenta, so the cross-section would
-    be vanishing for an exact momentum configuration.
+    This is because individual events at experiments
+    will be measured to be in a specific interval of
+    the differential quantity, meaning differential
+    cross-sections can be plotted as histograms. Predictions
+    can then be made on the theory side by binning simulated
+    events.
 

### 2022-12-28 ml.tex
for submission

@@
     in an attempt to disentangle these events. With these large
-    arrays, terabytes of data are harvested from
-    every collision. This presents a practical challenge for data
+    arrays, petabytes of data are recorded every year
+    across all LHC experiments.
+    This presents a practical challenge for data
     collection where machine learning has could provide a solution.
@@
     In a machine learning context, the fitting of a
-    model to training data is termed training the model.
+    model to data is termed training the model.
     Once the model is fitted, it is customary to use
@@
     \begin{equation}\label{eqn:neuron}
-        y = \phi(\mathbf{x}; \theta) = \phi(\mathbf{w}^{T}\mathbf{x} + b)
+        y = \phi(\mathbf{w}^{T}\mathbf{x} + b) = \phi_{\theta}(\mathbf{x})
     \end{equation}
@@
     by an activation function, $\phi$, which is carefully
-    chosen to perform a non-linear transformation. A neuron
-    is illustrated in Figure~\ref{fig:neuron}.
+    chosen to perform a non-linear transformation. The
+    notation $\phi_{\theta}$ represents an activation function
+    acting on $\mathbf{x}$ which has been combined with parameters $\theta$.
+    A neuron is illustrated in Figure~\ref{fig:neuron}.
 
@@
     \begin{equation}\label{eqn:nn_output}
-        f(\mathbf{x}; \theta) = \boldsymbol{\phi}^{(n)}(\ldots \boldsymbol{\phi}^{(2)}(\; \boldsymbol{\phi}^{(1)}(\mathbf{x}; \theta^{(1)})\; ; \theta^{(2)}) \; ; \theta^{(n)}) \, ,
+        % f(\mathbf{x}; \theta) = \boldsymbol{\phi}^{(n)}(\ldots \boldsymbol{\phi}^{(2)}(\; \boldsymbol{\phi}^{(1)}(\mathbf{x}; \theta^{(1)})\; ; \theta^{(2)}) \; ; \theta^{(n)}) \, ,
+        f(\mathbf{x}; \theta) = \boldsymbol{\phi}^{(n)}_{\theta}(\; \ldots \boldsymbol{\phi}^{(2)}_{\theta}\; (\boldsymbol{\phi}^{(1)}_{\theta}(\mathbf{x}) ) \; ) \, ,
     \end{equation}
-    where $\boldsymbol{\phi}^{(n)}$ and $\theta^{(n)}$
+    where $\boldsymbol{\phi}^{(n)}_{\theta}$
     denotes the outputs and parameters of layer $n$,
@@
     \begin{equation}\label{eqn:MSE}
-        L_{\mathrm{MSE}} = \dfrac{1}{N} \sum_{i=0}^{N} (g(\mathbf{x}_{i}) - y(\mathbf{x}_{i}; \theta))^{2} \, ,
+        L_{\mathrm{MSE}} = \dfrac{1}{N} \sum_{i=1}^{N} (g(\mathbf{x}_{i}) - y(\mathbf{x}_{i}; \theta))^{2} \, ,
     \end{equation}
@@
     the loss function is simply a proxy for maximising neural network
-    predictive power. Due to the training and testing datasets being finite,
+    predictive accuracy. Due to the training and testing datasets being finite,
     the generalisation of the neural network is not guaranteed for the
-    true underlying distribution from which the datasets are sampled
+    true underlying function from which the datasets are sampled
     from. This problem is generally referred to as overfitting.
@@
     gradient of the loss with respect to these parameters
-    with some step size (or learning rate) $\eta$. That is
+    with some step size (or learning rate) $\eta$. The update
+    rule is given by
     \begin{equation}\label{eqn:gradient_descent}
@@
     \subsection{Optimisation of hyperparameters}
+    \label{sec:hyperparameters}
     Another set of parameters that need to be optimised
@@
     Hyperparameters are parameters that are
-    not explicitly parameters of the model, instead they control
+    not explicitly trainable parameters of the model, instead they control
     the speed and quality of the training process.
@@
     hyperparameters is often computationally expensive for two
-    reasons. Firstly, the dimensionality of parameter space to choose from is large,
-    and so the curse of dimensionality kicks in rapidly.
+    reasons. Firstly, the dimensionality of parameter space to
+    choose from is large, meaning the number of samples required
+    to effectively explore the parameter space grows rapidly --
+    this is the so-called curse of dimensionality.
     Secondly, there is no exact a priori way to evaluate the
@@
     element generators to increase statistical power of the sample
-    by predicting on many more phase-space points. This
+    by evaluating on many more phase-space points efficiently. This
     requires the accumulated error in the surrogate model
@@
     for building an accurate surrogate model.
+
+    Related to the problem of modelling matrix elements, in recent
+    years there has been work in applying machine learning
+    methods to learn: event weights \cite{Danziger:2021eeg},
+    cross-sections \cite{Otten:2018kum,Buckley:2020bxg},
+    analytic expressions of squared amplitudes \cite{Alnuqaydan:2022ncd},
+    contour deformations for multi-loop integrals \cite{Winterhalder:2021ngy},
+    and simplifying polylogarithms \cite{Dersy:2022bym}.
 
@@
     now widely available so the acquisition of data is only
-    limited by time. Even then, this is a bottleneck only
+    limited by available computing time. Even then, this is a bottleneck only
     for the most expensive of processes as generation
@@
     the matrix element diverges.
-
-    Related to the problem of modelling matrix elements, in recent
-    years there has been work in applying machine learning
-    methods to learn: event weights \cite{Danziger:2021eeg},
-    cross-sections \cite{Otten:2018kum,Buckley:2020bxg},
-    analytic expressions of squared amplitudes \cite{Alnuqaydan:2022ncd},
-    contour deformations for multi-loop integrals \cite{Winterhalder:2021ngy},
-    and simplifying polylogarithms \cite{Dersy:2022bym}.
 
@@
     as factorisation-aware models.
-    In Chapter~\ref{chapter:fame1}, we show that by using
-    Catani-Seymour dipoles as a set of basis functions, tree-level
+    In Chapter~\ref{chapter:fame1}, we show that by forming an ansatz
+    out of a linear combination of Catani-Seymour dipoles with
+    fitted coefficients, tree-level
     matrix elements for $e^{+}e^{-} \rightarrow q \bar{q}$ + up to 3 gluons
@@
     where antenna functions replace Catani-Seymour dipoles as
-    the basis functions. Accuracy was observed to be of the
+    the ingredients in the ansatz. Accuracy was observed to be of the
     order 1\% for the 5 parton case, with good scaling
@@
     the uncertainty of the neural network prediction is to
-    bypass the usage of the neural network prediction itself
-    in production.
-    Instead it can be used as a proxy for the exact matrix
-    element in event unweighting. This was explored in
+    use the prediction in an intermediary step of a full
+    calculation, where the final prediction comes from the
+    accustomed matrix element providers. This was explored in
     Ref. \cite{Danziger:2021eeg},
@@
     In Chapter~\ref{chapter:fame3} we examine replacing
-    the original emulator in said work with a factorisation-aware
+    the original emulator in that work with a factorisation-aware
     model to study the potential gain factors with a more

### 2022-12-28 qcd.tex
for submission

@@
 
-    In this chapter we will discuss the problems that
+    In this chapter we will discuss the challenges that
     arise when evaluating these Feynman diagrams and outline
@@
     singularities. The divergences
-    associated with high energy modes are ultraviolet
+    associated with high energy modes are called ultraviolet
     (UV) divergences, and on the opposite end of the energy
@@
     \end{align}
-    where it conventional to define $Z_{1} = Z_{g}Z_{2}\sqrt{Z_{3}}$
+    where it is conventional to define $Z_{1} = Z_{g}Z_{2}\sqrt{Z_{3}}$
     such that we can set $Z_{n} = 1 + \delta_{n}$ for $n \in \{1, 2, 3\}$.
@@
     are determined by calculating $\delta_{n}$. There is freedom
-    in the choice of $\delta_{n}$ which gives rise to
+    in the choice of the finite part of $\delta_{n}$ which gives rise to
     different regularisation schemes. The minimal subtraction
@@
 
-    In an all orders calculation of an observable,
+    In an all-orders calculation of an observable,
     there would be no dependence on the renormalisation
@@
     (KLN) theorems \cite{Kinoshita:1962ur,Lee:1964is} which states
-    that at each order of perturbation theory, the virtual and real
+    that at each order of perturbation theory, the divergent parts of the real and virtual
     contributions exactly cancel out, leaving only finite corrections.
@@
     this criteria is an infrared and collinear (IRC) safe observable.
-    This leads to the concept of jets \cite{Salam:2010nqg} which are collimated
-    particles combined according to some jet algorithm.
+    This requirement of being insensitive to additional emissions
+    or arbitrarily soft emissions must also be met by jet definitions.
+    Jet definitions allow theorists and experimentalists to
+    systematically combine partons, hadrons, or energy deposits,
+    into collimated clusters called jets. Jets provide a way to relate
+    the produced hadrons back to the partons produced in the hard scattering
+    as they are broadly collimated in a similar direction.
+    For a review on jets see Ref.~\cite{Salam:2010nqg}.
 
@@
     which requires the integral to be in
-    integer dimensions. Therefore, numerical techniques
+    integer dimensions. Therefore, techniques
     have been developed to deal with these IR divergences as well. The
@@
     renormalisation scale, and $\bm{T}_{i}$ is the colour-charge
-    operator. It acts on the colour vectors in the reduced matrix element
-    to give factors of $C_{A}$ and $C_{F}$, hence this factorisation is not exact.
+    operator. These operators act on the colour space in the
+    reduced matrix element to give rise to matrices,
+    hence this factorisation is not exact. The properties
+    of these operators are given in more detail in
+    Appendix~\ref{appendix:catani_seymour}.
 
@@
     This is described by the splitting process $a \rightarrow ai + i$.
-    In this case their momenta can be decomposed as
+    In this case only momenta $p_{i}$ needs to be modified as
     \begin{align}\label{eqn:I_collinear_momenta}
@@
         p_{i}^{\mu} &= (1-x)p_{a}^{\mu} + k_{\perp}^{\mu} - \dfrac{k_{\perp}^{2}}{1-x}\dfrac{n^{\mu}}{2p_{a} \cdot n} \, , \\
-        p_{a}^{\mu} &= x p_{a}^{\mu} \, ,
         \end{split}
@@
     Similar to Eq. (\ref{eqn:soft_factorisation}) where the factorisation was
-    only true in the strict soft limit, Eqns. (\ref{eqn:F_collinear_factorisation})
+    only true in the strict soft limit, Eqs. (\ref{eqn:F_collinear_factorisation})
     and (\ref{eqn:I_collinear_factorisation}) are only
@@
 \subsection{Factorisation of colour-ordered amplitudes}\label{sec:CO_factorisation}
-    The factorisation formulae expressed in Eqns. (\ref{eqn:soft_factorisation}),
+    The factorisation formulae expressed in Eqs. (\ref{eqn:soft_factorisation}),
     (\ref{eqn:F_collinear_factorisation}), and (\ref{eqn:I_collinear_factorisation})
@@
     In general, any QCD amplitude can be
-    colour decomposed, that is the colour structure is
+    colour-decomposed, that is the colour structure is
     separated from the kinematics. Consider the process $e^{+}e^{-} \rightarrow q\bar{q} + n \; g$\footnote{For the scope of this thesis it is sufficient
@@
 
-    In Eqns. (\ref{eqn:CO_soft_factorisation}) and (\ref{eqn:CO_collinear_factorisation}),
+    In Eqs. (\ref{eqn:CO_soft_factorisation}) and (\ref{eqn:CO_collinear_factorisation}),
     the colour-ordered amplitude factorises exactly into a
@@
     It has been shown that one-loop colour-ordered amplitudes
-    also factorise in the soft and collinear limits \cite{Bern:1994zx,Bern:1998sc,Kosower:1999xi}.
+    also factorise in the soft and collinear limits \cite{Bern:1994zx,Bern:1995ix,Bern:1998sc,Kosower:1999xi}.
     At the one-loop level, there are new universal singular
@@
     structure of IR divergences and how the KLN theorem
-    necessitated the cancellation of IR divergences 
+    enforced the cancellation of IR divergences 
     arising from the virtual corrections and real-emission
@@
     pointwise approximation of $\mathcal{R}_{n+1}$ to cancel all IR
-    divergences such that $\mathcal{R}_{n+1} - \mathcal{C}_{n+1}$ is finite.
+    divergences such that $\mathcal{R}_{n+1} - \mathcal{C}_{n+1}$ is integrable finite.
     Additionally, since the pole structure in $\mathcal{V}_{n}$ is identical
@@
     necessary to construct all permutations of the dipole functions
-    since the dipoles only depends on three partons.
+    since each dipole function only depends on three partons.
 
@@
         where terms not singular in the limit $p_{i}p_{j} \rightarrow 0$
-        are omitted, and the dipole is
+        are denoted by the ellipsis, and the dipole is
         \begin{equation}\label{eqn:D_ijk}
@@
         \end{equation}
-        Eqns. (\ref{eqn:FF_mapping}) and (\ref{eqn:y_ijk})
+        Eqs. (\ref{eqn:FF_mapping}) and (\ref{eqn:y_ijk})
         are a $3 \rightarrow 2$ momenta mapping that maps
@@
         $\mathcal{D}_{ij,k}$ correctly reproduces the matrix element factorisation
-        behaviour seen in Eqns. (\ref{eqn:soft_factorisation}) and (\ref{eqn:F_collinear_factorisation}),
+        behaviour seen in Eqs. (\ref{eqn:soft_factorisation}) and (\ref{eqn:F_collinear_factorisation}),
         respectively. In particular, $\boldsymbol{V}_{ij,k}$ becomes
@@
         the matrix element in the different soft and collinear limits will be
-        reproduced by summing over all the relevant dipoles. Meaning that
-        the dipole functions can be used as a set of basis functions for
+        reproduced by summing over all the relevant dipoles, meaning that
+        the dipole functions can be used as a set of functional behaviours for
         approximating the matrix elements for arbitrary SM processes.
         This will be put into practice in Chapters \ref{chapter:fame1}
-        and \ref{chapter:fame3}. The omitted terms are non-divergent
+        and \ref{chapter:fame3}. The terms represented as ellipses in
+        the dipole factorisation formulae are non-divergent
         in all of phase-space.
@@
         treated all partons to be massless. Since quarks are massive
-        particles it is not possible in general to treat them
+        particles it is not always possible to treat them
         as massless. This is especially true for the top
@@
     The discussion in this section will be limited to the
-    situation of a massive colour-neutral boson decaying
+    situation of an off-shell colour-neutral boson decaying
     to massless QCD partons, suitable for multi-jet production
@@
     accounting for identical particles in the final-state
-    and for the presence of multiple antennae in the two
-    parton process. $I$ and $K$ are hard partons forming a colour
+    and for the presence of multiple antennae in the two-parton process.
+    $I$ and $K$ are hard partons forming a colour
     connected antenna that radiates particle $j$. The identities
@@
     to the one-loop singular functions mentioned in Section~\ref{sec:OL_factorisation}.
-    Since the one-loop factorisation formulae, Eqns. (\ref{eqn:1L_soft_factorisation})
+    Since the one-loop factorisation formulae, Eqs. (\ref{eqn:1L_soft_factorisation})
     and (\ref{eqn:1L_collinear_factorisation}), are of the form
     $|M^{(1)}_{n+1}|^{2}\rightarrow S^{(0)}|M^{(1)}_{n}|^{2} + S^{(1)}|M^{(0)}_{n}|^{2}$,
-    the term proportional to $S^(0)$, that is $X_{3}^{0}$, has to be removed
+    the term proportional to $S^{(0)}$, that is $X_{3}^{0}$, has to be removed
     in Eq. (\ref{eqn:X31}) to fulfil this requirement.
@@
     contributions from the antenna functions when trying
-    to use them as basis functions for emulating the
+    to use them as part of an approximation for the
     finite part of the one-loop matrix element.
@@
     different functions corresponding to the leading colour,
-    sub-leading colour, and closed quark look contributions.
+    sub-leading colour, and closed quark loop contributions.
     It is sufficient to consider the leading colour
@@
     sub-leading colour terms amounts to an
-    approximately 10\% effect.
+    approximately 10\% effect. (cite here)
 
@@
     limits due to being derived from physical matrix elements,
-    they can be used as a set of basis functions for approximating
-    matrix elements. The emulation of the NLO QCD K-factors
+    they can be combined in an ansatz for approximating
+    matrix elements. The emulation of the NLO QCD k-factors
     for electron-positron annihilation is explored in Chapter~\ref{chapter:fame2}.

### 2022-12-28 fame_appendex1.tex
acknowledgements

@@
 \section{Phase-space trajectories}\label{appendix:fame1_trajectory_generation}
-We can see the RAMBO algorithm as a map from the unit hypercube in some high dimension into the $n$-particle phase-space.
+We can see the {\RAMBO} algorithm as a map from the unit hypercube in some high dimension into the $n$-particle phase-space.
 A flat distribution in the hypercube maps to a flat distribution in the multi-particle phase-space.
-To generate our phase-space trajectories we pick two points at random in the unit hypercube and map the line between them using the RAMBO mapping.
+To generate our phase-space trajectories we pick two points at random in the unit hypercube and map the line between them using the {\RAMBO} mapping.
 As a result, each point on the resulting trajectory has equal probability density in phase-space.
-Any other phase-space generator that smoothly maps the unit hypercube to a multi-particle phase-space could replace RAMBO in this procedure and could lead to trajectories with very different characteristics.
+Any other phase-space generator that smoothly maps the unit hypercube to a multi-particle phase-space could replace {\RAMBO} in this procedure and could lead to trajectories with very different characteristics.
 One could for example imagine a
@@
 the end-points of the trajectories.
-We find that with RAMBO the trajectories tend not to avoid difficult phase-space configurations and are therefore a good test of the extrapolation properties of our method.
+We find that with {\RAMBO} the trajectories tend not to avoid difficult phase-space configurations and are therefore a good test of the extrapolation properties of our method.
 
@@
 
+Here we present another random phase-space trajectory similar to the one
+in Figure~\ref{fig:random_trajectory} to demonstrate that features of that
+figure are not unique.
+\begin{figure}
+    \includegraphics[width=\linewidth]{fame1/trajectory_1255.pdf}
+    \caption{Another random phase-space trajectory where the `double+single'
+    regions of phase-space are explicitly shown in blue.}
+    \label{fig:trajectory_1255}
+\end{figure}
 \end{document}

### 2022-12-28 main.tex
update main and pdf
acknowledgements

@@
 \begin{document}
-\title{Emulation of high multiplicity matrix elements for electron-positron and hadron-hadron colliders}
+\title{Machine learning high multiplicity matrix elements for electron-positron and hadron-hadron colliders}
 % \subtitle{An optional subtitle}
@@
 %
-	This is some abstract about this thesis.
+	The LHC is a large-scale particle collider experiment
+	collecting vast quantities of experimental data to study the fundamental
+	particles, and forces, of nature. Theoretical predictions made with
+	the SM can be compared with observables measured at experiments.
+	These predictions rely on the use of Monte Carlo event generators to 
+	simulate events which demand the evaluation of a matrix element.
+	For high multiplicity processes this can take up a significant
+	portion of the time spent simulating an event.
+	In this thesis, we explore the usage of machine learning
+	to accelerate the evaluation of matrix elements by introducing a
+	factorisation-aware neural network model. Matrix elements are plagued with singular
+	structures in regions of phase-space where particles become soft or collinear,
+	however, the behaviour of the matrix element in these limits are well-understood.
+	By exploiting the factorisation property of matrix elements in these limits,
+	the model can learn how to best represent the approximation
+	of the matrix elements as a linear combination of singular functions.
+	We examine the application of the model to $e^{-}e^{+}$
+	annihilation matrix elements at tree-level and one-loop level, as well
+	as to leading order $pp$ collisions where the acceleration of event generation
+	is critical for current experiments.
 %
@@
 
-	Research presented in this thesis is based on joint work
-	published in the articles:
+	Research presented in this thesis is based on joint work:
 	\begin{itemize}
@@
 066.
-		\item Chapter \ref{chapter:fame2} is based on ...
-		\item Chapter \ref{chapter:fame3} is based on ...
+		\item Chapter \ref{chapter:fame2} is based on research with Daniel Ma\^{i}tre in preparation to be published.
+		\item Chapter \ref{chapter:fame3} is based on research with Timo Janßen, Daniel Ma\^{i}tre, Steffen Schumann, and Frank Siegert in preparation to be published.
 	\end{itemize}
@@
 %
-	I would like to thank\ldots
+	The work carried out during this journey would not have been
+	possible without all the help and support from the people around me.
+	Firstly, I would like to express my gratitude to Michael Spannowsky, and
+	to my supervisor Daniel Ma\^{i}tre, who gave me a chance to undertake this
+	doctorate when nobody else did.
+	Daniel has always been supportive with his meticulous advice and constructive criticisms,
+	and I thank him for giving me the freedom to explore different avenues during our joint research projects.
+
+	All the people at the IPPP, especially Trudy and Joanne, have fostered a comfortable working
+	and social environment of which I was very glad to be a part of.
+	To all the students who have come and gone during my time at the IPPP, it was a pleasure
+	to get to know you all. Special mentions go to Alan, Asli, Dorian, Francesco, Joe, Kevin, Marian,
+	Ryan, and of course all the wonderful people in OC215 and beyond.
+	Outside of the particles, shout-out to Aidan, Cameron, Ed, Jack, and Vicky for being great friends
+	who I could always have a laugh with and chat about anything. I want to say thank you to Jack
+	in particular for being a friend I could always count on for a good time, and a serious time.
+
+	I want to express my deepest gratitude to my dear family and friends back home. Mum, dad, Annie,
+	and Tracy, Chris and Tyrone, your endless love and support has been more helpful than you can imagine
+	in completing this journey.
+
+	These past four years of my life have been tremendously formative in many ways. All the people
+	I have met, I have learned so much from, and I will cherish all
+	the memories that we shared together. I will leave Durham as a better person.
 %
 \end{acknowledgements*}
-
-\begin{epigraph*}
-%
-	This a place for some profound quote which sums up your work, feelings or
-	life.
-	\source{Some book}{A. Author}
-%
-\end{epigraph*}
-
-%\dedicationtext{This thesis is dedicated to}
-\begin{dedication*}
-%
-	Someone special
-	\also[and]{that other guy}
-%
-\end{dedication*}
 

### 2022-12-28 mc.tex
acknowledgements

@@
     to highly complex final-states with a large number
-    of particles. In a hadron collider, such as the LHC,
+    of particles. In any collider involving QCD processes,
     the particles we observe are not the constituents
@@
     numerical methods become more attractive. The current
-    prevailing method is Monte Carlo integration which
-    has no scaling with the dimensionality of the integral,
+    prevailing method is Monte Carlo integration as the
+    integration error has no scaling with the dimensionality of the integral,
     and so can be used for arbitrarily high multiplicity final-states
@@
     \begin{equation}\label{eqn:MC_integration}
-        I = \int_{V} \mathrm{d}^{d}x \; f(\mathbf{x}) \approx \langle I(f) \rangle_{x} = V \dfrac{1}{N} \sum_{i=1}^{N} f(\mathbf{x}_{i}) = V \langle f \rangle_{x} \, ,
+        I = \int_{V} \mathrm{d}\mathbf{x} \; f(\mathbf{x}) \approx I_{N} = V \dfrac{1}{N} \sum_{i=1}^{N} f(\mathbf{x}_{i}) = V \langle f \rangle_{\mathbf{x}} \, ,
     \end{equation}
     where the points $\mathbf{x}$ are randomly,
-    uniformly sampled in the volume $V$. $\langle f \rangle_{x}$
+    uniformly sampled in the volume $V$. Given a sample of $N$
+    points $\{\mathbf{x}_{1}, \ldots, \mathbf{x}_{N}\} \in V$,
+    $\langle f \rangle_{\mathbf{x}}$
     denotes the average of $f$ over these uniformly sampled
     points. The law of large numbers ensures that the
-    estimator $\langle I(f) \rangle_{x}$ will converge to
+    estimator $I_{N}$ will converge to
     the true value $I$ for $N \rightarrow \infty$. As a
@@
     $f(\mathbf{x}_{i})$ will follow a Gaussian distribution with
-    mean $\langle I(f) \rangle_{x}$,
+    mean $I_{N}$,
     meaning an error estimate on the estimator is given by
@@
     \begin{equation}\label{eqn:MC_error}
-        \sigma_{I} = V \sqrt{\dfrac{\langle f^{2} \rangle_{x} - \langle f \rangle_{x}^{2}}{N}}
+        \sigma_{I} = V \sqrt{\dfrac{\langle f^{2} \rangle_{\mathbf{x}} - \langle f \rangle_{\mathbf{x}}^{2}}{N-1}}
     \end{equation}
@@
     the random numbers to four-momenta. An example of a
-    simple integrator is the RAMBO algorithm \cite{Kleiss:1985gy,Platzer:2013esa}
-    which samples four-momenta space uniformly. Since the integrator
+    simple integrator is the {\RAMBO} algorithm \cite{Kleiss:1985gy,Platzer:2013esa}
+    which samples four-momenta space uniformly in a phase-space
+    integration. Since the integrator
     now effectively samples in momentum space, we refer to the points
@@
     reduction methods to more rapidly reduce the error
-    on the estimator.
+    on the estimator. These methods aim to reduce the numerator
+    in Eq.~(\ref{eqn:MC_error}).
 
@@
     This Jacobian, along with the PDF weights and flux factors,
-    are referred to as the phase-space weight, $J_{i}$. Taking
+    are referred to as the phase-space weight, $J_{\mathrm{PS}}$. Taking
     this phase-space weight together
@@
     \begin{equation}\label{eqn:event_weight}
-        w_{i} = |\mathcal{M}|_{i}^{2} J_{i} \, .
+        w = |\mathcal{M}|^{2} J_{\mathrm{PS}} \, .
     \end{equation}
@@
     of the event occurring at an experiment \cite{Plehn:2009nd}.
-    By generating a number of events in this manner,
-    and histogramming them, a sample of weighted events is created.
+    By generating a number of events in this manner
+    a sample of weighted events is created. Cross-sections
+    and distributions can then be calculated by histogramming
+    these weights.
 
-    At experiments, collision events are either observed or unobserved,
-    meaning each observed configuration contributes equally to
-    the measured cross-sections, and so are equally weighted.
-    Therefore, in order to generate events that
-    compare to experiments, it is necessary to translate
-    weighted events into unit-weight events.
-    This process is known as unweighting and is typically
-    carried out using a rejection sampling algorithm, as
-    outlined in Algorithm~\ref{alg:unweighting}. The rejection
-    sampling algorithm makes use of $w_{\mathrm{max}}$ which is the
-    maximal event weight in the integration volume.
+    The generation of weighted events is generally undesirable
+    as many events will have small weights with only a few events
+    making sizeable contributions to predictions. Furthermore,
+    these small weights still have to be run through expensive detector
+    simulations, making the usage of weighted events highly inefficient.
+    Therefore, it has become customary to generate events
+    with equal relative weight, so-called unweighted events.
 
-    \begin{algorithm}
+    An additional reason to generate unweighted events is that
+    experimentalists would like the probability of events occurring
+    at colliders to be described only by their frequency and not
+    with any additional weights. Since unweighted events have
+    equal relative weight, they satisfy this criteria.
+
+    The process of generating unweighted events is known as
+    unweighting and is typically carried out using a rejection
+    sampling algorithm.
+    Events are accepted or rejected by comparing the ratio
+    $w / w_{\mathrm{max}}$ to a uniform random number, $r$,
+    sampled in the interval $[0, 1]$, where $w_{\mathrm{max}}$ is
+    the maximal event weight in the integration volume.
+    Events are then accepted if $w / w_{\mathrm{max}} > r$.
+    This can be interpreted as converting the event weights
+    into a probability of being accepted or rejected as
+    small event weights will have a low probability of being
+    accepted, whereas the opposite is true for large event weights.
+    The rejection algorithm is outlined explicitly in
+    Algorithm~\ref{alg:unweighting}.
+
+    \begin{algorithm}[ht]
         \caption{Rejection sampling for unweighting events}\label{alg:unweighting}
@@
     \begin{equation}\label{eqn:unweighting_efficiency}
-        \varepsilon = \dfrac{N_{\mathrm{accepted}}}{N_{\mathrm{total}}} \, ,
+        \epsilon = \dfrac{N_{\mathrm{accepted}}}{N_{\mathrm{total}}} \, ,
     \end{equation}
@@
     was explored in Ref. \cite{Danziger:2021eeg} wherein a
-    novel two-stage unweighting procedure employing a surrogate
+    novel two-stage unweighting procedure employing a surrogate\footnote{The authors of Ref.~\cite{Danziger:2021eeg}
+    refer to their model approximating the event weight as a surrogate model, whereas in
+    this thesis we have used the term emulator. We will use these two terms interchangeably
+    henceforth.}
     model of the exact event weight was used to decrease the amount of
@@
     in NLO QCD and electroweak corrections.
-    Many of the modern matrix element generators are supplied
+    Many of the modern matrix element providers are supplied
     with a common generic interface to event generators,
@@
     of analytical expressions and their phase-space integrals.
-    They are also only accurate at high energy scales.
-    To evolve from the fixed-order
+    They are also only accurate at high energy scales. For low
+    transverse momentum emissions, large logarithms appearing at
+    every order breaks the $\alpha_{s}$ perturbative expansion.
+    To describe these kinematic regions, and to evolve from the fixed-order
     calculation to the non-perturbative regime where hadrons
@@
     The simulation of the formation of hadrons in the low-energy regime
-    of an event generator follows phenomenological, non-perturbative models.
+    of an event generator follows phenomenological models,
+    which are non-perturbative and consist of many
+    parameters fitted to data.
     These hadronisation models require the combination
@@
 
-    \section{A note on theoretical uncertainties}\label{sec:scale_variations}
+    \section{Estimation of theoretical uncertainties}\label{sec:scale_variations}
     The partonic cross-section, $\hat{\sigma}(\mu_{F}, \mu_{R})$, has
@@
     The rationale behind taking this approach is that $\delta$
-    is approximately the same order as the true missing
+    is approximately the same order of magnitude as the true missing
     higher order terms.

### 2022-12-28 ml.tex
acknowledgements

@@
 \begin{document}
-\chapter{Modern machine learning algorithms and their uses}
+\chapter{Machine learning in high energy physics}
 \label{chapter:ml}
-\section{Rise of machine learning in HEP}
+\section{Rise of machine learning in high energy physics}
     In high energy physics, where analyses

### 2022-12-28 paper1.tex
acknowledgements

@@
 \begin{document}
-\chapter{Emulation of tree-level $e^{+}e^{-}$ matrix elements}
+\chapter{Emulation of tree-level $e^{+}e^{-}$ to jets matrix elements}
 \label{chapter:fame1}
 
-\section{Introduction}
+\section{Motivation}
 \label{sec:intro}
@@
 behaviour on random trajectories in phase-space in Section~\ref{sec:trajectories}. 
-Our conclusions are presented in Section~\ref{sec:conclusion} 
-and some discussion of specific details are collected in the appendices so as not to distract from the main discussion.
+Finally, the findings of this chapter will be summarised in Section~\ref{sec:conclusion}.
+Some discussion of specific details are collected in Appendix~\ref{appendix:fame1}
+so as not to distract from the main discussion.
 
@@
 
-A neural network can be seen as a function $f(\vec{x}; \theta) = \vec{y}$, where $f: \mathbb{R}^{d} \rightarrow \mathbb{R}$ maps a 
-$d$-dimensional vector $\vec{x}$ of inputs onto the vector of outputs $\vec{y}$, and where $\theta$ are the parameters of the neural network which we aim to optimise 
-such that the outputs $\vec{y}$ of the neural network match the target as well as possible. 
+A neural network can be seen as a function $f(\mathbf{x}; \theta) = \mathbf{y}$, where $f: \mathbb{R}^{d} \rightarrow \mathbb{R}$ maps a 
+$d$-dimensional vector $\mathbf{x}$ of inputs onto the vector of outputs $\mathbf{y}$, and where $\theta$ are the parameters of the neural network which we aim to optimise 
+such that the outputs $\mathbf{y}$ of the neural network match the target as well as possible. 
 The simplest implementation of an emulator would be to take the input as the kinematic information of the phase-space point and the output to be the full matrix element. 
@@
 It is well known that in soft and collinear limits the matrix element in $(n+1)$-body phase-space factorises into a singular factor and a reduced matrix element in $n$-body phase-space \cite{Altarelli:1977zs,BASSETTO1983201} (Section~\ref{sec:me_factorisation}).
-This factorisation was used by Catani and Seymour \cite{Catani:1996vz} to construct subtraction terms for the real radiation part of a NLO calculation (Section~\ref{sec:CS_dipoles}).
+This factorisation was used by Catani and Seymour \cite{Catani:1996vz} to construct subtraction terms for the real radiation part of an NLO calculation (Section~\ref{sec:CS_dipoles}).
 They introduced a factorisation formula with universal dipoles that smoothly interpolates 
@@
 \subsection{Fitting coefficients of Catani-Seymour dipoles}\label{sec:ansatz}
-Instead of using a neural network to fit the matrix element directly, we use the dipole factorisation formula to build an ansatz of the colour and helicity summed $n+1$-body matrix element,
+Instead of using a neural network to fit the matrix element directly, we use the dipole factorisation formula to build an ansatz of the colour and helicity summed $(n+1)$-body matrix element,
 \begin{eqnarray}\label{eq:ansatz}
@@
 $C_{ijk}$ can be interpreted as the reduced matrix element in $n$-body phase-space. 
-Since the input for the $C_{ijk}$ function is the full $n+1$ phase-phase information, the neural network will also model the phase-space mappings usually introduced in the factorisation formula.
+Since the input for the $C_{ijk}$ function is the full $(n+1)$ phase-phase information, the neural network will also model the phase-space mappings usually introduced in the factorisation formula.
 A schematic diagram illustrating our ansatz is given in Figure~\ref{fig:nn_architecture}.
@@
 More detail on this is given in Section~\ref{sec:architecture}.
-The representation (\ref{eq:ansatz}) is not unique but 
+The representation in Eq. (\ref{eq:ansatz}) is not unique but 
 through appropriate training, the neural network takes advantage of the right ingredients to model 
@@
 \subsection{Data generation}\label{sec:data}
-For all multiplicities, phase-space is sampled uniformly using the RAMBO algorithm \cite{Kleiss:1985gy} with a centre-of-mass energy $\sqrt{s_{\mathrm{com}}} = 1000$ GeV.
+For all multiplicities, phase-space is sampled uniformly using the {\RAMBO} algorithm \cite{Kleiss:1985gy} with a centre-of-mass energy $\sqrt{s_{\mathrm{com}}} = 1000$ GeV.
 Phase-space points are subsequently clustered using {\FastJet} \cite{Cacciari:2011ma,noel_dawe_2021_4446849} with the $e^{+}e^{-}$ $k_{t}$ algorithm \cite{Catani:1991hj}.
@@
 This can be attributed to the fact that the coefficients $C_{ijk}$ rely heavily on the mapped momenta in $n$-body phase-space, where $y_{ij,k}$ is usually required when performing the momentum mapping, see e.g.
-\cite{Catani:1996vz}.
+Ref.~\cite{Catani:1996vz}.
 
-Following our ansatz (\ref{eq:ansatz}), we must provide spin-averaged Catani-Seymour dipoles to the neural network.
+Following our ansatz in Eq.~(\ref{eq:ansatz}), we must provide spin-averaged Catani-Seymour dipoles to the neural network.
 These dipoles are computed for all phase-space points before training begins, but they are not passed directly to the neural network as input features.
@@
 In Ref. \cite{Badger:2020uow}, the authors used 500k training samples to train their models.
-For a fair comparison of our respective methods, we follow their methodology closely by constructing our models to be as close as possible to theirs and generate training and testing data by using code from their project repository \cite{n3jet_repo}\footnote{For our main results we generate all data using our own code}.
+For a fair comparison of our respective methods, we follow their methodology closely by constructing our models to be as close as possible to theirs and generate training data by using code from their project repository \cite{n3jet_repo}\footnote{For our main results we generate all data using our own code}.
 Our model architectures will not be identical due to the difference in our methods, details of our model architecture are given in Section~\ref{sec:architecture}.
@@
 These choices of hyperparameters were not results of extensive scanning of parameter space and were chosen heuristically.
-Hyperparameters can be tuned more optimally using more sophisticated methods, but they all rely on training a large number of models which is computationally expensive and time-consuming.
+Hyperparameters can be tuned more optimally using more sophisticated methods as described in Section~\ref{sec:hyperparameters}, but they all rely on training a large number of models which is computationally expensive and time-consuming.
 
@@
 \begin{eqnarray}\label{eq:mse}
-    L_{\mathrm{MSE}} = \dfrac{1}{N} \sum_{i=1}^{N} \left(y_{i} - p(\vec{x}_{i}; \theta)\right)^{2} \, ,
+    L_{\mathrm{MSE}} = \dfrac{1}{N} \sum_{i=1}^{N} \left(y_{i} - p(\mathbf{x}_{i}; \theta)\right)^{2} \, ,
 \end{eqnarray}
-where $y_{i}$ is the target for the $i$-th sample out of $N$ samples, and $p(\vec{x}_{i}; \theta)$ is the corresponding prediction
+where $y_{i}$ is the target for the $i$-th sample out of $N$ samples, and $p(\mathbf{x}_{i}; \theta)$ is the corresponding prediction
 obtained by combining the dipole factors and the azimuthal dependency terms multiplied by their NN-learned coefficients. 
-In order to correct for the scale difference, we need to apply the transformation (\ref{eq:arcsinh}) on the neural network prediction first, with the same $S_{\mathrm{pred}}$ that was used for scaling the targets.
+In order to correct for the scale difference, we need to apply the transformation from Eq.~(\ref{eq:arcsinh}) on the neural network prediction first, with the same $S_{\mathrm{pred}}$ that was used for scaling the targets.
 We choose the mean squared error to measure the network's performance because it is sensitive to outliers in the target distribution.
@@
 In our case we would like the neural network to learn about the universal factorisation property in QCD by making it choose a minimal amount of dipoles to represent the matrix element in singular regions.
-In addition to preventing overfitting, (\ref{eq:penalty}) helps the neural network to extrapolate to more soft and collinear regions as it has learnt to choose which dipoles are relevant in specific configurations.
+In addition to preventing overfitting, Eq.~(\ref{eq:penalty}) helps the neural network to extrapolate to more soft and collinear regions as it has learnt to choose which dipoles are relevant in specific configurations.
 For non-singular configurations, the neural network is free to interpolate as there is not a clear set of dipoles that dominate, meaning $L_{\mathrm{pen}}$ is small.
@@
 constructing the neural networks at the centre of our model with a similar structure (e.g. same hidden layer structure, same activation functions, 
-same random initial weight distribution), training methodology (same {\EarlyStopping} criteria, same initial learning rate), and have generated training and testing datasets using code 
+same random initial weight distribution), training methodology (same {\EarlyStopping} criteria, same initial learning rate), and have generated training datasets using code 
 available on the authors' project repository with the relevant cuts.
+For the testing data, we thank the authors for providing the same testing set used in their publication.
 
@@
 We compare against the `single' and `ensemble' methods by training and testing using the corresponding datasets.
-For more detailed information on the differences please refer to \cite{Badger:2020uow}.
 Note that the height of the peak for the dipole histograms are not illustrated in the figure as it would not fit on the current axes but that is not important for this comparison.
@@
 The blue and orange dipole histograms representing our method
-    are cut off at the top on the main axes, but the most important feature is the narrowness of the peak centred around the ideal value.
+    are cut off at the top on the main axes, but the most important feature is the narrowness of the peak centred around the ideal value of 0.
     The insets show the detailed distribution of our result on a linear scale.}
@@
 In Figure~\ref{fig:error_500k_vs_40m}, we show the improvements in accuracy of our main NNs compared to the smaller NNs from Section~\ref{sec:n3jet_comparison}.
-Although the training and testing data for the smaller NNs are not identical due to differences in code used to generate the sets\footnote{Although RAMBO is used for phase-space generation in both works, the selection criteria is different (JADE algorithm vs $k_{t}$ algorithm).}, for this comparison it suffices to show that the larger NNs are orders of magnitude more accurate than the smaller NNs.
+Although the training and testing data for the smaller NNs are not identical due to differences in code used to generate the sets\footnote{Although {\RAMBO} is used for phase-space generation in both works, the selection criteria is different (JADE algorithm vs $k_{t}$ algorithm).}, for this comparison it suffices to show that the larger NNs are orders of magnitude more accurate than the smaller NNs.
 
@@
 We investigated 50 different random trajectories. For the discussion in this section
-we chose one that contains many interesting features, namely the matrix elements span many orders of magnitudes and there are distinct peaks in the trajectory.
+we chose one that contains many interesting features,
+namely the matrix elements span many orders of magnitudes and there are distinct peaks in the trajectory.
 In Figure~\ref{fig:random_trajectory} we show the predictions by the three 5-jet models trained on data with different global phase-space cuts for this trajectory.
@@
 Table~\ref{table:trajectory_predictions} shows the result for the regions of phase-space where training data was available, and those falling beyond the data available to the model.
-
-NN predictions are depicted as coloured scatter plots where the colour indicates the value of the minimum $s_{ij}$ between any pair of final-state particles at that phase-space point.
-To more easily visualise the extrapolation performance of the NNs we highlight the regions where the minimum $s_{ij}$ goes below the global phase-space cut applied to the training set the models were trained on,
-for each cut made. The regions of the plots where this occurs have been coloured in red.  
-With these trajectories being completely randomly selected in phase-space there is a possibility for there to be doubly singular points or worse.
-To check for this we used {\FastJet} to cluster the phase-space points in the same way we did for data generation, see Section~\ref{sec:data}.
-The pink regions indicate points which have two separate single unresolved limits, we label this configuration `Single+single'.
-The purple regions indicate points which have a double unresolved limit (i.e. three particles in one jet), we label this configuration `Double'.
-The blue regions indicate points which have both a double unresolved limit and a separate single unresolved limit, we label this configuration `Double+single'.
-We do not include the quark-anti-quark invariant in defining these regions as there is no associated infrared divergence.
-The pink, purple, and blue bands indicates regions of points which would have been discarded for our training and testing datasets.
-
-Accuracy is high when the minimum $s_{ij}$ of the trajectory is not below any $y_{\mathrm{cut}}$, i.e. when the NN prediction curve is blue.
-This is demonstrating that the NNs are interpolating well.
-Performance generally declines in the red, pink, purple, and blue regions, which is not unexpected as the NNs are extrapolating.
-Given that this trajectory has regions which go more collinear than any points the networks have been exposed to before, we would expect the networks which
-have been trained with the smallest $y_{\mathrm{cut}}$ parameter to perform best.
-We see that this is the case as accuracy is acceptable in the $y_{\mathrm{cut}} = 0.0001$ models, including in the regions where the NN is extrapolating.
 
@@
     Purple bands indicate double unresolved limits, i.e. three particles in one jet.
-    Blue bands indicate where there is both a double and a single unresolved limit.
-    The pink, purple, and blue bands represent regions of phase-space which would have been excluded by {\FastJet}.}
+    % Blue bands indicate where there is both a double and a single unresolved limit.
+    The pink, and purple bands represent regions of phase-space which would have been excluded by {\FastJet}.}
     \label{fig:random_trajectory}
 \end{figure}
+
+NN predictions are depicted as coloured scatter plots where the colour indicates the value of the minimum $s_{ij}$ between any pair of final-state particles at that phase-space point.
+To more easily visualise the extrapolation performance of the NNs we highlight the regions where the minimum $s_{ij}$ goes below the global phase-space cut applied to the training set the models were trained on,
+for each cut made. The regions of the plots where this occurs have been coloured in red.  
+With these trajectories being completely randomly selected in phase-space there is a possibility for there to be doubly singular points or worse.
+To check for this we used {\FastJet} to cluster the phase-space points in the same way we did for data generation, see Section~\ref{sec:data}.
+The pink regions indicate points which have two separate single unresolved limits, we label this configuration `Single+single'.
+The purple regions indicate points which have a double unresolved limit (i.e. three particles in one jet), we label this configuration `Double'.
+Although not seen in Figure~\ref{fig:random_trajectory}, there are points which have both a double unresolved limit and a separate single unresolved limit,
+we label this configuration `Double+single' and indicate is as a blue region\footnote{Another random phase-space trajectory is presented in Appendix~\ref{appendix:fame1_trajectory_generation} which contains these `double+single' regions.}.
+We do not include the quark-anti-quark invariant in defining these regions as there is no associated infrared divergence.
+The pink, purple, and blue bands indicates regions of points which would have been discarded for our training and testing datasets.
+
+Accuracy is high when the minimum $s_{ij}$ of the trajectory is not below any $y_{\mathrm{cut}}$, i.e. when the NN prediction curve is blue.
+This is demonstrating that the NNs are interpolating well.
+Performance generally declines in the red, pink, purple, and blue regions, which is not unexpected as the NNs are extrapolating.
+Given that this trajectory has regions which go more collinear than any points the networks have been exposed to before, we would expect the networks which
+have been trained with the smallest $y_{\mathrm{cut}}$ parameter to perform best.
+We see that this is the case as accuracy is acceptable in the $y_{\mathrm{cut}} = 0.0001$ models, including in the regions where the NN is extrapolating.
 
@@
     We present the percentage of points that lie outside 0.1\%/1.0\%/5.0\% errors.
-    Fraction of points indicates the percentage of points that lie in the region of interest, out of all phase-space points from the 50 trajectories we examine.
+    Fraction of points indicates the percentage of points that lie in the region of interest, out of all phase-space points from the 50 random trajectories we examined.
     }

### 2022-12-28 paper2.tex
initial draft of chapters
acknowledgements

--- 2022-08-30 ml.tex
@@
 \begin{document}
-\chapter{Machine learning}
-\label{chapter:ml}
-\section{Introduction to machine learning}
-\subsection{Supervised and unsupervised learning}
-\section{Neural networks}
-\subsection{Perceptron}
-\subsection{Densely connected neural networks}
-\subsection{Optimisation of weights}
-\subsection{ML applications in HEP}
+\chapter{Emulation of $e^{+}e^{-}$ to hadrons NLO k-factors}
+\label{chapter:fame2}
+
+\section{Motivation}
+In the previous chapter we described in detail the strategy for emulating
+tree-level $e^{+}e^{-}$ annihilation matrix elements by exploiting the factorisation
+property of matrix elements. The inclusion of the dipole functions in the ansatz
+meant that the infrared divergent behaviour did not have
+to be learnt solely by the neural network itself, instead the task was reduced
+to learning the more well-behaved coefficients of the dipole functions.
+
+A natural extension to the work presented in the previous chapter is
+to examine the emulation of the one-loop matrix elements for the same family of processes.
+In general, loop matrix elements are computationally much more expensive than
+their tree-level counterparts, meaning that the potential time saved by using
+a fast and accurate emulator over a more traditional one-loop provider is much higher
+than in the tree-level case.
+In this chapter we explore this avenue by using a similar procedure to Chapter~\ref{chapter:fame1}
+to construct a factorisation-aware model in order to emulate NLO QCD k-factors.
+
+The layout for this chapter is as follows. In Section~\ref{sec:1L_fitting} we detail
+the fitting procedure. Namely, we briefly recap antenna functions, as well as describe
+factorisation of matrix elements within the antenna function framework in order to
+arrive at an ansatz of the NLO k-factor. The construction of the neural network emulator
+is described in Section~\ref{sec:1L_nn}, along with the process of training, testing and
+deploying the model. In Section~\ref{sec:1L_results} we present the results of the
+emulator demonstrating its speed and accuracy, where we see significant speed gains
+whilst keeping the accuracy to the 1\% level. We conclude
+this chapter by summarising in Section~\ref{sec:1L_conclusion}.
+
+\section{Fitting framework for one-loop matrix element}\label{sec:1L_fitting}
+
+In this chapter we consider the emulation of one-loop 
+$e^+e^-\rightarrow Z / \gamma^{*} \rightarrow q\bar q +n_{g} g$
+matrix elements for $n_{g}$ up to and including 3, which corresponds
+to events with up to 5 partons. We denote the number of final-state
+partons as $n$. Instead of emulating the matrix element itself we build
+a surrogate for the related so-called $k$-factor
+%More specifically, we consider the emulation
+%of the one-loop k-factor
+\begin{equation}
+    k_{n} = \dfrac{2 \Re \left\{ \mathcal{M}^{(n, 0)} \mathcal{M}^{(n, 1)\, *} \right\} }{ |\mathcal{M}^{(n, 0)}|^{2} } \equiv \dfrac{|\mathcal{M}^{(n, 1)}|^{2}}{|\mathcal{M}^{(n, 0)}|^{2}} \, ,
+    \label{eq:k_factor}
+\end{equation}
+where $\mathcal{M}^{(n, \ell)}$ denotes the amplitude
+for a process with $n$ final-state partons, at loop-order
+$\ell$. We will refer to the interference term in the numerator
+as the one-loop matrix element henceforth and introduce
+this notation for brevity.
+The sum/averaging over colour and helicity is implicit for
+both the one-loop matrix element and the tree-level
+matrix element. The numerator in Eq. (\ref{eq:k_factor}) is the
+finite part of the interference between the tree-level and 
+one-loop level matrix element, where the conventional dimensional
+regularisation (CDR) scheme is used.
+
+We choose to emulate the k-factor instead of the one-loop matrix
+element directly because the division of the tree-level matrix
+element cancels the infrared divergences occurring for soft
+and collinear external particles. However, there are still
+logarithmic divergences that remain from the loop integral. Another 
+advantage is that the scale of the k-factors is naturally of
+the order unity, making it more amenable for emulation.
+
+In the following, we describe how we apply the same approach as the
+factorisation-aware formalism introduced in Chapter~\ref{chapter:fame1}
+to encapsulate the more complex structure of the one-loop
+matrix element to construct an accurate emulator for the
+one-loop k-factors that is robust against single collinear or soft divergences.
+
+An additional complication with one-loop matrix elements
+is that they are evaluated at a given renormalisation scale.
+This dependence can be derived from first principle, but we choose
+to instead incorporate this dependence into our neural network emulator
+as an input.
+This method, so-called parametric
+neural networks, has been utilised in other contexts
+\cite{Baldi:2016fzo,Ghosh:2021roe}.
+
+\subsection{Antenna functions}\label{sec:1L_antenna}
+In building an emulator for tree-level matrix elements in Chapter~\ref{chapter:fame1},
+Catani-Seymour dipoles \cite{Catani:1996vz} are sufficient
+to explain all single divergences arising in phase-space. For
+one-loop matrix elements we utilise antenna functions (see Section~\ref{sec:antenna_functions})
+which fulfil a similar purpose of providing a set of functional behaviours
+to build the matrix element out of.
+
+Antenna functions as given in Ref.~\cite{Gehrmann-DeRidder:2005btv}
+are derived from physical colour-ordered matrix elements and
+by construction have the correct infrared behaviour when specific sets of
+particles become unresolved. For our purposes we require the set of 
+antenna functions describing the scenario of one particle
+becoming unresolved at tree-level and one-loop level. Namely,
+following the notation of Ref.~\cite{Gehrmann-DeRidder:2005btv},
+these are the leading colour three-parton antenna functions
+$X_{ijk}^{0}$ and $X_{ijk}^{1}$, respectively. $X_{ijk}^{0}$
+describes all configurations where parton $j$ becomes unresolved,
+where $i$ and $k$ are the hard partons. The one-loop counterpart,
+$X_{ijk}^{1}$, correctly reproduces the single soft and collinear
+singularity structure appearing in the one-loop singular functions
+(see Section~\ref{sec:OL_factorisation}).
+From this description it is clear that the antenna functions
+depend on the momenta in the full $(n+1)$-body phase-space.
+
+In one antenna function, there are two hard partons
+which can both radiate off one unresolved parton. This is in
+contrast to the dipole function which only has an unresolved
+parton emitting from one parton. In that sense, a single antenna
+function is a linear combination of two dipole functions where
+the emitter and spectator are swapped. The advantage of this is
+that there are generally fewer antenna functions to consider,
+especially when the multiplicity increases.
+
+Although the one-loop matrix elements data that we use for fitting are not
+colour-ordered, the antenna functions nevertheless provide
+a set of useful functions that allow the neural network emulator
+to form accurate approximations of the one-loop k-factor.
+
+Since we are emulating the finite part of the one-loop matrix
+element we need to take care to extract all the finite parts
+from the one-loop antenna functions $X_{ijk}^{1}$. The
+full expression for the one-loop antenna we use is given as
+\begin{equation}
+    X_{ijk}^{1, F}  = \mathcal{F}inite(X_{ijk}^{1}) + \dfrac{11}{6}\log\left(\dfrac{\mu_{R}^{2}}{s_{ijk}} \right) X_{ijk}^{0} + \mathcal{F}(\mathbf{I}_{ij}^{(1)}(\epsilon, s_{ij})) X_{ijk}^{0} \, .
+    \label{eq:finite_X31}
+\end{equation}
+where the superscript $F$ in $X_{ijk}^{1,F}$ denotes the
+one-loop antenna function with all finite parts extracted\footnote{Henceforth we refer to $X_{ijk}^{1,F}$ as the
+one-loop antenna function, unless explicitly stated otherwise.}.
+Most of the finite parts of the antenna function are extracted in the term
+$\mathcal{F}inite(X_{ijk}^{1})$. The second term
+adjusts the renormalisation scale of the antenna function from
+the invariant mass of the antenna partons, $s_{ijk} = s_{ij} + s_{ik} + s_{jk}$,
+to the renormalisation scale the one-loop matrix element is
+evaluated at, $\mu_{R}^{2}$. The final term extracts the
+remaining finite parts from the singularity operators,
+where their expressions are explicitly given below for
+different partonic splittings.
+The finite part of the singularity operators can be written as
+\begin{equation}
+    \mathcal{F}(\mathbf{I}^{(1)}_{ij}(\epsilon, s_{ij})) = \epsilon_{0} + \epsilon_{1} \Re{(z)} + \dfrac{1}{2}\epsilon_{2} \Re{(z^{2})} \, ,
+\end{equation}
+where
+\begin{equation}
+    z = \mathrm{zlog}(\mu_{R}^{2}) - \mathrm{zlog}(-s_{ij})
+\end{equation}
+and coefficients are given as
+\begin{align}
+    \epsilon_{0} &= \dfrac{\pi^{2}}{24} \, , \nonumber \\
+    \epsilon_{1} &=
+    \begin{cases}
+        -\frac{5}{6}, & \text{if } ij = qg \; \text{or} \; gq \\
+        -\frac{3}{4}, & \text{if } ij = qq \\
+        -\frac{11}{12}, & \text{if } ij = gg 
+    \end{cases} \\
+    \epsilon_{2} &= -\dfrac{1}{2} \, . \nonumber
+\end{align}
+The $\mathrm{zlog}$ function extends the logarithm for all real-values
+\begin{align}
+    \mathrm{zlog}(x) = 
+    \begin{cases}
+        \log(x), & \text{if } x \geq 0 \\
+        \log(|x|) - i \pi , & \text{otherwise} \, .
+    \end{cases}
+\end{align}
+
+\subsection{Factorisation of matrix elements}\label{sec:1L_me_factorisation}
+In the following, we recap the factorisation of matrix
+elements as seen in Sections~\ref{sec:me_factorisation} and \ref{sec:OL_factorisation}
+within the antenna function framework.
+
+Tree-level matrix elements in $(n+1)$-body phase-space
+can be factorised in the single soft and collinear limits as
+\begin{equation}
+    |\mathcal{M}^{(n+1, 0)}|^{2} \longrightarrow X_{ijk}^{0} |\mathcal{M}^{(n, 0)}|^{2} \, ,
+    \label{eq:tree_factorisation}
+\end{equation}
+where $|\mathcal{M}^{(n, 0)}|^{2}$ is the reduced matrix element
+in $n$-body phase-space and $X_{ijk}^{0}$ is the three-parton tree-level
+antenna function
+introduced in Section~\ref{sec:antenna_functions}.
+The one-loop matrix element similarly
+exhibits factorisation in the soft and collinear limits.
+This has been extensively studied \cite{Bern:1994zx,Bern:1995ix,Bern:1999ry,Kosower:1999xi,Kosower:1997zr}
+with the splitting kernels computed \cite{Kosower:1997zr,Kosower:2002su,Gehrmann-DeRidder:2005btv}.
+Schematically, in the single soft and collinear limits, the one-loop
+matrix element can be deconstructed into
+\begin{equation}
+    |\mathcal{M}^{(n+1, 1)}|^{2} \longrightarrow X_{ijk}^{0} |\mathcal{M}^{(n, 1)}|^{2} + X_{ijk}^{1,F} |\mathcal{M}^{(n, 0)}|^{2} \, ,
+    \label{eq:ol_factorisation}
+\end{equation}
+where $X_{ijk}^{1,F}$ is the three-parton one-loop antenna function.
+This equation can be thought of as a tree-level splitting
+kernel multiplied by a one-loop reduced matrix element,
+plus a one-loop splitting kernel multiplied by a tree-level
+reduced matrix element. This is illustrated
+pictorially in Figure~\ref{fig:1L_ol_factorisation}.
+\begin{figure}
+    \includegraphics[width=\linewidth]{fame2/oneloop_factorisation.pdf}
+    \caption{Diagram illustrating factorisation of one-loop
+    matrix element.
+    In a singly unresolved limit, the $(n+1)$-body one-loop matrix
+    element tends to the sum of a $n$-body one-loop matrix element
+    multiplied by a tree-level splitting kernel and an $n$-body tree-level
+    matrix element multiplied by a one-loop splitting kernel.
+    The tree-level elements are drawn as a fully filled in
+    circle, while the one-loop elements are drawn as a donut shape.}
+    \label{fig:1L_ol_factorisation}
+\end{figure}
+
+\subsection{Ansatz for the k-factor}\label{sec:1L_ansatz}
+Given that both the tree-level and one-loop level matrix
+element factorise in the soft and collinear limits,
+we can rewrite the $(n+1)$-body k-factor in these limits as
+\begin{align}
+    k_{n+1} &\longrightarrow \dfrac{X_{ijk}^{0} |\mathcal{M}^{(n, 1)}|^{2} + X_{ijk}^{1,F} |\mathcal{M}^{(n, 0)}|^{2}}{X_{ijk}^{0} |\mathcal{M}^{(n, 0)}|^{2}} \nonumber \\
+    k_{n+1} &\longrightarrow \dfrac{|\mathcal{M}^{(n, 1)}|^{2}}{|\mathcal{M}^{(n, 0)}|^{2}} + \dfrac{X_{ijk}^{1,F}}{X_{ijk}^{0}} \\
+    k_{n+1} &\longrightarrow k_{n} + \dfrac{X_{ijk}^{1,F}}{X_{ijk}^{0}} \nonumber \, ,
+    \label{eq:k_factorisation}
+\end{align}
+where we see that the k-factor tends to a sum
+of the reduced k-factor, $k_{n}$, and a ratio
+of antenna functions. By summing over ratios of
+antenna functions for all limits of a
+given process, we can construct an ansatz for the
+k-factor over all of phase-space.
+
+This informs our ansatz for the k-factor, which can is given as
+\begin{equation}
+    k_{n+1} = C_{0} + \sum_{\{ijk\}} C_{ijk}\dfrac{X_{ijk}^{1,F}}{X_{ijk}^{0}} \,
+    \label{eq:1L_ansatz}
+\end{equation}
+where $C_{0}$ and $C_{ijk}$ are coefficients fitted by the neural network.
+$C_{0}$ is an additive term aiming to model the reduced
+k-factor, and $C_{ijk}$ are multiplying the ratio of
+antenna functions to fit the single collinear, and
+soft limits in multiple regions of phase-space.
+The sum over $\{ijk\}$ denotes the sum over
+the relevant permutations of final-state partons, accounting
+for all the relevant singular limits of the k-factor.
+This sum allows the neural network to make use of all
+the provided antenna functions to make an approximation
+of the colour-summed matrix element.
+The full set of antenna functions which we implement into our
+model is detailed below.
+
+The antenna functions are generically written as $X_{ijk}^{\ell}$,
+which in practice is replaced with specific antennae containing either
+$qg\bar{q}$, $qgg$ ($\bar{q}gg$), or $ggg$, which are
+referred to as $A$, $D$, and $F$ antennae, respectively. The antennae
+listed in Table~\ref{table:1L_antenna_functions} are sufficient to
+describe all infrared singularities in the partonic
+processes we consider.
+\begin{table}
+    \centering
+    \begin{tabular}{cccc}
+        \toprule
+        $n$ & Tree-level antenna & One-loop antenna & {\{i,j,k\}} permutations \\
+        \midrule
+        3 $(q \bar{q} g)$     & $A_{3}^{0}(q, g, \bar{q})$ & $A_{3}^{1}(q, g, \bar{q})$  & (1, 3, 2) \\
+        \midrule
+        4 $(q \bar{q} g g)$   & \makecell{$A_{3}^{0}(q, g, \bar{q})$ \\ $D_{3}^{0}(q, g, g)$} & \makecell{$A_{3}^{1}(q, g, \bar{q})$ \\ $D_{3}^{1}(q, g, g)$} & \makecell{(1, 3, 2), (1, 4, 2) \\ (1, 3, 4), (2, 3, 4)} \\
+        \midrule
+        5 $(q \bar{q} g g g)$ & \makecell{$A_{3}^{0}(q, g, \bar{q})$ \\ $D_{3}^{0}(q, g, g)$ \\ $F_{3}^{0}(g, g, g)$} & \makecell{$A_{3}^{1}(q, g, \bar{q})$ \\ $D_{3}^{1}(q, g, g)$ \\ $F_{3}^{1}(g, g, g)$} & \makecell{(1, 3, 2), (1, 4, 2), (1, 5, 2) \\ (1, 3, 4), (1, 3, 5), (1, 4, 5), \\ (2, 3, 4), (2, 3, 5), (2, 4, 5), \\ (3, 4, 5)} \\
+        \bottomrule
+    \end{tabular}
+    \caption{List of antenna functions we use for each process,
+    and the full list of {\{i,j,k\}} permutations, where $q=1$,
+    $\bar{q}=2$, $g=3,4,5$.}
+    \label{table:1L_antenna_functions}
+\end{table}
+
+Since the k-factor has infrared divergences arising from
+unresolved partons in the final-state being removed, and with
+the appropriate antenna functions accounting for the
+logarithmic divergences from the loop momenta, the
+challenging task of fitting a rapidly varying function
+over phase-space is reduced to fitting a group of
+well-behaved coefficients that dictate how to suitably utilise
+the antenna functions.
+
+\section{Building the neural network emulator}\label{sec:1L_nn}
+\subsection*{Dataset generation}
+Phase-space is sampled uniformly using the {\RAMBO} algorithm
+\cite{Kleiss:1985gy} with a centre-of-mass energy
+$\sqrt{s_{\mathrm{com}}} = 1000$ GeV. The global phase-space
+generation cut is set to $y_{\mathrm{cut}} = 0.0001$.
+%We have seen that the factorisation-aware method is
+%relatively agnostic to this choice of cut \cite{Maitre:2021uaa}.
+We have shown in \cite{Maitre:2021uaa} that the accuracy of the emulation
+is not greatly affected by the generation cut, but the extrapolation
+performance is increased with a more inclusive cut, so we have chosen this
+value. 
+{\FastJet} \cite{Cacciari:2011ma,noel_dawe_2021_4446849} is
+used to exclusively ($d_{\mathrm{cut}} = 0.01 \times s_{\mathrm{com}}$)
+cluster final-state jets with the $e^{+}e^{-}$ $k_{t}$
+algorithm such that there is at most a single unresolved parton.
+
+For each phase-space point generated, we sample a
+renormalisation scale, $\log(\mu_{R})$, from a uniform distribution
+with end points at $[\log(s_{\mathrm{com}}/4), \log(4s_{\mathrm{com}})]$.
+In other words, we sample the renormalisation scale logarithmically.
+We observe that the neural network manages to learn the
+renormalisation scale dependence well therefore opt to
+sample $\mu_{R}$ in a wider range than is usually used
+for the conventional scale variations which varies $\mu_{R}$
+up and down by factors 2.
+
+Generated phase-space points are fed to {\MadGraph} \cite{Alwall:2014hca,Hirschi:2011pa}
+to compute the tree-level and one-loop level matrix elements.
+For each phase-space point, we use the corresponding
+renormalisation scale sampled and evaluate the strong coupling
+constant at this scale using the NNPDF-4.0 NNLO PDF set
+\cite{NNPDF:2021njg} with the LHAPDF6 interface \cite{Buckley:2014ana}.
+All external particles are treated as massless and $m_{Z} = 91.188$ GeV.
+
+We generate 1100k data points in total, using 100k points for
+training and validation, leaving 1 million points
+for independent evaluation of model performance.\
+Training on a limited dataset is a realistic scenario
+for processes which are prohibitively expensive, and we show
+that it is possible to build an accurate emulator with the relatively
+small number of data points. Note that because we are
+sampling $\mu_{R}$ along with phase-space simultaneously,
+there is an extra dimension in the sampled space. This means that the 100k
+training points we have are not comparable to 100k training points
+if we had not sampled over $\mu_{R}$. In practice, we have found that
+sampling over $\mu_{R}$ has a small impact on accuracy but
+opt to go this route to have the flexibility to predict over
+a range of $\mu_{R}$.
+
+\subsection*{Inputs to model}
+As inputs to the neural network we provide the 4-momenta of all
+final-state partons. The renormalisation scale enters the network
+as $\log(\mu_{R})$ as we expect the dependence on $\mu_{R}$
+to be in the form of a logarithm. Following Ref.~\cite{Maitre:2021uaa}
+we include the phase-space mapping variables to aid the network in learning
+the reduced matrix element information. Namely, we include as
+inputs $r$ and $\rho$ from Ref.~\cite{Kosower:1997zr}
+\begin{align}
+    r_{ijk} &= \dfrac{s_{jk}}{s_{ij} + s_{jk}} \, , \\
+    \rho_{ijk} &= \sqrt{1 + \dfrac{4r(1-r)s_{ij}s_{jk}}{s_{ijk}s_{ik}}} \, ,
+    \label{eq:map_variables}
+\end{align}
+where $i$ and $k$ are the hard radiating partons, and $j$ is
+the unresolved parton. We include the subscript
+$_{ijk}$ on $r$ and $\rho$ to represent the explicit
+dependence on the specific set of momenta used to calculate them.
+To improve training, we transform these variables as
+$r \rightarrow \log(r)$ and $\rho \rightarrow \log(\rho - (1 - \varepsilon))$,
+where $\varepsilon$ is a small constant added to improve the
+numerical stability. It is taken to be $\varepsilon = 10^{-8}$.
+An additional input that we have observed to increase
+accuracy of the emulator are the Mandelstam invariants.
+These are fed into the model as $\log(s_{ij})$ for all
+pairs of final-state particles. All inputs are
+standardised to zero mean and unit variance.
+
+\subsection*{Outputs of model}
+The outputs of the neural network are the fitted coefficients $C_{0}$ and $C_{ijk}$
+in Eq. (\ref{eq:1L_ansatz}), which when combined with the antenna functions
+produces an approximation of the k-factor. We then recover the
+one-loop matrix element by multiplying by the corresponding
+tree-level matrix element. We do not provide the tree-level matrix
+element in the emulator as the evaluation time is generally
+much lower than that of the one-loop matrix element, and it is usual 
+for one-loop matrix elements to be evaluated at phase-space point sets
+where the tree-level matrix element has already been unweighted, so that 
+only the k-factor is required.
+
+To train the network we compare the target k-factors from {\MadGraph}
+to the predictions from the neural network in the loss function by
+providing the network with the antenna functions.
+The target distribution is standardised to zero
+mean and unit variance. Since the k-factors are of order unity
+we do not need to do any additional pre-processing to aid the network
+in training.
+
+\subsection*{Neural network architecture}
+\begin{figure}
+    \includegraphics[width=\linewidth]{fame2/nn_architecture.pdf}
+    \caption{A schematic diagram of the neural network emulator.
+    The emulator is a dense neural network with inputs:
+    phase-space points, $p$, momenta mapping variables,
+    $r_{ijk}$ and $\rho_{ijk}$, kinematic invariants, $s_{ij}$,
+    and renormalisation scale $\mu_{R}$.
+    The outputs of the network are the fitted coefficients, $C_0$ and $C_{ijk}$,
+    as given in Eq. (\ref{eq:1L_ansatz}).}
+    \label{fig:1L_nn}
+\end{figure}
+A schematic of the neural network model is given in Figure~\ref{fig:1L_nn}.
+We build the neural network emulator with {\Keras}
+\cite{chollet2015keras} and {\TensorFlow} \cite{tensorflow2015-whitepaper}.
+The emulator is a dense neural network with
+three hidden layers of 64 nodes each. This network
+size was chosen with consideration given to the number
+of training samples and to reduce the discrepancy between training
+and validation loss
+(i.e. a larger network is more prone to overfitting on the
+training set if there is insufficient data to fit the additional weights).
+For the remaining hyperparameters we summarise the network
+architecture in Table~\ref{table:parameters}.
+\begin{table}
+    \caption{Hyperparameters of the neural network and their values.}
+    \begin{center}
+        \begin{tabular}{ll}
+            \toprule
+            Parameter & Value \\
+            \midrule
+            Hidden layers & 3 \\
+            Nodes in hidden layers & [64, 64, 64] \\
+            Activation function & $\mathrm{swish}$ \cite{Hendrycks2016BridgingNA,DBLP:journals/corr/abs-1710-05941}\\
+            Weight initialiser & Glorot uniform \cite{pmlr-v9-glorot10a} \\
+            Loss function & MAE (k-factor), MSE (one-loop matrix element) \\
+            Batch size & 256 \\
+            Optimiser & Adam \cite{Kingma:2014vow} \\
+            Learning rate & $10^{-3}$ \\
+            Callbacks & \makecell[l]{{\EarlyStopping}, {\RatioEarlyStopping} \\ {\ReduceLROnPlateau}} \\
+            \bottomrule
+        \end{tabular}
+    \end{center}
+    \label{table:parameters}
+\end{table}
+
+We choose to use the mean absolute error (MAE) as the loss
+function for training because it is precisely the error measure we
+would like to minimise.
+The error for one prediction is given as
+\begin{equation}
+    k_{\mathrm{true}} - k_{\mathrm{pred}} = \dfrac{|\mathcal{M}^{(n, 1)}|^{2}_{\mathrm{true}} - |\mathcal{M}^{(n, 1)}|^{2}_{\mathrm{pred}}}{|\mathcal{M}^{(n, 0)}|^{2}_{\mathrm{true}}} = \Delta \, ,
+    \label{eq:mae}
+\end{equation}
+where the error in the one-loop matrix element normalised
+by the tree-level matrix element is what we want the
+neural network to minimise.
+Since k-factors are a ratio of matrix elements, the numerical
+values it can take are not unique for a given value of the
+tree-level matrix element and/or one-loop matrix element.
+For example, for two similar values of the k-factor, the scales
+of the matrix elements going into each ratio may be vastly different.
+To ensure that the network remains accurate for large values
+of the tree-level matrix element, where corresponding corrections
+contribute more to the total cross-section, we weight the training
+points by
+\begin{equation}
+    w_{i} = \log\left( \dfrac{|\mathcal{M}^{(n, 0)}|^{2}_{i}}{\min(|\mathcal{M}^{(n, 0)}|^{2}_{i})} \right) \, ,
+\end{equation}
+where the $i$ index denotes training samples.
+% as this minimises the total error in the combination
+% $|\mathcal{M}^{(n,0)}|^{2} + |\mathcal{M}^{(n,1)}|^{2}$.
+Although we use the MAE on the k-factors as the training loss,
+we terminate model training based on the one-loop matrix
+element accuracy. This is done by monitoring the mean squared
+error (MSE) between the model prediction and corresponding truth value
+at the end of each training epoch.
+This takes advantage of the compact k-factor distributions
+for training purposes, but bases model selection on the
+accuracy for the physical one-loop matrix elements.
+
+To reduce the effects of overfitting we have two {\EarlyStopping}
+criteria: the first is to stop training once the validation loss has
+not improved in 100 epochs, and the second is a {\RatioEarlyStopping}
+which terminates training if the ratio of training loss to validation
+loss drops below a certain threshold. We take this threshold
+to be $0.9$. We also use the {\ReduceLROnPlateau}
+callback as a way to adapt the learning rate during training.
+The learning rate is reduced by a factor of 0.7 whenever
+validation loss plateaus with a patience of 20 epochs. We find
+that with these hyperparameters we achieve a balance of reducing
+overfitting and quick training times. On average the models
+train in approximately 20 minutes on an Nvidia P100 GPU.
+
+Although we build and train our model using {\TensorFlow},
+we deploy the model using the Open Neural Network Exchange ({\ONNX})
+runtime \cite{onnxruntime}
+with the {\CUDA} execution provider to run predictions on an
+Nvidia P100 GPU. With the optimised operations in the {\ONNX}
+runtime, we see that compared to {\TensorFlow} the model inference
+time is reduced by an order of magnitude or more. Another advantage
+is that it gives users the flexibility to move the pipeline
+away from {\TensorFlow} on $\mathtt{Python}$ to a more generic interface
+to the neural network model. One example would be to integrate
+the {\ONNX} model into a \texttt{C++} workflow for use with current event
+generators to replace the one-loop provider with a neural
+network emulator.
+
+In addition to the {\CUDA} execution provider, we will use
+the {\ONNX} runtime CPU execution provider to compare with
+{\MadGraph} for a comparison of single CPU core performance.
+This will be the closest to a real world benchmark as event
+generators typically generate events on a single core.
+
+Although we find that the neural network models converge well,
+to account for stochasticity in the training, the random
+initialisation of model parameters, and to reduce
+variance on predictions, we initialise 20 models for training
+and use the mean of these models as our model prediction. This
+ensembling will also give a measure of the uncertainty
+due to the neural network optimisation, by using the standard
+deviation of the 20 replica model predictions.
+
+We plot in Figure~\ref{fig:1L_loss_curves} the losses of the 20 replica
+models for the 5 jet process, where we have plotted the loss for the
+k-factor (MAE) and the one-loop matrix element (MSE).
+We can see that the models have all
+converged to a similar point when training is terminated. The
+noise in the validation loss at the beginning of training can be
+attributed to the fitting of the coefficients: when the model is
+learning how to pick the relevant combination of antenna functions
+there can be large variations in the prediction, however, the
+variations become much smaller once the model learns the
+factorisation properties and converges. We can
+see the variations in the validation loss are small
+at the end of training, and that there is not a large discrepancy
+between the training and validation losses, as enforced by the
+{\RatioEarlyStopping} callback.
+
+\begin{figure}
+    \centering
+    \includegraphics[scale=0.57]{fame2/loss_curves.pdf}
+    \caption{Training and validation losses of the 20 independent
+    initialisations of the ensemble replicas plotted as individual
+    curves for the 5 parton model. We plot the MAE loss for the k-factor
+    and the MSE loss for the one-loop matrix element. The scale difference
+    between the MAE and MSE losses is a consequence of the form of the loss
+    functions and is not surprising.
+    We see that the replica models all converge
+    to a similar point with the validation loss being close to the training loss
+    when training is terminated. The step feature in the training loss
+    is due to the {\ReduceLROnPlateau} callback.}
+    \label{fig:1L_loss_curves}
+\end{figure}
+
+\section{Results}\label{sec:1L_results}
+\subsection{Error distributions}
+In this section we present results for our NLO QCD k-factor
+emulator for $e^{+}e^{-}$ annihilation into up to 5-jets.
+First we show a comparison between our model described in
+Section~\ref{sec:1L_nn} which we label `antenna',
+and a `naive' model with no factorisation properties built
+into the emulator: a densely connected neural network
+with the parameters given in Table~\ref{table:parameters}
+that directly predicts the k-factor.
+For the `naive' models we train with the MAE loss on the k-factors
+with no modifications and terminate training
+based on this loss. As with the `antenna' models, we also
+ensemble 20 individual replica models for the `naive' model predictions.
+In Figure~\ref{fig:1L_delta_distribution} we compare histograms
+of $\Delta$ for all final-state multiplicities between these two models.
+It is immediately clear that building
+in the factorisation structure of the matrix elements greatly increases
+accuracy, with increasing relative improvements
+for the higher multiplicity cases.
+The `antenna' error distributions are symmetric, strongly peaked around
+the ideal value of 0, and with tails falling off rapidly. We
+see the general trend of increasing multiplicity decreases accuracy,
+however we observe that the bulk of the 5 jet final-state is
+within percent accuracy and with the lower multiplicities well
+below this.
+\begin{figure}
+    \centering
+    \includegraphics[width=\linewidth]{fame2/all_jets_ratio_comparison_plots.pdf}
+    \caption{Error distribution in terms of $\Delta$ for all multiplicities. `Antenna'
+    model is as described in this article, and `naive' model is a simple densely connected
+    neural network model without any factorisation properties built in.
+    We keep the horizontal axis scale fixed for all subplots to make it easier to compare
+    accuracy across the different multiplicities.}
+    \label{fig:1L_delta_distribution}
+\end{figure}
+
+To show that we are accurate across the entire span of the tree-level
+matrix elements, and to have a closer inspection of the tails
+of the $\Delta$ distribution, we plot a 2d histogram of $\Delta$
+against the value of $|\mathcal{M}^{(5, 0)}|^{2}$ in
+Figure~\ref{fig:1L_delta_hexplot}. The bulk of phase-space points
+are contained in the high population bins depicted in yellow,
+representing the peaked distribution seen in Figure~\ref{fig:1L_delta_distribution},
+whereas the green to purple coloured bins represent the tails of the
+low population $\Delta$ distribution. We see that the accuracy
+stays contained inside a band and does not flare out
+as the magnitude of the tree-level matrix element increases.
+This shows that we manage to fit the k-factor even in
+the infrared and collinear limits where the tree-level
+matrix element becomes large. On the right-hand side subplot we plot
+the distribution of the tree-level matrix elements where
+we can see that that even with relatively few training
+points in the tails, the emulator is still able to predict
+these regions as well as where there is more abundant data.
+\begin{figure}
+    \centering
+    \includegraphics[scale=0.65]{fame2/5j_hexplot.pdf}
+    \caption{Left: $\Delta$ error distribution plotted against the
+    tree-level matrix element. Yellow bins indicates high density regions
+    of points and purple bins indicates single points.
+    Right: marginal distribution of tree-level matrix elements.
+    This illustrates that the network is able to reproduce
+    a good approximation across all sampled phase-space.}
+    \label{fig:1L_delta_hexplot}
+\end{figure}
+
+\subsection{Renormalisation scale dependence}
+\begin{figure}
+    \centering
+    \includegraphics[width=\linewidth]{fame2/renormalisation_scale_trajectories_logscale.pdf}
+    \caption{Renormalisation scale trajectories for all multiplicities.
+    These trajectories are predictions of the k-factor with the sum
+    of antenna functions subtracted to show renormalisation scale
+    dependence in the remainder. Each trajectory is at one phase-space point
+    sampled from {\RAMBO}, with $\mu_{R}$ spanning the range
+    $[s_{\mathrm{com}}/8, 8 s_{\mathrm{com}}]$. Each trajectory
+    is composed of 1000 points.
+    The region that $\mu_{R}$ was uniformly sampled from
+    for training is indicated as a red band. The error bands on the NN
+    predictions are too small to be seen.}
+    \label{fig:1L_mu_trajectory}
+\end{figure}
+
+In Figure~\ref{fig:1L_mu_trajectory} we show that our
+emulator has learned the renormalisation scale dependence,
+independent of the antenna functions. To produce
+a trajectory we first sample a phase-space point with
+the same cuts as described in Section~\ref{sec:1L_nn}, then we
+evaluate the k-factor at this phase-space point with $\mu_{R}$
+varying from $s_{\mathrm{com}}/8$ to $8s_{\mathrm{com}}$.
+We choose to sample from a wider range than used for training
+to examine the $\mu_{R}$ extrapolation performance.
+After subtracting the sum of the antenna functions we see that the
+remainder still has a dependence on the renormalisation
+scale that is accurately captured by the neural network.
+As with the $\Delta$ distribution plots in Figure~\ref{fig:1L_delta_distribution},
+there is a slight decrease in accuracy as we increase the
+multiplicity, however, the ratios and differences are well-behaved
+throughout the entirety of the trajectories inside the
+range of training data. The only anomaly occurs when the
+trajectories cross zero, causing spikes in the ratios,
+but the difference in truth and prediction remains well-behaved
+around these regions. Outside of the training
+range we see an acceptable extrapolation, but given that
+the training range is wider than the range in which the
+renormalisation scale is normally varied for scale
+variations, we do not find this problematic.
+
+\subsection{Total cross-section predictions}
+In Figure~\ref{fig:1L_cross_section} we show that we reproduce
+the total cross-sections over an integration of 1M phase-space
+points, for all multiplicities, at three
+different values of the renormalisation scale representing the
+nominal value ($\mu_{R} = \sqrt{s_{\mathrm{com}}} = 1000$ GeV)
+and the two variations usually used to estimate scale uncertainties.
+Note that these phase-space points were
+generated independently of those in Figure~\ref{fig:1L_delta_distribution},
+and that we are integrating the tree-level and one-loop interference, not the k-factor.
+We multiply our NLO k-factor prediction with the {\MadGraph} tree-level
+matrix element to reproduce the loop-matrix element for integration.
+The neural network errors are well
+below the statistical Monte Carlo integration error.
+
+\begin{figure}
+    \includegraphics[width=\linewidth]{fame2/cs_plots.pdf}
+    \caption{Error on the total cross-section across an integration
+    of 1 million phase-space points. For each multiplicity, we
+    evaluate the matrix elements at the three values of the 
+    renormalisation scale as reported.
+    The solid lines are the absolute percentage error in the true total
+    cross-section and the neural network predicted value. The dashed
+    line represents the statistical Monte Carlo integration error
+    which falls as $1/\sqrt{N}$. The jumps in error are due to large
+    values of the matrix element being integrated.}
+    \label{fig:1L_cross_section}
+\end{figure}
+
+By neural network error we are referring to the absolute percentage
+difference to the cross-section, and not the errors due to neural
+network optimisation. For that, we examine the variations in the 20
+replica model predictions in Figure~\ref{fig:1L_5j_cs} where we plot the
+total cross-section predictions as a scatter plot. The blue band
+illustrated is one standard deviation of the 20 predictions made.
+We see that the true value of the total cross-section is within
+the one standard deviation band. Not shown in the figure is the
+Monte Carlo statistical error which as seen in Figure~\ref{fig:1L_cross_section}
+dominates the absolute error between the NN predictions and
+the true value.
+
+\begin{figure}
+    \includegraphics[width=\linewidth]{fame2/5j_cross_sections.pdf}
+    \caption{Scatter plot showing the variation in predictions of
+    the 5 jet total cross-section for the 20 replica models, where
+    the plotted cross-section is normalised by the truth value
+    provided by {\MadGraph}.
+    Total cross-sections in each subplot are evaluated at the
+    renormalisation scale quoted in the upper right.
+    The mean prediction of the 20 replica models is drawn as the horizontal
+    blue line with one standard deviation illustrated by the blue band.
+    The total cross-section as predicted by {\MadGraph} is plotted
+    as the horizontal black line.}
+    \label{fig:1L_5j_cs}
+\end{figure}
+
+Since the discrepancy between the true total
+cross-section and NN predicted total cross-section is so small,
+this strongly indicates that once we fit on the relatively small
+training set, we can predict with good confidence on many more
+phase-space points to get a prediction of the cross-section before
+reaching the same level of error as the Monte Carlo integration error.
+This can also be seen from the shape of the NN error, it is relatively
+constant once enough points have been integrated.
+
+\subsection{Model evaluation time}
+In Figure~\ref{fig:1L_speed}, we plot the evaluation time of the emulator
+for both the CPU and {\CUDA} (GPU) execution providers in the {\ONNX} runtime,
+as well as the reference time from {\MadGraph}. For the NN (GPU)
+predictions we predict on 1M phase-space points concurrently, whereas
+for the NN (CPU) predictions, we predict on one point at a time.
+The times reported are then the mean of the total evaluation times.
+We see that compared to {\MadGraph}, our NN emulator
+is faster for all multiplicities, with the advantage being largest
+for the 5 jet case, with speed gains of over four orders of magnitude
+when utilising GPU acceleration. The advantage of using the GPU is
+not only from being able to batch process the predictions, it is
+also to leverage the auto-vectorisation tools provided
+by {\TensorFlow} to accelerate the model input computations.
+
+While the evaluation time of the GPU accelerated NNs are by far
+the quickest, and would be the ideal scenario for a NN emulator to be used,
+event generation typically occurs on CPUs where phase-space
+points are evaluated one at a time. This is precisely
+why our NN (CPU) predictions were made on single phase-points and
+not over batches, to showcase what the performance would be like
+when embedded in a typical production workflow. We observe that even with
+this constraint, the NN emulator is much quicker in the higher
+multiplicity cases.
+
+\begin{figure}
+    \includegraphics[width=\linewidth]{fame2/speed_plots.pdf}
+    \caption{Left: evaluation time in milliseconds of a single phase-space point.
+    Times quoted for {\MadGraph} are averaged over 1000 random phase-space
+    points. Times quoted for NN are averaged over 1M random phase-space
+    points. Right: ratio of evaluation times to {\MadGraph}. GPU used is
+    Nvidia P100 16GB, and CPU is Intel(R) Xeon(R) Gold 6126 CPU @ 2.60GHz.}
+    \label{fig:1L_speed}
+\end{figure}
+
+One of the main bottlenecks in the NN (CPU) prediction is using
+the NN ensemble to infer on single phase-space points, this is illustrated
+by the weak scaling in multiplicity
+in the left subplot of Figure~\ref{fig:1L_speed} and is illustrated
+explicitly in Figure~\ref{fig:1L_speed_breakdown}. Since the model
+architecture is identical for all multiplicities other than the
+final output layer, there will not be much difference in cost.
+Another bottleneck is the computation of model inputs which contain
+large, complex expressions with many evaluations of logarithms and
+dilogarithms. In our $\mathtt{Python}$ implementation,
+the computation time of these inputs is comparable to the
+model inference time in the 5 jet case, as shown in Figure~\ref{fig:1L_speed_breakdown}.
+Time taken to compute model input scales with final-state partons
+because of the increase in number of antenna functions
+in the ansatz to account for the larger number of singular configurations,
+as well as a larger number of the other input variables.
+
+By moving away from a \texttt{Python} oriented workflow
+to a \texttt{C++} workflow, we anticipate the model inputs
+can be computed much more quickly, potentially reducing the
+total prediction time by a factor of 2 for the 5 jet case.
+However, the model prediction time is already highly optimised by the {\ONNX} runtime,
+so we do not expect much improvement to be possible on that front.
+
+For the NN (GPU) predictions, the model inference time is negligible
+compared to the model input computations since we take advantage
+of predicting on the entire batch of 1M phase-space points at once,
+and so once averaged across this dataset each single point takes
+an insignificant amount of time.
+The model input computations are vectorised on the GPU and so we see
+an order of magnitude reduction in time taken to calculate them compared
+to computing them on a single core of a CPU.
+
+\begin{figure}
+    \centering
+    \includegraphics[width=\linewidth]{fame2/speed_breakdown.pdf}
+    \caption{Breakdown of the total time taken to predict on a phase-space
+    point (averaged over 1M predictions) into model input
+    computations and the actual model inference for both GPU and CPU
+    model deployments. The model inference
+    portions in the NN (GPU) subplot are very small.}
+    \label{fig:1L_speed_breakdown}
+\end{figure}
+
+We note that our emulator retains good performance even
+when we go to higher multiplicity. In Ref.~\cite{Badger:2022hwf}
+the authors %demonstrated that they were able to get per-mille
+%accuracy on a $2 \rightarrow 3$ process but the accuracy decreases
+%drastically when the multiplicity was increased to $2 \rightarrow 4$.
+expressed concern about the fact that the accuracy of the surrogate model
+is decreasing with the multiplicity of the process.
+We show that by incorporating suitable physically motivated
+functions in the ansatz that the network accuracy drops off much less
+rapidly when going to higher multiplicities.
+
+\section{Conclusion}\label{sec:1L_conclusion}
+In this chapter we presented the extension of Chapter~\ref{chapter:fame1},
+where we introduced the factorisation-aware model. By adapting the ingredients
+provided to the neural network model to a more suitable set,
+we have been able to move from the emulation of tree-level matrix element to
+NLO k-factors.
+We have shown that the philosophy of incorporating
+relevant physics information into the model greatly improves accuracy
+of predictions even for matrix elements with a more complex
+divergent structure.
+
+The results presented demonstrate that predictions are at the
+percent level for the most demanding process, with accuracy in the infrared regions
+of phase-space being well-behaved. We have also shown
+that for relatively few training points, the model is able to learn
+the target function well enough to evaluate many more phase-space points
+before the accumulated error matches the statistical integration error.
+Furthermore, the uncertainty of cross-section predictions
+associated with optimisation of model parameters
+was shown to contain the truth value. This provides evidence that
+the neural network model can be used to augment existing samples
+with additional phase-space points with confidence given that the
+absolute error of the cross-section prediction is relatively constant.
+
+In addition to being accurate, we have given evidence that the model,
+although optimally deployed on a GPU, is orders of magnitude quicker
+than traditional loop providers on a single CPU. By deploying the model
+with the {\ONNX} runtime a generic interface to the neural network model
+is available in many programming languages, allowing it to be embedded
+into modern event generators which are mainly written in \texttt{C++}.
 \end{document}

### 2022-12-28 paper3.tex
initial draft of chapters
acknowledgements

--- 2022-08-30 ml.tex
@@
 \begin{document}
-\chapter{Machine learning}
-\label{chapter:ml}
-\section{Introduction to machine learning}
-\subsection{Supervised and unsupervised learning}
-\section{Neural networks}
-\subsection{Perceptron}
-\subsection{Densely connected neural networks}
-\subsection{Optimisation of weights}
-\subsection{ML applications in HEP}
+\chapter{Emulation of hadron-hadron initiated matrix elements}
+\label{chapter:fame3}
+
+\section{Motivation}
+In the previous two chapters, we detailed the emulation of $e^{+}e^{-}$
+annihilation tree-level matrix elements and NLO QCD k-factors. We
+have shown that in $e^{+}e^{-}$ annihilation processes we are able to
+construct emulators that accurately reproduce the matrix elements,
+whilst keeping the evaluation time much lower than existing matrix
+element providers.
+
+The Large Electron Positron collider (LEP) was the last high energy
+$e^{+}e^{-}$ collider to be in operation with data collection
+terminating in 2000 to prepare for the LHC. Lepton-lepton colliders
+planned for the future \cite{Faus-Golfe:2022cgm} are not scheduled
+to begin operating for at least another decade. Therefore, we turn
+our focus to researching techniques to make predictions more efficient
+for the LHC, which will be collecting data until the end of the 2030s,
+namely, adapting the factorisation-aware model for hadronic collisions.
+To showcase the extension to hadronic collisions, we consider several
+partonic channels contributing to $Z+4$ and 5 jets production,
+and $t\bar{t}+3$ and 4 jets at leading order. These high-multiplicity
+processes are of particular importance for physics at the LHC as
+they contribute to the background of BSM searches. Furthermore,
+these processes typically have very low unweighting efficiencies.
+
+In Ref.~\cite{Danziger:2021eeg} a novel two-stage unweighting
+procedure was introduced that used a lightweight neural network
+as a surrogate model for the full event weight. This procedure utilises
+the rapid evaluations of the surrogate model to first unweight a trial event,
+then correct for the mismatch by re-weighting with the full event weight,
+where this second re-weighting step has a much higher efficiency.
+The authors showed that it was possible to accelerate event unweighting
+with this procedure but found that for some processes this procedure
+was slower than the standard one-step rejection sampling
+(see Section~\ref{sec:unweighting}). In this chapter
+we explore whether a more accurate emulation of matrix elements via
+the use of the factorisation-aware neural network model would lead
+to further acceleration of event unweighting and realise an improvement
+for the cases where it did not.
+
+The layout of this chapter is as follows. In Section~\ref{sec:pp_dipoles}
+we detail the extension of the model applied in Chapters~\ref{chapter:fame1}
+and \ref{chapter:fame2} to hadronic collisions. The novel two-step unweighting
+procedure introduced in Ref.~\cite{Danziger:2021eeg} is recapped in
+Section~\ref{sec:twostep}. The application of the factorisation-aware
+model in this two-step procedure, and implementation in the {\Sherpa} event
+generator framework will be discussed in Section~\ref{sec:fame_sherpa}.
+In Section~\ref{sec:pp_results} we will showcase the emulator accuracy
+and display the acceleration in the generation of unweighted events.
+Finally, we will conclude the chapter in Section~\ref{sec:pp_conclusion}
+by summarising our findings.
+
+\section{Neural network emulator framework}\label{sec:pp_dipoles}
+\subsection{Extension of ansatz to hadronic collisions}\label{sec:pp_ansatz}
+The methodology introduced in Chapter~\ref{chapter:fame1}
+for emulating $e^{+}e^{-}$ annihilation matrix elements can be straightforwardly
+adapted to hadronic collisions by accounting for the radiation
+coming from the initial-state partons. This is achieved by extending
+the set of dipole functions in the ansatz of the colour
+and helicity summed $(n+1)$-body matrix element. The ansatz
+is a sum over all permutations of dipole functions relevant
+for a given channel, each coming with a corresponding coefficient.
+It is given by
+\begin{equation}
+    |\mathcal{M}_{n+1}|^{2} = \sum_{\{ijk\}} C_{ijk}D_{ij,k} \, ,
+    \label{eqn:pp_ansatz}
+\end{equation}
+where emitter and spectator partons, denoted by $i$ and/or $k$,
+can now be initial- and final-state partons.
+Therefore, the set of dipoles $D_{ijk}$ now includes the final-final (FF),
+final-initial (FI), initial-final (IF), and initial-initial (II) dipoles,
+as opposed to just the FF dipoles for $e^{+}e^{-}$ annihilation.
+The coefficients $C_{ijk}$, which are more well-behaved than
+the matrix element in the soft and collinear limits, are
+fitted by the neural network and can be interpreted as reduced matrix
+elements. These coefficients are the outputs of the neural network,
+which forms an approximation of the matrix element once combined with
+the appropriate dipole functions.
+
+An additional extension that we study in this chapter is the
+inclusion of massive dipoles in the ansatz of the matrix element.
+This enables the examination of pure QCD processes with massive
+particles which is of particular importance for top quark pair
+production with jets. To that end, we include the massive FF, FI,
+and IF dipoles into the ansatz in Eq.~(\ref{eqn:pp_ansatz}).
+The massive dipole functions are
+generalisations of their massless counterparts, meaning it would
+be possible to only use the massive dipole functions for simplicity.
+However, only using the massive dipoles when a dipole contains a
+massive parton saves on computational costs as the massless dipole
+functions are cheaper to evaluate.
+
+To showcase these extensions, we emulate a selection of partonic
+channels at tree-level for $Z+4$ and 5 jets, and $t\bar{t}+3$ and 4 jets
+production at the LHC. We summarise the partonic channels considered
+in Table~\ref{table:pp_partonic_channels}.
+Together, these processes allow us to examine the performance
+of the extensions detailed above.
+We note that with all massless and massive dipoles implemented
+into the modelling framework, it is in principle possible to
+take advantage of the factorisation-aware model to cover
+QCD-enhanced behaviour at tree-level.
+
+\begin{table}
+    \centering
+    \begin{tabular}{cc}
+        \toprule
+        Process & Partonic channel(s) \\
+        \midrule
+        $Z+4j$ & $gg \rightarrow e^{-}e^{+}ggd\bar{d}$ \\
+        \midrule
+        $Z+5j$ & $gg \rightarrow e^{-}e^{+}gggd\bar{d}$ \\
+        \midrule
+        $t\bar{t}+3j$ & \makecell{$u\bar{u} \rightarrow t\bar{t}gd\bar{d}$ \\ $gg \rightarrow t\bar{t}ggg$} \\
+        \midrule
+        $t\bar{t}+4j$ & $ug \rightarrow t\bar{t}gggu$ \\
+        \bottomrule
+    \end{tabular}
+    \caption{List of partonic channels considered in this chapter.}
+    \label{table:pp_partonic_channels}
+\end{table}
+
+\subsection{Generation of data}\label{sec:pp_data}
+Data is generated using the {\Sherpa} event generator
+with the {\Amegic} matrix element generator.
+Below we list the selection criteria for the phase-space sampling.
+
+\subsubsection*{Z+jets}
+For $Z$ boson production in association with jets,
+we consider the partonic channels $g g \to e^- e^+ g g d \bar{d}$
+and $g g \to e^- e^+ g g g d \bar{d}$ at leading order, representing
+tree-level contributions to $Z+4$ jets and $Z+5$ jets production
+at the LHC, respectively. The phase-space sampled for generating training
+and testing data is constrained by requiring a dilepton invariant
+mass $m_{e^- e^+} > \SI{66}{\giga\electronvolt}$ and four or
+five anti-$k_t$ jets~\cite{Cacciari:2008gp} with radius parameter
+$R = 0.4$ and $p_{T,j} > \SI{20}{\giga\electronvolt}$.
+We consider a proton-proton centre-of mass energy of
+$\sqrt{s} = \SI{13}{\tera\electronvolt}$ and use the
+NNPDF-3.0 NNLO PDF set~\cite{NNPDF:2014otw}.
+
+\subsubsection*{$t\bar{t}$+jets}
+For top quark pair production, we consider three partonic channels that contribute to
+$t\bar{t}+3$ jets and $t\bar{t}+4$ jets in proton-proton
+collisions. These are pure QCD processes with massive particles
+which pose a challenge due to the top quarks carrying colour
+charge, meaning there is a significant proliferation of Feynman
+diagrams when considering their jet-associated production.
+For the processes contributing
+to $t\bar{t}+3$ jets we require three anti-$k_t$ jets
+with $R=0.4$ and $p_{T,j} > \SI{20}{\giga\electronvolt}$.
+The phase-space of the process $u g \to t \bar{t} g g g u$
+contributing to $t\bar{t}+4$ jets is constrained by requiring
+four staggered anti-$k_t$ jets with $R=0.4$, $p_{T,1} > \SI{100}{\giga\electronvolt}$,
+$p_{T,2} > \SI{50}{\giga\electronvolt}$, $p_{T,3} > \SI{40}{\giga\electronvolt}$,
+and $p_{T,4} > \SI{20}{\giga\electronvolt}$. We do not impose
+phase-space cuts for the external top quarks. They are
+treated as on-shell in the matrix element calculation,
+$p_t^2 = p_{\bar{t}}^2 = m_t^2$ with $m_t = \SI{173.4}{\giga\electronvolt}$,
+and only decayed a posteriori.
+
+\subsection{Constructing the emulator}\label{sec:pp_emulator}
+The emulator is a dense neural network built using the {\Keras} API
+to the {\TensorFlow} backend, and is deployed using the {\ONNX} runtime.
+The network consists of four hidden
+layers, each containing 128 nodes. We use the swish activation function
+for all layers, and initialise node weights according to the Glorot uniform
+distribution. We use the data generated from the {\Sherpa} event
+generator (described above in Section~\ref{sec:pp_data})
+to fit the neural network by minimising the mean squared error.
+Training is optimised with the Adam optimiser with an initial
+learning rate of $10^{-3}$. Learning rate is reduced by a factor
+of 0.7 during training when validation loss shows no improvement
+for 30 epochs, and model training is terminated once there is no
+improvement in the validation loss after 60 epochs. We summarise
+the neural network hyperparameters in Table~\ref{table:pp_hyperparameters}.
+
+\begin{table}
+    \caption{Summary of hyperparameters for the neural network
+    employed to emulate matrix elements for all partonic channels in
+    Table~\ref{table:pp_partonic_channels}.}
+    \begin{center}
+        \begin{tabular}{ll}
+            \toprule
+            Parameter & Value \\
+            \midrule
+            Hidden layers & 4 \\
+            Nodes in hidden layers & 128 \\
+            Activation function & swish \cite{Hendrycks2016BridgingNA,DBLP:journals/corr/abs-1710-05941}\\
+            Weight initialiser & Glorot uniform \cite{pmlr-v9-glorot10a} \\
+            Loss function & MSE \\
+            Batch size & 512 \\
+            Optimiser & Adam \cite{Kingma:2014vow} \\
+            Initial learning rate & $10^{-3}$ \\
+            Callbacks & {\EarlyStopping}, {\ReduceLROnPlateau} \\
+            \bottomrule
+        \end{tabular}
+    \end{center}
+    \label{table:pp_hyperparameters}
+\end{table}
+
+As inputs to the neural network, we feed in: the 4-momenta of initial-
+and final-state particles, the phase-space mapping variables for each
+class of dipole function\footnote{The mapping variables for FF, FI, IF, II dipoles are $y_{ij,k}$ (Eq.~(\ref{eqn:y_ijk})), $x_{ij,a}$ (Eq.~(\ref{eqn:x_ija})), $x_{ik,a}$ (Eq.~(\ref{eqn:x_ika})), and $x_{i,ab}$ (Eq.~(\ref{eqn:II_mapping})), respectively.},
+and the kinematic invariants $s_{ij}$ for all pairs of particles in
+the process. We refer to all phase-space mapping variables
+as $y_{ij,k}$ for brevity, following the notation of
+Eq.~(\ref{eqn:pp_ansatz}). To aid the network in training, we preprocess
+the phase-space mapping variables in the following manner\footnote{Since we do not
+have top quarks in the initial-state, we do not mention massive IF, or II dipoles in Eq.~(\ref{eqn:pp_mapping_transform}).}
+\begin{equation}
+    y_{ij,k} \rightarrow
+    \begin{cases}
+        \log(y_{ij,k})   & \text{if massive FI dipole} \, , \\
+        \log(1-y_{ij,k}) & \text{if massless FI, IF, or II dipole} \, ,\\
+        \log(y_{ij,k})   & \text{otherwise (massless FF dipole)} \, ,
+    \end{cases}
+    \label{eqn:pp_mapping_transform}
+\end{equation}
+such that their distributions have similar shape and width. The kinematic invariants
+are also transformed with the logarithm as they can span many orders
+of magnitude. All inputs are standardised to zero mean and unit variance,
+with each component of the 4-momenta being standardised separately.
+
+The target matrix elements are preprocessed by the transformation
+\begin{equation}
+    |\mathcal{M}_{n+1}|^{2} \rightarrow \mathrm{arcsinh}\left(\dfrac{|\mathcal{M}_{n+1}|^{2}}{S_{\mathrm{pred}}}\right) \, ,
+    \label{eqn:pp_me_transform}
+\end{equation}
+where $S_{\mathrm{pred}}$ is the prediction scale taken to be the
+minimum matrix element in the training set. The raw outputs of the
+neural network, $c_{ijk}$, are transformed to the coefficients appearing
+in the ansatz via the transformation
+\begin{equation}
+    C_{ijk} = S_{\mathrm{coef}} \times \sinh(c_{ijk})
+    \label{eqn:pp_coefs_transform}
+\end{equation}
+where $S_{\mathrm{coef}}$ is the coefficient scale, defined
+as $S_{\mathrm{pred}}/S_{\mathrm{dipole}}$. The dipole scale,
+$S_{\mathrm{dipole}}$, is the representative value of a dipole,
+which we take to be the median of all dipoles in the training set.
+A schematic diagram of the neural network is given in Figure~\ref{fig:pp_nn}.
+
+\begin{figure}
+    \includegraphics[scale=0.6]{fame3/nn_diagram.pdf}
+    \caption{Schematic diagram of neural network architecture
+     showing inputs: phase-space points $p$, phase-space mapping
+     variables $y_{ij,k}$ and kinematic invariants $s_{ij}$.
+     The outputs are the fitted coefficients $C_{ijk}$.}
+    \label{fig:pp_nn}
+\end{figure}
+
+In Chapters~\ref{chapter:fame1} and \ref{chapter:fame2}, the final
+predictions made for the matrix elements were the mean of an
+ensemble of independent replica models.
+Here we train 10 replica models to monitor the convergence of training,
+however, we select only the model with lowest validation loss for
+predictions. As a demonstration of model convergence, the training
+and validation loss of 10 replica models for the
+$gg \rightarrow e^{-}e^{+}ggd\bar{d}$ channel is shown in
+Figure~\ref{fig:pp_loss_plot}. This convergent behaviour is
+also seen in the other partonic channels.
+
+\begin{figure}
+    \includegraphics[width=\linewidth]{fame3/training_loss_curves.pdf}
+    \caption{Training and validation loss for the models emulating
+    $gg \rightarrow e^{-}e^{+}ggd\bar{d}$ matrix elements. We have
+    plotted the mean of the 10 losses as the solid line, with the bands
+    indicating the standard error across these 10 models.
+    The minimum training and validation loss quoted is for the best
+    performing model.}
+    \label{fig:pp_loss_plot}
+\end{figure}
+
+The rationale behind using an ensemble of models is to dampen the
+effects of random model initialisation, to reduce the effects of
+stochasticity of the training process, and to provide an estimation
+of the uncertainty of the neural network prediction. Additionally, the ensembling
+approach has the advantage of providing, in general, a more robust
+prediction than a single model, however, that comes at the price
+of greater evaluation times. In this chapter where we intend to use the
+trained network in an application where speed is a bottleneck, we
+have observed that using a single neural network is most performant.
+This can be attributed to the fact that there is diminishing returns
+in ensembling models, as there is overlapping information from each
+replica model. On the other hand, the evaluation time grows linearly
+with the number of models in the ensemble, meaning a single
+accurate model strikes a good balance between speed and accuracy.
+Additionally, the uncertainty associated with the network predictions
+is not relevant for the application presented in this chapter as the
+network is used in an intermediary step of a larger calculation.
+
+Since we are interested in reducing evaluation time of the neural
+network, it is natural to wonder whether it is worth trading
+off even more accuracy for speed in the form of reducing the size of the
+neural network. A smaller network with fewer nodes/layers would reduce
+the size of the weight matrices used during predictions, representing an
+increase in throughput.
+However, the neural network is deployed using the highly optimised
+{\ONNX} runtime where we have observed that more compact networks
+do not reduce evaluation time by very much at all, whereas there
+is an appreciable decrease in accuracy, leading to lower efficiency
+of the second unweighting step described below.
+
+\section{Novel two-stage unweighting algorithm}\label{sec:twostep}
+Generation of unweighted events is carried out via rejection sampling,
+as outlined in Section~\ref{sec:unweighting}, where event weights
+are normalised by the maximal event weight, $w_{\mathrm{max}}$,
+before being compared to a uniform random number $r \in [0,1]$. If
+$w / w_{\mathrm{max}} < r$ then events are accepted and assigned
+a unit-weight. This procedure means that small event weights
+have a lower probability of being accepted, whereas larger event weights
+are more likely to be accepted. An issue with unweighting is that
+the number of accepted events, $N$, is generally much lower than the
+number of trial events, $N^{\mathrm{trials}}$. This is encapsulated in
+the unweighting efficiency
+\begin{equation}
+    \epsilon = \dfrac{N}{N^{\mathrm{trials}}} \approx \dfrac{\langle w \rangle_{N^{\mathrm{trials}}}}{w_{\mathrm{max}}} \; \text{for large} \; N^{\mathrm{trials}} \, ,
+    \label{eqn:pp_unweighting_efficiency}
+\end{equation}
+where $\langle w \rangle_{N^{\mathrm{trials}}}$ is the mean event weight
+over all trials. The inverse $1 / \epsilon$ gives the average number of
+trial events required to accept one unit-weight event. The rejection
+sampling algorithm depends
+on the maximal event weight, $w_{\mathrm{max}}$, which cannot usually be
+exactly determined given finite statistics. Furthermore, it is possible to
+have large outlier event weights which when taken to be $w_{\mathrm{max}}$ would
+lead to excessively low unweighting efficiencies.
+
+One method to avoid this issue is to define a reduced $w_{\mathrm{max}}$
+and accept overweights, meaning weights with $w > w_{\mathrm{max}}$
+are accepted but are assigned with a correction factor
+$\tilde{w} = w / w_{\mathrm{max}}$.
+This process leads to partially
+unweighted events and is summarised in Algorithm~\ref{alg:partial_unweighting}.
+It is the default method to generate unweighted events in the {\Sherpa}
+framework.
+A systematic approach to reducing $w_{\mathrm{max}}$
+is the per-mille quantile reduction method which defines $w_{\mathrm{max}}^{\mathrm{p.m.}}$
+such that overweights contribute at most 0.1\% to the total cross-section.
+
+\begin{algorithm}
+    \caption{Rejection sampling for generating partially unweighted events.}
+    \label{alg:partial_unweighting}
+    \While{unweighting}{
+        generate random phase-space point $\mathbf{x}$\;
+        evaluate event weight $w \gets w(\mathbf{x})$\;
+        generate uniform random number $r \gets$ Random(0, 1)\;
+        \If{$w / w_{\mathrm{max}} > r$}{
+            \KwRet $\mathbf{x}$ and $\tilde{w} \gets \max(1,\; w / w_{\mathrm{max}})$\;
+        }
+    }
+\end{algorithm}
+
+Whilst the partial unweighting approach increases overall unweighting efficiency,
+it remains low for high-multiplicity processes.
+A novel two-step unweighting procedure was introduced in Ref.~\cite{Danziger:2021eeg}
+to accelerate unweighted event generation. This procedure is formed of two separate
+unweighting steps. The first unweighting step uses
+a lightweight surrogate neural network model, which approximates the event weight,
+to rapidly unweight the trial event. Given that the surrogate model is in general imperfect, the
+second unweighting step corrects for this by calculating the exact event weight
+and applying a correction factor, meaning the final event accepted is exact and has
+unit-weight. This second step is similar to the process of dealing with overweights described
+above.
+The advantage of this approach is that the expensive exact event weight is only
+evaluated once the first step accepts a trial event. This means that the number of 
+expensive exact event weight computations is drastically reduced for a sufficiently accurate
+surrogate model, hence accelerating unweighted event generation. In Ref.~\cite{Danziger:2021eeg},
+the authors used a surrogate model to approximate the full event weight, 
+$w = J_{\mathrm{PS}} |\mathcal{M}|^{2}$. In this chapter we explore the possibility
+to emulate only the matrix element using the factorisation-aware model detailed in the
+previous section, whilst keeping the phase-space weight computation exact.
+
+The two-stage algorithm is outlined in Algorithm~\ref{alg:twostep}, where we have
+denoted the matrix element approximation as $|M|^{2}$ in line number 3. This
+two-step approach introduces an additional weight, $x_{\mathrm{max}}$, analogous
+to the maximal event weight $w_{\mathrm{max}}$. It is the maximum truth-to-prediction
+ratio. This could in principle be a large value, so the quantile reduction method
+is used to define a more practical $x_{\mathrm{max}}^{\mathrm{p.m.}}$ for more efficient
+unweighting in the second step. This will be described in more detail in Section~\ref{sec:fame_sherpa}.
+\begin{algorithm}[ht]
+    \caption{Two-stage rejection sampling algorithm employing
+    a surrogate model to approximate exact event weight.}
+    \label{alg:twostep}
+    \While{unweighting}{
+      generate phase-space point $\mathbf{x}$\;
+      calculate approximate event weight $s \gets J_{\mathrm{PS}}|M|^{2}$\;
+      generate uniform random number $r_1 \gets$ Random(0, 1)\;
+        \tcp{first unweighting step}
+        \If{$s / w_{\mathrm{max}} > r_1$}{
+          calculate exact event weight $w \gets w(\mathbf{x})$\;
+          determine ratio $x \gets w/s$\;
+          generate uniform random number $r_2 \gets$ Random(0, 1)\;
+            \tcp{second unweighting step}
+            \If{$x / x_{\mathrm{max}} > r_2$}{
+            \KwRet $\mathbf{x}$ and $\widetilde{w} \gets \max(1, s/w_\text{max})\cdot\max(1, x/x_\text{max})$
+            }
+        }
+    }
+\end{algorithm}
+
+The acceleration of this approach can be quantified by the
+effective gain factor which is given by
+\begin{align}\label{eq:gain_factor}
+	f_{\text{eff}} & = \dfrac{T_{\text{standard}}}{T_{\text{surrogate}}} \nonumber \, ,\\
+	&= \dfrac{N_{\text{full}}^{\text{trials}} \times \langle t_{\text{full}}
+	\rangle}{N_{\text{1st}}^{\text{trials}} \times \left[\langle
+			t_{\text{surr}} \rangle + \langle t_{\text{PS}}
+		\rangle\right] + N_{\text{2nd}}^{\text{trials}} \times \langle
+	t_{\text{ME}} \rangle} \, , \\
+	&= \dfrac{1}{\frac{\langle t_{\text{surr}} \rangle + \langle
+			t_{\text{PS}} \rangle}{\langle t_{\text{full}} \rangle}
+			\times
+			\frac{\epsilon_{\text{full}}}{\epsilon_{\text{1st}}
+			\epsilon_{\text{2nd}}} + \frac{\langle
+		t_{\text{ME}} \rangle}{\langle t_{\text{full}} \rangle} \times
+	\frac{\epsilon_{\text{full}}}{\epsilon_{\text{2nd}}}} \nonumber \, .
+\end{align}
+This effective gain factor can be understood as the ratio of the
+average times of the standard unweighting approach, compared to the two-stage
+unweighting approach, such that a higher $f_{\mathrm{eff}}$ is desirable.
+The time spent in the standard approach is simply the number of trial events,
+$N_{\mathrm{full}}^{\mathrm{trials}}$, multiplied by the average evaluation
+time of the exact event weight in this sample of trials, $\langle t_{\mathrm{full}} \rangle$.
+The total time of the two-stage approach can be split up into two steps.
+The first step involves the evaluation of the surrogate model
+(including all input computations, pre- and post-processing) and the exact phase-space
+weight, $\langle t_{\mathrm{surr}} \rangle$ and $\langle t_{\mathrm{PS}} \rangle$,
+for the number of trial events $N_{\mathrm{1st}}^{\mathrm{trials}}$. The second
+step involves the evaluation of the exact matrix element (using the phase-space weight
+from the previous step to recover the exact event weight), $\langle t_{\mathrm{ME}} \rangle$,
+for $N_{\mathrm{2nd}}^{\mathrm{trials}}$ second step trial events.
+From the second to the final line of Eq.~(\ref{eq:gain_factor}), we divided the numerator and denominator by
+$N_{\mathrm{full}}^{\mathrm{trials}} \times \langle t_{\mathrm{full}} \rangle$
+and defined
+\begin{align}
+    \label{eq:unweighting_efficiencies}
+    \epsilon_\text{full} &= \dfrac{N}{N^\text{trials}_\text{full}} \approx \dfrac{\langle w \rangle}{w_{\mathrm{max}}^{\mathrm{p.m.}}} \, , \nonumber \\
+    \epsilon_\text{1st} &= \dfrac{N^\text{trials}_\text{2nd}}{N^\text{trials}_\text{1st}} \approx \dfrac{\langle s \rangle}{w_{\mathrm{max}}^{\mathrm{p.m.}}} \, , \\
+    \quad \epsilon_\text{2nd} &= \dfrac{N}{N^\text{trials}_\text{2nd}} \approx \dfrac{\langle x \rangle}{x_{\mathrm{max}}^{\mathrm{p.m.}}} \, , \nonumber
+\end{align}
+where $\langle w \rangle$ is the mean of the weights generated during {\Sherpa}'s
+integration run (see Section~\ref{sec:fame_sherpa}),
+$\langle s \rangle$ is the mean of the surrogate model predictions
+on the test set, and $\langle x \rangle$ is the mean of the corresponding truth-to-prediction ratios.
+These efficiencies represent the unweighting efficiencies of the standard
+approach, and of the two steps in the two-step approach, respectively.
+From the expression of the effective gain factor, Eq.~(\ref{eq:gain_factor}),
+we would expect large gains for processes where $\epsilon_{\mathrm{full}}$ is low.
+Indeed, this is the case for the high-multiplicity processes we consider. Furthermore,
+for a fast and accurate surrogate model, where
+$\langle t_{\mathrm{surr}} \rangle \ll \langle t_{\mathrm{full}} \rangle$,
+$\epsilon_{\mathrm{1st}} \approx \epsilon_{\mathrm{full}}$, and
+$\epsilon_{\mathrm{2nd}} \approx 1$, we would also expect large gains.
+
+\subsection{Implementation in {\Sherpa}}\label{sec:fame_sherpa}
+To implement the neural network emulator described in Section~\ref{sec:pp_emulator}
+as a surrogate model for the matrix elements in the two-step algorithm
+outlined in Section~\ref{sec:twostep},
+we need to generate data for training and testing the neural network,
+and to use the model to approximate the event weight during the unweighting
+process.
+Cross-section predictions in {\Sherpa} are made in two phases.
+Firstly, there is an optimisation phase which allows the integrator to adapt
+to the partonic channel of interest \cite{Kleiss:1994qy}. This is followed by an integration
+phase in which the integrator is used to determine the total cross-section.
+During this integration phase, weighted events are generated and $w_{\mathrm{max}}^{\mathrm{p.m.}}$
+is determined with the quantile reduction method which we describe below.
+In the standard partial unweighting procedure, the reduced maximal event weight,
+$w_{\mathrm{max}}^{\mathrm{p.m.}}$, is determined by first sorting weights
+generated during the integration phase, $\{w_{i}\}$, such that $w_{i} < w_{i+1}$, then requiring
+\begin{equation}
+    w_{\mathrm{max}}^{\mathrm{p.m.}} = \min\left(w_{j} \; \Bigg| \; \sum_{i=j+1}^{N} w_{i} < 0.001 \sum_{i=1}^{N} w_{i} \right) \, ,
+    \label{eq:w_max}
+\end{equation}
+i.e. find the first weight from $\{w_{i}\}$ such that the sum of all larger weights
+contributes at most 0.1\% to the total cross-section.
+
+During the integration phase, 1.5 million events are taken for training
+and testing data by saving the 4-momenta, phase-space weights, and matrix
+elements. In this chapter we study the emulation of colour and helicity
+summed matrix elements, which are evaluated using the {\Amegic} matrix
+element provider in {\Sherpa}. From these 1.5 million events, we take 800k events for
+training, 200k events for validation during training, and use the final 500k events
+for independent testing of the model once it has been trained. The testing
+dataset is used to determine the value of $x_{\mathrm{max}}$.
+The same quantile reduction method described above can be used to define
+a reduced $x_{\mathrm{max}}$, $x_{\mathrm{max}}^{\mathrm{p.m}}$, that allows
+for more efficient unweighting in the second step of the two-step procedure
+at the expense of overweights.
+By sorting the sequence of truth-to-prediction ratios, $\{x_{i}\}$, such that
+$x_{i} < x_{i+1}$ and using the same sorting order for $\{s_{i}\}$,
+$x_{\mathrm{max}}^{\mathrm{p.m.}}$ is defined as
+\begin{equation}
+    x_{\mathrm{max}}^{\mathrm{p.m.}} = \min\left(x_{j} \; \Bigg| \; \sum_{i=j+1}^{N} x_{i}s_{i} < 0.001 \sum_{i=1}^{N} x_{i}s_{i} \right) \, .
+    \label{eq:x_max}
+\end{equation}
+
+This process is repeated for each partonic channel that we study, where each channel
+has its own neural network surrogate model.
+The {\ONNX} runtime has a simple \texttt{C++} API, meaning it is straightforward to
+incorporate the model into a \texttt{C++} workflow. The interface to the {\Sherpa} framework
+is then a case of writing wrapper code that takes the phase-space point generated from
+{\Sherpa} to compute the model input variables $y_{ijk}$ and $s_{ij}$, make the model
+predictions, $C_{ijk}$, and then combines them with the appropriate dipole
+functions to form an approximation of the matrix element. This setup can then be
+used to perform the two-stage unweighting procedure in Algorithm~\ref{alg:twostep} by calling the
+surrogate model for every $|M|^{2}$ evaluation.
+
+\section{Results}\label{sec:pp_results}
+In this section, we evaluate the performance of the neural network model by
+examining the error distributions for the predictions made with the model.
+Following this, we display results for the application in the two-stage
+unweighting procedure by quoting effective gain factors, $f_{\mathrm{eff}}$.
+\subsection{Emulator error distributions}
+In Figure~\ref{fig:pp_err_dists}, we show error distributions for the neural network predictions,
+by plotting the prediction-to-truth ratio, $w/s$, for the 500k events in the independent testing
+dataset for each channel, as indicated by the legend labels. In the left-hand subplots, we
+plot $w/s$ on a linear scale to assess the model performance
+in terms of accuracy, whereas on the right-hand subplots we plot $\log_{10}(w/s)$. This allows us to
+examine the tails of the distribution which are important for the second unweighting step,
+as a large right-handed tail skews the value of $x_{\mathrm{max}}^{\mathrm{p.m.}}$.
+All error distributions are centred around the ideal values of 1 or 0, for the left- and
+right-hand subplots, respectively, with the distributions having a Gaussian-like distribution
+that is narrow.
+
+We see that performance on the channels associated with $Z$+jets production are
+generally worse than for the top quark pair production channels. This can be attributed
+to the multiplicity of the channels, as increasing multiplicity generally leads to a drop
+in accuracy as already discussed in Chapters~\ref{chapter:fame1} and \ref{chapter:fame2}.
+At the same multiplicity, the $gg \rightarrow t\bar{t}ggg$ channel presents a more difficult channel
+to model than $u\bar{u} \rightarrow t\bar{t}gd\bar{d}$ due to the presence of more
+infrared singularities, however, the associated accuracy with this gluon initiated channel
+remains higher than for the $t\bar{t}+4j$ case, in line with the assessment above.
+
+To inspect the tails on the right-hand subplots further, we plot 2d histograms of $w/s$
+against the value of the event weights. This allows
+us to see whether the tails emerge from large weight events, or from small weight events.
+This is illustrated in Figure~\ref{fig:pp_2d_hist_z5j} for the $Z+5j$ process as this
+is the channel with the lowest accuracy, meaning the other channels are better behaved.
+We see that the model is generally well-behaved
+for large event weights as the high population bins depicted in yellow, are centred around
+the ideal value of 0, and the error distribution is generally contained inside a band.
+From this plot, we also see that outlier events depicted in purple are predominately
+smaller event weights.
+
+\begin{figure}
+    \includegraphics[width=\linewidth]{fame3/ratio_plots.pdf}
+    \caption{Error distributions for all partonic channels. All histograms
+    are produced from 500k test events. Left: linear
+    truth-to-prediction ratio distributions, right: $\log_{10}$ truth-to-prediction
+    ratio distributions. The upper two subplots illustrate
+    the $Z+4j$ and $Z+5j$ processes, with the lower two subplots illustrating
+    the $t\bar{t}+3j$ and $t\bar{t}+4j$ processes.}
+    \label{fig:pp_err_dists}
+\end{figure}
+
+\begin{figure}
+    \includegraphics[scale=0.65]{fame3/z5j_hexbin_w.pdf}
+    \caption{Left: 2d histogram of $\log_{10}(w/s)$ against the respective
+    event weight, $w$ for the $gg \rightarrow e^{-}e^{+}gggd\bar{d}$ channel.
+    Yellow bins represent high population bins, and purple bins represent
+    single points. Right: marginal distribution of $gg \rightarrow e^{-}e^{+}gggd\bar{d}$
+    event weights.}
+    \label{fig:pp_2d_hist_z5j}
+\end{figure}
+
+\subsection{Unweighting gains}
+With the model performance validated, we move on to discussing
+the acceleration of unweighted event generation by using
+the neural network emulator as the surrogate model in the two-stage
+unweighting procedure.
+
+The amount of acceleration is quantified by the effective gain factor,
+$f_{\mathrm{eff}}$, which measures the average time saved to
+unweight a sample of events compared to the standard
+one-step rejection sampling approach. To determine $f_{\mathrm{eff}}$,
+we run through the procedure described in Algorithm~\ref{alg:twostep}
+for a number of events in the testing datasets and record
+$\langle t_{\mathrm{surr}} \rangle$, $\langle t_{\mathrm{PS}} \rangle$, and
+$\langle t_{\mathrm{ME}} \rangle$.
+We note that $\langle t_{\mathrm{full}} \rangle = \langle t_{\mathrm{PS}} \rangle + \langle t_{\mathrm{ME}} \rangle$.
+Depending on the complexity
+of the process, between 10 and 10000 events are unweighted to get a reliable
+estimate of these times. Eqs.~(\ref{eq:gain_factor}) and
+(\ref{eq:unweighting_efficiencies}) are then used to calculate $f_{\mathrm{eff}}$.
+
+In Table~\ref{table:gain_factors} we quote the timings, efficiencies,
+and the computed effective gain factors for all channels,
+where we can see the speed up ranges from 14 to up to 366 times
+compared to the standard approach. This is due to the rapid
+neural network evaluation times and good predictive accuracy leading
+to relatively high $\epsilon_{\mathrm{1st}}$ and $\epsilon_{\mathrm{2nd}}$.
+Interestingly, the largest
+gains are seen for the most computationally expensive processes,
+namely, the $Z+5j$ and $t\bar{t}+4j$ processes. This demonstrates
+that the surrogate model has managed to accurately model the matrix
+elements, and to produce the approximations rapidly. We can
+see that $\langle t_{\mathrm{surr}} \rangle$ only scales weakly
+with the multiplicity, whereas $\langle t_{\mathrm{ME}} \rangle$
+increases very rapidly for more complex final-states.
+
+\begin{table}
+    \sisetup{table-auto-round}
+    \centering
+    \small
+    \begin{tabular}{@{}
+        l
+        S[table-format=1.1e1,scientific-notation]
+        S[table-format=1.1e1,scientific-notation]
+        S[table-format=1.1e1,scientific-notation]
+        S[table-format=1.1e1,scientific-notation]
+        S[table-format=1.1e1,scientific-notation]
+        S[table-format=1.2]
+        l
+        @{}}
+        \toprule
+        \multirow{2}[3]{*}{Process} & \multicolumn{3}{c}{Timings / \unit{\second}} & \multicolumn{3}{c}{Efficiencies} & \multirow{2}
+        [2]{*}{$f_{\mathrm{eff}}$} \\
+        \cmidrule(lr){2-4} \cmidrule(lr){5-7}
+            & $\langle t_{\mathrm{ME}} \rangle$ & $\langle t_{\mathrm{PS}} \rangle$ & $\langle t_{\mathrm{surr}} \rangle$ & $\epsilon_{\mathrm{full}}$ & $\epsilon_{\mathrm{1st}}$ & $\epsilon_{\mathrm{2nd}}$ & \\
+        \midrule
+        $gg \rightarrow e^{-}e^{+}ggd\bar{d}$ & 0.05394600000000000 & 0.00039500000000000 & 0.00013900000000000 & 0.0155569 & 0.0140033 & 0.374251 & 14 \\
+        $gg \rightarrow e^{-}e^{+}gggd\bar{d}$ & 16.21580000000000155 & 0.00570000000000000 & 0.00020000000000000 & 0.0007556797419000464 & 0.0008515348336151582 & 0.2904892762892554 & 269 \\
+        \midrule
+        $u\bar{u} \rightarrow t\bar{t}gd\bar{d}$ & 0.00525400000000000 & 0.00003500000000000 & 0.00013800000000000 & 0.000872182 & 0.00247838 & 0.2849 & 23 \\
+        $gg \rightarrow t\bar{t}ggg$ & 3.26229999999999976 & 0.00090000000000000 & 0.00018300000000000 & 0.00884173 & 0.01514 & 0.5 & 55 \\
+        $ug \rightarrow t\bar{t}gggu$ & 51.20000000000000284 & 0.00400000000000000 & 0.00023900000000000 & 0.0014836499653593951 & 0.0015978362349985088 & 0.5704770330576479 & 366 \\
+        \bottomrule
+    \end{tabular}
+    \caption{Unweighting performance measures for all partonic channels.}
+    \label{table:gain_factors}
+\end{table}
+
+\section{Conclusion}\label{sec:pp_conclusion}
+In this chapter we have studied the extension of the factorisation-aware
+model from electron-positron annihilation matrix elements explored in
+Chapters~\ref{chapter:fame1} and \ref{chapter:fame2} to hadronic collisions.
+The model was readily adapted by changing the ingredients in the ansatz
+of the matrix element to include the full set of massless dipoles
+to account for radiation from the initial-state partons.
+Additionally, we included the full set of massive dipole functions to
+enable the accurate modelling of matrix elements with massive partons,
+in this case top quark pair production.
+We have demonstrated that with these additional dipole functions, it is
+possible to accurately approximate the matrix elements of high-multiplicity
+hadronic processes with well-behaved predictions for large event weights.
+
+We have also showcased an application of the neural network model in
+the form of accelerating event unweighting in the {\Sherpa} framework
+by using the neural network model in a novel two-stage unweighting procedure.
+The performance of this unweighting procedure relied on the fast and accurate
+predictions of the model.
+We showed that for all partonic channels considered, the unweighting procedure
+was accelerated, with the largest gains coming from the most
+computationally expensive channels where the effective gain factors
+were well over two orders of magnitude.
 \end{document}

### 2022-12-29 intro.tex
for submission pt2

@@
 for matrix elements.
-This construction of this neural network emulator is
+The construction of this neural network emulator is
 described in detail in Chapter~\ref{chapter:fame1} for tree-level
@@
     The explanation of gravity in a QFT is an open problem and cannot currently
-    be included in the SM as any Lagrangian including gravity cannot be non-renormalised.
+    be included in the SM as any Lagrangian including gravity cannot be renormalised.
     Fortunately, the effect of gravity is considered
@@
     All terms in the SM Lagrangian are invariant under local transformations
-    of the gauge group Eq. (\ref{eqn:SM_gauge}).
+    of the gauge group in Eq. (\ref{eqn:SM_gauge}).
 
@@
     QCD is a non-Abelian gauge theory with gauge group
-    $\mathrm{SU}(N_{c} = 3)_{c}$ where the charge is named colour.
+    $\mathrm{SU}(N_{c} = 3)$ where the charge is named colour.
     The gauge and fermion part of the QCD Lagrangian is
@@
     \begin{equation}\label{eqn:covariant_deriv}
-        (D_{\mu})_{ij} = \delta_{ij}\partial_{\mu} - ig_{s}T^{a}_{ij}A^{a}_{\mu} \, ,
+        (D_{\mu})_{ij} = \delta_{ij}\partial_{\mu} - \mathrm{i}g_{s}T^{a}_{ij}A^{a}_{\mu} \, ,
     \end{equation}
@@
     loop or leg. In general, we have terms
-    N$^{\mathrm{k}}$LO where each additional power of $\alpha_{s}$
+    N$^{k}$LO where each additional power of $\alpha_{s}$
     increases complexity of the computations, however,

### 2022-12-29 main.tex
for submission pt2

@@
 %
-	The work carried out during this journey would not have been
-	possible without all the help and support from the people around me.
+	The work carried out during this thesis would not have been
+	possible without all the support from the people around me.
 	Firstly, I would like to express my gratitude to Michael Spannowsky, and
@@
 	doctorate when nobody else did.
-	Daniel has always been supportive with his meticulous advice and constructive criticisms,
+	Daniel has always been supportive with his meticulous advice and guidance,
 	and I thank him for giving me the freedom to explore different avenues during our joint research projects.
 
-	All the people at the IPPP, especially Trudy and Joanne, have fostered a comfortable working
+	Everyone at the IPPP, especially Trudy and Joanne, has fostered a comfortable working
 	and social environment of which I was very glad to be a part of.
-	To all the students who have come and gone during my time at the IPPP, it was a pleasure
+	To all the students who have come and gone, it was a pleasure
 	to get to know you all. Special mentions go to Alan, Asli, Dorian, Francesco, Joe, Kevin, Marian,
-	Ryan, and of course all the wonderful people in OC215 and beyond.
+	Ryan, and of course all the wonderful people in OC215 and beyond. I would also like to thank
+	the proofreaders: Dan, Francesco, Guillaume, Hitham, Jack S, Joe AB, Oscar, Peter, Ryan, and Tommy.
 	Outside of the particles, shout-out to Aidan, Cameron, Ed, Jack, and Vicky for being great friends

### 2022-12-29 mc.tex
for submission pt2

@@
     An alternative to improving the unweighting efficiency
-    itself is to accelerate the process of unweighting itself. This
+    is to accelerate the process of unweighting itself. This
     was explored in Ref. \cite{Danziger:2021eeg} wherein a
@@
     forms an estimate of the uncertainty in the perturbatively
-    computed cross-section, when compared to the true cross-section.
+    computed cross-section, when compared to the central cross-section value.
     This can be written as

### 2022-12-29 ml.tex
for submission pt2

@@
     This presents a practical challenge for data
-    collection where machine learning has could provide a solution.
-    Indeed, boosted decision trees have been use
+    collection where machine learning could provide a solution.
+    Indeed, boosted decision trees have been used
     to enhance triggers \cite{CMS:2020cmk} to accept or
@@
     and Badger \cite{Badger:2020uow} where the authors emulated tree-level
-    matrix elements and NLO QCD K-factors for $e^{+}e^{-} \rightarrow q \bar{q}$ +
+    matrix elements and NLO QCD k-factors for $e^{+}e^{-} \rightarrow q \bar{q}$ +
     up to 3 gluons. With a higher final-state multiplicity, controlling

### 2022-12-29 qcd.tex
for submission pt2

@@
     built up from the Feynman rules of a theory (c.f. QCD Feynman rules
-    Figure \ref{fig:qcd_feynman_rules}). In evaluating certain
+    in Figure \ref{fig:qcd_feynman_rules}). In evaluating certain
     topologies of Feynman diagrams, integral expressions
@@
     schemes for IR divergences. Both of these
-    methods will be discussed in this section.
+    methods will be discussed in this chapter.
 
@@
     antenna functions are then applied in the context of NLO QCD
-    K-factor emulation for electron-positron annihilation in Chapter~\ref{chapter:fame2}.
+    k-factor emulation for electron-positron annihilation in Chapter~\ref{chapter:fame2}.
 
@@
 
-        A key feature of these kernels are that they smoothly interpolate
+        A key feature of these kernels is that they smoothly interpolate
         between the soft and collinear limits, meaning they do not double count
@@
         extends to massive quarks \cite{Catani:2000ef}.
-        In Reference \cite{Catani:2002hc} the authors derive dipole
+        In Ref. \cite{Catani:2002hc} the authors derive dipole
         functions for massive partons which
@@
     sub-leading colour terms amounts to an
-    approximately 10\% effect. (cite here)
+    approximately 10\% effect.
 

### 2023-03-13 intro.tex
improve ghost description

@@
     where $c^{a}$ ($\bar{c}^{a}$) are Faddeev-Popov ghosts (antighosts).
-    The ghosts are unphysical particles and so cannot appear as
-    external states. However, they must be included in internal lines for
-    a gauge invariant theory as the ghost loops cancel the unphysical
-    degrees of freedom in gluon loops.
+    Ghosts are particles that subtract the longitudinal
+    degree of freedom from gluons, and as such it is usual to
+    use them to cancel the longitudinal degree of freedom of gluon propagators.
+    It should be noted that ghosts could in principle be included as external states as well
+    for the same purpose, however, it is more conventional to use the usual physical polarisation sum
+    instead.
 

### 2023-03-13 qcd.tex
clearer confinement description

@@
 
-    Another consequence of the running coupling
-    is colour confinement, in which quarks and
-    gluons cannot exist as free particles. Due
-    to the increasing strength of coupling at low
-    energies, quarks and gluons are forced to
-    form composite, colourless particles. This means
-    that at low energies, perturbative QCD
+    At low energies, quarks and gluons cannot exist as free
+    particles and are forced to form composite, colourless
+    particles. This phenomenon, named confinement, is compatible
+    with the observation of the running coupling, but the
+    perturbative calculation predicting the running coupling
+    does not necessarily lead to the non-perturbative phenomenon
+    of confinement.
+    This means that at low energies, perturbative QCD
     is not an accurate description of nature,

### 2023-03-13 qcd.tex
include "inclusive" for KLN

@@
     (KLN) theorems \cite{Kinoshita:1962ur,Lee:1964is} which states
-    that at each order of perturbation theory, the divergent parts of the real and virtual
+    that for sufficiently inclusive quantities, at each order of
+    perturbation theory, the divergent parts of the real and virtual
     contributions exactly cancel out, leaving only finite corrections.

### 2023-03-13 paper1.tex
footnote about alpha_s scale

@@
 All external legs have been considered to be massless. 
-The strong coupling constant has been set to $\alpha_{s} = 0.118$, and the electromagnetic coupling constant has been set to $\alpha_{e} = 1.0 / 132.5070$.
+The strong coupling constant has been set to $\alpha_{s} = 0.118$\footnote{This choice of
+the coupling is only valid for $\mu_{R}^{2} = m_{Z}^{2}$ and ideally,
+$\alpha_{s}$ should have been evaluated at the scale $\mu_{R} = 1000$ GeV. This change would result
+in slightly different plots, but does not affect the validity of the key findings
+of this chapter.}, and the electromagnetic coupling constant has been set to $\alpha_{e} = 1.0 / 132.5070$.
 The mass of the $Z$-boson is taken to be $m_{Z} = 91.188$ GeV.

### 2023-03-13 main.tex
fix abstract

@@
 	structures in regions of phase-space where particles become soft or collinear,
-	however, the behaviour of the matrix element in these limits are well-understood.
+	however, the behaviour of the matrix element in these limits is well-understood.
 	By exploiting the factorisation property of matrix elements in these limits,

### 2023-03-13 intro.tex
gravity negligible at colliders
P_{ba} -> P_{ab}

@@
     Fortunately, the effect of gravity is considered
-    to be negligible on the scales considered in the SM, and so it is ignored.
+    to be negligible on the scales considered at high energy colliders, and so it is ignored.
     Each of the three fundamental forces described by the SM is mediated
@@
     \end{equation}
-    where $P_{ba}(\frac{x}{z})$ are splitting functions
+    where $P_{ab}(\frac{x}{z})$ are splitting functions
     representing parton $b$ emitting a parton $a$, that carries

### 2023-03-13 qcd.tex
wording for perturbative treatment

@@
     asymptotic freedom \cite{Gross:1973id,Politzer:1973fx},
-    and is justification for treating QCD perturbatively
+    and is the justification for treating QCD perturbatively
     when the energy scale is high, such as at collider experiments.

### 2023-03-13 qcd.tex
improve introduction of ijk indices
k_perp typos

@@
     between the singular function and the reduced matrix element.
-    This singular factor
-    is not unique and can be represented by different
-    approximations as long as they reproduce the
+    The indices of $\bm{S}_{ijk}$ already hints at the fact
+    that the singular functions only depends on three partons,
+    and not all of the external states.
+    This singular factor is not unique and can be represented by
+    different approximations as long as they reproduce the
     correct IR behaviour. In this thesis we discuss
@@
     in Section~\ref{sec:CS_dipoles} and antenna functions \cite{Gehrmann-DeRidder:2005btv}
-    in Section~\ref{sec:antenna_functions}. The indices
-    of $\bm{S}_{ijk}$ already hints at the fact that the singular
-    functions only depends on three partons, and not all
-    of the external states.
+    in Section~\ref{sec:antenna_functions}. The indices $ijk$
+    will be given specific assignments in these two different
+    approximations.
 
@@
 
-    The collinear limit can be then be defined as
+    The collinear limit can then be defined as
     \begin{equation}\label{eqn:F_collinear limit}
@@
     \end{equation}
-    where terms less singular than $k_{\perp}^{2}$ have been neglected.
+    where terms less singular than $1 / k_{\perp}^{2}$ have been neglected.
     This reduced matrix element is obtained by replacing the partons

### 2023-03-13 qcd.tex
rework superscript (0) and (1)

@@
     functions and the factorisation formulae are modified.
+    The structure of these modified formulae are of the form:
+    tree-level splitting function multiplied by a one-loop
+    amplitude, plus a one-loop splitting function multiplied
+    by a tree-level amplitude. The loop level of
+    amplitude and splitting functions are denoted by superscripts,
+    $^{(0)}$ for tree-level and $^{(1)}$ for one-loop level.
 
@@
     \begin{equation}\label{eqn:1L_soft_factorisation}
-        M_{n+1}^{(1)}(\ldots, i, j, k, \ldots) \rightarrow S_{ijk}^{(0)} \, M_{n}^{1}(\ldots, i, k, \ldots) + S_{ijk}^{(1)}(\epsilon) \, M_{n}^{0}(\ldots, i, k, \ldots) \, ,
+        M_{n+1}^{(1)}(\ldots, i, j, k, \ldots) \rightarrow S_{ijk}^{(0)} \, M_{n}^{(1)}(\ldots, i, k, \ldots) + S_{ijk}^{(1)}(\epsilon) \, M_{n}^{(0)}(\ldots, i, k, \ldots) \, ,
     \end{equation}
@@
     are the one-loop splitting functions \cite{Bern:1999ry}.
-
-    In both these formulae the structure is of the form:
-    tree-level splitting function multiplied by a one-loop
-    amplitude, plus a one-loop splitting function multiplied
-    by a tree-level amplitude.
 

### 2023-03-13 qcd.tex
missing + signs in dipole formulae
improve spacing

@@
                 &\sum_{i,j}\sum_{k\neq i,j} \mathcal{D}_{ij,k}(p_{1},\ldots,p_{n+1};p_{a},\ldots) \\
-                &\sum_{i,j}\sum_{a} \mathcal{D}_{ij}^{a}(p_{1},\ldots,p_{n+1};p_{a},\ldots) + \ldots \, ,
+                + &\sum_{i,j}\sum_{a} \mathcal{D}_{ij}^{a}(p_{1},\ldots,p_{n+1};p_{a},\ldots) + \ldots \, ,
             \end{split}
@@
         \end{equation}
-
         The spin-averaged splitting functions are
@@
         is an additional dipole contribution compared to the initial-final case.
+
         The dipole factorisation is given by
@@
                 &\sum_{a,i}\sum_{k \neq i} \mathcal{D}_{k}^{ai}(p_{1}, \ldots, p_{n+1};p_{a},p_{b}) \\
-                &\sum_{a,i}\sum_{b \neq a} \mathcal{D}^{ai,b}(p_{1}, \ldots, p_{n+1};p_{a},p_{b}) + \ldots \, ,
+                + &\sum_{a,i}\sum_{b \neq a} \mathcal{D}^{ai,b}(p_{1}, \ldots, p_{n+1};p_{a},p_{b}) + \ldots \, ,
             \end{split}
@@
             &\sum_{i,j}\sum_{k \neq i, j} \mathcal{D}_{ij,k}(p_{1},\ldots,p_{n+1};p_{a},p_{b}) + \sum_{i,j}\sum_{a} \mathcal{D}_{ij}^{a}(p_{1},\ldots,p_{n+1};p_{a},p_{b}) \\
-            &\sum_{a,i}\sum_{k \neq i} \mathcal{D}_{k}^{ai}(p_{1},\ldots,p_{n+1};p_{a},p_{b}) + \sum_{a,i}\sum_{b \neq a} \mathcal{D}^{ai,b}(p_{1},\ldots,p_{n+1};p_{a},p_{b}) + \ldots \, ,
+            + &\sum_{a,i}\sum_{k \neq i} \mathcal{D}_{k}^{ai}(p_{1},\ldots,p_{n+1};p_{a},p_{b}) + \sum_{a,i}\sum_{b \neq a} \mathcal{D}^{ai,b}(p_{1},\ldots,p_{n+1};p_{a},p_{b}) + \ldots \, ,
             \end{split}

### 2023-03-13 mc.tex
briefly describe FOAM algorithm

@@
     Unweighting efficiencies for high-multiplicity processes
-    are often well below 1\% \cite{Hoche:2019flt,Gao:2020zvv} and the development
-    of methods to improve the efficiency are an active area of research
-    \cite{Jadach:1999sf,Jadach:2002kn}.
+    are often well below 1\% \cite{Hoche:2019flt,Gao:2020zvv}.
+    The development of methods to improve the efficiency are an active
+    area of research \cite{Jadach:1999sf,Jadach:2002kn}, wherein the
+    authors explore an algorithm that adaptively splits up the integration
+    region in order to systematically reduce the ratio of
+    the maximum event weight to the average event weight. This leads to
+    increased unweighting efficiency when carrying out the rejection sampling
+    algorithm.
     These low efficiencies lead to most of the

### 2023-03-13 paper2.tex
add process in figure caption

@@
     \caption{Scatter plot showing the variation in predictions of
-    the 5 jet total cross-section for the 20 replica models, where
+    the $e^{+}e^{-} \rightarrow q\bar{q}ggg$ total cross-section for the 20 replica models, where
     the plotted cross-section is normalised by the truth value

### 2023-03-13 conclusion.tex
change madloop to madgraph

@@
 could be predicted with accuracy at the 1\% level, representing a speed up in
-evaluation of over 3 orders of magnitude compared to {\MadLoop} on
+evaluation of over 3 orders of magnitude compared to {\MadGraph} on
 a single CPU core.

### 2023-03-13 fame1_appendex.tex
fix fame1 appendix

--- 2022-12-28 fame_appendex1.tex
@@
 \label{appendix:fame1}
+
+\section{Phase-space trajectories}\label{appendix:fame1_trajectory_generation}
+We can see the {\RAMBO} algorithm as a map from the unit hypercube in some high dimension into the $n$-particle phase-space.
+A flat distribution in the hypercube maps to a flat distribution in the multi-particle phase-space.
+To generate our phase-space trajectories we pick two points at random in the unit hypercube and map the line between them using the {\RAMBO} mapping.
+As a result, each point on the resulting trajectory has equal probability density in phase-space.
+Any other phase-space generator that smoothly maps the unit hypercube to a multi-particle phase-space could replace {\RAMBO} in this procedure and could lead to trajectories with very different characteristics.
+One could for example imagine a
+sophisticated algorithm that only maps points close to the boundary of the hypercube to soft or collinear configurations.
+Using such an algorithm would have trajectories that avoid configurations with many particles more collinear or soft than
+the end-points of the trajectories.
+We find that with {\RAMBO} the trajectories tend not to avoid difficult phase-space configurations and are therefore a good test of the extrapolation properties of our method.
+Figure~\ref{fig:detectorpath} shows the trajectory we chose in Section~\ref{sec:trajectories}.
+
+\begin{figure}
+    \centering
+    \includegraphics[width=0.45\textwidth]{fame1/yphi.png}
+    \includegraphics[width=0.45\textwidth]{fame1/E.png}
+    \caption{Left: rapidity and azimuthal angle trajectories for the five final state particles in the trajectories used in Section~\ref{sec:trajectories}.
+    Right: Evolution of the same particle energies as a function of the position along the segment between the two random points in the unit hypercube.
+    }
+    \label{fig:detectorpath}
+\end{figure}
 
@@
 
-\section{Phase-space trajectories}\label{appendix:fame1_trajectory_generation}
-We can see the {\RAMBO} algorithm as a map from the unit hypercube in some high dimension into the $n$-particle phase-space.
-A flat distribution in the hypercube maps to a flat distribution in the multi-particle phase-space.
-To generate our phase-space trajectories we pick two points at random in the unit hypercube and map the line between them using the {\RAMBO} mapping.
-As a result, each point on the resulting trajectory has equal probability density in phase-space.
-Any other phase-space generator that smoothly maps the unit hypercube to a multi-particle phase-space could replace {\RAMBO} in this procedure and could lead to trajectories with very different characteristics.
-One could for example imagine a
-sophisticated algorithm that only maps points close to the boundary of the hypercube to soft or collinear configurations.
-Using such an algorithm would have trajectories that avoid configurations with many particles more collinear or soft than
-the end-points of the trajectories.
-We find that with {\RAMBO} the trajectories tend not to avoid difficult phase-space configurations and are therefore a good test of the extrapolation properties of our method.
-
-Figure~\ref{fig:detectorpath} shows the trajectory we chose in Section~\ref{sec:trajectories}.
-
-\begin{figure}
-    \centering
-    \includegraphics[width=0.45\textwidth]{fame1/yphi.png}
-    \includegraphics[width=0.45\textwidth]{fame1/E.png}
-    \caption{Left: rapidity and azimuthal angle trajectories for the five final state particles in the trajectories used in Section~\ref{sec:trajectories}.
-    Right: Evolution of the same particle energies as a function of the position along the segment between the two random points in the unit hypercube.
-    }
-    \label{fig:detectorpath}
-\end{figure}
-
+\section{Additional random phase-space trajectory}\label{appendix:fame1_trajectory2}
 Here we present another random phase-space trajectory similar to the one
@@
 figure are not unique.
-\begin{figure}
+\begin{figure}[ht]
     \includegraphics[width=\linewidth]{fame1/trajectory_1255.pdf}
@@
 \end{figure}
+
 \end{document}

### 2023-03-13 paper1.tex
column wording
fix fame1 appendix

@@
 In Figure~\ref{fig:random_trajectory} we show the predictions by the three 5-jet models trained on data with different global phase-space cuts for this trajectory.
-Left column shows the actual matrix element prediction and right column shows the ratios of the prediction to {\NJet}.
+The left column shows the actual matrix element prediction and the right column shows the ratios of the prediction to {\NJet}.
 We analysed the predictions for the 50 random trajectories and measured the fraction of their length where the accuracy falls within given intervals.
@@
 Although not seen in Figure~\ref{fig:random_trajectory}, there are points which have both a double unresolved limit and a separate single unresolved limit,
-we label this configuration `Double+single' and indicate is as a blue region\footnote{Another random phase-space trajectory is presented in Appendix~\ref{appendix:fame1_trajectory_generation} which contains these `double+single' regions.}.
+we label this configuration `Double+single' and indicate is as a blue region\footnote{Another random phase-space trajectory is presented in Appendix~\ref{appendix:fame1_trajectory2} which contains these `double+single' regions.}.
 We do not include the quark-anti-quark invariant in defining these regions as there is no associated infrared divergence.

### 2023-03-13 mc.tex
chapter 3 typo
extra bit about scale variation

@@
     NNLO and N$^{3}$LO QCD corrections have been computed
-    for select processes but have yet to be reach the
+    for select processes but have yet to reach the
     level of automation of NLO.
@@
     is approximately the same order of magnitude as the true missing
-    higher order terms.
+    higher order terms. This is true if the size of the perturbative
+    coefficients remains similar, which would be violated if, for example,
+    a new important channel opens up by increasing the perturbative
+    order.
 

### 2023-03-13 paper2.tex
chapter 6 typo
mu_{R} range typo

@@
 
-This informs our ansatz for the k-factor, which can is given as
+This informs our ansatz for the k-factor, which is given as
 \begin{equation}
@@
 renormalisation scale, $\log(\mu_{R})$, from a uniform distribution
-with end points at $[\log(s_{\mathrm{com}}/4), \log(4s_{\mathrm{com}})]$.
+with end points at [$\log(\sqrt{s_{\text{com}}}/4)$, $\log(4\sqrt{s_{\text{com}}})$].
 In other words, we sample the renormalisation scale logarithmically.

### 2023-03-13 paper1.tex
chapter 5 parameters

@@
 The mass of the $Z$-boson is taken to be $m_{Z} = 91.188$ GeV.
-These parameters and those not listed, are consistent with those in the Standard Model mode of \aNLO \cite{Alwall:2014hca}.
+Below we quote the full set of parameters entering the computer code
+used to evaluate matrix elements. We do not quote the renormalisation
+scale as it does not affect the value of tree-level matrix elements.
+\begin{table}[ht]
+    \centering
+    \begin{tabular}{ll}
+        \toprule
+        Parameters & \\
+        \midrule
+        Masses & \makecell[l]{$m_{b} = 4.7$ GeV \\ $m_{t} = 173$ GeV \\ $m_{\tau} = 1.777$ GeV \\ $m_{Z} = 91.188$ GeV \\ $m_{W} = 80.419$ GeV \\ $m_{H} = 125$ GeV} \\
+        \midrule
+        Couplings & \makecell[l]{$\alpha_{s} = 0.118$ \\ $\alpha_{e} = 1/132.507$ \\ $G_{F} = 1.166390 \times 10 ^{-5}$ GeV$^{-2}$} \\
+        \midrule
+        Decay widths & \makecell[l]{$\Gamma_{t} = 1.4915$ GeV \\ $\Gamma_{Z} = 2.441404$ GeV \\ $\Gamma_{W} = 2.0467$ GeV \\ $\Gamma_{H} = 6.382339 \times 10^{-3}$ GeV} \\
+        \bottomrule
+        \end{tabular}
+    \caption{List of parameters for the matrix element generation. Masses and decay widths not listed explicitly have been set to zero.}
+    \label{table:fame1_params}
+\end{table}
 

### 2023-03-13 paper2.tex
chapter 6 parameters

@@
 \cite{NNPDF:2021njg} with the LHAPDF6 interface \cite{Buckley:2014ana}.
-All external particles are treated as massless and $m_{Z} = 91.188$ GeV.
+The full set of parameters used to evaluate the matrix elements are
+quoted in Table~\ref{table:fame2_params}.
+\begin{table}[ht]
+    \centering
+    \begin{tabular}{ll}
+        \toprule
+        Parameters & \\
+        \midrule
+        Scales & \makecell[l]{$\mu_{R}$ sampled logarithmically in range [$\log(\sqrt{s_{\text{com}}}/4)$, $\log(4\sqrt{s_{\text{com}}})$]} \\
+        \midrule
+        Masses & \makecell[l]{$m_{b} = 4.7$ GeV \\ $m_{t} = 173$ GeV \\ $m_{\tau} = 1.777$ GeV \\ $m_{Z} = 91.188$ GeV \\ $m_{W} = 80.419$ GeV \\ $m_{H} = 125$ GeV} \\
+        \midrule
+        Couplings & \makecell[l]{$\alpha_{s} = \alpha_{s}(\mu_{R}^{2})$ \\ $\alpha_{e} = 1/132.507$ \\ $G_{F} = 1.166390 \times 10 ^{-5}$ GeV$^{-2}$} \\
+        \midrule
+        Decay widths & \makecell[l]{$\Gamma_{t} = 1.4915$ GeV \\ $\Gamma_{Z} = 2.441404$ GeV \\ $\Gamma_{W} = 2.0467$ GeV \\ $\Gamma_{H} = 6.382339 \times 10^{-3}$ GeV} \\
+        \bottomrule
+        \end{tabular}
+    \caption{List of parameters for the matrix element generation. Masses and decay widths not listed explicitly have been set to zero.}
+    \label{table:fame2_params}
+\end{table}
 

### 2023-03-13 paper3.tex
chapter 7 parameters

@@
 with the {\Amegic} matrix element generator.
-Below we list the selection criteria for the phase-space sampling.
+Below we list the selection criteria for the phase-space sampling and
+document the parameters used for matrix element evaluation in Table~\ref{table:fame3_params}.
 
@@
 and only decayed a posteriori.
+
+\begin{table}[t]
+    \centering
+    \begin{tabular}{ll}
+        \toprule
+        Parameters & \\
+        \midrule
+        Scales & \makecell[l]{$\mu_{R} = \mu_{F} = 91.188$ GeV for $Z+$jets \\ $\mu_{R} = \mu_{F} = 175$ GeV for $t\bar{t}+$jets} \\
+        \midrule
+        Masses & \makecell[l]{$m_{d} = 0.01$ GeV \\ $m_{u} = 0.005$ GeV \\ $m_{s} = 0.2$ GeV \\ $m_{c} = 1.42$ GeV \\ $m_{b} = 4.8$ GeV \\ $m_{t} = 173.21$ GeV \\ $m_{e} = 0.511$ MeV \\ $m_{\mu} = 0.105$ GeV \\ $m_{\tau} = 1.777$ GeV \\ $m_{Z} = 91.1876$ GeV \\ $m_{W} = 80.385$ GeV \\ $m_{H} = 125$ GeV} \\
+        \midrule
+        Couplings & \makecell[l]{$\alpha_{s}$ set by PDF \\ $\alpha_{e} = 1/137.03599976$} \\
+        \midrule
+        Decay widths & \makecell[l]{$\Gamma_{t} = 2$ GeV \\ $\Gamma_{\tau} = 2.26735 \times 10^{-12}$ GeV \\ $\Gamma_{Z} = 2.4952$ GeV \\ $\Gamma_{W} = 2.085$ GeV \\ $\Gamma_{H} = 4.07 \times 10^{-3}$ GeV} \\
+        \bottomrule
+        \end{tabular}
+    \caption{List of parameters for the matrix element generation. Masses and decay widths not listed explicitly have been set to zero.}
+    \label{table:fame3_params}
+\end{table}
 

### 2023-03-13 paper2.tex
change parameter table captions
footnote for finite part

@@
 matrix element. The numerator in Eq. (\ref{eq:k_factor}) is the
-finite part of the interference between the tree-level and 
+finite part\footnote{See Section 3.2 and Appendix A.1 in Ref.~\cite{Hirschi:2011pa} for explicit definitions.}
+of the interference between the tree-level and 
 one-loop level matrix element, where the conventional dimensional
@@
         \end{tabular}
-    \caption{List of parameters for the matrix element generation. Masses and decay widths not listed explicitly have been set to zero.}
+    \caption{List of parameters used for matrix element generation for all processes.
+$\alpha_{s}$ is evaluated with the NNPDF-4.0 NNLO PDF set.
+Masses and decay widths not listed explicitly have been set to zero.}
     \label{table:fame2_params}

### 2023-03-13 paper2.tex
fix mu_R typos

@@
     sampled from {\RAMBO}, with $\mu_{R}$ spanning the range
-    $[s_{\mathrm{com}}/8, 8 s_{\mathrm{com}}]$. Each trajectory
+    $[\sqrt{s_{\mathrm{com}}}/8, 8 \sqrt{s_{\mathrm{com}}}]$. Each trajectory
     is composed of 1000 points.
@@
 evaluate the k-factor at this phase-space point with $\mu_{R}$
-varying from $s_{\mathrm{com}}/8$ to $8s_{\mathrm{com}}$.
+varying from $\sqrt{s_{\mathrm{com}}}/8$ to $8\sqrt{s_{\mathrm{com}}}$.
 We choose to sample from a wider range than used for training
